### 2023-07-19 code/analysis-pilot.Rmd
first commit

---
title: "analysis-pilot"
output: html_document
date: "2023-07-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-07-19 code/analysis-pilot.Rmd
data cleaning initial

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(here)
+library(jsonlite)
+
+ParseJSONColumn <- function(x) {
+  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
+    fromJSON(flatten = T)
+}
+
+pilot_start=lubridate::ymd('2023-7-07') 
+pilot_end=lubridate::ymd('2023-7-11')
 ```
 
-## R Markdown
+```{r}
+raw_data <- read.csv(here("data/rounds7.19.csv"), header = TRUE) %>% 
+  filter(createdAt >= pilot_start) %>%
+  filter(createdAt <= pilot_end) %>% 
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
+  select(-chat) |> 
+  pivot_longer(cols=starts_with('player'), names_to=c("playerId", "info"), names_prefix="player_", names_sep="_",
+               values_to="values", values_transform=as.character, values_drop_na=T) %>% 
+  filter(values!="") %>% 
+  filter(playerId!=speaker) %>% 
+  pivot_wider(names_from=info, values_from=values)
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+  write_csv(here(data_location, 'raw_results.csv'))
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-07-21 code/analysis-pilot.Rmd
data cleaning 14:55

@@
 
-  write_csv(here(data_location, 'raw_results.csv'))
+# write_csv(here(data, 'rounds7.19.csv'))
 ```
+
+# remove columns: numPlayers, sound, activePlayerCount, trialNum, distractor, tangramURLs, response, target, submitted, X_id, repNum, correct, targetNum, stageIds
+```{r}
+filtered_data <- raw_data %>% 
+  select(-numPlayers) %>% 
+  select(-sound) %>%  
+  select(-activePlayerCount) %>% 
+  select(-trialNum) %>% 
+  select(-distractor) %>% 
+  select(-tangramURLs) %>% 
+  select(-response) %>% 
+  select(-target) %>% 
+  select(-submitted) %>% 
+  select(-X_id) %>% 
+  select(-repNum) %>% 
+  select(-correct) %>% 
+  select(-targetNum) %>% 
+  select(-stageIds)
+
+```
+
+#remove NA cells in "time"
+```{r}
+filtered_data <- filtered_data %>% 
+  drop_na(time)
+
+View(filtered_data)
+```
+
+#relocate columns
+```{r}
+filtered_data <- filtered_data %>% 
+  select(createdAt, everything()) %>% 
+  relocate(countCorrect, .before = time)
+
+View(filtered_data)
+```
+
+#rename columns
+```{r}
+colnames(filtered_data)[1] = "Date" %>% 
+colnames(filtered_data)[]
+
+View(filtered_data)
+```

### 2023-07-25 code/analysis-pilot.Rmd
accuracy 14:19

@@
 
+#accuracy
+```{r}
+accuracy <- filtered_data %>% 
+  group_by(gameId) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>% 
+  mutate(fraction = correct/total*100)
+
+View(accuracy)
+```
+
+#graph accuracy to time
+#graph accuracy
+
 #rename columns
@@
 colnames(filtered_data)[1] = "Date" %>% 
-colnames(filtered_data)[]
+#colnames(filtered_data)[]
 

### 2023-07-26 code/analysis-pilot.Rmd
accuracy bar plot 17:26

--- 2023-07-21 code/analysis-pilot.Rmd
@@
 library(jsonlite)
+library(ggplot2)
 
@@
 
-#rename columns
+#accuracy graph
 ```{r}
-colnames(filtered_data)[1] = "Date" %>% 
-colnames(filtered_data)[]
+accuracy <- filtered_data %>% 
+  group_by(gameId) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>% 
+  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
 
-View(filtered_data)
+ggplot(data = accuracy, aes(y = gameId, x = percent)) +
+  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#228B22", width = .7) +
+  geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  theme_classic() +
+  xlab("Percentage Accuracy") +
+  ylab("Game ID") +
+  theme(aspect.ratio = 4/5,
+        axis.title = element_text(size=14,face="bold"),
+        axis.text = element_text(size=12),
+        legend.position = "none")
 ```
+
+time=time
+
+#graph accuracy to time (scatterplot maybe?)
+#graph accuracy (bar graph)
+
+```{r}
+ggplot(data=accuracy, aes(x = Age.months, y=Laughs, color=Experimenter)) +
+  geom_point()
+```
+```

### 2023-07-31 code/analysis-pilot.Rmd
add clean data, wrangle time partway

@@
 
-pilot_start=lubridate::ymd('2023-7-07') 
-pilot_end=lubridate::ymd('2023-7-11')
+pilot_start=lubridate::ymd('2023-07-07') 
+pilot_end=lubridate::ymd('2023-07-12')
+
+data_source <- "data/rounds.csv"
+player_source <- "data/players.csv"
+
+player_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test 
+```
+
+# Veronica works on wrangling transcripts here
+
+```{r}
+
+players <- read_csv(here(player_source)) |> filter(id %in% player_ids) |> select(id, playerId=`_id`)
+
+games <- read_csv(here("data/games.csv"), col_types="ccc") |> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+
 ```
@@
 ```{r}
-raw_data <- read.csv(here("data/rounds7.19.csv"), header = TRUE) %>% 
+
+raw_data <- read.csv(here(data_source), header = TRUE) %>% 
   filter(createdAt >= pilot_start) %>%
@@
   filter(playerId!=speaker) %>% 
-  pivot_wider(names_from=info, values_from=values)
+  pivot_wider(names_from=info, values_from=values) |> 
+  left_join(players, by=c("playerId")) |> 
+  select(-playerId) |> 
+  rename(playerId=id) |> 
+  left_join(players, by=c("speaker"="playerId")) |> 
+  select(-speaker) |> 
+  rename(speaker=id) |> 
+  left_join(games) |> 
+  rename(game=gameConfig) |> 
+  select(-gameId, -offset_sec) |> 
+  filter(!is.na(game)) |> 
+  mutate(time=as.numeric(time))
 
 # write_csv(here(data, 'rounds7.19.csv'))
+```
+
+```{r}
+#just for pilot, need to fix offset
+fix <- raw_data |> filter(trialNum==0) |> select(game, time)
+
+real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)
+
+timing <- raw_data |> group_by(game) |> mutate(cum_time=cumsum(time), cum_sec=cum_time/1000+3*trialNum)
+
+```
+
+```{r}
+
+wrangle_transcript <- function(game_num){
+readxl::read_xlsx(here("data/clean_transcripts", str_c(game_num,".xlsx"))) |> 
+  select(start, end, text, speaker) |> 
+  mutate(start=start/1000, end=end/1000) |> 
+    nest(data=c(start, end, text, speaker))
+}
+
+#wrangle_transcript("game32")
+
+real_offset$transcripts <- map_df(real_offset$gameConfig, wrangle_transcript)
+
+transcripts <- real_offset |> unnest(transcripts) |> unnest(data)
+
+# this section is janky as hell, but it does work
+
+timed_transcripts <- transcripts |> mutate(start=start-offset_sec, end=end-offset_sec)
 ```
@@
 
-View(filtered_data)
+#View(filtered_data)
 ```

### 2023-07-31 code/analysis-pilot.Rmd
having accurate looking timing date on transcripts!

@@
 
-timing <- raw_data |> group_by(game) |> mutate(cum_time=cumsum(time), cum_sec=cum_time/1000+3*trialNum)
-
 ```
@@
 wrangle_transcript <- function(game_num){
-readxl::read_xlsx(here("data/clean_transcripts", str_c(game_num,".xlsx"))) |> 
+read_csv(here("data/clean_transcripts", str_c(game_num,".csv"))) |> 
   select(start, end, text, speaker) |> 
@@
 
-timed_transcripts <- transcripts |> mutate(start=start-offset_sec, end=end-offset_sec)
+```
+
+```{r, include=F, eval=F}
+# a very derpy approach to aligning trial boundaries with transcripts
+
+timing <- raw_data |> group_by(game) |> mutate(cum_time=cumsum(time), cum_sec=cum_time/1000+3*trialNum) |> select(trialNum, game, cum_sec) |> mutate(start=ifelse(trialNum==0, 0,lag(cum_sec)), end=cum_sec, trialNumStart=trialNum, trialNumEnd=trialNum, speaker=NA, text=NA) |> select(-cum_sec, -trialNum)
+
+timed_transcripts <- transcripts |>
+  mutate(start=start-offset_sec, end=end-offset_sec) |> 
+  select(-offset_sec, -time, -gameId) |> 
+  rename(game=gameConfig) |> 
+  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
+  bind_rows(timing) |> 
+  group_by(game) |> 
+  arrange(start) |> 
+  fill(trialNumStart) |> 
+  arrange(end) |> 
+  fill(trialNumEnd, .direction="up") |> 
+  arrange(game) |> 
+  filter(start>0)
+
+# turns out this isn't good enough for timing, let's try stages instead!
+
+```
+
+```{r}
+
+stages <- read_csv(here("data/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)
+
+gameStarts <- stages |> filter(index==0) |> rename(first=startTimeAt) |> select(-index)
+
+stage_timing <- stages |> left_join(gameStarts) |> mutate(rel_time=startTimeAt-first) |> mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> filter(index%%2==0) |> mutate(trialNum=index/2) |> select(gameId, trialNum, start=rel_time, end=rel_time) |> left_join(games) |> mutate(trialNumStart=trialNum, trialNumEnd=lag(trialNum)) |> ungroup() |> select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
+```
+
+```{r}
+timed_transcripts <-  transcripts |>
+  mutate(start=start-offset_sec, end=end-offset_sec) |> 
+  select(-offset_sec, -time, -gameId) |> 
+  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
+  rename(game=gameConfig) |> 
+  bind_rows(stage_timing) |> 
+  group_by(game) |> 
+  arrange(start) |> 
+  fill(trialNumStart) |> 
+  arrange(end) |> 
+  fill(trialNumEnd, .direction="up") |> 
+  arrange(game) |> 
+  filter(!is.na(text)) |> 
+  write_csv(here("data/timed_transcript.csv"))
 ```

### 2023-08-01 code/analysis-pilot.Rmd
new
new1

@@
 accuracy <- filtered_data %>% 
-  group_by(gameId) %>% 
+  group_by(game) %>% 
   summarize(total=n(),
@@
 
-ggplot(data = accuracy, aes(y = gameId, x = percent)) +
+ggplot(data = accuracy, aes(y = game, x = percent)) +
   geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#228B22", width = .7) +
@@
 
-time=time
-
 #graph accuracy to time (scatterplot maybe?)
-#graph accuracy (bar graph)
 

### 2023-08-01 code/analysis-pilot.Rmd
15:46 update

@@
 
-ggplot(data = accuracy, aes(y = game, x = percent)) +
+ggplot(data = accuracy, aes(x = percent, y = game)) +
   geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#228B22", width = .7) +

### 2023-08-15 code/analysis-pilot.Rmd
update code

@@
 
+real_start=lubridate::ymd('2023-07-17')
+real_end=lubridate::ymd('2023-08-09')
+
 data_source <- "data/rounds.csv"
@@
 
-player_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test 
+pilot_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test
+
+real_ids <-str_c("id", c(13:16, 19:36, 39:54, 57:58)) # skipping kids who didn't complete enough trials
 ```
@@
 
-players <- read_csv(here(player_source)) |> filter(id %in% player_ids) |> select(id, playerId=`_id`)
+players <- read_csv(here(player_source)) |> filter(id %in% real_ids) %>% select(id, playerId=`_id`)
 
-games <- read_csv(here("data/games.csv"), col_types="ccc") |> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+games <- read_csv(here("data/games.csv"), col_types="ccc") %>% mutate(offset_sec=0)
+#|> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
 
@@
 raw_data <- read.csv(here(data_source), header = TRUE) %>% 
-  filter(createdAt >= pilot_start) %>%
-  filter(createdAt <= pilot_end) %>% 
+  filter(createdAt >= real_start) %>%
+  filter(createdAt <= real_end) %>% 
   rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
@@
   select(-speaker) |> 
-  rename(speaker=id) |> 
+  rename(speaker=id) %>% 
   left_join(games) |> 
@@
 # write_csv(here(data, 'rounds7.19.csv'))
+real_offset <- raw_data
 ```
 
-```{r}
+```{r, eval=F}
 #just for pilot, need to fix offset
@@
 
-ggplot(data = accuracy, aes(x = percent, y = game)) +
-  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#228B22", width = .7) +
+ggplot(data = accuracy, aes(x = percent, y = reorder(game, -percent))) +
+  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
   geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
@@
         legend.position = "none")
+ggsave("AccuracySorted.png")
+
 ```

### 2023-08-15 code/analysis-pilot.Rmd
work on timing

@@
 # write_csv(here(data, 'rounds7.19.csv'))
-real_offset <- raw_data
 ```
 
-```{r, eval=F}
+```{r}
 #just for pilot, need to fix offset
-fix <- raw_data |> filter(trialNum==0) |> select(game, time)
+#fix <- raw_data |> filter(trialNum==0) |> select(game, time)
 
-real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)
+#real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)
+
+real_offset <- games
 

### 2023-08-16 code/analysis-pilot.Rmd
passable timing data

@@
 
-# Veronica works on wrangling transcripts here
+# Data prep
 
@@
 
-games <- read_csv(here("data/games.csv"), col_types="ccc") %>% mutate(offset_sec=0)
-#|> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+games <- read_csv(here("data/games.csv"), col_types="ccc")|> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
 
@@
 
-# write_csv(here(data, 'rounds7.19.csv'))
 ```
@@
 wrangle_transcript <- function(game_num){
-read_csv(here("data/clean_transcripts", str_c(game_num,".csv"))) |> 
+read_csv(here("data/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
   select(start, end, text, speaker) |> 
@@
 
-```{r, include=F, eval=F}
-# a very derpy approach to aligning trial boundaries with transcripts
-
-timing <- raw_data |> group_by(game) |> mutate(cum_time=cumsum(time), cum_sec=cum_time/1000+3*trialNum) |> select(trialNum, game, cum_sec) |> mutate(start=ifelse(trialNum==0, 0,lag(cum_sec)), end=cum_sec, trialNumStart=trialNum, trialNumEnd=trialNum, speaker=NA, text=NA) |> select(-cum_sec, -trialNum)
-
-timed_transcripts <- transcripts |>
-  mutate(start=start-offset_sec, end=end-offset_sec) |> 
-  select(-offset_sec, -time, -gameId) |> 
-  rename(game=gameConfig) |> 
-  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
-  bind_rows(timing) |> 
-  group_by(game) |> 
-  arrange(start) |> 
-  fill(trialNumStart) |> 
-  arrange(end) |> 
-  fill(trialNumEnd, .direction="up") |> 
-  arrange(game) |> 
-  filter(start>0)
-
-# turns out this isn't good enough for timing, let's try stages instead!
-
-```
-
 ```{r}
@@
 
-stage_timing <- stages |> left_join(gameStarts) |> mutate(rel_time=startTimeAt-first) |> mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> filter(index%%2==0) |> mutate(trialNum=index/2) |> select(gameId, trialNum, start=rel_time, end=rel_time) |> left_join(games) |> mutate(trialNumStart=trialNum, trialNumEnd=lag(trialNum)) |> ungroup() |> select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
+stage_timing <- stages |> left_join(gameStarts) |>
+  mutate(rel_time=startTimeAt-first) |> 
+  mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> 
+   mutate(trialNum=index/2,
+          start=rel_time,
+          end=lead(rel_time)) |> 
+  select(gameId, trialNum,start, end) |>  left_join(games) |>  ungroup() |> 
+  mutate(trialNumStart=trialNum, trialNumEnd=trialNum) |> 
+  select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
 ```
@@
   mutate(start=start-offset_sec, end=end-offset_sec) |> 
-  select(-offset_sec, -time, -gameId) |> 
+  select(-offset_sec, -gameId) |> 
   mutate(trialNumStart=NA, trialNumEnd=NA) |> 
@@
   write_csv(here("data/timed_transcript.csv"))
+```
+
+```{r}
+# how many trial mismatches are there?
+timed_transcripts |> filter(end>0) |> filter(!is.na(trialNumEnd)) |> #it's in the game region, not rules or post game
+ mutate(mismatch=trialNumEnd-trialNumStart) |> 
+  group_by(mismatch) |> 
+  tally()
+
+#so like a third mismatch, and only a couple off by more than 1
 ```

### 2023-08-16 code/analysis-pilot.Rmd
tag excluded games

@@
 
-games <- read_csv(here("data/games.csv"), col_types="ccc")|> mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+games <- read_csv(here("data/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
 
@@
 ```
+# Prep accuracy
 
-# remove columns: numPlayers, sound, activePlayerCount, trialNum, distractor, tangramURLs, response, target, submitted, X_id, repNum, correct, targetNum, stageIds
+## remove columns: numPlayers, sound, activePlayerCount, trialNum, distractor, tangramURLs, response, target, submitted, X_id, repNum, correct, targetNum, stageIds
 ```{r}
@@
 
-#remove NA cells in "time"
+## remove NA cells in "time"
 ```{r}
@@
 
-#relocate columns
+## relocate columns
 ```{r}
@@
 
-#accuracy graph
+# accuracy graph
 ```{r}

### 2023-08-16 code/analysis-pilot.Rmd
14:58

@@
 
+#scatterplot
+ggplot(data=data_long, aes(x = time, y= percent, color=Experimenter)) +
+  geom_point(size=1.5, alpha=.7) +
+  geom_smooth(se=F) +
+  theme_classic() +
+  xlab("Age in months") +
+  ylab("Number of Laughs") +
+  theme(aspect.ratio = 3/5,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=12),
+        legend.title=element_text(size=14,face="bold"),
+        legend.text=element_text(size=12),
+        legend.direction="horizontal",
+        legend.position="bottom") +
+  facet_wrap(~time)
+
 ```
@@
 ```{r}
-ggplot(data=accuracy, aes(x = Age.months, y=Laughs, color=Experimenter)) +
-  geom_point()
+
 ```

### 2023-08-16 code/analysis-pilot.Rmd
new
15:26

@@
 ```
+# timed_transcripts editing:
+```{r}
+timed_new <- timed_transcripts %>% 
+  filter(!speaker %in% c("I", "M", "i")) %>% 
+  mutate(wordCount = str_count(text, "\\S+"))
+
+# filter trials, filter out practice trials, round trialNumStart up for 0.5
+# geom line - color = game OR group = game
+
+```
+
 # Prep accuracy

### 2023-08-16 code/analysis-pilot.Rmd
first spaghetti

--- 2023-08-16 code/analysis-pilot.Rmd
@@
 ```
+# timed_transcripts editing
+```{r}
+timed_new <- timed_transcripts %>% 
+  filter(!speaker %in% c("I", "M", "i")) %>% 
+  filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
+  mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart))
+
+View(timed_new)
+
+```
+
+# timed_new graphing
+```{r}
+# filter trials
+# geom line - color = game OR group = game
+ggplot(timed_new, aes(x = roundStart, y = wordCount, group = game, color = game)) + 
+  geom_line() +
+  labs(color = "Game ID", x = "Round #", y = "Utterance Count") +
+  #scale_x_continuous(#breaks=seq(1,10,1), limits = c(2,13)) +
+  #ylim(120, 200) + 
+  theme_classic()
+
+```
+
 # Prep accuracy
@@
   mutate(fraction = correct/total*100, percent = signif(fraction, 2)) 
+  View(accuracy)
 

### 2023-08-17 code/analysis-pilot.Rmd
line plots

--- 2023-08-16 code/analysis-pilot.Rmd
@@
 ```
+partner 1:
+id13
+id15
+id19
+id21
+id24
+id26
+id28
+id30
+id32
+id33
+id35
+id39
+id42
+id43
+id45
+id47
+id50
+id52
+id54
+id58
+
+partner 2:
+id14
+id16
+id20
+id22
+id23
+id25
+id27
+id29
+id31
+id34
+id36
+id40
+id41
+id44
+id46
+id48
+id49
+id51
+id53
+id57
+
+# timed_transcripts editing for partner1
+```{r}
+timed_new <- timed_transcripts %>% 
+  filter(!speaker %in% c("I", "M", "i", "id14", "id16", "id20", "id22", "id23", "id25", "id27", "id29", "id31", "id34", "id36", "id40", "id41", "id44", "id46", "id48", "id49", "id51", "id53", "id57")) %>%
+  filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
+  mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>% 
+  filter(!roundStart %in% c(2.0, 4.0, 6.0, 8.0, 10.0, 12.0)) %>% 
+  group_by(game, roundStart) %>% 
+    summarize(totwordCount = sum(wordCount))
+
+View(timed_new)
+
+```
+
+# timed_new graphing - left in speaker1listener
+```{r}
+ggplot(timed_new, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
+  geom_line() +
+  #geom_smooth() +
+  labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
+  scale_x_continuous(breaks=seq(3,13,2), limits = c(3,13)) +
+  scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
+  theme_classic() +
+  theme(aspect.ratio = 9/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+        legend.position="right")
+```
+# timed_transcripts editing for partner2
+```{r}
+timed_neww <- timed_transcripts %>% 
+  filter(!speaker %in% c("I", "M", "i", "id13", "id15", "id19", "id21", "id24", "id26", "id28", "id30", "id32", "id33", "id35", "id39", "id42", "id43", "id45", "id47", "id50", "id52", "id54", "id58")) %>%
+  filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
+  mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>% 
+  filter(!roundStart %in% c(3.0, 5.0, 7.0, 9.0, 11.0, 13.0)) %>% 
+  group_by(game, roundStart) %>% 
+    summarize(totwordCount = sum(wordCount))
+
+View(timed_neww)
+
+```
+
+# timed_neww graphing - left in speaker2listener
+```{r}
+ggplot(timed_neww, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
+  geom_line() +
+  #geom_smooth() +
+  labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
+  scale_x_continuous(breaks=seq(2,12,2), limits = c(2,12)) +
+  scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
+  theme_classic() +
+  theme(aspect.ratio = 9/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+        legend.position="right")
+```
+
 # Prep accuracy

### 2023-08-18 code/analysis-pilot.Rmd
data analysis

--- 2023-08-16 code/analysis-pilot.Rmd
@@
 ```
-# timed_transcripts editing:
+<!--
+partner 1:
+id13
+id15
+id19
+id21
+id24
+id26
+id28
+id30
+id32
+id33
+id35
+id39
+id42
+id43
+id45
+id47
+id50
+id52
+id54
+id58
+
+partner 2:
+id14
+id16
+id20
+id22
+id23
+id25
+id27
+id29
+id31
+id34
+id36
+id40
+id41
+id44
+id46
+id48
+id49
+id51
+id53
+id57
+-->
+
+# timed_transcripts editing for partner1
 ```{r}
-timed_new <- timed_transcripts %>% 
-  filter(!speaker %in% c("I", "M", "i")) %>% 
-  mutate(wordCount = str_count(text, "\\S+"))
+# timed_new <- timed_transcripts %>%
+#   filter(!speaker %in% c("I", "M", "i", "id14", "id16", "id20", "id22", "id23", "id25", "id27", "id29", "id31", "id34", "id36", "id40", "id41", "id44", "id46", "id48", "id49", "id51", "id53", "id57")) %>%
+#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>%
+#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>%
+#   filter(!roundStart %in% c(2.0, 4.0, 6.0, 8.0, 10.0, 12.0)) %>%
+#   group_by(game, roundStart) %>%
+#     summarize(totwordCount = sum(wordCount))
+# 
+# View(timed_new)
 
-# filter trials, filter out practice trials, round trialNumStart up for 0.5
-# geom line - color = game OR group = game
+```
 
+# timed_new graphing - left in speaker1listener
+```{r}
+# ggplot(timed_new, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
+#   geom_line() +
+#   #geom_smooth() +
+#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
+#   scale_x_continuous(breaks=seq(3,13,2), limits = c(3,13)) +
+#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
+#   theme_classic() +
+#   theme(aspect.ratio = 9/9,
+#         axis.title=element_text(size=14,face="bold"),
+#         axis.text=element_text(size=9, color = "black"),
+#         legend.title=element_text(size=13,face="bold"),
+#         legend.text=element_text(size=9),
+#         legend.position="right")
+```
+# timed_transcripts editing for partner2
+```{r}
+# timed_neww <- timed_transcripts %>% 
+#   filter(!speaker %in% c("I", "M", "i", "id13", "id15", "id19", "id21", "id24", "id26", "id28", "id30", "id32", "id33", "id35", "id39", "id42", "id43", "id45", "id47", "id50", "id52", "id54", "id58")) %>%
+#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
+#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>% 
+#   filter(!roundStart %in% c(3.0, 5.0, 7.0, 9.0, 11.0, 13.0)) %>% 
+#   group_by(game, roundStart) %>% 
+#     summarize(totwordCount = sum(wordCount))
+# 
+# View(timed_neww)
+
+```
+
+# timed_neww graphing - left in speaker2listener
+```{r}
+# ggplot(timed_neww, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
+#   geom_line() +
+#   #geom_smooth() +
+#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
+#   scale_x_continuous(breaks=seq(2,12,2), limits = c(2,12)) +
+#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
+#   theme_classic() +
+#   theme(aspect.ratio = 9/9,
+#         axis.title=element_text(size=14,face="bold"),
+#         axis.text=element_text(size=9, color = "black"),
+#         legend.title=element_text(size=13,face="bold"),
+#         legend.text=element_text(size=9),
+#         legend.position="right")
 ```
@@
   select(-response) %>% 
-  select(-target) %>% 
   select(-submitted) %>% 
@@
 accuracy <- filtered_data %>% 
+  filter(!index %in% c(0, 1)) %>%
   group_by(game) %>% 
@@
   theme_classic() +
-  xlab("Percentage Accuracy") +
-  ylab("Game ID") +
+  xlab("\nPercentage Accuracy") +
+  ylab("Game ID\n") +
+  scale_x_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
   theme(aspect.ratio = 4/5,
-        axis.title = element_text(size=14,face="bold"),
-        axis.text = element_text(size=12),
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
         legend.position = "none")
@@
 
-#scatterplot
-ggplot(data=data_long, aes(x = time, y= percent, color=Experimenter)) +
-  geom_point(size=1.5, alpha=.7) +
-  geom_smooth(se=F) +
+```
+#accuracy through time
+```{r}
+accuracy <- filtered_data %>% 
+  filter(!index %in% c(0, 1)) %>%
+  group_by(index) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>% 
+  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
+
+ggplot(data = accuracy, aes(x = index-1, y = percent)) +
+  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  geom_smooth(method="lm")+
+  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
   theme_classic() +
-  xlab("Age in months") +
-  ylab("Number of Laughs") +
-  theme(aspect.ratio = 3/5,
+  xlab("\nTrial") +
+  ylab("Percent accuracy\n") +
+  scale_x_continuous(breaks=seq(1,12,1)) +
+  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
+  theme(aspect.ratio = 4/5,
         axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=12),
-        legend.title=element_text(size=14,face="bold"),
-        legend.text=element_text(size=12),
-        legend.direction="horizontal",
-        legend.position="bottom") +
-  facet_wrap(~time)
+        axis.text=element_text(size=9, color = "black"),
+        legend.position = "none")
+ggsave("AccuracybyTrial.png")
+```
+#accuracy by target
+```{r}
+accuracy <- filtered_data %>% 
+  mutate(image = case_when(
+  target == "/experiment/B1.jpg" ~ "Walking",
+  target == "/experiment/D1.jpg" ~ "Lying",
+  target == "/experiment/E1.jpg" ~ "Jumping",
+  target == "/experiment/I1.jpg" ~ "Holding",
+)) %>% 
+  filter(!index %in% c(0, 1)) %>%
+  group_by(image) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>% 
+  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
+
+ggplot(data = accuracy, aes(x = image, y = percent)) +
+  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  geom_smooth(method="lm")+
+  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  theme_classic() +
+  xlab("\nTarget") +
+  ylab("Percent accuracy\n") +
+  #scale_x_continuous(breaks=seq(1,12,1)) +
+  scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100), expand=c(0,0)) +
+  theme(aspect.ratio = 4/5,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.position = "none")
+ggsave("AccuracybyTarget.png")
+```
+
+#reduction
+```{r}
+join_data <- filtered_data %>% 
+  mutate(image = case_when(
+  target == "/experiment/B1.jpg" ~ "Walking",
+  target == "/experiment/D1.jpg" ~ "Lying",
+  target == "/experiment/E1.jpg" ~ "Jumping",
+  target == "/experiment/I1.jpg" ~ "Holding",
+)) %>% 
+  rename(listener = playerId, trialNum = index) %>% 
+  select(-createdAt, -exclude, -time) %>% 
+  pivot_longer(cols = "listener":"speaker", names_to = "role", values_to = "speaker")
+```
+
+```{r}
+timed_new <- timed_transcripts %>% 
+  drop_na(trialNumStart, trialNumEnd) %>% 
+  filter(!speaker %in% c("I", "M", "i")) %>% 
+  mutate(wordCount = str_count(text, "\\S+"), trialNum = ceiling(trialNumStart)) %>% 
+  full_join(
+    join_data,
+    by = c("speaker", "game", "trialNum")
+  ) %>%
+  mutate(wordCount=ifelse(is.na(wordCount), 0, wordCount)) %>% 
+  select(speaker, game, trialNum, role, countCorrect, wordCount, image) %>% 
+  filter(!trialNum %in% c(0.0, 1.0)) %>% 
+   group_by(game, trialNum, role, image) %>% 
+    summarize(totwordCount = sum(wordCount))
+```
+
+#speaker graphing
+```{r}
+timed_speaker <- timed_new %>% 
+  filter(role == "speaker")
+
+ggplot(timed_speaker, aes(x = trialNum-1, y = totwordCount)) + 
+  geom_line(aes(group=game)) +
+  geom_smooth() +
+  labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
+  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
+  theme_classic() +
+  theme(aspect.ratio = 9/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+        legend.position="right")
+ggsave("SpeakerWord.png")
+```
+
+#listener graphing
+```{r}
+timed_listener <- timed_new %>% 
+  filter(role == "listener")
+
+ggplot(timed_listener, aes(x = trialNum-1, y = totwordCount)) + 
+  geom_line(aes(group=game)) +
+  geom_smooth() +
+  labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
+  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
+  theme_classic() +
+  theme(aspect.ratio = 9/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+        legend.position="right")
+ggsave("ListenerWord.png")
+```
+
+<!-- #speakerbytarget -->
+```{r}
+# speaker_target <- timed_new %>% 
+#   mutate(block=(trialNum-2)%/%4+1) %>% 
+#   filter(role == "speaker") %>% 
+#   mutate(gameximage=str_c(game,image))
+# 
+# ggplot(speaker_target, aes(x = block, y = totwordCount, color=image)) + 
+#   #geom_line(aes(group=gameximage), alpha=.5, color="gray") +
+#   #geom_jitter()+
+#   geom_smooth(method="lm") +
+#   labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
+#   #scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
+#   theme_classic() +
+#   #facet_wrap(~image)+
+#   theme(aspect.ratio = 9/9,
+#         axis.title=element_text(size=14,face="bold"),
+#         axis.text=element_text(size=9, color = "black"),
+#         legend.title=element_text(size=13,face="bold"),
+#         legend.text=element_text(size=9),
+#         legend.position="right")
 
 ```
-
-#graph accuracy to time (scatterplot maybe?)
-
-```{r}
-
-```
-```

### 2023-08-18 code/analysis-pilot.Rmd
labels on graphs

@@
   theme_classic() +
-  xlab("\nPercentage Accuracy") +
+  xlab("\nPercentage accuracy") +
   ylab("Game ID\n") +
@@
   geom_smooth(method="lm")+
-  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
   theme_classic() +
@@
   scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100), expand=c(0,0)) +
+  geom_text(aes(label = percent), vjust = 1.7, color = "white", size = 3) +
   theme(aspect.ratio = 4/5,

### 2023-08-19 code/analysis-pilot.Rmd
new labels

@@
   geom_smooth() +
-  labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
+  labs(color = "Game ID", x = "\nTrial", y = "Word Count\n") +
   scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +

### 2023-08-19 code/analysis-pilot.Rmd
dimensions of graph

@@
   theme_classic() +
-  theme(aspect.ratio = 9/9,
+  theme(aspect.ratio = 7/9,
         axis.title=element_text(size=14,face="bold"),

### 2023-08-19 code/analysis-pilot.Rmd
image accuracy

@@
 ```
+#accuracy of each image over time
+```{r}
+accuracy <- filtered_data %>% 
+  mutate(image = case_when(
+  target == "/experiment/B1.jpg" ~ "Walking",
+  target == "/experiment/D1.jpg" ~ "Lying",
+  target == "/experiment/E1.jpg" ~ "Jumping",
+  target == "/experiment/I1.jpg" ~ "Holding",
+)) %>% 
+  mutate(trialNum = index-1) %>% 
+  filter(!index %in% c(0, 1)) %>%
+  group_by(image, index) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>%
+  mutate(fraction = correct/total*100, percent = signif(fraction, 2)) %>% 
+  filter(image == "Walking")
+
+ggplot(accuracy, aes(x = index-1, y = percent)) +
+  geom_line(aes(group=image)) +
+  labs(color = "Tangram", x = "\nTrial", y = "Percent accuracy\n") +
+  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
+  scale_y_continuous(breaks=seq(0,100,10), limits = c(0,100)) +
+  theme_classic() +
+  theme(aspect.ratio = 7/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+        legend.position="right")
+
+ggsave("WalkingAccuracybyTarget.png")
+```
 

### 2023-09-27 code/data_prep.Rmd
start processing for tagged transcript

--- 2023-08-19 code/analysis-pilot.Rmd
@@
 ---
-title: "analysis-pilot"
-output: html_document
-date: "2023-07-19"
+title: "Wrangle transcripts"
+output: html_notebook
 ---
@@
 
-# Data prep
+# Empirica output prep
 
@@
 
+# Transcript wrangle - new
 ```{r}
 
+real_offset <- real_offset %>% filter(gameConfig=="game41")
 wrangle_transcript <- function(game_num){
 read_csv(here("data/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
-  select(start, end, text, speaker) |> 
+  select(start, end, text, speaker, trial, role, echo, description, backchannel) |> 
+    fill(trial) %>% 
   mutate(start=start/1000, end=end/1000) |> 
-    nest(data=c(start, end, text, speaker))
+    nest(data=c(start, end, text, speaker, trial, role, echo, description, backchannel))
 }
 
-#wrangle_transcript("game32")
+#wrangle_transcript("game41") %>% unnest(data) %>% View()
 
@@
 ```{r}
+# Quick check
+
+timed_transcripts %>% filter(!is.na(trialNumStart)) %>%
+
+```
+
+```{r}
 # how many trial mismatches are there?
@@
 ```
-<!--
-partner 1:
-id13
-id15
-id19
-id21
-id24
-id26
-id28
-id30
-id32
-id33
-id35
-id39
-id42
-id43
-id45
-id47
-id50
-id52
-id54
-id58
-
-partner 2:
-id14
-id16
-id20
-id22
-id23
-id25
-id27
-id29
-id31
-id34
-id36
-id40
-id41
-id44
-id46
-id48
-id49
-id51
-id53
-id57
--->
-
-# timed_transcripts editing for partner1
-```{r}
-# timed_new <- timed_transcripts %>%
-#   filter(!speaker %in% c("I", "M", "i", "id14", "id16", "id20", "id22", "id23", "id25", "id27", "id29", "id31", "id34", "id36", "id40", "id41", "id44", "id46", "id48", "id49", "id51", "id53", "id57")) %>%
-#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>%
-#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>%
-#   filter(!roundStart %in% c(2.0, 4.0, 6.0, 8.0, 10.0, 12.0)) %>%
-#   group_by(game, roundStart) %>%
-#     summarize(totwordCount = sum(wordCount))
-# 
-# View(timed_new)
-
-```
-
-# timed_new graphing - left in speaker1listener
-```{r}
-# ggplot(timed_new, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
-#   geom_line() +
-#   #geom_smooth() +
-#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
-#   scale_x_continuous(breaks=seq(3,13,2), limits = c(3,13)) +
-#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
-#   theme_classic() +
-#   theme(aspect.ratio = 9/9,
-#         axis.title=element_text(size=14,face="bold"),
-#         axis.text=element_text(size=9, color = "black"),
-#         legend.title=element_text(size=13,face="bold"),
-#         legend.text=element_text(size=9),
-#         legend.position="right")
-```
-# timed_transcripts editing for partner2
-```{r}
-# timed_neww <- timed_transcripts %>% 
-#   filter(!speaker %in% c("I", "M", "i", "id13", "id15", "id19", "id21", "id24", "id26", "id28", "id30", "id32", "id33", "id35", "id39", "id42", "id43", "id45", "id47", "id50", "id52", "id54", "id58")) %>%
-#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
-#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>% 
-#   filter(!roundStart %in% c(3.0, 5.0, 7.0, 9.0, 11.0, 13.0)) %>% 
-#   group_by(game, roundStart) %>% 
-#     summarize(totwordCount = sum(wordCount))
-# 
-# View(timed_neww)
-
-```
-
-# timed_neww graphing - left in speaker2listener
-```{r}
-# ggplot(timed_neww, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
-#   geom_line() +
-#   #geom_smooth() +
-#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
-#   scale_x_continuous(breaks=seq(2,12,2), limits = c(2,12)) +
-#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
-#   theme_classic() +
-#   theme(aspect.ratio = 9/9,
-#         axis.title=element_text(size=14,face="bold"),
-#         axis.text=element_text(size=9, color = "black"),
-#         legend.title=element_text(size=13,face="bold"),
-#         legend.text=element_text(size=9),
-#         legend.position="right")
-```
-
-# Prep accuracy
-
-## remove columns: numPlayers, sound, activePlayerCount, trialNum, distractor, tangramURLs, response, target, submitted, X_id, repNum, correct, targetNum, stageIds
-```{r}
-filtered_data <- raw_data %>% 
-  select(-numPlayers) %>% 
-  select(-sound) %>%  
-  select(-activePlayerCount) %>% 
-  select(-trialNum) %>% 
-  select(-distractor) %>% 
-  select(-tangramURLs) %>% 
-  select(-response) %>% 
-  select(-submitted) %>% 
-  select(-X_id) %>% 
-  select(-repNum) %>% 
-  select(-correct) %>% 
-  select(-targetNum) %>% 
-  select(-stageIds)
-
-```
-
-## remove NA cells in "time"
-```{r}
-filtered_data <- filtered_data %>% 
-  drop_na(time)
-
-View(filtered_data)
-```
-
-## relocate columns
-```{r}
-filtered_data <- filtered_data %>% 
-  select(createdAt, everything()) %>% 
-  relocate(countCorrect, .before = time)
-
-#View(filtered_data)
-```
-
-# accuracy graph
-```{r}
-accuracy <- filtered_data %>% 
-  filter(!index %in% c(0, 1)) %>%
-  group_by(game) %>% 
-  summarize(total=n(),
-            correct=sum(countCorrect)) %>% 
-  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
-
-ggplot(data = accuracy, aes(x = percent, y = reorder(game, -percent))) +
-  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nPercentage accuracy") +
-  ylab("Game ID\n") +
-  scale_x_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
-ggsave("AccuracySorted.png")
-
-```
-#accuracy through time
-```{r}
-accuracy <- filtered_data %>% 
-  filter(!index %in% c(0, 1)) %>%
-  group_by(index) %>% 
-  summarize(total=n(),
-            correct=sum(countCorrect)) %>% 
-  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
-
-ggplot(data = accuracy, aes(x = index-1, y = percent)) +
-  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  geom_smooth(method="lm")+
-  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nTrial") +
-  ylab("Percent accuracy\n") +
-  scale_x_continuous(breaks=seq(1,12,1)) +
-  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
-ggsave("AccuracybyTrial.png")
-```
-#accuracy by target
-```{r}
-accuracy <- filtered_data %>% 
-  mutate(image = case_when(
-  target == "/experiment/B1.jpg" ~ "Walking",
-  target == "/experiment/D1.jpg" ~ "Lying",
-  target == "/experiment/E1.jpg" ~ "Jumping",
-  target == "/experiment/I1.jpg" ~ "Holding",
-)) %>% 
-  filter(!index %in% c(0, 1)) %>%
-  group_by(image) %>% 
-  summarize(total=n(),
-            correct=sum(countCorrect)) %>% 
-  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
-
-ggplot(data = accuracy, aes(x = image, y = percent)) +
-  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  geom_smooth(method="lm")+
-  theme_classic() +
-  xlab("\nTarget") +
-  ylab("Percent accuracy\n") +
-  #scale_x_continuous(breaks=seq(1,12,1)) +
-  scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100), expand=c(0,0)) +
-  geom_text(aes(label = percent), vjust = 1.7, color = "white", size = 3) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
-ggsave("AccuracybyTarget.png")
-```
-
-#reduction
-```{r}
-join_data <- filtered_data %>% 
-  mutate(image = case_when(
-  target == "/experiment/B1.jpg" ~ "Walking",
-  target == "/experiment/D1.jpg" ~ "Lying",
-  target == "/experiment/E1.jpg" ~ "Jumping",
-  target == "/experiment/I1.jpg" ~ "Holding",
-)) %>% 
-  rename(listener = playerId, trialNum = index) %>% 
-  select(-createdAt, -exclude, -time) %>% 
-  pivot_longer(cols = "listener":"speaker", names_to = "role", values_to = "speaker")
-```
-
-```{r}
-timed_new <- timed_transcripts %>% 
-  drop_na(trialNumStart, trialNumEnd) %>% 
-  filter(!speaker %in% c("I", "M", "i")) %>% 
-  mutate(wordCount = str_count(text, "\\S+"), trialNum = ceiling(trialNumStart)) %>% 
-  full_join(
-    join_data,
-    by = c("speaker", "game", "trialNum")
-  ) %>%
-  mutate(wordCount=ifelse(is.na(wordCount), 0, wordCount)) %>% 
-  select(speaker, game, trialNum, role, countCorrect, wordCount, image) %>% 
-  filter(!trialNum %in% c(0.0, 1.0)) %>% 
-   group_by(game, trialNum, role, image) %>% 
-    summarize(totwordCount = sum(wordCount))
-```
-
-#speaker graphing
-```{r}
-timed_speaker <- timed_new %>% 
-  filter(role == "speaker")
-
-ggplot(timed_speaker, aes(x = trialNum-1, y = totwordCount)) + 
-  geom_line(aes(group=game)) +
-  geom_smooth() +
-  labs(color = "Game ID", x = "\nTrial", y = "Word Count\n") +
-  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
-  theme_classic() +
-  theme(aspect.ratio = 7/9,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.title=element_text(size=13,face="bold"),
-        legend.text=element_text(size=9),
-        legend.position="right")
-ggsave("SpeakerWord.png")
-```
-
-#listener graphing
-```{r}
-timed_listener <- timed_new %>% 
-  filter(role == "listener")
-
-ggplot(timed_listener, aes(x = trialNum-1, y = totwordCount)) + 
-  geom_line(aes(group=game)) +
-  geom_smooth() +
-  labs(color = "Game ID", x = "\nTrial", y = "Word Count\n") +
-  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
-  theme_classic() +
-  theme(aspect.ratio = 7/9,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.title=element_text(size=13,face="bold"),
-        legend.text=element_text(size=9),
-        legend.position="right")
-ggsave("ListenerWord.png")
-```
-
-<!-- #speakerbytarget -->
-```{r}
-# speaker_target <- timed_new %>% 
-#   mutate(block=(trialNum-2)%/%4+1) %>% 
-#   filter(role == "speaker") %>% 
-#   mutate(gameximage=str_c(game,image))
-# 
-# ggplot(speaker_target, aes(x = block, y = totwordCount, color=image)) + 
-#   #geom_line(aes(group=gameximage), alpha=.5, color="gray") +
-#   #geom_jitter()+
-#   geom_smooth(method="lm") +
-#   labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
-#   #scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
-#   theme_classic() +
-#   #facet_wrap(~image)+
-#   theme(aspect.ratio = 9/9,
-#         axis.title=element_text(size=14,face="bold"),
-#         axis.text=element_text(size=9, color = "black"),
-#         legend.title=element_text(size=13,face="bold"),
-#         legend.text=element_text(size=9),
-#         legend.position="right")
-
-```

### 2023-09-29 code/analysis_1.Rmd
start data prep and analyses

---
title: "Analysis - expt 1"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)

transcript <- read_csv(here("data/timed_transcript.csv"))

dat <- read_csv(here("data/clean_data.csv"))

time_palette <- c("practice"="grey","block 1"="blue","block 2"="green", "block 3"="red")#TODO FIX
```

Trials were excluded if the speaker did not provide any description before the listener clicked.
# Accuracy over time

```{r}
# want to know which trials to exclude b/c no description given
exclusion <- transcript %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trialNum=trial, echo) %>% unique() %>% mutate("S_talked"=1)

dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```
# Rate of echoing

will need to check more closely, may have missed some
```{r}
dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  mutate(is.echo=ifelse(is.na(echo),0,1)) %>% 
  ggplot(aes(x = trialNum+1, y = is.echo, fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```

# Speed to start of description

```{r}

transcript %>% filter(role=="S") %>% filter(!is.na(description)) %>% group_by(game, trial) %>% 
  mutate(diff=end_time-trial_start) %>% summarize(first=min(diff)) %>% 
  mutate(type=case_when(trial<2 ~ "practice",
                        trial<6 ~ "block 1",
                        trial<10 ~ "block 2",
                        trial<14 ~ "block 3")) %>% 
  ggplot(aes(x = trial+1, y = first, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Time to start of description") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

TO DO try to understand negative outliers?

# Speed to response

```{r}

 dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  ggplot(aes(x = trialNum+1, y = time/1000, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  coord_cartesian(ylim=c(0,100))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

# Total kid words said in at least vague reference

# Descriptive something something
- qualitative on descriptions and similarities of descriptions (within/between kids) (maybe cut when kids say the same thing thrice down to once?)

- what descriptions "worked" -- some sort of qualitative range

### 2023-09-29 code/data_prep.Rmd
start data prep and analyses

@@
 
+label_image <- function(string){
+  new_string <- case_when(
+    string=="/experiment/B1.jpg" ~ "walk",
+    string=="/experiment/D1.jpg" ~ "swim",
+    string=="/experiment/E1.jpg" ~ "jump",
+    string=="/experiment/I1.jpg" ~ "hold",
+    string=="/experiment/banana.png"~ "banana",
+    string=="/experiment/bird.jpeg" ~ "bird",
+    string=="/experiment/dog.png" ~ "dog",
+    string =="/experiment/apple.png" ~ "apple"
+  )
+  return(new_string)
+}
+
+less_raw_data <- raw_data %>% select(game, target, distractor, trialNum, targetNum, repNum, response, correct, time, listener=playerId, speaker) %>% 
+  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>% write_csv(here("data/clean_data.csv"))
 ```
@@
 # Transcript wrangle - new
+
 ```{r}
 
-real_offset <- real_offset %>% filter(gameConfig=="game41")
 wrangle_transcript <- function(game_num){
@@
   arrange(game) |> 
-  filter(!is.na(text)) |> 
+  filter(!is.na(text))
+
+echoes <- timed_transcripts %>% select(game,trial, echo) %>% filter(echo=="Y") %>% unique()
+
+trial_timing <- stage_timing %>% select(game, trial=trialNumStart, trial_start=start)
+
+transcript_full <- timed_transcripts %>% select(-echo, -exclude) %>% left_join(echoes) %>% 
+  left_join(trial_timing) %>% 
+  rename(start_time=start, end_time=end) %>% 
   write_csv(here("data/timed_transcript.csv"))
@@
 
+# Checks
 ```{r}
-# Quick check
+# Quick check -- for kid talking, so auto-trial and hand-trial align?
 
-timed_transcripts %>% filter(!is.na(trialNumStart)) %>%
+transcript_full %>% filter(!is.na(trialNumStart)) %>% filter(str_detect(speaker, "id")) %>% filter(!is.na(role)) %>% filter(trial!=trialNumStart)
+
+# we'll go with the by-hand method -- these mostly seem to be when there's a wide span of time! possibly due to whisper being silly or us hand filling due to whisper missing things
 
 ```
-
-```{r}
-# how many trial mismatches are there?
-timed_transcripts |> filter(end>0) |> filter(!is.na(trialNumEnd)) |> #it's in the game region, not rules or post game
- mutate(mismatch=trialNumEnd-trialNumStart) |> 
-  group_by(mismatch) |> 
-  tally()
-
-#so like a third mismatch, and only a couple off by more than 1
-```

### 2023-10-02 code/analysis_1.Rmd
checks and fixes on tagging quality

@@
 will need to check more closely, may have missed some
+
+or we'll just say that the rate is high and we can't subset but need to deal with the fact there just is echoing (in this experiment)
+
 ```{r}

### 2023-10-02 code/data_prep.Rmd
checks and fixes on tagging quality

@@
 ```
+
+checks on data:
+* anything that has a description should be from a speaker and have Y for backchannel
+* anything that has a Y for backchannel should be from a child
+
+```{r}
+transcript_full %>% filter(!is.na(description)) %>% filter(is.na(backchannel))
+transcript_full %>% filter(!is.na(description)) %>% ungroup() %>% select(role) %>% unique()
+transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% select(role) %>% unique() 
+transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% filter(!str_detect(speaker,"id"))
+```

### 2023-10-04 code/analysis_1.Rmd
work on graphs

@@
 title: "Analysis - expt 1"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(ggplot2)
+library(lme4)
 
+knitr::opts_chunk$set(
+                      echo=F, warning=F, cache=T, 
+                      message=F, sanitize = T)
+theme_set(theme_classic())
 transcript <- read_csv(here("data/timed_transcript.csv"))
@@
 dat <- read_csv(here("data/clean_data.csv"))
+exclusion <- transcript %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trialNum=trial, echo) %>% unique() %>% mutate("S_talked"=1)
 
-time_palette <- c("practice"="grey","block 1"="blue","block 2"="green", "block 3"="red")#TODO FIX
+time_palette <- c("practice"="grey","block 1"="#E41A1C","block 2"="#377EB8", "block 3"="#4DAF4A")#TODO FIX
+
 ```
@@
 Trials were excluded if the speaker did not provide any description before the listener clicked.
-# Accuracy over time
 
-```{r}
-# want to know which trials to exclude b/c no description given
-exclusion <- transcript %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trialNum=trial, echo) %>% unique() %>% mutate("S_talked"=1)
-
-dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
-  mutate(type=case_when(trialNum<2 ~ "practice",
-                        trialNum<6 ~ "block 1",
-                        trialNum<10 ~ "block 2",
-                        trialNum<14 ~ "block 3")) %>% 
-  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
-  stat_summary(fun.data="mean_cl_boot", geom="bar") +
-    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
-  scale_fill_manual(values=time_palette)+
-  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  #geom_smooth(method="lm")+
-  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nTrial") +
-  ylab("Percent accuracy\n") +
-  #scale_x_continuous(breaks=seq(1,12,1)) +
-  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
-```
 # Rate of echoing
 
-will need to check more closely, may have missed some
-
-or we'll just say that the rate is high and we can't subset but need to deal with the fact there just is echoing (in this experiment)
+I may be missing some echoes, although the rate of echoing is high enough we can't subset based on it (not to mention it's not independent of child behavior). 
+So we're not going to exclude based on this and it'll just be a caveat to interpretation.
 
@@
 
-# Speed to start of description
+# Accuracy over time
+
+```{r}
+dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+  mutate(type=case_when(trialNum<2 ~ "practice",
+                        trialNum<6 ~ "block 1",
+                        trialNum<10 ~ "block 2",
+                        trialNum<14 ~ "block 3")) %>% 
+  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
+  stat_summary(fun.data="mean_cl_boot", geom="bar") +
+    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
+  scale_fill_manual(values=time_palette)+
+  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  #geom_smooth(method="lm")+
+  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  theme_classic() +
+  xlab("\nTrial") +
+  ylab("Percent accuracy\n") +
+  #scale_x_continuous(breaks=seq(1,12,1)) +
+  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
+  theme(aspect.ratio = 4/5,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.position = "none")
+```
+
+The practice trials are great -- probably some improvement over time in regular trials?
 
@@
 
-transcript %>% filter(role=="S") %>% filter(!is.na(description)) %>% group_by(game, trial) %>% 
+data_for_mods <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+  filter(trialNum>1) %>% 
+  mutate(trialNum=trialNum-2,
+         correct.num=ifelse(correct, 1,0),
+         time_sec= time/1000)
+
+```
+
+```{r}
+acc_mod <- glmer(correct.num ~ trialNum + (trialNum|game) + (1|target), family=binomial(), data=data_for_mods)
+
+summary(acc_mod)
+confint(acc_mod, parm=c("(Intercept)","trialNum"))
+
+```
+
+Accuracy is probably increasing, but interval overlaps 0.
+
+# Speed to start of description
+
+I was trying to look at whether speakers initiated descriptions faster later on.
+Some weird negative outliers suggest a timing glitch in at least one expt.
+But also, it doesn't look like this is true (also requires relying on more layers of timing accuracy & alignment).
+
+```{r}
+transcript %>% filter(role=="S") %>% filter(backchannel=="Y") %>% group_by(game,speaker, trial) %>% 
   mutate(diff=end_time-trial_start) %>% summarize(first=min(diff)) %>% 
@@
   ggplot(aes(x = trial+1, y = first, color=type)) +
-  geom_jitter(width=.2, height=0, alpha=.5)+
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+  #geom_jitter(width=.2, height=0, alpha=.5)+
+  geom_line(aes(group=speaker), color="black")+
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange") +
   scale_color_manual(values=time_palette)+
@@
 ```
+```{r}
+transcript %>% filter(role=="S") %>% filter(backchannel=="Y") %>% group_by(game,speaker, trial) %>% 
+  mutate(diff=end_time-trial_start) %>% summarize(first=min(diff)) %>% filter(first<0)
+
+```
 
@@
 # Speed to response
+
+How long do trials take?
+
+Note, some high outliers cut out of view.
 
@@
                         trialNum<14 ~ "block 3")) %>% 
+  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
   ggplot(aes(x = trialNum+1, y = time/1000, color=type)) +
   geom_jitter(width=.2, height=0, alpha=.5)+
+  #geom_line(aes(group=game))+
     stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
@@
   xlab("\nTrial") +
-  coord_cartesian(ylim=c(0,100))+
+  ylab("Time to answer in seconds")+
+  coord_cartesian(ylim=c(0,60))+
   #ylab("Percent trials with echoing \n by experimenter\n") +
   #scale_x_continuous(breaks=seq(1,12,1)) +
-  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
+  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
+  theme(aspect.ratio = 4/5,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.position = "none")
+
+```
+speed to response does get faster over time!
+
+```{r}
+
+speed_mod <- lmer(time_sec ~ trialNum + (trialNum|game) + (1|target),  data=data_for_mods)
+
+summary(speed_mod)
+confint(speed_mod, parm=c("(Intercept)","trialNum"))
+
+```
+Is getting faster over time at about 1 second / trial.
+
+# Length of description from speaker
+
+```{r}
+
+correctness <- dat %>% select(game, trial=trialNum, target, distractor, correct)
+transcript %>% filter(!is.na(description)) %>% group_by(game,speaker, trial) %>% 
+  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% 
+  left_join(correctness) %>% 
+  mutate(word_count=str_count(words, "\\S+")) %>% 
+  mutate(type=case_when(trial<2 ~ "practice",
+                        trial<6 ~ "block 1",
+                        trial<10 ~ "block 2",
+                        trial<14 ~ "block 3")) %>% 
+  ggplot(aes(x = trial+1, y = word_count, color=type)) +
+    geom_jitter(width=.2, height=0, alpha=.5)+
+  #geom_line(aes(group=game))+
+ # facet_wrap(.~correct) + 
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+  scale_color_manual(values=time_palette)+
+  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  #geom_smooth(method="lm")+
+  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  theme_classic() +
+  xlab("\nTrial") +
+  #coord_cartesian(ylim=c(0,60))+
+  #ylab("Percent trials with echoing \n by experimenter\n") +
+  #scale_x_continuous(breaks=seq(1,12,1)) +
+  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
   theme(aspect.ratio = 4/5,
@@
 
-# Total kid words said in at least vague reference
+Going up slightly, if anything, not down. (Although for this task, not sure I'd expect adults to go down rather than to start at fast and stay there).  But it's still different!
 
-# Descriptive something something
-- qualitative on descriptions and similarities of descriptions (within/between kids) (maybe cut when kids say the same thing thrice down to once?)
+Could also look at total words that are at least vaguely game related, although this will have "it looks like", repetition, and inconsistently tagged "Yes" in response to "do you see it" so idk if that's useful
 
-- what descriptions "worked" -- some sort of qualitative range
+```{r}
+
+words_for_mod <- transcript %>% filter(!is.na(description)) %>% rename(trialNum=trial) %>%  left_join(dat) %>% 
+  group_by(game,speaker, trialNum, target) %>% 
+  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% 
+  mutate(word_count=str_count(words, "\\S+")) %>%  
+  filter(trialNum>1) %>% 
+  mutate(trialNum=trialNum-2)
+
+word_mod  <- lmer(word_count ~ trialNum + (trialNum|game) + (1|target), data=words_for_mod)
+
+summary(word_mod)
+confint(word_mod, parm=c("(Intercept)","trialNum"))
+```
+If anything, a slight positive relationship.
+
+# Sbert
+
+```{r}
+sims <- read_rds(here("data/similarities.rds"))
+
+labels <- read_csv(here("data/pre_sbert.csv"))
+
+labels1 <- labels %>% rename(trial1=trial, game1=game, words1=words, target1=target, distractor1=distractor)
+
+labels2 <- labels %>% rename(trial2=trial, game2=game, words2=words, target2=target, distractor2=distractor)
+```
+
+## Check practice trials
+
+```{r}
+
+sims %>% filter(trial1 %in% c("0","1")) %>% filter(trial2 %in% c("0","1")) %>% left_join(labels1) %>% left_join(labels2) %>% mutate(same=target1==target2) %>% ggplot(aes(x=sim, fill=same))+geom_density(alpha=.5)
+```
+We expect when it's the same, high agreement and when they're different, low agreement. This checks out.
+
+## Within a game 
+Descriptions from different kids of the same item are more similar than diff items.
+Same item is described more like partner describes it than like random other kid does.
+(Note that within game we can only compare targets across blocks, so doing cross-block for everything)
+
+```{r}
+
+sims %>%  filter(!is.na(sim)) %>% 
+  filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
+  left_join(labels1) %>% left_join(labels2) %>%
+  mutate(same_target=ifelse(target1==target2, "same_target", "diff_target"),  #label if target is same or not
+  same_child=case_when(
+    game1!=game2 ~ "diff game",
+    (as.numeric(trial1)-as.numeric(trial2))%%2==1~"diff_child", #children alternate
+    T ~"same_child"),
+  type1=case_when(trial1<2 ~ "practice",
+                        trial1<6 ~ "block 1",
+                        trial1<10 ~ "block 2",
+                        trial1<14 ~ "block 3"),
+         type2=case_when(trial2<2 ~ "practice",
+                        trial2<6 ~ "block 1",
+                        trial2<10 ~ "block 2",
+                        trial2<14 ~ "block 3")) %>% 
+  filter(type1!=type2) %>% # can only compare across blocks in some cases, so only cross ever
+    mutate(blocks=case_when(
+      type1=="block 1" & type2=="block 2"~ "1-2",
+      type1=="block 2" & type2=="block 1"~ "1-2",
+      type1=="block 1" & type2=="block 3"~ "1-3",
+      type1=="block 3" & type2=="block 1"~ "1-3",
+      type1=="block 2" & type2=="block 3"~ "2-3",
+      type1=="block 3" & type2=="block 2"~ "2-3"
+      )) %>% 
+  ggplot(aes(x=str_c(same_target,"\n",same_child), y=sim, color=blocks)) +
+  geom_jitter(alpha=.01, color="black") +
+      stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
+  labs(x="Conditions compared")
+```
+
+## Do descriptions get more different?
+
+Not seeing noticeable change over time.
+
+```{r}
+
+sims %>%  filter(!is.na(sim)) %>% 
+  filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
+  left_join(labels1) %>% left_join(labels2) %>%
+  mutate(same_target=ifelse(target1==target2, "same_target", "diff_target"),  #label if target is same or not
+  same_child=case_when(
+    game1!=game2 ~ "diff game",
+    (as.numeric(trial1)-as.numeric(trial2))%%2==1~"diff_child", #children alternate
+    T ~"same_child"),
+  type1=case_when(trial1<2 ~ "practice",
+                        trial1<6 ~ "block 1",
+                        trial1<10 ~ "block 2",
+                        trial1<14 ~ "block 3"),
+         type2=case_when(trial2<2 ~ "practice",
+                        trial2<6 ~ "block 1",
+                        trial2<10 ~ "block 2",
+                        trial2<14 ~ "block 3")) %>% 
+  filter(type1==type2) %>% # can only compare across blocks in some cases, so only cross ever
+  ggplot(aes(x=str_c(same_target,"\n",same_child), y=sim, color=type1)) +
+  geom_jitter(alpha=.01, color="black") +
+      stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
+  labs(x="Conditions compared")
+```
+
+For above, could try subsetting by successful utterances or something.
+
+# The fun part
+
+What sorts of wacky descriptions do kids use successfully?
+
+```{r, out.width="25%"}
+
+knitr::include_graphics(here("experiments/preschool_tangram/public/experiment/E1.jpg"))
+
+labels %>% left_join(dat %>% rename(trial=trialNum)) %>% filter(correct) %>% filter(target=="jump") %>% select(words)
+```
+
+```{r, out.width="25%"}
+
+knitr::include_graphics(here("experiments/preschool_tangram/public/experiment/B1.jpg"))
+
+labels %>% left_join(dat %>% rename(trial=trialNum)) %>% filter(correct) %>% filter(target=="walk") %>% select(words)
+```
+
+```{r, out.width="25%"}
+
+knitr::include_graphics(here("experiments/preschool_tangram/public/experiment/D1.jpg"))
+
+labels %>% left_join(dat %>% rename(trial=trialNum)) %>% filter(correct) %>% filter(target=="swim") %>% select(words)
+```
+
+```{r, out.width="25%"}
+
+knitr::include_graphics(here("experiments/preschool_tangram/public/experiment/I1.jpg"))
+
+labels %>% left_join(dat %>% rename(trial=trialNum)) %>% filter(correct) %>% filter(target=="hold") %>% select(words)
+```
+
+# Pre-reg
+In addition to graphical analyses (above), we said
+We plan to run the model:
+DV  ~ trial_num + (trial_num|dyad) + (1|target) 
+for the DVs: accuracy (logistic model), speed (linear model) and number of words of speaker description (linear model ).
+
+# Pretty for CAMP
+
+```{r, fig.height=8, fig.width=5}
+library(cowplot)
+
+acc_w_legend <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+  mutate(type=case_when(trialNum<2 ~ "practice",
+                        trialNum<6 ~ "block 1",
+                        trialNum<10 ~ "block 2",
+                        trialNum<14 ~ "block 3"),
+         type=factor(type, levels=c("practice", "block 1", "block 2", "block 3"))) %>% 
+  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
+  stat_summary(fun.data="mean_cl_boot", geom="bar") +
+    stat_summary(fun.data="mean_cl_boot", geom="linerange") +
+  scale_fill_manual(values=time_palette)+
+  theme_bw() +
+  xlab("Trial") +
+  ylab("Accuracy") +
+  labs(fill="")+
+  geom_hline(yintercept=.5, linetype="dashed")+
+    coord_cartesian(ylim=c(0,1), expand=F)+
+  theme(axis.text=element_text(size=10, color = "black"),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        #axis.title.x=element_blank(),
+                axis.title=element_text(size=12),
+        legend.text=element_text(size=12),
+        legend.position = "bottom")
+
+acc_legend <- cowplot::get_legend(acc_w_legend)
+acc_plot <- acc_w_legend + theme(legend.position="none")
+
+speed_plot <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+  mutate(type=case_when(trialNum<2 ~ "practice",
+                        trialNum<6 ~ "block 1",
+                        trialNum<10 ~ "block 2",
+                        trialNum<14 ~ "block 3")) %>% 
+  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
+  ggplot(aes(x = trialNum+1, y = time/1000, color=type)) +
+  geom_jitter(width=.2, height=0, alpha=.7)+
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+  scale_color_manual(values=time_palette)+
+  theme_bw() +
+  ylab("Speed (seconds)")+
+  xlab("Trial")+
+  coord_cartesian(ylim=c(0,60), expand=F)+
+  theme(axis.text=element_text(size=10, color = "black"),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+                axis.title=element_text(size=12),
+        #axis.title.x=element_blank(),
+        legend.position = "none")
+
+word_plot <- transcript %>% filter(!is.na(description)) %>% group_by(game,speaker, trial) %>% 
+  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% 
+  mutate(word_count=str_count(words, "\\S+")) %>% 
+  mutate(type=case_when(trial<2 ~ "practice",
+                        trial<6 ~ "block 1",
+                        trial<10 ~ "block 2",
+                        trial<14 ~ "block 3")) %>% 
+  ggplot(aes(x = trial+1, y = word_count, color=type)) +
+    geom_jitter(width=.2, height=0, alpha=.7)+
+  #geom_line(aes(group=game))+
+ # facet_wrap(.~correct) + 
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+  scale_color_manual(values=time_palette)+
+  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  #geom_smooth(method="lm")+
+  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  theme_bw() +
+  xlab("Trial") +
+  ylab("Description length (words)")+
+  coord_cartesian(ylim=c(0,20), expand=F)+
+  #ylab("Percent trials with echoing \n by experimenter\n") +
+  #scale_x_continuous(breaks=seq(1,12,1)) +
+  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
+  theme(axis.text=element_text(size=10, color = "black"),
+        axis.title=element_text(size=12),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        legend.position = "none")
+
+plot_grid(acc_legend,acc_plot,speed_plot,word_plot, ncol=1, rel_heights = c(.2,.7,1,1), labels=c("","A", "B","C"), label_size = 18, label_x=0.12,label_y=.97 )
+
+ggsave(here("plot1.pdf"))
+```

### 2023-10-04 write-ups/CAMP6/abstract.tex
start camp abstract

\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\pagestyle{empty}
\usepackage[kerning=true]{microtype}
\usepackage{parskip}
\usepackage{sansmath}
\usepackage[font={small, bf}]{caption}
\usepackage[font={small}]{subcaption}
\usepackage{graphicx}
\usepackage{multicol}
\setlength{\abovecaptionskip}{0pt}
\setlength{\floatsep}{10pt}
\setlength{\textfloatsep}{0pt}
\setlength{\intextsep}{0pt}
\setlength{\belowcaptionskip}{0pt}
\setlength{\parindent}{5ex}
\setlength{\parskip}{0pt}
% Feel free to use additional packages for glosses, figures, whatnot.

% The next bit is for reserving sufficient space for authors,
% affiliations, and e-mail address.  No need to change for initial
% anonymous version.  For the final version, replace the
% \toggletrue{anonymous} with \togglefalse{anonymous} to de-anonymize.
\usepackage{etoolbox}
\newtoggle{anonymous}
\toggletrue{anonymous}

\renewcommand{\title}[1]{\textbf{#1}\\}
\newcommand{\authors}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}\\}
\newcommand{\email}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}}

\begin{document}

% First page:

% Insert title, authors, affiliations, and e-mail address in the next three lines:

\title{TODO TITLE}
\authors{Veronica Boyce, Ilaria Chen, Bobby Sparks, Malia Perez, Michael C. Frank} 
\email{vboyce@stanford.edu;  Stanford University}
\newline
%

% Intro
%TODO citations

\textbf{Methods:}

% Results

\textbf{Results:}

\textbf{Conclusion:}

\newpage

\begin{figure}

	\caption{Schematic of methods}
\end{figure}

	\begin{figure}
		\begin{minipage}{.5\textwidth}
			\captionof{figure}{Results}
			{	\includegraphics[width=\textwidth]{plot1.pdf}} 
			\begin{small}
				FOOBAR

			\end{small}

		\end{minipage}
		~~~
		\begin{minipage}{.5\textwidth}	
		\captionof{figure}{Similarity Results}
		{\includegraphics[width=\textwidth]{sims.pdf}} 
		\begin{small}
		Descriptions given by children were compared using sentence embeddings. Descriptions were most similar when provided by the same child for the same image. However, children described images more similarly to their partners than to children in other games.  
		\end{small}
		\end{minipage}
	\end{figure}

\begin{figure}\textbf{References:}

\end{figure}
\end{document}

### 2023-10-04 code/analysis_1.Rmd
write text

@@
 
-```{r}
+```{r, fig.width=4, fig.height=3}
 
@@
   left_join(labels1) %>% left_join(labels2) %>%
-  mutate(same_target=ifelse(target1==target2, "same_target", "diff_target"),  #label if target is same or not
+  mutate(same_target=ifelse(target1==target2, "Same target", "Different target"),  #label if target is same or not
   same_child=case_when(
-    game1!=game2 ~ "diff game",
-    (as.numeric(trial1)-as.numeric(trial2))%%2==1~"diff_child", #children alternate
-    T ~"same_child"),
+    game1!=game2 ~ "Different\n games",
+    (as.numeric(trial1)-as.numeric(trial2))%%2==1~"Different\n speaker", #children alternate
+    T ~"Same\n speaker"),
   type1=case_when(trial1<2 ~ "practice",
@@
     mutate(blocks=case_when(
-      type1=="block 1" & type2=="block 2"~ "1-2",
-      type1=="block 2" & type2=="block 1"~ "1-2",
-      type1=="block 1" & type2=="block 3"~ "1-3",
-      type1=="block 3" & type2=="block 1"~ "1-3",
-      type1=="block 2" & type2=="block 3"~ "2-3",
-      type1=="block 3" & type2=="block 2"~ "2-3"
+      type1=="block 1" & type2=="block 2"~ "Block 1 v 2   ",
+      type1=="block 2" & type2=="block 1"~ "Block 1 v 2   ",
+      type1=="block 1" & type2=="block 3"~ "Block 1 v 3   ",
+      type1=="block 3" & type2=="block 1"~ "Block 1 v 3   ",
+      type1=="block 2" & type2=="block 3"~ "Block 2 v 3   ",
+      type1=="block 3" & type2=="block 2"~ "Block 2 v 3   "
       )) %>%
-  ggplot(aes(x=str_c(same_target,"\n",same_child), y=sim, color=blocks)) +
-  geom_jitter(alpha=.01, color="black") +
+  ggplot(aes(x=str_c(same_child), y=sim, color=blocks)) +
+  facet_grid(.~same_target) +
+  theme_bw()+
+  #geom_jitter(alpha=.01, color="black") +
+  labs(x="", y="Cosine similarity", color="")+
+  scale_color_brewer(type="qual", palette="Dark2")+
+  scale_y_continuous(limits=c(0,1), expand=c(0,0))+
       stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
-  labs(x="Conditions compared")
+    guides( color = guide_legend( override.aes = list(linetype=0, size=1)))+
+  theme(strip.background =element_rect(fill="white"),
+        legend.position="bottom",
+        axis.title.x=element_blank())
+
+ggsave(here("write-ups/CAMP6/sims.pdf"))
 ```
@@
 
-ggsave(here("plot1.pdf"))
+ggsave(here("write-ups/CAMP6/plot1.pdf"))
 ```
+
+```{r}
+
+```

### 2023-10-04 write-ups/CAMP6/abstract.tex
write text

@@
 
-\textbf{Methods:}
+%TODO math
 
+%TODO crossref figures!
+
+As children acquire language, they must master many communication skills such as using metaphor and tracking the knowledge state of their partner.
+Adults demonstrate their communicative abilities in iterated reference games, where they describe novel figures to communicative partners with high accuracy.
+Over repetitions, description length reduces as partners come to share a nickname for the figure (Clark \& Wilkes-Gibbs 1986, Hawkins et al.
+2020).
+This paradigm is commonly studied with adults, but very little research on iterated communication games has been done with pre-literate children.
+
+Glucksberg et al.
+(1966) found that 5 year old children were not able to successfully communicate with each other about novel figures in a more complicated paradigm involving stacking blocks.
+However, children as young as 4 were successful at a two-choice, child-friendly iterated reference game when playing with their parent (Leung et al., 2020).
+Here, we test child-child dyads using a simplified version of Leung et al.
+2020 to determine if children can communicate with each other about novel figures and whether they show partner-specificity.
+
+\textbf{Methods:} Pairs of children played a matching game tablet game while sitting across the table from each other.
+On each trial, the "speaker" child would see two images, one in a box, and describe the boxed image (orally) to their partner.
+The "listener" child saw the same two images and was tasked with helping Smurfy (a stuffy) guess the matching target image by touching it on the screen.
+Children alternated roles trial by trial; to help children keep track of role, they passed Smurfy back and forth.
+We had 4 target images (a subset of stimuli from Leung et al.
+2020); distractors were drawn from the set of other targets.
+Children completed 2 practice trials with familiar shapes, and then 12 target trials (3 blocks of the 4 target images).
+We recruited 4-5 year old children from a university preschool; 20 pairs completed at least 8 out of 12 target trials and were included in the analysis.
+Children's utterances were transcribed from video recording, using Whisper and hand-checked.
 % Results
@@
 \textbf{Results:} 
+Children were above chance at selecting the target image, and accuracy barely increased over time ($\beta=$0.13 [-.04, .31]).
+Children's speed to selecting a target increased over time ($\beta=$-1.24 [-1.89, -.59]).
+Children tended to produce fairly short descriptions, and the length of descriptions did not decrease over time ($\beta=$0.11 [-.03, .25]).
 
-\textbf{Conclusion:}
+To compare the similarities of children's descriptions to those of their partners and other children; we embedded the utterances using S-BERT (Reimers \& Gurevych 2019) and used pairwise cosine similarity of embeddings as a proxy for utterance similarity.
+Children's utterances for an image are most similar to their own descriptions of the same image in different blocks.
+However, children's descriptions for an image are more similar to their partner's descriptions than the descriptions of children in other games.
+This indicates some level of partner specificity -- children are not just talking past each other, but are influenced by the descriptions used by their partner.
 
+\textbf{Caveats and Future Directions:} 
+Transcripts of the sessions revealed that the research assistant echoed the children's descriptions on the majority of trials (ex: "Mary says she sees a bear.
+Do you see a bear?").
+While this does not add any description not given by the children, repetition and prosody could influence children, and it does lend some "authority" to the description.
+Additionally, we have a small sample size of only 20 pairs of children, 4 images, and 3 repetitions.
+In a planned experiment 2, we will replicate this experiment, with a clearer experimenter script to avoid echoing, as well as additional tweaks to clarify the experiment for children.
+
+This preliminary study suggests that (contra Glucksberg et al.) 4-5 year old children *can* describe and select novel figures with each other at above chance accuracy.
+However, the initially short utterances and lack of reduction offers intriguing suggestions about possible differences between children's and adults approaches.
+These could be due to different linguistic abilities, social expectations, or different task difficulty.
 \newpage
@@
 \begin{figure}
-
+%TODO open materials
 	\caption{Schematic of methods}
+	TODO schematic; stim pictures ; repetition structure 
 \end{figure}
@@
 		\end{small}
+
+		TODO examples of the cute things kids say
 		\end{minipage}
@@
 \begin{figure}\textbf{References:}
-
+Leung, Hawkins, Yurovsky.
+Cogsci, 2020.
+$\cdot$ Glucksberg, Krauss, Weisberg.
+J of Expt Child Psychology, 1966.
+$\cdot$ Clark \& Wilkes-Gibbs.
+Cognition, 1986.
+$\cdot$ Hawkins, Frank, Goodman.
+Cognitive Science, 2020.
+$\cdot$ Reimers \& Gurevych.
+arXiv 2019.
 \end{figure}

### 2023-10-05 write-ups/CAMP6/abstract.tex
abstract
draft of abstract

@@
 \newtoggle{anonymous}
-\toggletrue{anonymous}
+\togglefalse{anonymous}
 
@@
 
-\textbf{Methods:} Pairs of children played a matching game tablet game while sitting across the table from each other.
+\textbf{Methods:} Pairs of children played a matching game on tablets while sitting across the table from each other.
 On each trial, the "speaker" child would see two images, one in a box, and describe the boxed image (orally) to their partner.
@@
 Children's speed to selecting a target increased over time ($\beta=$-1.24 [-1.89, -.59]).
-Children tended to produce fairly short descriptions, and the length of descriptions did not decrease over time ($\beta=$0.11 [-.03, .25]).
+Children generally produced short descriptions, and the length of descriptions tended to slightly increase over time ($\beta=$0.11 [-.03, .25]).
 
@@
 However, children's descriptions for an image are more similar to their partner's descriptions than the descriptions of children in other games.
-This indicates some level of partner specificity -- children are not just talking past each other, but are influenced by the descriptions used by their partner.
+This indicates some level of partner specificity -- children are not just talking past each other, but are influenced by what their partner has said.
 
@@
 Additionally, we have a small sample size of only 20 pairs of children, 4 images, and 3 repetitions.
-In a planned experiment 2, we will replicate this experiment, with a clearer experimenter script to avoid echoing, as well as additional tweaks to clarify the experiment for children.
+In a planned experiment 2, we will replicate this experiment, with a clearer experimenter script to avoid echoing and additional tweaks to clarify the procedure.
 
@@
 	\caption{Schematic of methods}
-	TODO schematic; stim pictures ; repetition structure 
+ \begin{center}{ \includegraphics[width=.8\textwidth, trim={8cm 6cm 1cm 6.4cm}, clip]{exp-diagram.pdf}}
+ 	\end{center}
 \end{figure}
-
 	\begin{figure}
@@
 			\begin{small}
-				FOOBAR
+			\textbf{	A.} Children's accuracy is at ceiling on practice trials and above chance on target trials.\textbf{ B.} Children speed up over time on target trials.
+\textbf{C.} Children's descriptions are usually short, but do not reduce in length over time.
 
@@
 However, children described images more similarly to their partners than to children in other games.
+Across games, the same image is described more similarly than different images.
 		\end{small}
 
-		TODO examples of the cute things kids say
+		\bigskip
+		\begin{small}
+\textbf{\centering Example successful descriptions:\\}
+		a person $\bullet$ a person walking $\bullet$ sad walking $\bullet$ a man with two foot and a diamond shaped tummy and a square head
+
+		a flying person $\bullet$
+		a superhero$\bullet$
+		somebody jumping $\bullet$ two wings with a square in the back with two feet
+
+		someone building a funnel $\bullet$
+		a server$\bullet$
+		a person holding up a plate $\bullet$ a man with one foot and a line and a square and a triangle
+
+		a crocodile tasting an animal in the water$\bullet$
+		a whale$\bullet$
+		someone flying a plane without a cover$\bullet$
+		a person lying on the ground
+
+		\end{small}
 		\end{minipage}
@@
 
-\begin{figure}\textbf{References:}
+\begin{figure}
+	\begin{small} \textbf{References:}
 Leung, Hawkins, Yurovsky.
 Cogsci, 2020.
-$\cdot$ Glucksberg, Krauss, Weisberg.
+$\bullet$ Glucksberg, Krauss, Weisberg.
 J of Expt Child Psychology, 1966.
-$\cdot$ Clark \& Wilkes-Gibbs.
+$\bullet$ Clark \& Wilkes-Gibbs.
 Cognition, 1986.
-$\cdot$ Hawkins, Frank, Goodman.
+$\bullet$ Hawkins, Frank, Goodman.
 Cognitive Science, 2020.
-$\cdot$ Reimers \& Gurevych.
+$\bullet$ Reimers \& Gurevych.
 arXiv 2019.
+\end{small}
 \end{figure}

### 2023-10-10 write-ups/CAMP6/abstract.tex
tweaked abstract

--- 2023-10-04 write-ups/CAMP6/abstract.tex
@@
 
-\title{TODO TITLE}
+\title{Ad-hoc referential expressions for novel shapes in 4-5 year old children}
 \authors{Veronica Boyce, Ilaria Chen, Bobby Sparks, Malia Perez, Michael C. Frank} 
@@
 
-\textbf{Methods:}
+%TODO math
 
+%TODO crossref figures!
+
+As children acquire language, they must master many communication skills such as using metaphor and tracking the knowledge state of their partner.
+Adults demonstrate these communicative abilities in iterated reference games, where they describe novel figures to partners with high accuracy.
+Over repetitions, description length reduces as pairs come to share nicknames for the figures (Clark \& Wilkes-Gibbs 1986, Hawkins et al.
+2020).
+This paradigm is commonly used with adults, but little used with pre-literate children.
+
+In  Glucksberg et al.
+(1966), 5 year old children were not able to successfully communicate with each other about novel figures in a more complicated paradigm involving stacking blocks.
+However, children as young as 4 were successful at a two-choice, child-friendly iterated reference game when playing with their parent (Leung et al., 2020).
+Here, we test child-child dyads using a simplified version of Leung et al.
+2020 to determine if children can communicate with each other and create partner-specific nicknames for novel figures in a simple iterated paradigm.
+
+\textbf{Methods:} Pairs of children played a matching game on tablets.
+On each trial, the ``speaker'' child would see two images, one in a box, and describe the boxed image (orally) to their partner (Fig 1).
+The ``listener'' child saw the same two images and was tasked with helping Smurfy (a stuffy) guess the matching target image by touching it on the screen.
+Children alternated roles trial by trial; the children passed Smurfy back and forth to track their roles.
+The 4 target images were a subset of stimuli from Leung et al.
+2020; distractors were drawn from the set of other targets.
+Children completed 2 practice trials with familiar shapes, and then 12 target trials (3 blocks of the 4 target images; Fig 1).
+We recruited 4-5 year old children from a university preschool; 20 pairs completed at least 8 out of 12 target trials and were included in the analysis.
+Utterances were transcribed from video recording using Whisper and then hand-checked.
 % Results
@@
 \textbf{Results:} 
+Children uses a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page).
+Children were 84\% accurate at selecting the target image, and accuracy barely increased over time ($\beta=$0.13 [-.04, .31]; Fig 2A).
+Children's speed to select a target increased over time ($\beta=$-1.24 [-1.89, -.59]; Fig 2B).
+Children generally produced short descriptions, and the length of descriptions tended to slightly increase over time ($\beta=$0.11 [-.03, .25]; Fig 2C).
 
-\textbf{Conclusion:}
+To compare the similarities of children's descriptions to those of their partners and other children, we embedded the utterances using S-BERT (Reimers \& Gurevych 2019) and used pairwise cosine similarity of embeddings as a proxy for utterance similarity.
+Children's utterances for an image are most similar to their own descriptions of the same image in different blocks (Fig 3).
+However, children's descriptions for an image are more similar to their partner's descriptions than the descriptions of children in other games.
+This indicates some level of partner specificity: children are not just talking past each other but are influenced by their partners prior descriptions.
 
+\textbf{Caveats and Future Directions:} 
+Transcripts of the sessions revealed that the research assistant echoed the children's descriptions on the majority of trials (ex: ``Mary says she sees a bear.
+Do you see a bear?'').
+While this did not introduce any new descriptions, the repetition could potentially lend ``authority'' to the description and influence children.
+Additionally, we have a limited sample of only 20 pairs of children, 4 images, and 3 repetitions.
+In a planned experiment 2, we will replicate this experiment, with a clearer experimenter script to avoid echoing and additional tweaks to clarify the procedure.
+
+This preliminary study suggests that (contra Glucksberg et al.) 4-5 year old children \textbf{can} describe and select novel figures with each other at above chance accuracy.
+However, the initially short utterances and lack of reduction offers intriguing suggestions about possible differences between children's and adult's approaches to to iterated reference games.
+These could be due to different linguistic abilities, social expectations, or differential task difficulty.
 \newpage
@@
 \begin{figure}
+%TODO open materials
+	\caption{Schematic of methods}
+ \begin{center}{ \includegraphics[width=.8\textwidth, trim={8cm 6cm 1cm 6.4cm}, clip]{exp-diagram.pdf}}
+ 	\end{center}
+ 	\begin{small}
+ 	20 pairs of children played a matching game on tablets while facing each other across a table. Each game consisted of 2 practice trials and 12 critical trials. 
+ 	\end{small}
 
-	\caption{Schematic of methods}
 \end{figure}
-
 	\begin{figure}
@@
 However, children described images more similarly to their partners than to children in other games.
+Across games, the same image is described more similarly than different images.
+		\end{small}
+
+		\bigskip
+		\begin{small}
+\textbf{\centering Example successful descriptions:\\}
+\smallskip
+		a person $\bullet$ a person walking $\bullet$ sad walking $\bullet$ a man with two foot and a diamond shaped tummy and a square head
+
+		\smallskip
+		a flying person $\bullet$
+		a superhero$\bullet$
+		somebody jumping $\bullet$ two wings with a square in the back with two feet
+		\smallskip
+
+		someone building a funnel $\bullet$
+		a server$\bullet$
+		a person holding up a plate $\bullet$ a man with one foot and a line and a square and a triangle
+		\smallskip
+
+		a crocodile tasting an animal in the water$\bullet$
+		a whale$\bullet$
+		someone flying a plane without a cover$\bullet$
+		a person lying on the ground
+
 		\end{small}
@@
 
-\begin{figure}\textbf{References:}
-
+\begin{figure}
+	\begin{small} \textbf{References:}
+Leung, Hawkins, Yurovsky.
+Cogsci, 2020.
+$\bullet$ Glucksberg, Krauss, Weisberg.
+J of Expt Child Psychology, 1966.
+$\bullet$ Clark \& Wilkes-Gibbs.
+Cognition, 1986.
+$\bullet$ Hawkins, Frank, Goodman.
+Cognitive Science, 2020.
+$\bullet$ Reimers \& Gurevych.
+arXiv 2019.
+\end{small}
 \end{figure}

### 2023-10-10 write-ups/CAMP6/abstract.tex
post mike comments

@@
 % Insert title, authors, affiliations, and e-mail address in the next three lines:
-
-\title{Ad-hoc referential expressions for novel shapes in 4-5 year old children}
+\noindent\title{4-5 year old children can successfully communicate using ad-hoc referential expressions}
 \authors{Veronica Boyce, Ilaria Chen, Bobby Sparks, Malia Perez, Michael C. Frank} 
@@
 2020).
-This paradigm is commonly used with adults, but little used with pre-literate children.
+This paradigm is commonly used with adults, but the consensus has been that tracking ad-hoc references in this way is too difficult for young children (in part due to notable failures in the literature; Glucksberg et al., 1966).
+We revisit this question, using a simplified method adapted from Leung et al.
+2020.
 
-In  Glucksberg et al.
-(1966), 5 year old children were not able to successfully communicate with each other about novel figures in a more complicated paradigm involving stacking blocks.
-However, children as young as 4 were successful at a two-choice, child-friendly iterated reference game when playing with their parent (Leung et al., 2020).
-Here, we test child-child dyads using a simplified version of Leung et al.
-2020 to determine if children can communicate with each other and create partner-specific nicknames for novel figures in a simple iterated paradigm.
-
+\bigskip
 \textbf{Methods:} Pairs of children played a matching game on tablets.
@@
 
+\bigskip
+
 \textbf{Results:} 
 Children uses a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page).
-Children were 84\% accurate at selecting the target image, and accuracy barely increased over time ($\beta=$0.13 [-.04, .31]; Fig 2A).
+Children were 84\% accurate at selecting the target image, and accuracy was relatively stable over time ($\beta=$0.13 [-.04, .31]; Fig 2A).
 Children's speed to select a target increased over time ($\beta=$-1.24 [-1.89, -.59]; Fig 2B).
@@
 This indicates some level of partner specificity: children are not just talking past each other but are influenced by their partners prior descriptions.
+
+\bigskip
 

### 2023-10-16 write-ups/CAMP6/abstract.tex
anonymize and submit!

@@
 \newtoggle{anonymous}
-\togglefalse{anonymous}
+\toggletrue{anonymous}
 
@@
 \textbf{Results:} 
-Children uses a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page).
+Children use a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page).
 Children were 84\% accurate at selecting the target image, and accuracy was relatively stable over time ($\beta=$0.13 [-.04, .31]; Fig 2A).
@@
 		a flying person $\bullet$
-		a superhero$\bullet$
+		a superhero $\bullet$
 		somebody jumping $\bullet$ two wings with a square in the back with two feet
@@
 		someone building a funnel $\bullet$
-		a server$\bullet$
+		a server $\bullet$
 		a person holding up a plate $\bullet$ a man with one foot and a line and a square and a triangle
@@
 
-		a crocodile tasting an animal in the water$\bullet$
-		a whale$\bullet$
-		someone flying a plane without a cover$\bullet$
+		a crocodile tasting an animal in the water $\bullet$
+		a whale $\bullet$
+		someone flying a plane without a cover $\bullet$
 		a person lying on the ground

### 2024-01-10 code/analysis_1.Rmd
tweaked abstract
graph tweaks for talk

@@
 
+```
+
+```{r}
+
+data_for_mods %>% ungroup() %>% summarize(correct=sum(correct.num)/n())
 ```
@@
 
-acc_w_legend <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+acc_dat <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
   mutate(type=case_when(trialNum<2 ~ "practice",
@@
                         trialNum<14 ~ "block 3"),
-         type=factor(type, levels=c("practice", "block 1", "block 2", "block 3"))) %>% 
-  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
-  stat_summary(fun.data="mean_cl_boot", geom="bar") +
-    stat_summary(fun.data="mean_cl_boot", geom="linerange") +
-  scale_fill_manual(values=time_palette)+
+         type=factor(type, levels=c("practice", "block 1", "block 2", "block 3")))
+  acc_w_legend <- ggplot(acc_dat, aes(x = trialNum+1, y = as.numeric(correct), color=type)) +
+  stat_summary(data=acc_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange") +
+  scale_color_manual(values=time_palette)+
   theme_bw() +
@@
   geom_hline(yintercept=.5, linetype="dashed")+
-    coord_cartesian(ylim=c(0,1), expand=F)+
+    coord_cartesian(ylim=c(0,1), expand=T)+
   theme(axis.text=element_text(size=10, color = "black"),
@@
 
-speed_plot <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
+speed_dat <- dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
   mutate(type=case_when(trialNum<2 ~ "practice",
@@
                         trialNum<14 ~ "block 3")) %>% 
-  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
-  ggplot(aes(x = trialNum+1, y = time/1000, color=type)) +
+  mutate(time=ifelse(is.na(time),180000,time)) #treat the one that timed out as being max time
+ speed_plot <-  ggplot(speed_dat, aes(x = trialNum+1, y = time/1000, color=type)) +
   geom_jitter(width=.2, height=0, alpha=.7)+
+  stat_summary(data=speed_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
     stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
@@
 
-word_plot <- transcript %>% filter(!is.na(description)) %>% group_by(game,speaker, trial) %>% 
+word_dat <- transcript %>% filter(!is.na(description)) %>% group_by(game,speaker, trial) %>% 
   summarize(words=str_c(description, sep=" ", collapse=", ")) %>% 
@@
                         trial<10 ~ "block 2",
-                        trial<14 ~ "block 3")) %>% 
-  ggplot(aes(x = trial+1, y = word_count, color=type)) +
-    geom_jitter(width=.2, height=0, alpha=.7)+
+                        trial<14 ~ "block 3"))
+word_plot <- 
+  ggplot(word_dat, aes(x = trial+1, y = word_count)) +
+    geom_jitter(aes(color=type), width=.2, height=0, alpha=.7)+
   #geom_line(aes(group=game))+
  # facet_wrap(.~correct) + 
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+      stat_summary(data=word_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange") +
   scale_color_manual(values=time_palette)+
@@
 
-plot_grid(acc_legend,acc_plot,speed_plot,word_plot, ncol=1, rel_heights = c(.2,.7,1,1), labels=c("","A", "B","C"), label_size = 18, label_x=0.12,label_y=.97 )
+plot_grid(acc_plot,speed_plot,word_plot, ncol=1, rel_heights = c(1,1,1)) #labels=c("","A", "B","C"), label_size = 18, label_x=0.12,label_y=.97 )
 
-ggsave(here("write-ups/CAMP6/plot1.pdf"))
+ggsave(here("write-ups/CAMP6/plot1.png"))
 ```
@@
 ```{r}
+plot_grid(acc_legend)
+
+ggsave(here("write-ups/CAMP6/legend.pdf"), width=6, height=2)
 

### 2024-08-08 code/expt_1/analysis-pilot.Rmd
reorg more

--- 2023-08-19 code/analysis-pilot.Rmd
@@
 ```
+#accuracy of each image over time
+```{r}
+accuracy <- filtered_data %>% 
+  mutate(image = case_when(
+  target == "/experiment/B1.jpg" ~ "Walking",
+  target == "/experiment/D1.jpg" ~ "Lying",
+  target == "/experiment/E1.jpg" ~ "Jumping",
+  target == "/experiment/I1.jpg" ~ "Holding",
+)) %>% 
+  filter(!index %in% c(0, 1)) %>%
+  mutate(block=(index-2)%/%4+1) %>% 
+  group_by(block, image) %>% 
+  summarize(total=n(),
+            correct=sum(countCorrect)) %>%
+  mutate(fraction = correct/total*100, percent = signif(fraction, 2))
+
+ggplot(accuracy, aes(x = block, y = percent, color = image)) +
+  geom_line(aes(group=image)) +
+  #geom_smooth()+
+  labs(color = "Tangram", x = "\nOccurence", y = "Percent accuracy\n") +
+  scale_x_continuous(breaks=seq(1,3,1)) +
+  scale_y_continuous(breaks=seq(0,100,10), limits = c(0,100)) +
+  theme_classic() +
+  theme(aspect.ratio = 9/9,
+        axis.title=element_text(size=14,face="bold"),
+        axis.text=element_text(size=9, color = "black"),
+        legend.title=element_text(size=13,face="bold"),
+        legend.text=element_text(size=9),
+legend.position="right")
+
+ggsave("AccuracyOccurrence.png")
+```
 

### 2024-08-08 code/expt_1/analysis_1.Rmd
reorg more

--- 2024-01-10 code/analysis_1.Rmd
@@
          type=factor(type, levels=c("practice", "block 1", "block 2", "block 3")))
-  acc_w_legend <- ggplot(acc_dat, aes(x = trialNum+1, y = as.numeric(correct), color=type)) +
-  stat_summary(data=acc_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange") +
+  acc_w_legend <- ggplot(acc_dat, aes(x = trialNum+1, y = as.numeric(correct))) +
+ # stat_summary(data=acc_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
+    geom_smooth(data=acc_dat |> filter(type!="practice"), method="lm", color="black", se=F)+
+    stat_summary(aes(color=type), fun.data="mean_cl_boot", geom="pointrange") +
   scale_color_manual(values=time_palette)+
@@
  speed_plot <-  ggplot(speed_dat, aes(x = trialNum+1, y = time/1000, color=type)) +
-  geom_jitter(width=.2, height=0, alpha=.7)+
-  stat_summary(data=speed_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
+  geom_jitter(width=.2, height=0, alpha=.3, color="grey")+
+       geom_smooth(data=acc_dat |> filter(type!="practice"), method="lm", color="black", se=F)+
+
+    stat_summary(fun.data="mean_cl_boot", geom="pointrange") +
   scale_color_manual(values=time_palette)+
@@
 word_plot <- 
-  ggplot(word_dat, aes(x = trial+1, y = word_count)) +
-    geom_jitter(aes(color=type), width=.2, height=0, alpha=.7)+
+  ggplot(word_dat, aes(x = trial+1, y = word_count, color=type)) +
+    geom_jitter(aes(color=type), width=.2, height=0, alpha=.3, color="grey")+
   #geom_line(aes(group=game))+
  # facet_wrap(.~correct) + 
-      stat_summary(data=word_dat |> filter(type!="practice"), fun.data="mean_cl_boot", geom="line", color="black") +
+       geom_smooth(data=word_dat |> filter(type!="practice"), method="lm", color="black", se=F)+
+
     stat_summary(fun.data="mean_cl_boot", geom="pointrange") +

### 2024-08-08 code/expt_1/data_prep.Rmd
reorg more

--- 2023-10-02 code/data_prep.Rmd
@@
 ```
+
+# Sbert data prep
+
+```{r}
+transcript <- read_csv(here("data/timed_transcript.csv"))
+
+dat <- read_csv(here("data/clean_data.csv")) %>% select(trial=trialNum, game, target, distractor)
+
+transcript %>% filter(!is.na(description)) %>% group_by(trial, game) %>% 
+  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% left_join(dat) %>% write_csv(here("data/pre_sbert.csv"))
+
+```
+
+```{r}
+library(reticulate)
+np <- import("numpy")
+mat = np$load(here('data/post_sbert.npy'))
+saveRDS(mat,here('data/post_sbert.RData'))
+```
+
+```{r}
+sbert <- read_csv(here("data/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/post_sbert.RData")) %>% 
+  as_tibble())
+
+### helper funcs
+get_sim_matrix = function(df, F_mat, method = 'cosine') {
+  feats = F_mat[df$feature_ind,]
+  if(method == 'cor') {
+    return(cor(t(feats), method = 'pearson'))
+  } else if (method == 'euclidean') {
+    return(as.matrix(dist(feats, method = 'euclidean')))
+  } else if (method == 'cosine') {
+    return(as.matrix(lsa::cosine(t(feats))))
+  } else {
+    stop(paste0('unknown method', method))
+  }
+}
+
+# note this does de-duplicated version
+flatten_sim_matrix <- function(cormat, ids) {
+  ut <- upper.tri(cormat)
+  data.frame(
+    dim1 = ids[row(cormat)[ut]],
+    dim2 = ids[col(cormat)[ut]],
+    sim  = as.numeric(cormat[ut])
+  ) %>%
+    mutate(dim1 = as.character(dim1),
+           dim2 = as.character(dim2))
+}
+
+make_across_df <- function(M_mat, F_mat, method) {
+  M_mat %>%
+    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
+                          as.character(.$combinedId)))
+}
+
+### funcs
+  F_mat <- sbert%>% select(starts_with("V")) %>% as.matrix() #Features
+  M_mat <- sbert %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
+
+similarities <- M_mat %>%
+    mutate(combinedId=str_c(game,trial,sep="_")) %>%
+    make_across_df(F_mat, 'cosine') %>%
+    separate(dim1, into=c("game1","trial1"), convert=T, sep="_") %>%
+    separate(dim2, into=c("game2","trial2"), convert=T, sep="_") %>%
+    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("data/similarities.rds"))
+
+```

### 2024-08-08 code/expt_1/leung_comparison.Rmd
reorg more

---
title: "Leung data comparison"
output:
  html_document:
    df_print: paged
---
```{r prep, echo=F, message=F, warning=F}
library(here)
library(tidyverse)

knitr::opts_chunk$set(
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
theme_set(theme_bw())
```

```{r, eval=F}
all_4yos <- list.files(here("leung_data/4yos")) |> map_df(~read_csv(here("leung_data/4yos/",.))) |> bind_rows() |> 
  write_csv(here("leung_data/4yos_uncoded.csv"))

all_adults <- list.files(here("leung_data/adult")) |> map_df(~read_csv(here("leung_data/adult/",.))) |> bind_rows() |> 
  write_csv(here("leung_data/adults_uncoded.csv"))

```

** note **: in the adult data, for sub id 104, trial 9 (target D1), it had been that two lines had selections marked. I believe the first should have had selection NA b/c it has text as well and this has been fixed so we don't have duplicates.

```{r}
kids <- read_csv(here("leung_data/4yos_coded_filtered.csv")) |> 
  filter(target %in% c("B1", "D1", "E1", "I1"))

adults <- read_csv(here("leung_data/adults_coded_filtered.csv"))|> 
  filter(target %in% c("B1", "D1", "E1", "I1"))

kids_selection <- kids |> filter(!is.na(selection)) |> select(listener=selection, target, trial, subid, type)

adults_selection <- adults |> filter(!is.na(selection)) |> select(listener=person, target, trial, subid, type)

adults_tagged <- adults |> left_join(adults_selection) |> 
  mutate(role=ifelse(person==listener, "listener", "speaker")) |> 
  mutate(words=str_count(description, "\\S+"),
         words=ifelse(is.na(words), 0, words),
         all_words=str_count(utterance, "\\S+"),
         all_words=ifelse(is.na(all_words), 0, all_words),
         block=(trial-1)%/%10)

kids_tagged <- kids |> 
  left_join(kids_selection) |> 
  mutate(role=ifelse(person==listener, "listener", "speaker")) |> 
   mutate(words=str_count(description, "\\S+"),
         words=ifelse(is.na(words), 0, words),
         all_words=str_count(utterance, "\\S+"),
         all_words=ifelse(is.na(all_words), 0, all_words),
         block=(trial-1)%/%10)

```

```{r, fig.height=3, fig.width=6}

adults_summary <- adults_tagged  |> 
  group_by(subid, role, block, target) |> 
  summarize(words=sum(words),
            all_words=sum(all_words))

  # ggplot(adults_summary, aes(x=block, y=words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,20))+
  # theme_classic()

  # ggplot(adults_summary, aes(x=block, y=all_words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,40))+
  # theme_classic()

  kids_summary <- kids_tagged  |> 
    filter(!is.na(role)) |> 
  group_by(subid, person, role, block, target) |> 
  summarize(words=sum(words),
            all_words=sum(all_words))

  #   ggplot(kids_summary, aes(x=block, y=words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", color="black")+
  #   coord_cartesian(ylim=c(0,20))+facet_grid(~ person)
  # 
  # 
  # ggplot(kids_summary, aes(x=block, y=all_words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,40))+facet_grid(~person)

  all <- adults_summary |> mutate(person="adult-to-adult") |> 
    bind_rows(kids_summary) |> 
    mutate(person=case_when(
      person=="child" ~ "child-to-parent",
      person=="parent" ~ "parent-to-child",
      T ~ person
    ))

  time_palette <- c("practice"="grey","1"="#E41A1C","2"="#377EB8", "3"="#4DAF4A", "4"="purple")#TODO FIX
      ggplot(all |> filter(role=="speaker"), aes(x=block+1, y=words, color=as.factor(block+1)))+
    geom_jitter(width=.2, height=0, alpha=.4)+
    stat_summary(fun.data="mean_cl_boot", color="black")+
                stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
          scale_color_manual(values=time_palette)+
        theme(legend.position="none")+
    coord_cartesian(ylim=c(0,20), expand=F)+facet_grid(~ person)+
        labs(title="Re-analysis of Leung et al 2023 data", x="Block", y="Description length")

ggsave(here("foo.png"))

  ggplot(all |> filter(role=="speaker"), aes(x=block+1, y=all_words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
        stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,40))+facet_grid(~person)+
    labs(title="Total Words from speaker", x="Block", y="Total words")

        ggplot(all |> filter(role=="listener"), aes(x=block+1, y=words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", color="black")+
                stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,20))+facet_grid(~ person)+
        labs(title="Referential words from listener", x="Block", y="Referential words")

  ggplot(all |> filter(role=="listener"), aes(x=block+1, y=all_words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
        stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,40))+facet_grid(~person)+
    labs(title="Total Words from listener", x="Block", y="Total words")

```

### SBERT

```{r eval=F}
all <- adults_tagged |> bind_rows(kids_tagged) |> filter(!is.na(description)) |> filter(role=="speaker") |> 
  select(subid, target, person, description, block) %>%
  group_by(subid, block, target, person) |> filter(!is.na(description)) %>% 
  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% write_csv(here("leung_data/pre_sbert.csv"))
```

```{r eval=F}
library(reticulate)
np <- import("numpy")
mat = np$load(here('leung_data/post_sbert.npy'))
saveRDS(mat,here('leung_data/post_sbert.RData'))
```

```{r}
sbert <- read_csv(here("leung_data/pre_sbert.csv")) %>% bind_cols(read_rds(here("leung_data/post_sbert.RData")) %>% 
  as_tibble())

### helper funcs
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}

### funcs
  F_mat <- sbert%>% select(starts_with("V")) %>% as.matrix() #Features
  M_mat <- sbert %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())

similarities <- M_mat %>%
    mutate(combinedId=str_c(subid,block,person,target,sep="_")) %>%
    make_across_df(F_mat, 'cosine') |> 
    separate(dim1, into=c("game1","block1","person1", "target1"), convert=T, sep="_") %>%
    separate(dim2, into=c("game2","block2","person2", "target2"), convert=T, sep="_") %>%
    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("leung_data/similarities.rds"))

```

```{r, fig.width=5, fig.height=3}
sims <- read_rds(here("leung_data/similarities.rds")) |> as_tibble()

sims %>%  filter(!is.na(sim)) %>% 
  mutate(same_target=ifelse(target1==target2, "Same target", "Different target"),  
  same_child=case_when(
    game1!=game2 ~ "Different\n games",
    person1!=person2 ~ "Different\n speakers",
    T ~"Same\n speaker"),
  roles=case_when(
    person1 %in%c("left", "right") & person2 %in% c("left", "right") ~ "adult-adult",
    person1 == "parent" & person2=="parent" ~ "parent-parent",
    person1 == "child" & person2=="child" ~ "child-child",
    person1 == "child" & person2=="parent" ~ "parent-child",
    person1 == "parent" & person2=="child" ~ "parent-child",
    T ~ "bad"
  ) )|> 
    filter(roles!="bad") |> # don't want mix between adult & parent-child
  filter(block1!=block2) |> # can only compare across blocks in some cases, so only cross ever
  ggplot(aes(x=str_c(same_child), y=sim, color=roles)) +
  facet_grid(.~same_target) +
  theme_bw()+
  labs(x="", y="Cosine similarity", color="")+
  scale_color_brewer(type="qual", palette="Dark2")+
  scale_y_continuous(limits=c(0,1), expand=c(0,0))+
      stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
    guides( color = guide_legend( override.aes = list(linetype=0, size=1)))+
  theme(strip.background =element_rect(fill="white"),
        legend.position="bottom",
        axis.title.x=element_blank())

ggsave(here("bar.png"))

```

### 2024-12-11 code/expt_1/analysis_1.Rmd
start cogsci

@@
 correctness <- dat %>% select(game, trial=trialNum, target, distractor, correct)
+
 transcript %>% filter(!is.na(description)) %>% group_by(game,speaker, trial) %>% 

### 2024-12-11 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
start cogsci

---
title: "TODO"
bibliography: library.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, \\Stanford University \And {\large \bf Bobby Sparks (TODO email) } Department of Psychology, \\Stanford University
    \AND {\large \bf Yannick Mofor (TODO email)} \\ TODO affiliation, Stanford University \And {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, \\ Stanford University}

abstract: >
    TODO abstract
    Communication is important and stuff and kids need to learn it to. One task that illustrates the rich communicative capabilities is iterated reference games, where adults jointly converge on mutually understandable names for novel objects. Some prominent early experiments where children did not succeed led to the belief that children could not do this task. TODO FIX. However, more recent work with lessened task demands suggests that children may be more capable than previously thought. Here, we test children's communicative adaption to a partner with iterated reference games played by pairs of 4-5 year old preschoolers. Across a total of N pairs of children, we find children can accurately pick out referents for each other and that they are sensitive to the descriptions of their partners. TODO wrap up sentence.

keywords: >
    TODO keywords

output: cogsci2024::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)

library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(here)
library(brms)
library(rstan)
library(rstanarm)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

expt_1 <- "data/expt_1"
expt_2 <- "data/expt_2"
mod_loc <- "code/models"
time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

# experiment 1 imports
transcript_1 <- read_csv(here(expt_1,"timed_transcript.csv"))
dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial=trialNum)
exclusion_1 <- transcript_1 %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game,  echo,) %>% unique() %>% mutate("S_talked"=1)

words_1 <- transcript_1 |> filter(!is.na(description)) |> mutate(words=description |> str_count("\\S+")) |> 
  filter(role=="S") |> 
  group_by(game, trial) |> 
  summarize(words=sum(words))

clean_1 <- dat_1 %>% left_join(exclusion_1) %>% filter(S_talked==1) |> left_join(words_1) |>  mutate(type=case_when(trial<2 ~ "practice",
                        trial<6 ~ "block 1",
                        trial<10 ~ "block 2",
                        trial<14 ~ "block 3"))

# sims_1 <- read_rds(here(expt_1,"similarities.rds")) |> 
#     as_tibble() 
#   inner_join(clean_1|> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
#   inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
#   filter(target1 == target2) |>
#   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
#   mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
#          same_game=(game1==game2) |> as.numeric(),
#         later = ifelse(block1 > block2, block1, block2), 
#         earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)

# expt 2 imports
transcript_2 <- read_csv(here(expt_2, "transcripts.csv"))
link <- read_csv(here(expt_2, "link_transcripts.csv"))

exclude_2 <- transcript_2 |>
  filter(!is.na(echo)) |>
  select(gameId, trial)

data_2 <- read_csv(here(expt_2, "clean_data.csv")) |>
  rename(trial = trialNum) |>
  anti_join(exclude_2)

transcript_2 <- transcript_2 |>
  anti_join(exclude_2) |>
  filter(!is.na(description))

words_2 <- transcript_2 |> mutate(words=description |> str_count("\\S+")) |> 
  filter(role=="S") |> 
  group_by(gameId, gameConfig, trial) |> 
  summarize(words=sum(words))

clean_2 <- data_2 |> left_join(words_2) |>
  filter(!is.na(words)) |> 
   mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |> 
  mutate(correct=as.numeric(correct))

# sims_2 <- read_rds(here(expt_2, "similarities.rds")) |>
#   as_tibble() |>
#   inner_join(clean_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
#   inner_join(clean_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
#   filter(target1 == target2) |>
#   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
#   mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
#          same_game=(game1==game2) |> as.numeric(),
#         later = ifelse(block1 > block2, block1, block2), 
#         earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
```

# TODOs

* Check transcription for expt 1 and 2 especially for ref expression

* Rerun sbert stuff

* Run models

* Double check pre-regs and note deviations

* Include fun examples of the things kiddos said

# Pre-regs
https://osf.io/kcv8j

https://osf.io/y2dax

Analyses -- Expt 1

This is exploratory research, and we will explore the following questions graphically:
* how accurate are children over time
* how many words of description does the speaker produce over time
* do children speed up in their selections over time
* do children match the descriptions their partner uses

We will also analyze the transcripts to look at children’s turn-taking structure and how they understand each other.

We plan to run the model:
DV  ~ trial_num + (trial_num|dyad) + (1|target) 
for the DVs: 
[x] accuracy (logistic model), 
[ ] speed (linear model) and 
[ ] number of words of speaker description (linear model ).

Analyses -- Expt 2

Primarily, we aim to replicate the results of experiment 1, using the following models.

[x] Logistic model of accuracy: correct ~ trialNum + (trialNum|game) + (1|target) (prior normal(0,1) for beta and sd, lkj(1) for correlation)

[ ] Linear model of description length: words ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(5,10), beta and sd priors of normal(0,5), and correlation of lkj(1))

[ ] Speed: time_seconds ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(60,100), beta and sd priors of (0,20) and correlation of lkj(1))

We will interpret the fixed effect of trial number as being about changes in referring over time.

For utterance similarities, we will embed the describer’s descriptions using S-BERT and use cosine similarity as a proxy for utterance similarity.

We are unsure whether 4 blocks will be sufficient to see potential change over time, so we will 
[ ] * compare similarities of utterances between different games v within one game v within one speaker: sim ~ same_game + same_speaker + (1|target) 
[ ] * compare similarities of utterances to the last utterance within a game: sim_to_last ~ earlier_block + same_speaker + (1|game) + (1|target)
[ ] * compare similarities of utterances to the next utterance: sim_to_next ~ earlier_block + same_speaker + (1|game) + (1|target)
[ ] * compare the similarity level of descriptions from the same block in different games: sim ~ block + (1|target)
Priors for the above similarity models: normal(.5,.2) for intercept, normal(0,.1) for
beta, and normal(0,.05) for sd.

We will also qualitatively explore children’s turn-taking structure and how they understand each other.

# Introduction

TODO introduction

# Experiment 1 Methods

Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand. Young children can be very sensitive to task demands and cognitive load that can hide early abilities (TODO cite something for this), so we wanted to use a simple paradigm with experimenter scaffolding. We adapted the experimental framework from @leung, but further simplified it by reducing the total pool of targets and the number of trials children completed.

TODO link to pre-registration

## Participants
4 and 5 year old children were recruited from a university nursery school during the school day. Children played with another child from the same class. TODO any exclusions Experiment 1 was conducted between June and August 2023. Children were only included in analysis if they completed at least 8 of the critical trials. We had XX complete games and YY incomplete, but included games.

## Materials
For the target stimuli, we used four of the ten tangram images from @leung, chosen based on visual dissimilarity. The matching game was coded as an Empirica experiment (CITATION) and run on two tablets locked in a kiosk mode so children could access other websites or apps.

## Procedure
Once a pair of children agreed to play the game, they were taken to a quiet testing room where procedure was explained to them. Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game. Children sat across a table from each other, each with a tablet in front of them. On each trial, one child saw two images, one of them in a black box, and was asked to "say what they saw" in the black box so their partner (and Smurfy) could tap the corresponding image. The guesser saw the same two images (in a randomized order), but with neither boxed. Upon tapping an image, both children recieved feedback in form of a smiley or frowney face and an audible sound. After each trial, children's roled switched, demarcated by passing Smurfy back and forth.

Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 targets. Targets were randomly paired with another of the critical images as the foil.

The experimenters running the game did not volunteer descriptions, but did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements. The entire interaction was video-recorded.

TODO DIAGRAM OF EXPERIMENT + Picture of children playing the game as top bar

## Data processing

Children's selections and the time to selection were recorded from the experiment software. Children's descriptions were transcribed from the video-recording, using Whisper (CITE) for the first pass and then hand-corrected by experimenters. Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used.

# Experiment 1 Results

## Accuracy

```{r}
 clean_1 |> 
  ggplot(aes(x = trial+1, y = as.numeric(correct), fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

```{r}
acc_priors <- c(
set_prior("normal(0,1)", class="b"),
set_prior("normal(0,1)", class="sd"),
set_prior("lkj(1)", class="cor")
)

acc_mod_data <- clean_1 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
                                  trial.num=trial-2) |> filter(type!="practice")
acc_mod <- brm(correct.num ~ trial.num + (trial.num|game) + (1|target),
               family = bernoulli(link = "logit"),
               data=acc_mod_data,
               file = here(mod_loc, "acc_1.rds"),
               prior=acc_priors,
               control = list(adapt_delta=.95))
```

## Speed

```{r}
 clean_1 |> 
  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
  ggplot(aes(x = trial+1, y = time/1000, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
  #geom_line(aes(group=game))+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Time to answer in seconds")+
  coord_cartesian(ylim=c(0,60))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```

## Description length

```{r}

clean_1 |> 
  ggplot(aes(x = trial+1, y = words, color=type)) +
    geom_jitter(width=.2, height=0, alpha=.5)+
  #geom_line(aes(group=game))+
 # facet_wrap(.~correct) + 
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  #coord_cartesian(ylim=c(0,60))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

## Convergence

```{r, fig.width=4, fig.height=3}

# sims_1 %>%  filter(!is.na(sim)) %>% 
#   filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
#   left_join(labels1) %>% left_join(labels2) %>%
#   mutate(same_target=ifelse(target1==target2, "Same target", "Different target"),  #label if target is same or not
#   same_child=case_when(
#     game1!=game2 ~ "Different\n games",
#     (as.numeric(trial1)-as.numeric(trial2))%%2==1~"Different\n speaker", #children alternate
#     T ~"Same\n speaker"),
#   type1=case_when(trial1<2 ~ "practice",
#                         trial1<6 ~ "block 1",
#                         trial1<10 ~ "block 2",
#                         trial1<14 ~ "block 3"),
#          type2=case_when(trial2<2 ~ "practice",
#                         trial2<6 ~ "block 1",
#                         trial2<10 ~ "block 2",
#                         trial2<14 ~ "block 3")) %>% 
#   filter(type1!=type2) %>% # can only compare across blocks in some cases, so only cross ever
#     mutate(blocks=case_when(
#       type1=="block 1" & type2=="block 2"~ "Block 1 v 2   ",
#       type1=="block 2" & type2=="block 1"~ "Block 1 v 2   ",
#       type1=="block 1" & type2=="block 3"~ "Block 1 v 3   ",
#       type1=="block 3" & type2=="block 1"~ "Block 1 v 3   ",
#       type1=="block 2" & type2=="block 3"~ "Block 2 v 3   ",
#       type1=="block 3" & type2=="block 2"~ "Block 2 v 3   "
#       )) %>%
#   ggplot(aes(x=str_c(same_child), y=sim, color=blocks)) +
#   facet_grid(.~same_target) +
#   theme_bw()+
#   #geom_jitter(alpha=.01, color="black") +
#   labs(x="", y="Cosine similarity", color="")+
#   scale_color_brewer(type="qual", palette="Dark2")+
#   scale_y_continuous(limits=c(0,1), expand=c(0,0))+
#       stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
#     guides( color = guide_legend( override.aes = list(linetype=0, size=1)))+
#   theme(strip.background =element_rect(fill="white"),
#         legend.position="bottom",
#         axis.title.x=element_blank())

```

```{r}
# sims_1 %>%  filter(!is.na(sim)) %>% 
#   filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
#   left_join(labels1) %>% left_join(labels2) %>%
#   mutate(same_target=ifelse(target1==target2, "same_target", "diff_target"),  #label if target is same or not
#   same_child=case_when(
#     game1!=game2 ~ "diff game",
#     (as.numeric(trial1)-as.numeric(trial2))%%2==1~"diff_child", #children alternate
#     T ~"same_child"),
#   type1=case_when(trial1<2 ~ "practice",
#                         trial1<6 ~ "block 1",
#                         trial1<10 ~ "block 2",
#                         trial1<14 ~ "block 3"),
#          type2=case_when(trial2<2 ~ "practice",
#                         trial2<6 ~ "block 1",
#                         trial2<10 ~ "block 2",
#                         trial2<14 ~ "block 3")) %>% 
#   filter(type1==type2) %>% # can only compare across blocks in some cases, so only cross ever
#   ggplot(aes(x=str_c(same_target,"\n",same_child), y=sim, color=type1)) +
#   geom_jitter(alpha=.01, color="black") +
#       stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
#   labs(x="Conditions compared")
```

# Experiment 1 Discussion

# Experiment 2 Methods

After the first experiment, we wanted to run another study using the same paradigm, but fixing issues that arose during experiment 1. We also tightened the experimenter script to reduce the possibility that the experimenters were influencing children's descriptions and understanding of descriptions. Given that most pairs in experiment 1 were able to play the game fairly quickly, we also added a 4th experimental block to be able to look for change over time.

As Experiment 2 was very similar to Experiment 1, here we note the differences.

## Participants
Experiment 2 was run between March and August of 2024, at the same university preschool. No children participated in both experiments. ZZ pairs of children completed all 16 critical trials, and YY pairs of children completed between 8 and 16 critical trials.

## Materials
The same 4 critical images were used as in experiment 1, although this time, children saw these images 4 times. In response to some children struggling with the abrupt switch from nameable to non-nameable shapes, we introduced more practice trials for experiment 2. Additionally, before the practice trials, we added a "bubble popping" exercise to familiarize children with how to tap the tablet appropriately (a problem in experiment 1).

## Procedure
The procedure was much the same. The smurfy puppet was swapped out for a more attractive smurfy stuffed animal. The experimental script was tightened. TODO link to materials for where the script is.

## Data processing
Data was processed in the same way as experiment 1.

# Experiment 2 Results

## Accuracy

```{r}
clean_2 |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_fill_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, .1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  theme(
    aspect.ratio = 4 / 5,
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```

```{r}
acc_priors <- c(
set_prior("normal(0,1)", class="b"),
set_prior("normal(0,1)", class="sd"),
set_prior("lkj(1)", class="cor")
)

acc_mod_data <- clean_2 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
                                  trial.num=trial-4) |> filter(type!="practice")
acc_mod <- brm(correct.num ~ trial.num + (trial.num|gameId) + (1|target),
               family = bernoulli(link = "logit"),
               data=acc_mod_data,
               file = here(mod_loc, "acc_2.rds"),
               prior=acc_priors,
               control = list(adapt_delta=.95))
```

## Speed

```{r}
clean_2 |> ggplot( aes(x = trial + 1, y = time / 1000, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  theme(legend.position = "none")
```

## Description length

```{r}

clean_2 |> ggplot( aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = clean_2|> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  theme(legend.position = "none")
```

## Convergence

```{r}
# sims_2 |>
#   mutate(source = case_when(
#     speaker1 == speaker2 ~ "same speaker",
#     game1 == game2 ~ "same game",
#     T ~ "different games"
#   )) |>
#   ggplot(aes(x = source, y = sim, color = target1)) +
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
#   labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
#   theme(legend.position = "bottom")

```

```{r}
# sims_2 |>
#   filter(game1 == game2) |>
#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
#   filter(later == 4) |>
#   mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
#   ggplot(aes(as.character(earlier), sim, color = target1)) +
#   geom_point(alpha = .1, position = position_dodge(width = .2)) +
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
#   facet_wrap(~same_speaker) +
#   labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
#   theme(legend.position = "bottom")
```

```{r}
# sims_2 |>
#   filter(game1 == game2) |>
#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
#   filter(later == earlier + 1) |>
#   mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
#   ggplot(aes(as.character(earlier), sim, color = target1)) +
#   geom_point(alpha = .1, position = position_dodge(width = .2)) +
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
#   facet_wrap(~same_speaker) +
#   labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
#   theme(legend.position = "bottom")
```

```{r}
# sims_2 |>
#   filter(game1 != game2) |>
#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
#   filter(later == earlier) |>
#   ggplot(aes(as.character(earlier), sim, color = target1)) +
#   geom_point(alpha = .01, position = position_dodge(width = .2)) +
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
#   labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
#   theme(legend.position = "bottom")
```

# General Discussion

Summary of experiments

Limitations. The population of children at university nursery schools is non-representative, and the set of materials we used was also not that varied. This set of tangram images may be easier to distinguish and refer to than some sets used with adults, leading to overall shorter utterances. Probably shouldn't say based on this that children can "do reference" at age 4, but it is evidence that under supportive circumstances, a number of children at this age are able to.

We also specifically target the construction of referring expressions that can be jointly understood. There are other parts of the coordination where help was provided in children seemed stuck or confused, such as when to make a choice or ask for more information.

Broader implications

This work (along with other work on children's referential communicaiton) suggests that there's a more gradual development
Has implications for how we think about children's language development. There's debate over how that is ordered and whether communication/pragmatics is a final stage, or how all the stages are boostrapped (CITATIONS). This early ability to use language for communicative purposes is more consistent with the early pragmatics viewpoint.

Suggestive of a gradual development where children's capabilities are increasing for a wide amount of childhood, as their working memory capacity and executive function improve and they are able to better track other's states of knowledge and keep track of wider arrays of images.

<!--# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.-->

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2024-12-13 code/expt_1/data_prep.Rmd
check all transcripts and rerun pre-processing

@@
 
-data_source <- "data/rounds.csv"
-player_source <- "data/players.csv"
+data_source <- "data/expt_1/rounds.csv"
+player_source <- "data/expt_1/players.csv"
 
@@
 
-games <- read_csv(here("data/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+games <- read_csv(here("data/expt_1/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
 
@@
 wrangle_transcript <- function(game_num){
- read_csv(here("data/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
+ read_csv(here("data/expt_1/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
   select(start, end, text, speaker, trial, role, echo, description, backchannel) |> 
@@
 
-stages <- read_csv(here("data/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)
+stages <- read_csv(here("data/expt_1/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)
 
@@
   rename(start_time=start, end_time=end) %>% 
-  write_csv(here("data/timed_transcript.csv"))
+  write_csv(here("data/expt_1/timed_transcript.csv"))
 ```
@@
 ```{r}
-transcript <- read_csv(here("data/timed_transcript.csv"))
+transcript <- read_csv(here("data/expt_1/timed_transcript.csv"))
 
-dat <- read_csv(here("data/clean_data.csv")) %>% select(trial=trialNum, game, target, distractor)
+dat <- read_csv(here("data/expt_1/clean_data.csv")) %>% select(trial=trialNum, game, target, distractor)
 
 transcript %>% filter(!is.na(description)) %>% group_by(trial, game) %>% 
-  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% left_join(dat) %>% write_csv(here("data/pre_sbert.csv"))
+  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% left_join(dat) %>% write_csv(here("data/expt_1/pre_sbert.csv"))
 
@@
 np <- import("numpy")
-mat = np$load(here('data/post_sbert.npy'))
-saveRDS(mat,here('data/post_sbert.RData'))
+mat = np$load(here('data/expt_1/post_sbert.npy'))
+saveRDS(mat,here('data/expt_1/post_sbert.RData'))
 ```
@@
 ```{r}
-sbert <- read_csv(here("data/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/post_sbert.RData")) %>% 
+sbert <- read_csv(here("data/expt_1/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/expt_1/post_sbert.RData")) %>% 
   as_tibble())
@@
     separate(dim2, into=c("game2","trial2"), convert=T, sep="_") %>%
-    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("data/similarities.rds"))
+    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("data/expt_1/similarities.rds"))
 

### 2024-12-13 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
check all transcripts and rerun pre-processing

@@
 
-* Check transcription for expt 1 and 2 especially for ref expression
-
-* Rerun sbert stuff
+* Rerun [re-processing 
+sbert stuff
 
@@
 * Include fun examples of the things kiddos said
+
+* fix transcripts to have kid ids that are more consistent (not pear123)
+
+* are there analyses to link success with description sticking (do successful descriptions stick more)
+
+* qual description of different ways kids do coordinate with each other
+
+* do we want to meta-analyze at all
+
+* get age range mean and gender distributions!
+
+* mention that we post-hoc excluded trials where we couldn't transcript any speaker description (either didn't happen or completely unintelligible)
 
@@
 ZZ pairs of children completed all 16 critical trials, and YY pairs of children completed between 8 and 16 critical trials.
+TODO mention that one kid was actually too young
 

### 2024-12-13 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
ran models

@@
 options(mc.cores = parallel::detectCores())
+theme_set(theme_bw())
 
@@
 dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial=trialNum)
-exclusion_1 <- transcript_1 %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game,  echo,) %>% unique() %>% mutate("S_talked"=1)
+exclusion_1 <- transcript_1 %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trial) %>% unique() %>% mutate("S_talked"=1)
 
@@
 
-# sims_1 <- read_rds(here(expt_1,"similarities.rds")) |> 
-#     as_tibble() 
-#   inner_join(clean_1|> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
-#   inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
-#   filter(target1 == target2) |>
-#   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
-#   mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
-#          same_game=(game1==game2) |> as.numeric(),
-#         later = ifelse(block1 > block2, block1, block2), 
-#         earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
+sims_1 <- read_rds(here(expt_1,"similarities.rds")) |>
+    as_tibble() |> 
+  inner_join(clean_1|> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
+  inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
+  filter(target1 == target2) |>
+  filter(trial1>1) |> 
+  filter(trial2>1) |> 
+  mutate(trial1=trial1+2,
+         trial2=trial2+2) |> 
+  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
+  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
+         same_game=(game1==game2) |> as.numeric(),
+        later = ifelse(block1 > block2, block1, block2),
+        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
 
@@
 
-# sims_2 <- read_rds(here(expt_2, "similarities.rds")) |>
-#   as_tibble() |>
-#   inner_join(clean_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
-#   inner_join(clean_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
-#   filter(target1 == target2) |>
-#   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
-#   mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
-#          same_game=(game1==game2) |> as.numeric(),
-#         later = ifelse(block1 > block2, block1, block2), 
-#         earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
+sims_2 <- read_rds(here(expt_2, "similarities.rds")) |>
+  as_tibble() |>
+  inner_join(clean_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
+  inner_join(clean_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
+  filter(target1 == target2) |>
+  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
+  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
+         same_game=(game1==game2) |> as.numeric(),
+        later = ifelse(block1 > block2, block1, block2),
+        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
 ```
@@
 
-* Rerun [re-processing 
-sbert stuff
-
-* Run models
-
-* Double check pre-regs and note deviations
+* save model summaries
 
@@
 
-* get age range mean and gender distributions!
+* Experiment 1: do children match the descriptions their partner uses
 
-* mention that we post-hoc excluded trials where we couldn't transcript any speaker description (either didn't happen or completely unintelligible)
-
-# Pre-regs
-https://osf.io/kcv8j
-
-https://osf.io/y2dax
-
-Analyses -- Expt 1
-
-This is exploratory research, and we will explore the following questions graphically:
-* how accurate are children over time
-* how many words of description does the speaker produce over time
-* do children speed up in their selections over time
-* do children match the descriptions their partner uses
-
-We will also analyze the transcripts to look at children’s turn-taking structure and how they understand each other.
-
-We plan to run the model:
-DV  ~ trial_num + (trial_num|dyad) + (1|target) 
-for the DVs: 
-[x] accuracy (logistic model), 
-[ ] speed (linear model) and 
-[ ] number of words of speaker description (linear model ).
-
-Analyses -- Expt 2
-
-Primarily, we aim to replicate the results of experiment 1, using the following models.
-
-[x] Logistic model of accuracy: correct ~ trialNum + (trialNum|game) + (1|target) (prior normal(0,1) for beta and sd, lkj(1) for correlation)
-
-[ ] Linear model of description length: words ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(5,10), beta and sd priors of normal(0,5), and correlation of lkj(1))
-
-[ ] Speed: time_seconds ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(60,100), beta and sd priors of (0,20) and correlation of lkj(1))
-
-We will interpret the fixed effect of trial number as being about changes in referring over time.
-
-For utterance similarities, we will embed the describer’s descriptions using S-BERT and use cosine similarity as a proxy for utterance similarity.
-
-We are unsure whether 4 blocks will be sufficient to see potential change over time, so we will 
-[ ] * compare similarities of utterances between different games v within one game v within one speaker: sim ~ same_game + same_speaker + (1|target) 
-[ ] * compare similarities of utterances to the last utterance within a game: sim_to_last ~ earlier_block + same_speaker + (1|game) + (1|target)
-[ ] * compare similarities of utterances to the next utterance: sim_to_next ~ earlier_block + same_speaker + (1|game) + (1|target)
-[ ] * compare the similarity level of descriptions from the same block in different games: sim ~ block + (1|target)
-Priors for the above similarity models: normal(.5,.2) for intercept, normal(0,.1) for
-beta, and normal(0,.05) for sd.
-
-We will also qualitatively explore children’s turn-taking structure and how they understand each other.
+* Experiment 1 + experiment 2: We will also analyze the transcripts to look at children’s turn-taking structure and how they understand each other.
 
@@
 
-TODO link to pre-registration
+This experiment was pre-registered at https://osf.io/kcv8j.
 
@@
 We had XX complete games and YY incomplete, but included games.
+TODO children's age range + median and gender split.
 
@@
 
+We excluded trials where the child in the teller role did not produce a description, or where all description was unintelligible and impossible to transcribe.
+
 # Experiment 1 Results
@@
 
-```{r}
+```{r, eval=F}
 acc_priors <- c(
@@
 
-acc_mod_data <- clean_1 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
+acc_mod_data_1 <- clean_1 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
                                   trial.num=trial-2) |> filter(type!="practice")
-acc_mod <- brm(correct.num ~ trial.num + (trial.num|game) + (1|target),
+acc_mod_1 <- brm(correct.num ~ trial.num + (trial.num|game) + (1|target),
                family = bernoulli(link = "logit"),
-               data=acc_mod_data,
+               data=acc_mod_data_1,
                file = here(mod_loc, "acc_1.rds"),
@@
 ## Speed
+
+```{r, eval=F}
+
+speed_priors <- c(
+  set_prior("normal(60,100)", class="Intercept"),
+set_prior("normal(0,20)", class="b"),
+set_prior("normal(0,20)", class="sd"),
+set_prior("lkj(1)", class="cor")
+)
+
+speed_mod_data_1 <- clean_1 |> mutate(time.sec=time/1000,
+                                  trial.num=trial-2) |> filter(type!="practice")
+
+speed_mod_1 <- brm(time.sec ~ trial.num + (trial.num|game) + (1|target),
+               data=speed_mod_data_1,
+               file = here(mod_loc, "speed_1.rds"),
+               prior=speed_priors,
+               control = list(adapt_delta=.95))
+```
 
@@
 ## Description length
+
+```{r, eval=F}
+
+description_priors <- c(
+  set_prior("normal(5,10)", class="Intercept"),
+set_prior("normal(0,5)", class="b"),
+set_prior("normal(0,5)", class="sd"),
+set_prior("lkj(1)", class="cor")
+)
+
+description_mod_data_1 <- clean_1 |> mutate(
+                                  trial.num=trial-2) |> filter(type!="practice")
+
+description_mod_1 <- brm(words ~ trial.num + (trial.num|game) + (1|target),
+               data=description_mod_data_1,
+               file = here(mod_loc, "description_1.rds"),
+               prior=speed_priors,
+               control = list(adapt_delta=.95))
+
+```
 
@@
 
-```{r, fig.width=4, fig.height=3}
-
-# sims_1 %>%  filter(!is.na(sim)) %>% 
-#   filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
-#   left_join(labels1) %>% left_join(labels2) %>%
-#   mutate(same_target=ifelse(target1==target2, "Same target", "Different target"),  #label if target is same or not
-#   same_child=case_when(
-#     game1!=game2 ~ "Different\n games",
-#     (as.numeric(trial1)-as.numeric(trial2))%%2==1~"Different\n speaker", #children alternate
-#     T ~"Same\n speaker"),
-#   type1=case_when(trial1<2 ~ "practice",
-#                         trial1<6 ~ "block 1",
-#                         trial1<10 ~ "block 2",
-#                         trial1<14 ~ "block 3"),
-#          type2=case_when(trial2<2 ~ "practice",
-#                         trial2<6 ~ "block 1",
-#                         trial2<10 ~ "block 2",
-#                         trial2<14 ~ "block 3")) %>% 
-#   filter(type1!=type2) %>% # can only compare across blocks in some cases, so only cross ever
-#     mutate(blocks=case_when(
-#       type1=="block 1" & type2=="block 2"~ "Block 1 v 2   ",
-#       type1=="block 2" & type2=="block 1"~ "Block 1 v 2   ",
-#       type1=="block 1" & type2=="block 3"~ "Block 1 v 3   ",
-#       type1=="block 3" & type2=="block 1"~ "Block 1 v 3   ",
-#       type1=="block 2" & type2=="block 3"~ "Block 2 v 3   ",
-#       type1=="block 3" & type2=="block 2"~ "Block 2 v 3   "
-#       )) %>%
-#   ggplot(aes(x=str_c(same_child), y=sim, color=blocks)) +
-#   facet_grid(.~same_target) +
-#   theme_bw()+
-#   #geom_jitter(alpha=.01, color="black") +
-#   labs(x="", y="Cosine similarity", color="")+
-#   scale_color_brewer(type="qual", palette="Dark2")+
-#   scale_y_continuous(limits=c(0,1), expand=c(0,0))+
-#       stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
-#     guides( color = guide_legend( override.aes = list(linetype=0, size=1)))+
-#   theme(strip.background =element_rect(fill="white"),
-#         legend.position="bottom",
-#         axis.title.x=element_blank())
+```{r}
+sims_1 |>
+  mutate(source = case_when(
+    speaker1 == speaker2 ~ "same speaker",
+    game1 == game2 ~ "same game",
+    T ~ "different games"
+  )) |>
+  ggplot(aes(x = source, y = sim, color = target1)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
+  theme(legend.position = "bottom")
 
@@
 ```{r}
-# sims_1 %>%  filter(!is.na(sim)) %>% 
-#   filter(!trial1 %in% c("0","1")) %>% filter(!trial2 %in% c("0","1"))%>% #not practice
-#   left_join(labels1) %>% left_join(labels2) %>%
-#   mutate(same_target=ifelse(target1==target2, "same_target", "diff_target"),  #label if target is same or not
-#   same_child=case_when(
-#     game1!=game2 ~ "diff game",
-#     (as.numeric(trial1)-as.numeric(trial2))%%2==1~"diff_child", #children alternate
-#     T ~"same_child"),
-#   type1=case_when(trial1<2 ~ "practice",
-#                         trial1<6 ~ "block 1",
-#                         trial1<10 ~ "block 2",
-#                         trial1<14 ~ "block 3"),
-#          type2=case_when(trial2<2 ~ "practice",
-#                         trial2<6 ~ "block 1",
-#                         trial2<10 ~ "block 2",
-#                         trial2<14 ~ "block 3")) %>% 
-#   filter(type1==type2) %>% # can only compare across blocks in some cases, so only cross ever
-#   ggplot(aes(x=str_c(same_target,"\n",same_child), y=sim, color=type1)) +
-#   geom_jitter(alpha=.01, color="black") +
-#       stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
-#   labs(x="Conditions compared")
+sims_1 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == 3) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |> 
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .1, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  facet_wrap(~same_speaker) +
+  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
 ```
 
+```{r}
+sims_1 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier + 1) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .1, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  facet_wrap(~same_speaker) +
+  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+sims_1 |>
+  filter(game1 != game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .01, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
+```
 # Experiment 1 Discussion
@@
 
-As Experiment 2 was very similar to Experiment 1, here we note the differences.
+As Experiment 2 was very similar to Experiment 1, here we note the differences. Experiment 2 was pre-registered at https://osf.io/y2dax.
 
@@
 ZZ pairs of children completed all 16 critical trials, and YY pairs of children completed between 8 and 16 critical trials.
-TODO mention that one kid was actually too young
+TODO mention that one kid was actually too young TODO mention age range / median and gender distribution.
 
@@
 ```{r}
+
+speed_priors <- c(
+  set_prior("normal(60,100)", class="Intercept"),
+set_prior("normal(0,20)", class="b"),
+set_prior("normal(0,20)", class="sd"),
+set_prior("lkj(1)", class="cor")
+)
+
+speed_mod_data_2<- clean_2 |> mutate(time.sec=time/1000,
+                                  trial.num=trial-2) |> filter(type!="practice")
+
+speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num|gameId) + (1|target),
+               data=speed_mod_data_2,
+               file = here(mod_loc, "speed_2.rds"),
+               prior=speed_priors,
+               control = list(adapt_delta=.95))
+```
+
+```{r}
 clean_2 |> ggplot( aes(x = trial + 1, y = time / 1000, color = type)) +
@@
 ## Description length
+
+```{r, eval=F}
+
+description_priors <- c(
+  set_prior("normal(5,10)", class="Intercept"),
+set_prior("normal(0,5)", class="b"),
+set_prior("normal(0,5)", class="sd"),
+set_prior("lkj(1)", class="cor")
+)
+
+description_mod_data_2 <- clean_2 |> mutate(
+                                  trial.num=trial-2) |> filter(type!="practice")
+
+description_mod_2 <- brm(words ~ trial.num + (trial.num|gameId) + (1|target),
+               data=description_mod_data_2,
+               file = here(mod_loc, "description_2.rds"),
+               prior=speed_priors,
+               control = list(adapt_delta=.95))
+```
 
@@
 ```{r}
-# sims_2 |>
-#   mutate(source = case_when(
-#     speaker1 == speaker2 ~ "same speaker",
-#     game1 == game2 ~ "same game",
-#     T ~ "different games"
-#   )) |>
-#   ggplot(aes(x = source, y = sim, color = target1)) +
-#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-#   labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
-#   theme(legend.position = "bottom")
+conv_priors <- c(
+  set_prior("normal(.5,.2)", class="Intercept"),
+set_prior("normal(0,.1)", class="b"),
+set_prior("normal(0,.05)", class="sd")
+)
+
+sim_across_between_data_2 <- sims_2 |> 
+mutate(same_game = case_when(
+    game1 == game2 ~ 1,
+    T ~ 0
+  )) |>
+    mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
+
+sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker+ (1|target),
+               data=sim_across_between_data_2,
+               file = here(mod_loc, "sim_across_between_2.rds"),
+               prior=conv_priors,
+               control = list(adapt_delta=.95))
+
+sim_to_last_data_2 <- sims_2 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == 4) |> 
+    mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+           earlier_block.num=earlier-1)
+
+sim_to_last_mod_2 <- brm(sim ~ earlier_block.num+ same_speaker+ (1|game1)+(1|target),
+               data=sim_to_last_data_2,
+               file = here(mod_loc, "sim_to_last_2.rds"),
+               prior=conv_priors,
+               control = list(adapt_delta=.95))
+
+sim_to_next_data_2 <- sims_2 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier+1) |> 
+    mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+           earlier_block.num=earlier-1)
+
+sim_to_next_mod_2 <- brm(sim ~ earlier_block.num+ same_speaker+ (1|game1)+(1|target),
+               data=sim_to_next_data_2,
+               file = here(mod_loc, "sim_to_next_2.rds"),
+               prior=conv_priors,
+               control = list(adapt_delta=.95))
+
+sim_across_data_2 <- sims_2 |>
+  filter(game1 != game2) |>
+  filter(block1 == block2) |> 
+    mutate(block.num=block1-1)
+
+sim_across_mod_2 <- brm(sim ~ block.num+(1|target),
+               data=sim_across_data_2,
+               file = here(mod_loc, "sim_across_2.rds"),
+               prior=conv_priors,
+               control = list(adapt_delta=.95))
 
@@
 ```{r}
-# sims_2 |>
-#   filter(game1 == game2) |>
-#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-#   filter(later == 4) |>
-#   mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-#   ggplot(aes(as.character(earlier), sim, color = target1)) +
-#   geom_point(alpha = .1, position = position_dodge(width = .2)) +
-#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-#   facet_wrap(~same_speaker) +
-#   labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
-#   theme(legend.position = "bottom")
+sims_2 |>
+  mutate(source = case_when(
+    speaker1 == speaker2 ~ "same speaker",
+    game1 == game2 ~ "same game",
+    T ~ "different games"
+  )) |>
+  ggplot(aes(x = source, y = sim, color = target1)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
+  theme(legend.position = "bottom")
+
 ```
@@
 ```{r}
-# sims_2 |>
-#   filter(game1 == game2) |>
-#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-#   filter(later == earlier + 1) |>
-#   mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-#   ggplot(aes(as.character(earlier), sim, color = target1)) +
-#   geom_point(alpha = .1, position = position_dodge(width = .2)) +
-#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-#   facet_wrap(~same_speaker) +
-#   labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
-#   theme(legend.position = "bottom")
+sims_2 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == 4) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .1, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  facet_wrap(~same_speaker) +
+  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
 ```
@@
 ```{r}
-# sims_2 |>
-#   filter(game1 != game2) |>
-#   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-#   filter(later == earlier) |>
-#   ggplot(aes(as.character(earlier), sim, color = target1)) +
-#   geom_point(alpha = .01, position = position_dodge(width = .2)) +
-#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-#   labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
-#   theme(legend.position = "bottom")
+sims_2 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier + 1) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .1, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  facet_wrap(~same_speaker) +
+  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+sims_2 |>
+  filter(game1 != game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .01, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
 ```

### 2024-12-13 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
save model summaries

@@
 mod_loc <- "code/models"
+mod_results <- "code/models/summary"
+mod_form <- "code/models/formulae"
 time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX
@@
         earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
+```
+
+```{r, eval=F}
+
+library(tidybayes)
+
+save_summary <- function(model) {
+  intervals <- gather_draws(model, `b_.*`, regex = T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex = T) %>%
+    mutate(above_0 = ifelse(.value > 0, 1, 0)) %>%
+    group_by(.variable) %>%
+    summarize(pct_above_0 = mean(above_0)) %>%
+    left_join(intervals, by = ".variable") %>%
+    mutate(
+      lower = .lower,
+      upper = .upper,
+      Term = str_sub(.variable, 3, -1),
+      Estimate = .value
+    ) %>%
+    select(Term, Estimate, lower, upper)
+
+  stats
+}
+
+form <- function(model) {
+  dep <- as.character(model$formula[2])
+  ind <- as.character(model$formula[3])
+
+  str_c(dep, " ~ ", ind) %>%
+    str_replace_all(" ", "") %>%
+    str_replace_all("\\*", " $\\\\times$ ") %>%
+    str_replace_all("\\+", "&nbsp;+ ") %>%
+    str_replace_all("~", "$\\\\sim$ ")
+}
+
+do_model <- function(path) {
+  model <- read_rds(here(mod_loc, path))
+  save_summary(model) |> write_rds(here(mod_loc, "summary", path))
+  model$formula |> write_rds(here(mod_loc, "formulae", path))
+  print(summary(model))
+}
+
+mods <- list.files(path = here(mod_loc), pattern = ".*rds") |> walk(~ do_model(.))
+
 ```

### 2024-12-13 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
draft stats and figures

@@
 
-form <- function(model) {
-  dep <- as.character(model$formula[2])
-  ind <- as.character(model$formula[3])
-
-  str_c(dep, " ~ ", ind) %>%
-    str_replace_all(" ", "") %>%
-    str_replace_all("\\*", " $\\\\times$ ") %>%
-    str_replace_all("\\+", "&nbsp;+ ") %>%
-    str_replace_all("~", "$\\\\sim$ ")
-}
-
 do_model <- function(path) {
@@
 
+```{r}
+stats <- function(model, row, decimal = 2) {
+  model <- model |>
+    mutate(
+      Estimate = round(Estimate, digits = decimal),
+      Lower = round(lower, digits = decimal),
+      Upper = round(upper, digits = decimal),
+      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
+    ) |>
+    select(Term, Estimate, `Credible Interval`)
+  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
+}
+
+stats_text <- function(model, row, decimal = 2) {
+  model <- model |>
+    mutate(
+      Estimate = round(Estimate, digits = decimal),
+      Lower = round(lower, digits = decimal),
+      Upper = round(upper, digits = decimal),
+      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
+    ) |>
+    select(Term, Estimate, `Credible Interval`)
+  str_c(model[row, 2], "  ", model[row, 3])
+}
+
+form <- function(model_form) {
+  dep <- as.character(model_form$formula[2])
+  ind <- as.character(model_form$formula[3])
+
+  str_c(dep, " ~ ", ind) |>
+    str_replace_all(" ", "") |>
+    str_replace_all("\\*", " $\\\\times$ ") |>
+    str_replace_all("\\+", "&nbsp;+ ") |>
+    str_replace_all("~", "$\\\\sim$ ")
+}
+
+```
+
 # TODOs
-
-* save model summaries
 
@@
 
+```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TODO \\label{game}", cache=FALSE}
+
+```
+
 ## Data processing
@@
 
+## Statistical analysis
+
+Statistical analyses were run in brms with weakly informative priors. We present the estimate and 95% credible intervals.
+
 # Experiment 1 Results
@@
 
-```{r}
- clean_1 |> 
-  ggplot(aes(x = trial+1, y = as.numeric(correct), fill=type)) +
-  stat_summary(fun.data="mean_cl_boot", geom="bar") +
-    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
-  scale_fill_manual(values=time_palette)+
-  theme_classic() +
-  xlab("\nTrial") +
-  ylab("Percent accuracy\n") +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
-
+```{r accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+clean_2 |>
+    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
+  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> 
+  ggplot(aes(x = trial, y = correct)) +
+  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
+  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
+  scale_fill_manual(values = time_palette) +
+  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
+  geom_smooth(method = "lm", color = "black") +
+  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  xlab("Trial") +
+  ylab("Accuracy") +
+  scale_x_continuous(breaks = seq(1, 16, 4), expand = c(0, 0)) +
+  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0)) +
+  geom_hline(yintercept = .5) +
+  facet_grid(expt~.)+
+  theme(legend.position = "none",
+        panel.grid = element_blank(),
+        strip.background = element_blank())
 ```
@@
 ```
+
+```{r}
+acc_mod_1 <- read_rds(here(mod_results, "acc_1.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))
+
+acc_form <- read_rds(here(mod_form,"acc_1.rds"))
+```
+
+We fit a mixed effects model of accuracy (`r acc_form |> form()`).
+Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
 
@@
 ```{r}
- clean_1 |> 
-  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
-  ggplot(aes(x = trial+1, y = time/1000, color=type)) +
-  geom_jitter(width=.2, height=0, alpha=.5)+
-  #geom_line(aes(group=game))+
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
-  scale_color_manual(values=time_palette)+
-  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  #geom_smooth(method="lm")+
-  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nTrial") +
-  ylab("Time to answer in seconds")+
-  coord_cartesian(ylim=c(0,60))+
-  #ylab("Percent trials with echoing \n by experimenter\n") +
-  #scale_x_continuous(breaks=seq(1,12,1)) +
-  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
+speed_mod_1 <- read_rds(here(mod_results, "speed_1.rds"))
+
+speed_form <- read_rds(here(mod_form, "speed_1.rds"))
 ```
+
+We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
+The first trial was `r stats_text(speed_mod_1, 1)`, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
 
@@
 
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+
+clean_2 |>
+    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
+  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> ggplot( aes(x = trial, y = words, color = type)) +
+  geom_jitter(alpha = .5, color = "grey") +
+  stat_summary(fun.data = "mean_cl_boot") +
+  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
+  scale_color_manual(values = time_palette) +
+  geom_smooth(method = "lm", color = "black") +
+  xlab("Trial") +
+  ylab("Words") +
+  facet_grid(expt~.)+
+  theme(legend.position = "none",
+        panel.grid = element_blank(),
+        strip.background = element_blank())
+```
+
 ```{r}
+description_mod_1 <- read_rds(here(mod_results, "description_1.rds"))
 
-clean_1 |> 
-  ggplot(aes(x = trial+1, y = words, color=type)) +
-    geom_jitter(width=.2, height=0, alpha=.5)+
-  #geom_line(aes(group=game))+
- # facet_wrap(.~correct) + 
-    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
-  scale_color_manual(values=time_palette)+
-  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  #geom_smooth(method="lm")+
-  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nTrial") +
-  #coord_cartesian(ylim=c(0,60))+
-  #ylab("Percent trials with echoing \n by experimenter\n") +
-  #scale_x_continuous(breaks=seq(1,12,1)) +
-  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
-  theme(aspect.ratio = 4/5,
-        axis.title=element_text(size=14,face="bold"),
-        axis.text=element_text(size=9, color = "black"),
-        legend.position = "none")
+description_form <- read_rds(here(mod_form, "description_1.rds"))
 
@@
 
+We ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
+The initial length was `r stats_text(description_mod_1, 1)` and description length was relatively stable over time (`r stats_text(description_mod_1, 2)`).
+
 ## Convergence
 
-```{r}
-sims_1 |>
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+
+sims_2 |>
   mutate(source = case_when(
@@
     T ~ "different games"
-  )) |>
+  )) |> mutate(expt="Expt 2") |> 
+  bind_rows(sims_1 |>
+  mutate(source = case_when(
+    speaker1 == speaker2 ~ "same speaker",
+    game1 == game2 ~ "same game",
+    T ~ "different games"
+  )) |> mutate(expt="Expt 1")) |> 
   ggplot(aes(x = source, y = sim, color = target1)) +
@@
   labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
-  theme(legend.position = "bottom")
+  facet_grid(expt~.)+
+  theme(legend.position = "bottom",
+        panel.grid = element_blank(),
+        strip.background = element_blank())
 
@@
 
-```{r}
+```{r, include=F}
 sims_1 |>
@@
 
-```{r}
+```{r include=F}
 sims_1 |>
@@
   theme(legend.position = "bottom")
+
 ```
 
-```{r}
+```{r, include=F}
 sims_1 |>
@@
 ```
+
+There were too few trials to really look at change over time in this short experiment. However, we were able to look at ...
+
+We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT
+
 # Experiment 1 Discussion
@@
 
-```{r}
-clean_2 |>
-  ggplot(aes(x = trial + 1, y = correct)) +
-  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
-  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
-  scale_fill_manual(values = time_palette) +
-  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
-  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
-  theme_classic() +
-  xlab("\nTrial") +
-  ylab("Percent accuracy\n") +
-  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
-  scale_y_continuous(breaks = seq(0, 1, .1), expand = c(0, 0)) +
-  geom_hline(yintercept = .5) +
-  theme(
-    aspect.ratio = 4 / 5,
-    axis.title = element_text(size = 14, face = "bold"),
-    axis.text = element_text(size = 9, color = "black"),
-    legend.position = "none"
-  )
-```
-
 ```{r, eval=F}
@@
 ```
+
+```{r}
+acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))
+
+```
+
+ Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`).
 
@@
 ```{r}
-clean_2 |> ggplot( aes(x = trial + 1, y = time / 1000, color = type)) +
-  geom_jitter(alpha = .5, color = "grey") +
-  stat_summary(fun.data = "mean_cl_boot") +
-  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
-  scale_color_manual(values = time_palette) +
-  theme_classic() +
-  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
-  xlab("Trial") +
-  ylab("Time (seconds)") +
-  theme(legend.position = "none")
+speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))
+
 ```
+
+ The first trial was `r stats_text(speed_mod_2, 1)`, and children got faster over time (`r stats_text(speed_mod_2, 2)`).
 
@@
 ```{r}
+description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))
 
-clean_2 |> ggplot( aes(x = trial + 1, y = words, color = type)) +
-  geom_jitter(alpha = .5, color = "grey") +
-  stat_summary(fun.data = "mean_cl_boot") +
-  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
-  scale_color_manual(values = time_palette) +
-  theme_classic() +
-  geom_smooth(data = clean_2|> filter(trial > 3), method = "lm", color = "black") +
-  xlab("Trial") +
-  ylab("Words") +
-  theme(legend.position = "none")
 ```
+
+ The initial length was `r stats_text(description_mod_2, 1)` and description length was relatively stable over time (`r stats_text(description_mod_2, 2)`).
 
@@
 
-```{r}
-sims_2 |>
-  mutate(source = case_when(
-    speaker1 == speaker2 ~ "same speaker",
-    game1 == game2 ~ "same game",
-    T ~ "different games"
-  )) |>
-  ggplot(aes(x = source, y = sim, color = target1)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
-  theme(legend.position = "bottom")
-
-```
-
-```{r}
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
 sims_2 |>
@@
 
-```{r}
+```{r, include=F}
 sims_2 |>
@@
 ```
+
+```{r}
+sim_across_between_2_results <- read_rds(here(mod_results, "sim_across_between_2.rds"))
+
+sim_across_between_form<- read_rds(here(mod_form, "sim_across_between_2.rds"))
+
+sim_across_2_results <- read_rds(here(mod_results, "sim_across_2.rds"))
+
+sim_across_form<- read_rds(here(mod_form, "sim_across_2.rds"))
+
+sim_to_next_2_results <- read_rds(here(mod_results, "sim_to_next_2.rds"))
+
+sim_to_next_form<- read_rds(here(mod_form, "sim_to_next_2.rds"))
+
+sim_to_last_2_results <- read_rds(here(mod_results, "sim_to_last_2.rds"))
+
+sim_to_last_form<- read_rds(here(mod_form, "sim_to_last_2.rds"))
+```
+
+We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT
+
+As a coarse comparison, we repeated the analysis from Experiment 1, comparing the similarity of descriptions to the same target for the same-speaker, same-game, or different-game.
+We modeled this as `r sim_across_between_form |> form()`.
+Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
+The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partners than to another child.
+
+With the greater length of experiment 2, we are also able to look for change over time, that could be indicative of convergence to shared descriptions within a game and divergence between games.
+
+To look for convergence, we compare the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`.
+Over the first three blocks, descriptions do become increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+Descriptions are more similar to the last block if they come from the same child who gaves the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
+
+Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
+Although over time descriptions do get more similar to the last block utterance, the distance between adjacent block utterances is relatively constant `r sim_to_next_2_results |> stats_text(2,3)`.
+
+Finally, we can see if descriptions in different games diverge across time: `r sim_across_form |> form()`.
+As the games progress, descriptions from different games become slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
 

### 2025-01-03 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
work on methods and results

@@
 library(rstanarm)
+library(ggthemes)
 rstan_options(auto_write = TRUE)
@@
 
-# TODOs
-
-* Include fun examples of the things kiddos said
-
-* fix transcripts to have kid ids that are more consistent (not pear123)
-
-* are there analyses to link success with description sticking (do successful descriptions stick more)
-
-* qual description of different ways kids do coordinate with each other
-
-* do we want to meta-analyze at all
-
-* Experiment 1: do children match the descriptions their partner uses
-
-* Experiment 1 + experiment 2: We will also analyze the transcripts to look at children’s turn-taking structure and how they understand each other.
-
 # Introduction
@@
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load that can hide early abilities (TODO cite something for this), so we wanted to use a simple paradigm with experimenter scaffolding.
+Young children can be very sensitive to task demands and cognitive load that can hide early abilities (TODO cite something for this), so we used a simple paradigm where an experimenter could scaffold the children's iteraction as needed.
 We adapted the experimental framework from @leung, but further simplified it by reducing the total pool of targets and the number of trials children completed.
@@
 Children played with another child from the same class.
-TODO any exclusions Experiment 1 was conducted between June and August 2023.
-Children were only included in analysis if they completed at least 8 of the critical trials.
-We had XX complete games and YY incomplete, but included games.
+Experiment 1 was conducted between June and August 2023.
+Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials.
+We had 19 complete games and 1 incomplete, but included game.
 TODO children's age range + median and gender split.
@@
 For the target stimuli, we used four of the ten tangram images from @leung, chosen based on visual dissimilarity.
-The matching game was coded as an Empirica experiment (CITATION) and run on two tablets locked in a kiosk mode so children could access other websites or apps.
+We coded the matching game using Empirica and hosted it on a lab server (CITATION).
+We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate to other websites or applications during the game.
 
 ## Procedure
-Once a pair of children agreed to play the game, they were taken to a quiet testing room where procedure was explained to them.
+Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room where the game was explained to them.
 Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game.
@@
 The guesser saw the same two images (in a randomized order), but with neither boxed.
-Upon tapping an image, both children recieved feedback in form of a smiley or frowney face and an audible sound.
-After each trial, children's roled switched, demarcated by passing Smurfy back and forth.
+Upon tapping an image, both children received feedback in form of a smiley or frowny face and an audible sound.
+After each trial, children's roles switched.
+Children passed Smurfy back and forth to help them keep track of whether they were the "guesser" or "teller" on a given trial.
 
@@
 
-We excluded trials where the child in the teller role did not produce a description, or where all description was unintelligible and impossible to transcribe.
+We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe.
+After exclusions, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials remaining.
 
@@
 
-```{r accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time.
+Experiment 1 had 3 blocks (12 total critical trials) and experiment 2 had 4 blocks (16 critical trials).
+Error bars are bootstrapped 95% CIs with a linear trend line overlaid.
+\\label{acc}" }
 clean_2 |>
@@
 
-We fit a mixed effects model of accuracy (`r acc_form |> form()`).
+Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully (CITATION).
+As shown in Figure \ref{acc}, children were above chance in their selections.
+To confirm this and test for any changes in accuracy over time, we fit a mixed effects model of accuracy (`r acc_form |> form()`).
 Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
@@
 
+As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
 We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
-The first trial was `r stats_text(speed_mod_1, 1)`, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
+The first trial critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
 
@@
 
-```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of referential expression (in words) produced by the speaker each trial.
+Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs.
+\\label{length}" }
 
@@
 
+In iterated reference games in adults, a canonical finding is that the length of descriptions goes from long to short over repeated references to the initially hard to describe shapes.
+We looked at how long the descriptions the children used were to see if the same trend occurred.
 We ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
-The initial length was `r stats_text(description_mod_1, 1)` and description length was relatively stable over time (`r stats_text(description_mod_1, 2)`).
+The initial length was `r stats_text(description_mod_1, 1)` and description length was relatively stable over time (`r stats_text(description_mod_1, 2)`, shown in Figure \ref{length}).
+How long of a description is initially warranted depends on how iconic or easily describable the shape is, as well as how large and close the contrast set is.
+In this case, the low number of options to distinguish and relatively easy to describe shapes may mean long initial descriptions are less necessary (todo mention how long the similar things are in leung).
+However, young children may also choose to provide shorter descriptions than adults would.
+
+TODO add anecdotes here with some examples of children's descriptions.
 
@@
 
-```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between referential descriptions from the speaker to the same target under different circumstances.
+Different games refers to similarities across two speakers from different games, same game to similarities across the two participants in one game, and same speaker to descriptions from the same person in different blocks.
+Dots are means and lines are bootstrapped 95% CIs.
+\\label{sim}" }
 
@@
   facet_grid(expt~.)+
+  scale_color_solarized()+
   theme(legend.position = "bottom",
@@
 
-There were too few trials to really look at change over time in this short experiment. However, we were able to look at ...
+While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length (cite boyce 2024).
+With only 3 repetitions for each target, we do not have enough data to look at convergence over time.
+However, we can still analalyze the semantics of the descriptions to see if children are influenced by their partner's descriptions or not.
 
-We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT
+Following cite Boyce et al 2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions.
+We compare the semantic similarities between descriptions of the same target based on who produced the description.
+Our question is whether the two children within the same game produce descriptions that are more similar to each other than two children in different games.
+As shown in Figure \ref{sim}, different children in the same game do produce more similar descriptions than children in different games, although the descriptions are less similar than descriptions from the same child in different rounds.
+This provides weak evidence for some coordination between chidren.
+
+TODO could provide anecdotes about children's coordination approaches
+
+TODO should we run the brms model on this?
 
@@
 
-After the first experiment, we wanted to run another study using the same paradigm, but fixing issues that arose during experiment 1.
-We also tightened the experimenter script to reduce the possibility that the experimenters were influencing children's descriptions and understanding of descriptions.
-Given that most pairs in experiment 1 were able to play the game fairly quickly, we also added a 4th experimental block to be able to look for change over time.
+In experiment 1, children were above chance accuracy in their selections.
+In experiment 2, we aimed to repeat the same paradigm with a tighter experimental script to reduce possible influence of experimenters.
+Additionally, we aimed to fix sources of confusion and frustration in experiment 1.
+As most pairs in experiment 1 completed the game fairly quickly, we added a 4th experimental block to allow for more analyses of change over time.
 
@@
 ## Participants
-Experiment 2 was run between March and August of 2024, at the same university preschool.
+Experiment 2 was run between March and August of 2024, at the same university preschool as experiment 1.
 No children participated in both experiments.
-ZZ pairs of children completed all 16 critical trials, and YY pairs of children completed between 8 and 16 critical trials.
+30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
 TODO mention that one kid was actually too young TODO mention age range / median and gender distribution.
@@
 In response to some children struggling with the abrupt switch from nameable to non-nameable shapes, we introduced more practice trials for experiment 2.
-Additionally, before the practice trials, we added a "bubble popping" exercise to familiarize children with how to tap the tablet appropriately (a problem in experiment 1).
-
+We used a total of 4 practice trials, designed to transition from easily recognizable shapes to slightly blockier black and wide shapes to smooth the transition to the critical trials.
 ## Procedure
-The procedure was much the same. The smurfy puppet was swapped out for a more attractive smurfy stuffed animal. The experimental script was tightened. TODO link to materials for where the script is.
+The procedure was much the same.
+We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in the first experiment).
+The smurfy puppet was swapped out for a more attractive smurfy stuffed animal.
+The experimental script was fully written out so children all recieved the same instructions.
+To prevent experimenter's influencing children's descriptions or understanding of descriptions, we wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
+TODO link to materials for where the script is.
 
@@
 Data was processed in the same way as experiment 1.
+After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
 
 # Experiment 2 Results
+
+We report the same set of analyses and model results as in Experiment 1 as well as additional analyses of how the semantic content of children's descriptions changes over time.
 
@@
 
- Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`).
+As in Experiment 1, in Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
+This confirms that children are able to communicate with each other about these abstract shapes.
 
@@
 
- The first trial was `r stats_text(speed_mod_2, 1)`, and children got faster over time (`r stats_text(speed_mod_2, 2)`).
+In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)`).
+Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
 
@@
 
- The initial length was `r stats_text(description_mod_2, 1)` and description length was relatively stable over time (`r stats_text(description_mod_2, 2)`).
+The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (`r stats_text(description_mod_2, 2)`, Figure \ref{length}).
+This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
+
+TODO could include examples of cute things kiddos said here
 
@@
 
-```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between earlier blocks (1-3) with last block (4) for descriptions to the same image within the same group.
+SImilarity measured as cosine similarity between S-BERT embeddings of referential descriptions.
+Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
+\\label{tolast}" }
 sims_2 |>
@@
   facet_wrap(~same_speaker) +
+  scale_color_solarized()+
   labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
@@
 
-We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT
+To look at semantic distance between utterances, we again operationalize similarity between pairs of utterances as the cosine similarity between their SBERT embeddings (CITE SBERT).
 
@@
 Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
-The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partners than to another child.
+The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partner's than to another child's.
 
-With the greater length of experiment 2, we are also able to look for change over time, that could be indicative of convergence to shared descriptions within a game and divergence between games.
+The greater number of trials in Experiment 2 makes it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
 
-To look for convergence, we compare the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`.
-Over the first three blocks, descriptions do become increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions are more similar to the last block if they come from the same child who gaves the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
+To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`.
+Over the first three blocks, descriptions become increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+Descriptions are more similar to the last block if they come from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
 
 Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
-Although over time descriptions do get more similar to the last block utterance, the distance between adjacent block utterances is relatively constant `r sim_to_next_2_results |> stats_text(2,3)`.
+Although over time descriptions do get more similar to the last block utterance, the distance between adjacent block utterances is relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
 
-Finally, we can see if descriptions in different games diverge across time: `r sim_across_form |> form()`.
-As the games progress, descriptions from different games become slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
+As each partnership converges to their shared nicknames, partnerships often diverge from one another as groups focus on distinct aspects of the image.
+We tested whether descriptions in different games to the same target diverged over time: `r sim_across_form |> form()`.
+As the games progress, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
 

### 2025-01-03 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
added age ranges

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
-                      fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
-                      message=F, sanitize = T)
+knitr::opts_chunk$set(
+  fig.width = 3, fig.height = 3, fig.crop = F,
+  fig.pos = "tb", fig.path = "figs/",
+  echo = F, warning = F, cache = F,
+  message = F, sanitize = T
+)
 
@@
 # experiment 1 imports
-transcript_1 <- read_csv(here(expt_1,"timed_transcript.csv"))
-dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial=trialNum)
-exclusion_1 <- transcript_1 %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trial) %>% unique() %>% mutate("S_talked"=1)
+transcript_1 <- read_csv(here(expt_1, "timed_transcript.csv"))
+dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial = trialNum)
+exclusion_1 <- transcript_1 %>%
+  filter(!is.na(description)) %>%
+  filter(role == "S") %>%
+  select(game, trial) %>%
+  unique() %>%
+  mutate("S_talked" = 1)
 
-words_1 <- transcript_1 |> filter(!is.na(description)) |> mutate(words=description |> str_count("\\S+")) |> 
-  filter(role=="S") |> 
+words_1 <- transcript_1 |>
+  filter(!is.na(description)) |>
+  mutate(words = description |> str_count("\\S+")) |>
+  filter(role == "S") |>
   group_by(game, trial) |>
-  summarize(words=sum(words))
+  summarize(words = sum(words))
 
-clean_1 <- dat_1 %>% left_join(exclusion_1) %>% filter(S_talked==1) |> left_join(words_1) |>  mutate(type=case_when(trial<2 ~ "practice",
-                        trial<6 ~ "block 1",
-                        trial<10 ~ "block 2",
-                        trial<14 ~ "block 3"))
+clean_1 <- dat_1 %>%
+  left_join(exclusion_1) %>%
+  filter(S_talked == 1) |>
+  left_join(words_1) |>
+  mutate(type = case_when(
+    trial < 2 ~ "practice",
+    trial < 6 ~ "block 1",
+    trial < 10 ~ "block 2",
+    trial < 14 ~ "block 3"
+  ))
 
-sims_1 <- read_rds(here(expt_1,"similarities.rds")) |>
+sims_1 <- read_rds(here(expt_1, "similarities.rds")) |>
   as_tibble() |>
-  inner_join(clean_1|> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
+  inner_join(clean_1 |> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
   inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
   filter(target1 == target2) |>
-  filter(trial1>1) |> 
-  filter(trial2>1) |> 
-  mutate(trial1=trial1+2,
-         trial2=trial2+2) |> 
+  filter(trial1 > 1) |>
+  filter(trial2 > 1) |>
+  mutate(
+    trial1 = trial1 + 2,
+    trial2 = trial2 + 2
+  ) |>
   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
-  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
-         same_game=(game1==game2) |> as.numeric(),
+  mutate(
+    same_speaker = (speaker1 == speaker2) |> as.numeric(),
+    same_game = (game1 == game2) |> as.numeric(),
     later = ifelse(block1 > block2, block1, block2),
-        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
+    earlier = ifelse(block1 > block2, block2, block1)
+  ) |>
+  mutate(target = target1)
 
@@
 
-words_2 <- transcript_2 |> mutate(words=description |> str_count("\\S+")) |> 
-  filter(role=="S") |> 
+words_2 <- transcript_2 |>
+  mutate(words = description |> str_count("\\S+")) |>
+  filter(role == "S") |>
   group_by(gameId, gameConfig, trial) |>
-  summarize(words=sum(words))
+  summarize(words = sum(words))
 
-clean_2 <- data_2 |> left_join(words_2) |>
+clean_2 <- data_2 |>
+  left_join(words_2) |>
   filter(!is.na(words)) |>
@@
   filter(!is.na(response)) |>
-  mutate(correct=as.numeric(correct))
+  mutate(correct = as.numeric(correct))
 
@@
   mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
-  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
-         same_game=(game1==game2) |> as.numeric(),
+  mutate(
+    same_speaker = (speaker1 == speaker2) |> as.numeric(),
+    same_game = (game1 == game2) |> as.numeric(),
     later = ifelse(block1 > block2, block1, block2),
-        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
+    earlier = ifelse(block1 > block2, block2, block1)
+  ) |>
+  mutate(target = target1)
 ```
@@
 ```{r, eval=F}
-
 library(tidybayes)
@@
 mods <- list.files(path = here(mod_loc), pattern = ".*rds") |> walk(~ do_model(.))
-
 ```
@@
 }
-
 ```
@@
 We had 19 complete games and 1 incomplete, but included game.
-TODO children's age range + median and gender split.
+Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
+
+```{r, eval=F}
+# demo <- read_csv(here("demographics.csv")) |>
+#   filter(included == "x") |>
+#   mutate(date_test = mdy(date_test), date_birth = mdy(date_birth)) |>
+#   select(date_test, date_birth, gender) |>
+#   mutate(
+#     age = date_test - date_birth,
+#     age = as.numeric(age) / 365.25 * 12
+#   ) |> 
+#   mutate(expt=ifelse(date_test>ymd("2024-01-01"), "expt2", "expt1")) |> 
+#   group_by(expt) |> 
+#   summarize(female=sum(gender=="female"),
+#             total=n(),
+#             median=median(age),
+#             min=min(age),
+#             max=max(age))
+```
 
@@
 ```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TODO \\label{game}", cache=FALSE}
-
 ```
@@
 clean_2 |>
-    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
-  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> 
+  filter(type != "practice") |>
+  mutate(trial = trial - 3, expt = "Expt 2") |>
+  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
   ggplot(aes(x = trial, y = correct)) +
@@
   geom_hline(yintercept = .5) +
-  facet_grid(expt~.)+
-  theme(legend.position = "none",
+  facet_grid(expt ~ .) +
+  theme(
+    legend.position = "none",
     panel.grid = element_blank(),
-        strip.background = element_blank())
+    strip.background = element_blank()
+  )
 ```
@@
 acc_priors <- c(
-set_prior("normal(0,1)", class="b"),
-set_prior("normal(0,1)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(0,1)", class = "b"),
+  set_prior("normal(0,1)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-acc_mod_data_1 <- clean_1 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
-                                  trial.num=trial-2) |> filter(type!="practice")
-acc_mod_1 <- brm(correct.num ~ trial.num + (trial.num|game) + (1|target),
+acc_mod_data_1 <- clean_1 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
+acc_mod_1 <- brm(correct.num ~ trial.num + (trial.num | game) + (1 | target),
   family = bernoulli(link = "logit"),
-               data=acc_mod_data_1,
+  data = acc_mod_data_1,
   file = here(mod_loc, "acc_1.rds"),
-               prior=acc_priors,
-               control = list(adapt_delta=.95))
+  prior = acc_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 ```{r}
-acc_mod_1 <- read_rds(here(mod_results, "acc_1.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))
+acc_mod_1 <- read_rds(here(mod_results, "acc_1.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
 
-acc_form <- read_rds(here(mod_form,"acc_1.rds"))
+acc_form <- read_rds(here(mod_form, "acc_1.rds"))
 ```
@@
 ```{r, eval=F}
-
 speed_priors <- c(
-  set_prior("normal(60,100)", class="Intercept"),
-set_prior("normal(0,20)", class="b"),
-set_prior("normal(0,20)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(60,100)", class = "Intercept"),
+  set_prior("normal(0,20)", class = "b"),
+  set_prior("normal(0,20)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-speed_mod_data_1 <- clean_1 |> mutate(time.sec=time/1000,
-                                  trial.num=trial-2) |> filter(type!="practice")
+speed_mod_data_1 <- clean_1 |>
+  mutate(
+    time.sec = time / 1000,
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
 
-speed_mod_1 <- brm(time.sec ~ trial.num + (trial.num|game) + (1|target),
-               data=speed_mod_data_1,
+speed_mod_1 <- brm(time.sec ~ trial.num + (trial.num | game) + (1 | target),
+  data = speed_mod_data_1,
   file = here(mod_loc, "speed_1.rds"),
-               prior=speed_priors,
-               control = list(adapt_delta=.95))
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 ```{r, eval=F}
-
 description_priors <- c(
-  set_prior("normal(5,10)", class="Intercept"),
-set_prior("normal(0,5)", class="b"),
-set_prior("normal(0,5)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(5,10)", class = "Intercept"),
+  set_prior("normal(0,5)", class = "b"),
+  set_prior("normal(0,5)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-description_mod_data_1 <- clean_1 |> mutate(
-                                  trial.num=trial-2) |> filter(type!="practice")
+description_mod_data_1 <- clean_1 |>
+  mutate(
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
 
-description_mod_1 <- brm(words ~ trial.num + (trial.num|game) + (1|target),
-               data=description_mod_data_1,
+description_mod_1 <- brm(words ~ trial.num + (trial.num | game) + (1 | target),
+  data = description_mod_data_1,
   file = here(mod_loc, "description_1.rds"),
-               prior=speed_priors,
-               control = list(adapt_delta=.95))
-
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 \\label{length}" }
-
 clean_2 |>
-    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
-  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> ggplot( aes(x = trial, y = words, color = type)) +
+  filter(type != "practice") |>
+  mutate(trial = trial - 3, expt = "Expt 2") |>
+  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
+  ggplot(aes(x = trial, y = words, color = type)) +
   geom_jitter(alpha = .5, color = "grey") +
@@
   ylab("Words") +
-  facet_grid(expt~.)+
-  theme(legend.position = "none",
+  facet_grid(expt ~ .) +
+  theme(
+    legend.position = "none",
     panel.grid = element_blank(),
-        strip.background = element_blank())
+    strip.background = element_blank()
+  )
 ```
@@
 description_form <- read_rds(here(mod_form, "description_1.rds"))
-
 ```
@@
 
+```{r}
+conv_priors <- c(
+  set_prior("normal(.5,.2)", class = "Intercept"),
+  set_prior("normal(0,.1)", class = "b"),
+  set_prior("normal(0,.05)", class = "sd")
+)
+
+sim_across_between_data_1 <- sims_1 |>
+  mutate(same_game = case_when(
+    game1 == game2 ~ 1,
+    T ~ 0
+  )) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
+
+sim_across_between_mod_1 <- brm(sim ~ same_game + same_speaker + (1 | target),
+  data = sim_across_between_data_1,
+  file = here(mod_loc, "sim_across_between_1.rds"),
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
+```
+
 ```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between referential descriptions from the speaker to the same target under different circumstances.
@@
 \\label{sim}" }
-
 sims_2 |>
@@
     T ~ "different games"
-  )) |> mutate(expt="Expt 2") |> 
+  )) |>
+  mutate(expt = "Expt 2") |>
   bind_rows(sims_1 |>
@@
       T ~ "different games"
-  )) |> mutate(expt="Expt 1")) |> 
+    )) |> mutate(expt = "Expt 1")) |>
   ggplot(aes(x = source, y = sim, color = target1)) +
@@
   labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
-  facet_grid(expt~.)+
-  scale_color_solarized()+
-  theme(legend.position = "bottom",
+  facet_grid(expt ~ .) +
+  scale_color_solarized() +
+  theme(
+    legend.position = "bottom",
     panel.grid = element_blank(),
-        strip.background = element_blank())
+    strip.background = element_blank()
+  )
+```
 
+```{r}
+sim_across_between_1_results <- read_rds(here(mod_results, "sim_across_between_1.rds"))
+
+sim_across_between_form <- read_rds(here(mod_form, "sim_across_between_1.rds"))
 ```
@@
   theme(legend.position = "bottom")
-
 ```
@@
 As shown in Figure \ref{sim}, different children in the same game do produce more similar descriptions than children in different games, although the descriptions are less similar than descriptions from the same child in different rounds.
+We modeled this as `r sim_across_between_form |> form()`.
+Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
+The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partner's than to another child's.
 This provides weak evidence for some coordination between chidren.
@@
 TODO could provide anecdotes about children's coordination approaches
-
-TODO should we run the brms model on this?
 
@@
 30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
-TODO mention that one kid was actually too young TODO mention age range / median and gender distribution.
+Our target age range was 4 and 5 year olds, but one almost 4-year-old was unintentionally included.
+Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
 
@@
 acc_priors <- c(
-set_prior("normal(0,1)", class="b"),
-set_prior("normal(0,1)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(0,1)", class = "b"),
+  set_prior("normal(0,1)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-acc_mod_data <- clean_2 |> mutate(correct.num=ifelse(correct==TRUE, 1,0),
-                                  trial.num=trial-4) |> filter(type!="practice")
-acc_mod <- brm(correct.num ~ trial.num + (trial.num|gameId) + (1|target),
+acc_mod_data <- clean_2 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 4
+  ) |>
+  filter(type != "practice")
+acc_mod <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
   family = bernoulli(link = "logit"),
-               data=acc_mod_data,
+  data = acc_mod_data,
   file = here(mod_loc, "acc_2.rds"),
-               prior=acc_priors,
-               control = list(adapt_delta=.95))
+  prior = acc_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 ```{r}
-acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))
-
+acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
 ```
@@
 ```{r}
-
 speed_priors <- c(
-  set_prior("normal(60,100)", class="Intercept"),
-set_prior("normal(0,20)", class="b"),
-set_prior("normal(0,20)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(60,100)", class = "Intercept"),
+  set_prior("normal(0,20)", class = "b"),
+  set_prior("normal(0,20)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-speed_mod_data_2<- clean_2 |> mutate(time.sec=time/1000,
-                                  trial.num=trial-2) |> filter(type!="practice")
+speed_mod_data_2 <- clean_2 |>
+  mutate(
+    time.sec = time / 1000,
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
 
-speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num|gameId) + (1|target),
-               data=speed_mod_data_2,
+speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num | gameId) + (1 | target),
+  data = speed_mod_data_2,
   file = here(mod_loc, "speed_2.rds"),
-               prior=speed_priors,
-               control = list(adapt_delta=.95))
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))
-
 ```
@@
 ```{r, eval=F}
-
 description_priors <- c(
-  set_prior("normal(5,10)", class="Intercept"),
-set_prior("normal(0,5)", class="b"),
-set_prior("normal(0,5)", class="sd"),
-set_prior("lkj(1)", class="cor")
+  set_prior("normal(5,10)", class = "Intercept"),
+  set_prior("normal(0,5)", class = "b"),
+  set_prior("normal(0,5)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
 )
 
-description_mod_data_2 <- clean_2 |> mutate(
-                                  trial.num=trial-2) |> filter(type!="practice")
+description_mod_data_2 <- clean_2 |>
+  mutate(
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
 
-description_mod_2 <- brm(words ~ trial.num + (trial.num|gameId) + (1|target),
-               data=description_mod_data_2,
+description_mod_2 <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
+  data = description_mod_data_2,
   file = here(mod_loc, "description_2.rds"),
-               prior=speed_priors,
-               control = list(adapt_delta=.95))
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
 description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))
-
 ```
@@
 conv_priors <- c(
-  set_prior("normal(.5,.2)", class="Intercept"),
-set_prior("normal(0,.1)", class="b"),
-set_prior("normal(0,.05)", class="sd")
+  set_prior("normal(.5,.2)", class = "Intercept"),
+  set_prior("normal(0,.1)", class = "b"),
+  set_prior("normal(0,.05)", class = "sd")
 )
@@
 
-sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker+ (1|target),
-               data=sim_across_between_data_2,
+sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker + (1 | target),
+  data = sim_across_between_data_2,
   file = here(mod_loc, "sim_across_between_2.rds"),
-               prior=conv_priors,
-               control = list(adapt_delta=.95))
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
 
@@
   filter(later == 4) |>
-    mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-           earlier_block.num=earlier-1)
+  mutate(
+    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+    earlier_block.num = earlier - 1
+  )
 
-sim_to_last_mod_2 <- brm(sim ~ earlier_block.num+ same_speaker+ (1|game1)+(1|target),
-               data=sim_to_last_data_2,
+sim_to_last_mod_2 <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
+  data = sim_to_last_data_2,
   file = here(mod_loc, "sim_to_last_2.rds"),
-               prior=conv_priors,
-               control = list(adapt_delta=.95))
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
 
@@
   mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier+1) |> 
-    mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-           earlier_block.num=earlier-1)
+  filter(later == earlier + 1) |>
+  mutate(
+    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+    earlier_block.num = earlier - 1
+  )
 
-sim_to_next_mod_2 <- brm(sim ~ earlier_block.num+ same_speaker+ (1|game1)+(1|target),
-               data=sim_to_next_data_2,
+sim_to_next_mod_2 <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
+  data = sim_to_next_data_2,
   file = here(mod_loc, "sim_to_next_2.rds"),
-               prior=conv_priors,
-               control = list(adapt_delta=.95))
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
 
@@
   filter(block1 == block2) |>
-    mutate(block.num=block1-1)
+  mutate(block.num = block1 - 1)
 
-sim_across_mod_2 <- brm(sim ~ block.num+(1|target),
-               data=sim_across_data_2,
+sim_across_mod_2 <- brm(sim ~ block.num + (1 | target),
+  data = sim_across_data_2,
   file = here(mod_loc, "sim_across_2.rds"),
-               prior=conv_priors,
-               control = list(adapt_delta=.95))
-
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
 ```
@@
   facet_wrap(~same_speaker) +
-  scale_color_solarized()+
+  scale_color_solarized() +
   labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
@@
 
-sim_across_between_form<- read_rds(here(mod_form, "sim_across_between_2.rds"))
+sim_across_between_form <- read_rds(here(mod_form, "sim_across_between_2.rds"))
 
@@
 
-sim_across_form<- read_rds(here(mod_form, "sim_across_2.rds"))
+sim_across_form <- read_rds(here(mod_form, "sim_across_2.rds"))
 
@@
 
-sim_to_next_form<- read_rds(here(mod_form, "sim_to_next_2.rds"))
+sim_to_next_form <- read_rds(here(mod_form, "sim_to_next_2.rds"))
 
@@
 
-sim_to_last_form<- read_rds(here(mod_form, "sim_to_last_2.rds"))
+sim_to_last_form <- read_rds(here(mod_form, "sim_to_last_2.rds"))
 ```
@@
 As the games progress, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
+
+TODO question for the group: do we want to include any meta analysis across the two expts?
+
+TODO do we want to try to look at whether successful descriptions are more likely to "stick" (might have low sample b/c accuracy is high)
 

### 2025-01-07 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
add diagram and do work on incorporating intro

@@
 
-TODO introduction
+<!-- intro --> 
+Learning language requires not just learning the facts of a language, but also learning how to use that language to communicate.
+One common test environment for language use is referential communication, the ability to give a description so that an interlocuter can pick out a referent (object, picture, idea) from a set of possibilities.
+Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the type of description and the amount of information they provide to their beliefs about the interlocuters knowledge state.
+
+One particular skill in referential communication is the ability to build up a new shared name for an object that initially does not have a canonical name, such as an abstract shape.
+This ability is tested in *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner.
+Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
+Success at this task requires mastery of a number of linguistic and communicative skills: the ability to produce initial descriptions of abstract shapes, monitoring for comprehension and asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
+(TODO any other skills we want to call out?)
+
+All of these are skills that children must learn as part of their acquisition of language.
+There's been debate on the relative order of different linguistic competencies, with some arguing that pragmatics and tailoring language to accomplish goals is a late addition, and others arguing that sensitivity to communicative intent is an early-arising tool that helps children acquire words.
+How well do young children do at creating appropriate referential descriptions in an iterated reference game?
+what's the developmental trajectory of producing expressions to achieve joint understanding?
+TODO fix
+
+An study on children's referential communication declared young (AGE) children incapable of the task of child-child referential communication (CITATION).
+In their paradigm, one child was given blocks in a specific order from a dispenser.
+Their task was to describe the drawing on the block so their their partner could pick out the corresponding block from an unordered set of 6 blocks.
+Then each child would put their block on a peg, and repeat, until they had stacked all 6 blocks.
+Children recieved feedback (?) and then played the game repeatedly with the same sets of blocks.
+When the drawings on the blocks were recognizable farm animals, children as young as TODO were successful.
+However, when the drawings were abstract shapes, children were not able to produce descriptions.
+TODO describe what the trajectory was as children got older: In similar experiments with older children, CITE found a gradual slope of improvement with TODO whatever.
+CITE attributed young children's failures to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
+The researchers thus declare children converse in a manner that is too egocentric to accommodate the cooperation required of referential communication.
+However, the complexity of the stacking paradigm and the large number of potential targets makes this a poor test of children's ability to produce adequate referent expressions as the task demands could tax children's limited executive function and mask children's abilities.
+
+TODO need to figure out which exactly to cite for what of: Krauss & Glucksberg 1977 ; Glucksberg et al 1966 ; Glucksberg & Krauss 1967; Krauss & Glucksberg 1969
+
+Since then, a number of studies have targeted specific skills needed for referential communication, including expectations around consistent labeling ("conversational pacts"), awareness of other's perspectives, and sensitivity to descriptive adequacy in varying contexts.
+These skills are intertwined, but studies targeting each of them suggest that children show the roots of understanding during the preschool years, albeit expressed inconsistently, and often in non-adultlike ways.
+HAHA maybe this just isn't true for near/far context stuff and they're just bad at it
+
+One piece of cooperative communication is the expectation that one person will re-use a description that worked before, and not imposing this expectation on other people who were not present.
+As listeners, preschoolers (3-5 year olds) show an expectation that the same term will be repeated as they are faster to select a target when it is referred to in a consistent way CITATIONS.
+However, this preference for familiar descriptions is applied regardless of whether they are said by the same person or a new person.
+Preschoolers are said to be "hyperconventional", often verbally protesting the use of a new term ("That's a horse, not a pony!").
+As the producers of descriptions, children show some partner-specificity, but are far from consistent.
+TODO Koyman et al had 4 and 6 year old children do a version of the paradigm from Brennan & Clark 1996.
+In a near referring context of different shoes, children used specific referring expressions ("woman's shoe") with one partner.
+The question was in later trials, when only one shoe was present in the context, would children keep the entrained term or switch back to a more general term ("shoe").
+Both 4 and 6 year olds did some of each, but 6 year olds were somewhat more likely to continue using the specific term with the same partner than with a new partner, whereas 4 year olds used expressions at the same rate regardless of partner.
+
+Expectations of consistency are entwined with knowing who knows what and being able to take the perspective of an interlocuter.
+Young children are notoriously ego-centric (Epley & Morewedge, ), but even preschoolers are able to take other's perspectives, especially when the are not under too much cognitive load (Juan, Khu, Graham).
+Across the 2-4 year olds range, children are developing sensitivity to what is given their interlocuter based on the discourse history and visual context available to their interlocuter (Matthews 2006).
+Older preschoolers are more likely to use more given forms such as pronouns when their interlocuter has more context and indefinite forms when their interlocuter has less context.
+5-year-old children can also account for privileged ground as they gaze for shorter periods at objects they know their interlocutor cannot view in a retrieval task.
+(Nilsen, Graham) An ability to overcome the egocentrism is linked to greater levels of executive functioning as measured independently.
+ 5 and 6-year-old children show sensitivity to their adult partner’s perspective in the privileged ground when tasked to help them choose objects of varying sizes while some are occluded to their interlocutor.
+(Nadig, Seviy).
+
+Another skill children need to tailoring the specificity of their utterances to the context.
+Children struggle with this, with 4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency (CITE Matthews 2012) TODO CITE LEUNG STUDY 3 as well When children must pick out one of two very similar images, 4 and 5 year olds aren't good at mentioning the relevant features, although they do better when playing in an interactive game than when not.
+This suggests that understanding the task and having the task be more similar to children's everyday experiences may be important.
+- Grigoroglou & Papafragou (2019) 
+(Morisseau, Davies, Matthews) find that 3 and 5-year-old children can accurately discriminate descriptions that are ambiguous or overly informative by lengthening their response times and asking for clarification when trying to match an expression to two similar images.
+This discrimination increases in age as 5-year-olds are more sensitive to differences in ambiguity.
+Children also show an awareness of the information their listeners need by differing their expressions when the situation demands it, as they use longer, complex descriptors when presented with two similar referents, and more vague descriptors when presented with dissimilar ones (Nadig Sedivy).
+
+In addition to testing these 3 skills mentioned above, iterated reference games also pose a couple additional challenges.
+Producing descriptions is harder since there are not canonical names for the target objects or their features.
+Additionally, depending on the set up of the game and the number of options, children's limited cognitive resources may impair their performance.
+A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e.
+fewer distractor referents.
+(Abbot-Smith, Nursoo, Croll).
+This suggests that tailored referential expression production is possible in children, but it is likely to be masked in situations that involve cognitive load.
+TODO not sure where to mention but somewhere we need to address that iterated ref games have the abstract referent thing, so producing a description is less trivial (not just choosing from a set)
+
+TODO Since CITE FOOBAR, only a few studies have revisited the question of how well children can put together different skills to communicate in an iterated reference game.
+CITE Branigan 2016 tested 72 8-10 year olds on iterated referential communication using a paradigm based on CITE Wilkes-Gibbes & Clark 1992.
+In the training period, pairs of children matched tangrams in order from a pool of 8 images.
+Qualitatively, children on average showed the classic patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions.
+However, children's level of accuracy was far below what is typical for adults and varied dramatically from pair to pair.
+
+Some of the skills required in iterated reference games are communicative rather than purely linguistic.
+In CITE BOhn 2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link (no audio).
+Children's comprehension was good and increased over repetitions, and children showed signs of agreeing on conventionalized gestures with the partner tending to use the same gestures back when roles switched.
+Conventionalization was stronger for 6 year olds than 4 year olds, based on the rating of the similarity of gestures within a dyad versus gestures between dyads.
+
+A recent study on younger children examines children's abilities to play an iterated reference game with a parent.
+CITE LEUNG had TODO child pairs of 4, 6, 8 year olds play a matching game with their parent.
+Each participant had a tablet that showed two images.
+The describer had a box around one image and their job was to pick out that image for the other participant.
+TODO describe this study better!
+Overall, there were a total of 10 target images, and the images repeated 4 times.
+In this task, even 4 year olds had an accuracy of TODO and showed ?
+an increase over repetitions.
+
+Given the potential for task demands in the TODO study and the level of success young children had with their parents at a simplified, less demanding paradigm, we revisit the question of children's ability to converge on appropriate referential expressions with each other.
+In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified paradigm to reduce cognitive demands.
+
+```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental Setup and Procedure.
+Panel A shows the experimental setup with the teller and guesser across the table from each other.
+Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
+Panel C shows the procedure for Experiment 1; within in critical block targets were ordered randomly.
+Panel D shows the procedure for Experiment 2.
+\\label{game}", cache=FALSE}
+
+knitr::include_graphics("diagram.pdf")
+
+# edit diagram at 
+# https://docs.google.com/drawings/d/1gio1CNSedECJGrkrNDUi_AydPZcYzESfl4WhZD2l9Bc/edit?usp=sharing
+```
 
@@
 ## Materials
-For the target stimuli, we used four of the ten tangram images from @leung, chosen based on visual dissimilarity.
+For the target stimuli, we used four of the ten tangram images from @leung, chosen based on visual dissimilarity (Figure \ref{game}B).
 We coded the matching game using Empirica and hosted it on a lab server (CITATION).
@@
 Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game.
-Children sat across a table from each other, each with a tablet in front of them.
+Children sat across a table from each other, each with a tablet in front of them (Figure \ref{game}A).
 On each trial, one child saw two images, one of them in a black box, and was asked to "say what they saw" in the black box so their partner (and Smurfy) could tap the corresponding image.
@@
 
-Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 targets.
+Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 targets (Figure \ref{game}C).
 Targets were randomly paired with another of the critical images as the foil.
@@
 The entire interaction was video-recorded.
-
-TODO DIAGRAM OF EXPERIMENT + Picture of children playing the game as top bar
-
-```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TODO \\label{game}", cache=FALSE}
-```
 
@@
 We used a total of 4 practice trials, designed to transition from easily recognizable shapes to slightly blockier black and wide shapes to smooth the transition to the critical trials.
+
 ## Procedure
-The procedure was much the same.
+The procedure was much the same (Figure \ref{game}D).
 We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in the first experiment).
 The smurfy puppet was swapped out for a more attractive smurfy stuffed animal.
-The experimental script was fully written out so children all recieved the same instructions.
+The experimental script was fully written out so children all received the same instructions.
 To prevent experimenter's influencing children's descriptions or understanding of descriptions, we wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
 TODO link to materials for where the script is.
+The idea was that the experimenter would help with understanding of game mechanics such as who's turn it was to tell and who should press the screen to move the game along if children stalled, but would not provide or repeat any content about the images, what to ask, or whether a description was adequate.
 

### 2025-01-08 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
add more models, work on discussion a bit

@@
 ---
-title: "TODO"
-bibliography: library.bib
+title: "TODO Preschoolers form conventional pacts with each other to communicate about novel shapes"
+bibliography: preschool-tangrams.bib
 csl: apa7.csl
@@
 abstract: >
-    TODO abstract
-    Communication is important and stuff and kids need to learn it to.
-One task that illustrates the rich communicative capabilities is iterated reference games, where adults jointly converge on mutually understandable names for novel objects.
-Some prominent early experiments where children did not succeed led to the belief that children could not do this task.
-TODO FIX.
-However, more recent work with lessened task demands suggests that children may be more capable than previously thought.
-Here, we test children's communicative adaption to a partner with iterated reference games played by pairs of 4-5 year old preschoolers.
-Across a total of N pairs of children, we find children can accurately pick out referents for each other and that they are sensitive to the descriptions of their partners.
-TODO wrap up sentence.
+  Learning language requires not just learning the facts of a language, but also learning how to use that language to communicate.
+Iterated reference games require rich communicative skills as participants jointly converge on mutually understandable names for initially novel objects.
+Some classical experiments with young children are interpreted as showing that 4-5 year old children were incapable of succeeding at iterated references games with eac other.
+More recent work on young children and their parents using simpler paradigms reveal greater abilities even in young children.
+Here, we revisit the question of young children's referential communicative abilitlies with iterated reference games played between pairs of 4-5 year old preschoolers.
+Across 51 pairs of children (TODO double check N), we find that children were Y% accurate, and that they often used descriptions similar to their partner's.
+Preschoolers' capacity to construct effective referring expressions in novel contexts is consistent with a larger picture of early emerging ability to use language in pragmatic and communicative ways.
+TODO fix wrap up sentence!
 
 keywords: >
-    TODO keywords
+    development; iterated reference game; TODO keywords
 
@@
 mod_form <- "code/models/formulae"
-time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX
+time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple")
 
@@
 One particular skill in referential communication is the ability to build up a new shared name for an object that initially does not have a canonical name, such as an abstract shape.
-This ability is tested in *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner.
+This ability is tested in *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
 Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
 Success at this task requires mastery of a number of linguistic and communicative skills: the ability to produce initial descriptions of abstract shapes, monitoring for comprehension and asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
-(TODO any other skills we want to call out?)
-
-All of these are skills that children must learn as part of their acquisition of language.
-There's been debate on the relative order of different linguistic competencies, with some arguing that pragmatics and tailoring language to accomplish goals is a late addition, and others arguing that sensitivity to communicative intent is an early-arising tool that helps children acquire words.
-How well do young children do at creating appropriate referential descriptions in an iterated reference game?
-what's the developmental trajectory of producing expressions to achieve joint understanding?
-TODO fix
-
-An study on children's referential communication declared young (AGE) children incapable of the task of child-child referential communication (CITATION).
+
+All of these are skills that children must learn as part of their acquisition of language, but there is debate on the acquisition trajectories of different linguistic and communicative skills.
+Studying how children play iterated reference games can provide inisght into the developmental trajectories of the ability to produce expressions to achieve joint understanding.
+TODO this paragraph needs help
+(I'm unsure if we want to explicitly connect to the pragmatics-last v pragmatics-always debate in how ordered kid language acquisition is -- maybe we leave it and mention it in the discussion?)
+
+A study on children's referential communication declared 4-5 year old preschoolers incapable of the task of child-child referential communication [@glucksberg1966].
 In their paradigm, one child was given blocks in a specific order from a dispenser.
@@
 Then each child would put their block on a peg, and repeat, until they had stacked all 6 blocks.
-Children recieved feedback (?) and then played the game repeatedly with the same sets of blocks.
-When the drawings on the blocks were recognizable farm animals, children as young as TODO were successful.
-However, when the drawings were abstract shapes, children were not able to produce descriptions.
-TODO describe what the trajectory was as children got older: In similar experiments with older children, CITE found a gradual slope of improvement with TODO whatever.
-CITE attributed young children's failures to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
-The researchers thus declare children converse in a manner that is too egocentric to accommodate the cooperation required of referential communication.
-However, the complexity of the stacking paradigm and the large number of potential targets makes this a poor test of children's ability to produce adequate referent expressions as the task demands could tax children's limited executive function and mask children's abilities.
-
-TODO need to figure out which exactly to cite for what of: Krauss & Glucksberg 1977 ; Glucksberg et al 1966 ; Glucksberg & Krauss 1967; Krauss & Glucksberg 1969
+Children then played the game repeatedly with the same sets of blocks.
+The 4-5 year old children were successful on practice trials where the blocks had recognizable drawings and they were able to see each other.
+Younger children (3-4 years old) were not able to understand the instructions and complete the practice trials.
+However, even children who were able to do the practice trials were not successful when they could not see the other child's blocks and the drawings were abstract shapes; children produced referential descriptions that were not clearly pick out the target shape to the other child.
+In similar experiments with older children, children improved gradually across age (up to the teen years) both at their initial accuracy and on their improvement over repetitions, although even the 9th grade sample was noticeably worse than their adult college student sample [@glucksberg1967; @krauss1969].
+@glucksberg1966 attributed the young children's failures to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
+The researchers thus declared children converse in a manner that is too egocentric to accommodate the cooperation required of referential communication.
+However, the complexity of the stacking paradigm and the large number of potential targets meant that this was not a clean test of children's ability to produce adequate referent expressions as the experiment had substantial task demands that could mask children's abilities.
 
@@
 These skills are intertwined, but studies targeting each of them suggest that children show the roots of understanding during the preschool years, albeit expressed inconsistently, and often in non-adultlike ways.
-HAHA maybe this just isn't true for near/far context stuff and they're just bad at it
+HAHA maybe this just isn't true for near/far context stuff and they're just bad at it TODO fix
 
 One piece of cooperative communication is the expectation that one person will re-use a description that worked before, and not imposing this expectation on other people who were not present.
-As listeners, preschoolers (3-5 year olds) show an expectation that the same term will be repeated as they are faster to select a target when it is referred to in a consistent way CITATIONS.
+As listeners, preschoolers (3-5 year olds) show an expectation that the same term will be repeated as they are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010].
 However, this preference for familiar descriptions is applied regardless of whether they are said by the same person or a new person.
-Preschoolers are said to be "hyperconventional", often verbally protesting the use of a new term ("That's a horse, not a pony!").
+Preschoolers are said to be "hyperconventional", often verbally protesting the use of a new term ("That's a horse, not a pony!") [@matthews2010].
 As the producers of descriptions, children show some partner-specificity, but are far from consistent.
-TODO Koyman et al had 4 and 6 year old children do a version of the paradigm from Brennan & Clark 1996.
+@koymen2014 had 4 and 6 year old children do a version of the paradigm from @brennan1996.
 In a near referring context of different shoes, children used specific referring expressions ("woman's shoe") with one partner.
@@
 Expectations of consistency are entwined with knowing who knows what and being able to take the perspective of an interlocuter.
-Young children are notoriously ego-centric (Epley & Morewedge, ), but even preschoolers are able to take other's perspectives, especially when the are not under too much cognitive load (Juan, Khu, Graham).
-Across the 2-4 year olds range, children are developing sensitivity to what is given their interlocuter based on the discourse history and visual context available to their interlocuter (Matthews 2006).
+Young children are notoriously ego-centric [@epley2004; TODO there are probably better things to cite here instead], but even preschoolers are able to take other's perspectives, especially when the are not under too much cognitive load [@sanjuan2015].
+Across the 2-4 year olds range, children are developing sensitivity to what is given their interlocuter based on the discourse history and visual context available to their interlocuter [@matthews2006].
 Older preschoolers are more likely to use more given forms such as pronouns when their interlocuter has more context and indefinite forms when their interlocuter has less context.
-5-year-old children can also account for privileged ground as they gaze for shorter periods at objects they know their interlocutor cannot view in a retrieval task.
-(Nilsen, Graham) An ability to overcome the egocentrism is linked to greater levels of executive functioning as measured independently.
- 5 and 6-year-old children show sensitivity to their adult partner’s perspective in the privileged ground when tasked to help them choose objects of varying sizes while some are occluded to their interlocutor.
-(Nadig, Seviy).
+4 to 6 year old children show sensitivity to what their adult partner can or cannot see when picking out objects where there is a size competitor [@nadig2002, @nilsen2009].
+Children are more likely to use a more specific reference expression with a size modifier when their partner can see the size competitor than when their partner cannot see the competitor, but children produce the modifier inconsistently even when it is necessary to distinguish the target.
 
 Another skill children need to tailoring the specificity of their utterances to the context.
-Children struggle with this, with 4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency (CITE Matthews 2012) TODO CITE LEUNG STUDY 3 as well When children must pick out one of two very similar images, 4 and 5 year olds aren't good at mentioning the relevant features, although they do better when playing in an interactive game than when not.
+Children struggle with this, with 4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency [@matthews2012]  Children ages 4-8 often produce the same label for a target object regardless of whether the context is close or far, leading to descriptions that are underinformative for close contexts [@leung2024].
+When children must pick out one of two very similar images, 4 and 5 year olds aren't good at mentioning the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
 This suggests that understanding the task and having the task be more similar to children's everyday experiences may be important.
-- Grigoroglou & Papafragou (2019) 
-(Morisseau, Davies, Matthews) find that 3 and 5-year-old children can accurately discriminate descriptions that are ambiguous or overly informative by lengthening their response times and asking for clarification when trying to match an expression to two similar images.
-This discrimination increases in age as 5-year-olds are more sensitive to differences in ambiguity.
-Children also show an awareness of the information their listeners need by differing their expressions when the situation demands it, as they use longer, complex descriptors when presented with two similar referents, and more vague descriptors when presented with dissimilar ones (Nadig Sedivy).
+By age 5, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances, although children still do not ask for clarification all the time [@morisseau2013].
+Their receptive sensitivity to ambiguity is sometimes mirrored in production.
 
@@
 A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e.
-fewer distractor referents.
-(Abbot-Smith, Nursoo, Croll).
+fewer distractor referents [@abbot-smith2016].
 This suggests that tailored referential expression production is possible in children, but it is likely to be masked in situations that involve cognitive load.
-TODO not sure where to mention but somewhere we need to address that iterated ref games have the abstract referent thing, so producing a description is less trivial (not just choosing from a set)
-
-TODO Since CITE FOOBAR, only a few studies have revisited the question of how well children can put together different skills to communicate in an iterated reference game.
-CITE Branigan 2016 tested 72 8-10 year olds on iterated referential communication using a paradigm based on CITE Wilkes-Gibbes & Clark 1992.
+
+Since @glucksberg1966, only a few studies have revisited the question of how well children can put together different skills to communicate in an iterated reference game.
+@branigan2016 tested 72 8-10 year olds on iterated referential communication using a paradigm based on @wilkes-gibbs1992.
 In the training period, pairs of children matched tangrams in order from a pool of 8 images.
@@
 Some of the skills required in iterated reference games are communicative rather than purely linguistic.
-In CITE BOhn 2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link (no audio).
+In @bohn2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link (no audio).
 Children's comprehension was good and increased over repetitions, and children showed signs of agreeing on conventionalized gestures with the partner tending to use the same gestures back when roles switched.
@@
 A recent study on younger children examines children's abilities to play an iterated reference game with a parent.
-CITE LEUNG had TODO child pairs of 4, 6, 8 year olds play a matching game with their parent.
-Each participant had a tablet that showed two images.
-The describer had a box around one image and their job was to pick out that image for the other participant.
-TODO describe this study better!
-Overall, there were a total of 10 target images, and the images repeated 4 times.
-In this task, even 4 year olds had an accuracy of TODO and showed ?
-an increase over repetitions.
-
-Given the potential for task demands in the TODO study and the level of success young children had with their parents at a simplified, less demanding paradigm, we revisit the question of children's ability to converge on appropriate referential expressions with each other.
+@leung2024 had 63 children ages 4-8 year olds play a matching game with their parents.
+This experiment was implemented on tablets, so on each trial, participants saw two imiages on the tablet.
+The describer's tablet had a box around one image and their task was to verbally describe the image so their partner could select the corresponding image on their tablet.
+The target images were tangrams, with a total of 10 target images, which repeated across 4 blocks for a total of 40 trials per pair.
+In this task, even 4 year olds had an initial accuracy above 80% which rose to above 90% on the later repetitions.
+
+Given the potential for task demands in @glucksberg1966 and the level of success young children had with their parents at a simplified, less demanding paradigm, we revisit the question of children's ability to converge on appropriate referential expressions with each other.
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified paradigm to reduce cognitive demands.
@@
 
+(Note: we should talk about whether it makes more sense to do all the methods and then all the results versus the current expt 1 than expt 2 layout.)
+
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load that can hide early abilities (TODO cite something for this), so we used a simple paradigm where an experimenter could scaffold the children's iteraction as needed.
-We adapted the experimental framework from @leung, but further simplified it by reducing the total pool of targets and the number of trials children completed.
+Young children can be very sensitive to task demands and cognitive load that can hide early abilities [CITATION WANTED], so we used a simple paradigm where an experimenter could scaffold the children's iteraction as needed.
+We adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
 
@@
 ## Materials
-For the target stimuli, we used four of the ten tangram images from @leung, chosen based on visual dissimilarity (Figure \ref{game}B).
-We coded the matching game using Empirica and hosted it on a lab server (CITATION).
+For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B).
+We coded the matching game using Empirica and hosted it on a lab server [@almaatouq2020].
 We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate to other websites or applications during the game.
@@
 Children's selections and the time to selection were recorded from the experiment software.
-Children's descriptions were transcribed from the video-recording, using Whisper (CITE) for the first pass and then hand-corrected by experimenters.
+Children's descriptions were transcribed from the video-recording, using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters.
 Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used.
@@
 How long of a description is initially warranted depends on how iconic or easily describable the shape is, as well as how large and close the contrast set is.
-In this case, the low number of options to distinguish and relatively easy to describe shapes may mean long initial descriptions are less necessary (todo mention how long the similar things are in leung).
-However, young children may also choose to provide shorter descriptions than adults would.
-
-TODO add anecdotes here with some examples of children's descriptions.
+In this case, the low number of options to distinguish and relatively easy to describe shapes may mean long initial descriptions are less necessary.
+(The adult controls in @leung2024 use shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.) However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
 
@@
 
-While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length (cite boyce 2024).
+While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length [@boyce2024].
 With only 3 repetitions for each target, we do not have enough data to look at convergence over time.
@@
 
-Following cite Boyce et al 2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions.
+Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions.
 We compare the semantic similarities between descriptions of the same target based on who produced the description.
@@
 
-TODO could provide anecdotes about children's coordination approaches
-
 # Experiment 1 Discussion
+
+In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter.
+As our goal here was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures, children recieved a lot of scaffolding around the experimental interaction.
+Sometimes, this included experimenter echoing children's descriptions "Child A says they see a ...
+Do you see a ...?" which could potentially influence children's responses.
+In Experiment 2, we repeat the same paradigm, with tighter experimenter controls and a larger sample size.
 
@@
 
-To look at semantic distance between utterances, we again operationalize similarity between pairs of utterances as the cosine similarity between their SBERT embeddings (CITE SBERT).
+To look at semantic distance between utterances, we again operationalize similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019].
 
@@
 
+# Post hoc results
+
+## TODO Meta-analytic placeholder 
 TODO question for the group: do we want to include any meta analysis across the two expts?
 
-TODO do we want to try to look at whether successful descriptions are more likely to "stick" (might have low sample b/c accuracy is high)
+```{r, eval=F}
+
+acc_priors <- c(
+  set_prior("normal(0,1)", class = "b"),
+  set_prior("normal(0,1)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+acc_mod_data_meta <-  clean_1 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice") |> mutate(gameId=game) |> bind_rows(clean_2 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 4
+  ) |>
+  filter(type != "practice"))
+
+acc_mod_both <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
+  family = bernoulli(link = "logit"),
+  data = acc_mod_data_meta,
+  file = here(mod_loc, "acc_meta.rds"),
+  prior = acc_priors,
+  control = list(adapt_delta = .95)
+)
+
+```
+
+```{r, eval=F}
+description_priors <- c(
+  set_prior("normal(5,10)", class = "Intercept"),
+  set_prior("normal(0,5)", class = "b"),
+  set_prior("normal(0,5)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+description_mod_data_meta <- clean_1 |>
+  mutate(
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")|> mutate(gameId=game) |> bind_rows(
+ clean_2 |>
+  mutate(
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice"))
+
+description_mod_meta <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
+  data = description_mod_data_meta,
+  file = here(mod_loc, "description_meta.rds"),
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
+
+```
+
+```{r, eval=F}
+conv_priors <- c(
+  set_prior("normal(.5,.2)", class = "Intercept"),
+  set_prior("normal(0,.1)", class = "b"),
+  set_prior("normal(0,.05)", class = "sd")
+)
+
+sim_across_between_data_2 <- sims_2 |>
+  mutate(same_game = case_when(
+    game1 == game2 ~ 1,
+    T ~ 0
+  )) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
+
+sim_across_between_data_1 <- sims_1 |>
+  mutate(same_game = case_when(
+    game1 == game2 ~ 1,
+    T ~ 0
+  )) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
+
+sim_across_between_data_meta <- sim_across_between_data_2 |> bind_rows(sim_across_between_data_1)
+
+sim_across_between_mod_meta <- brm(sim ~ same_game + same_speaker + (1 | target),
+  data = sim_across_between_data_meta,
+  file = here(mod_loc, "sim_across_between_meta.rds"),
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
+
+sim_to_next_data_meta <- sims_2 |> bind_rows(sims_1) |> 
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier + 1) |>
+  mutate(
+    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+    earlier_block.num = earlier - 1
+  )
+
+sim_to_next_mod_meta <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
+  data = sim_to_next_data_meta,
+  file = here(mod_loc, "sim_to_next_meta.rds"),
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
+
+```
+
+```{r}
+acc_mod_meta <- read_rds(here(mod_results, "acc_meta.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
+
+description_mod_meta <- read_rds(here(mod_results, "description_meta.rds"))
+
+sim_across_between_mod_meta <- read_rds(here(mod_results, "sim_across_between_meta.rds"))
+
+sim_to_next_mod_meta <- read_rds(here(mod_results, "sim_to_next_meta.rds"))
+```
+
+As the two experiments were similar to one another, we re-run models pooling across the two experiments. 
+Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`.
+
+The average length of descriptions on the first trial was `r stats_text(description_mod_meta, 1)` words and description length was relatively stable over time (`r stats_text(description_mod_meta, 2)`, Figure \ref{length}).
+
+Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
+Since the two expeirments had different length games, we do not want to pool them in comparing to the last block, but we can pool them comparing to the next block.
+
+ The distance between adjacent block utterances is relatively constant: `r sim_to_next_mod_meta |> stats_text(2,3)`.
+
+Query -- I did not include experiment # as a random effect here, and maybe should have?
+I can rerun if we decide we want this analysis and think expt # as a random effect with game nested within it would be useful.
+
+## using accuracy to predict similarities
+
+```{r}
+# try visual first
+
+acc_sim_1 <- clean_1 |>
+  filter(type != "practice") |> mutate(game1=game) |> 
+  mutate(correct=ifelse(correct, 1,0),
+         earlier_trial=trial+2) |> 
+  select(earlier_trial, target, correct, game1)
+
+acc_sim_2 <- clean_2 |>
+  filter(type != "practice") |> select(earlier_trial=trial, target, correct, game1=gameId)
+
+sim_acc_2 <- sims_2 |> 
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
+  filter(later == earlier + 1) |>
+  mutate(
+    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+    earlier_block.num = earlier - 1
+  ) |> 
+  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker) |> 
+  left_join(acc_sim_2)
+
+sim_acc_1 <- sims_1 |> 
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
+  filter(later == earlier + 1) |>
+  mutate(
+    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
+    earlier_block.num = earlier - 1
+  ) |> 
+  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker) |> 
+  left_join(acc_sim_1)
+
+#ggplot(sim_acc_1, aes(x=earlier, y=sim, color=as.factor(correct)))+geom_point(alpha=.4, position=position_dodge(.4))+stat_summary()
+
+#ggplot(sim_acc_2, aes(x=earlier, y=sim, color=as.factor(correct)))+geom_point(alpha=.4, position=position_dodge(.4))+stat_summary()
+
+```
+
+```{r, eval=F}
+conv_priors <- c(
+  set_prior("normal(.5,.2)", class = "Intercept"),
+  set_prior("normal(0,.1)", class = "b"),
+  set_prior("normal(0,.05)", class = "sd")
+)
+sim_acc_to_next_data <- sim_acc_1 |> bind_rows(sim_acc_2)
+sim_acc_to_next_mod_meta <- brm(sim ~ earlier_block.num*correct + same_speaker*correct + (1 | game1) + (1 | target),
+  data = sim_acc_to_next_data,
+  file = here(mod_loc, "sim_acc_to_next_meta.rds"),
+  prior = conv_priors,
+  control = list(adapt_delta = .95)
+)
+
+# more stickiness if correct, no clear interactions 
+# being correct has a similar size effect to same speaker
+```
+
+```{r}
+sim_acc_to_next_mod_meta <- read_rds(here(mod_results, "sim_acc_to_next_meta.rds"))
+sim_acc_to_next_form <- read_rds(here(mod_form, "sim_acc_to_next_meta.rds"))
+
+```
+
+As a post-hoc measure, we look at whether descriptions that elicit correct selections have more staying power than descriptions that do not elicit correct selections.
+Here we operationalize "staying power" as similarity to the next block utterance, and we model with `r form(sim_acc_to_next_form)`.
+Correct utterances have an increase in similarity with the next block by `r stats_text(sim_acc_to_next_mod_meta, 2, 3)` with no substantial interaction with block number or same speaker.
+Numerically, the size of the boost for being correct is similar to the size of the boost from coming form the same speaker `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
+
+## Not sure where to put, cute things kids said
+
+Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
+game 62 is example of played well and fast (also kids bffs) Children sometimes asked clarifying questions of their partner (for instance, when the partner described the shape as "human" or "person", children sometimes sought clarification and arrived at descriptions such as "...
+see game 81, 79?)
+
+game 66 triangle feet
+
+game 61 generally good
+
+For utterances just look through for examples TODO
 
@@
 
-Summary of experiments
+Across 2 experiments and 51 pairs of 4 and 5 year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
+Children varied substantially in how easily they understood the task and what sorts of descriptions they produced, but overall accuracy was high (NUMBER), indicating that children were able to produce adequate descriptions.
+Additionally, children's utterances were more similar to the descriptions their partner used than to other children, and in experiment 2, children's utterances showed signs of converging toward conceptual pacts.
+This provides evidence that at least some preschoolers, in an environment with low task demands and support for understanding the structure of the task, can succeed at iterated referential communication.
+While this task is substantially scaled down relative to measures used for adult competence, and children do not display all the typical trends (i.e no sign of shortening of referential expression), it is a sign that the relevant skills are continuously improving across childhood, rather than only appearing later on.
 
@@
 There are other parts of the coordination where help was provided in children seemed stuck or confused, such as when to make a choice or ask for more information.
+We think this is reasonable, because our goal is on coordination of referring expressions, not larger coordination on tasks or joint instruction following.
 
@@
 
-This work (along with other work on children's referential communicaiton) suggests that there's a more gradual development
+This work (along with other work on children's referential communicaiton) suggests that there's a more gradual development of children's communicative and linguistic skills.
+Rather than an abrupt acquisition of a skill, children seem to be refining their abilities across childhood, leading to to the presence of skills early on that become more consistently expressed later.
 Has implications for how we think about children's language development.
-There's debate over how that is ordered and whether communication/pragmatics is a final stage, or how all the stages are boostrapped (CITATIONS).
+There's debate over how that is ordered and whether communication/pragmatics is a final stage, or how all the stages are boostrapped (CITATIONS WANTED).
 This early ability to use language for communicative purposes is more consistent with the early pragmatics viewpoint.
-
 Suggestive of a gradual development where children's capabilities are increasing for a wide amount of childhood, as their working memory capacity and executive function improve and they are able to better track other's states of knowledge and keep track of wider arrays of images.

### 2025-01-08 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
full draft

@@
 Iterated reference games require rich communicative skills as participants jointly converge on mutually understandable names for initially novel objects.
-Some classical experiments with young children are interpreted as showing that 4-5 year old children were incapable of succeeding at iterated references games with eac other.
+Some classical experiments with young children are interpreted as showing that 4-5 year old children were incapable of succeeding at iterated references games with each other.
 More recent work on young children and their parents using simpler paradigms reveal greater abilities even in young children.
-Here, we revisit the question of young children's referential communicative abilitlies with iterated reference games played between pairs of 4-5 year old preschoolers.
+Here, we revisit the question of young children's referential communicative abilites with iterated reference games played between pairs of 4-5 year old preschoolers.
 Across 51 pairs of children (TODO double check N), we find that children were Y% accurate, and that they often used descriptions similar to their partner's.
@@
 TODO fix wrap up sentence!
+TODO possibly shorten!
 
@@
 output: cogsci2024::cogsci_paper
-#final-submission: \cogscifinalcopy
+final-submission: \cogscifinalcopy
 ---
@@
 library(ggthemes)
+library(kableExtra)
 rstan_options(auto_write = TRUE)
@@
 <!-- intro -->
+
+[way to long, please chop down]
 Learning language requires not just learning the facts of a language, but also learning how to use that language to communicate.
-One common test environment for language use is referential communication, the ability to give a description so that an interlocuter can pick out a referent (object, picture, idea) from a set of possibilities.
+One common test environment for language use is referential communication, the ability to give a description so that an interlocuter can pick out a target referent from a set of possibilities.
 Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the type of description and the amount of information they provide to their beliefs about the interlocuters knowledge state.
 
-One particular skill in referential communication is the ability to build up a new shared name for an object that initially does not have a canonical name, such as an abstract shape.
-This ability is tested in *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
+One test of referential communication skill is *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
 Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
-Success at this task requires mastery of a number of linguistic and communicative skills: the ability to produce initial descriptions of abstract shapes, monitoring for comprehension and asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
+Success at this task requires mastery of a number of linguistic and communicative skills including producing adequate initial descriptions, monitoring for comprehension and asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
 
 All of these are skills that children must learn as part of their acquisition of language, but there is debate on the acquisition trajectories of different linguistic and communicative skills.
-Studying how children play iterated reference games can provide inisght into the developmental trajectories of the ability to produce expressions to achieve joint understanding.
-TODO this paragraph needs help
-(I'm unsure if we want to explicitly connect to the pragmatics-last v pragmatics-always debate in how ordered kid language acquisition is -- maybe we leave it and mention it in the discussion?)
+Studying how children play iterated reference games can provide insight into the developmental trajectories of the ability to produce expressions to achieve joint understanding.
 
 A study on children's referential communication declared 4-5 year old preschoolers incapable of the task of child-child referential communication [@glucksberg1966].
-In their paradigm, one child was given blocks in a specific order from a dispenser.
-Their task was to describe the drawing on the block so their their partner could pick out the corresponding block from an unordered set of 6 blocks.
-Then each child would put their block on a peg, and repeat, until they had stacked all 6 blocks.
-Children then played the game repeatedly with the same sets of blocks.
-The 4-5 year old children were successful on practice trials where the blocks had recognizable drawings and they were able to see each other.
-Younger children (3-4 years old) were not able to understand the instructions and complete the practice trials.
-However, even children who were able to do the practice trials were not successful when they could not see the other child's blocks and the drawings were abstract shapes; children produced referential descriptions that were not clearly pick out the target shape to the other child.
-In similar experiments with older children, children improved gradually across age (up to the teen years) both at their initial accuracy and on their improvement over repetitions, although even the 9th grade sample was noticeably worse than their adult college student sample [@glucksberg1967; @krauss1969].
+In their paradigm, one child was given a set of 6 blocks in a specific order.
+Their task was to describe the images on the block so their partner could pick out their corresponding block.
+As children described and selected blocks, they stacked them on pegs, so that the order of blocks on the peg was a measure of correct matching.
+4-5 year old children were successful on practice trials where the blocks had drawings of farm animals and they had visual access to each others blocks; however, younger children (3-4 year olds) were not even able to understand the instructions and complete the practice trials.
+On the critcal trials, the 4-5 year old children were tested with abstract shapes on the blocks and no visual access to the other child's blocks.
+In this condition, even after multiple rounds with the same images, children were not able to produce referential descriptions that allowed for success at the matching task.
+
+Similar experiments with older children indicated a gradual improvement across age (up to the teen years) both at their initial accuracy and on their improvement over repetitions, although even the 9th grade sample was noticeably worse than their adult college student sample [@glucksberg1967; @krauss1969].
 @glucksberg1966 attributed the young children's failures to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
-The researchers thus declared children converse in a manner that is too egocentric to accommodate the cooperation required of referential communication.
 However, the complexity of the stacking paradigm and the large number of potential targets meant that this was not a clean test of children's ability to produce adequate referent expressions as the experiment had substantial task demands that could mask children's abilities.
@@
 Since then, a number of studies have targeted specific skills needed for referential communication, including expectations around consistent labeling ("conversational pacts"), awareness of other's perspectives, and sensitivity to descriptive adequacy in varying contexts.
-These skills are intertwined, but studies targeting each of them suggest that children show the roots of understanding during the preschool years, albeit expressed inconsistently, and often in non-adultlike ways.
-HAHA maybe this just isn't true for near/far context stuff and they're just bad at it TODO fix
-
-One piece of cooperative communication is the expectation that one person will re-use a description that worked before, and not imposing this expectation on other people who were not present.
-As listeners, preschoolers (3-5 year olds) show an expectation that the same term will be repeated as they are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010].
-However, this preference for familiar descriptions is applied regardless of whether they are said by the same person or a new person.
-Preschoolers are said to be "hyperconventional", often verbally protesting the use of a new term ("That's a horse, not a pony!") [@matthews2010].
+A number of these studies have found roots of understanding emerging during hte preschool years, albeit with inconsistent expression.
+
+Young children show some sensitivity to the expectation that during cooperative communicate a person will likely re-use their own descriptions.
+As listeners, preschoolers (3-5 year olds) are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010].
+However, this preference for familiar descriptions is applied regardless of whether they are said by the same person or a new person, and they sometimes verbally protesting the use of a new term ("That's a horse, not a pony!") [@matthews2010].
 As the producers of descriptions, children show some partner-specificity, but are far from consistent.
 @koymen2014 had 4 and 6 year old children do a version of the paradigm from @brennan1996.
-In a near referring context of different shoes, children used specific referring expressions ("woman's shoe") with one partner.
-The question was in later trials, when only one shoe was present in the context, would children keep the entrained term or switch back to a more general term ("shoe").
-Both 4 and 6 year olds did some of each, but 6 year olds were somewhat more likely to continue using the specific term with the same partner than with a new partner, whereas 4 year olds used expressions at the same rate regardless of partner.
-
-Expectations of consistency are entwined with knowing who knows what and being able to take the perspective of an interlocuter.
+Children used specific referring expressions ("woman's shoe") with one partner in a visual context of different types of shes.
+In later trials when only one shoe was present and thus a general expression like "shoe" was adequate, 6 year olds were somewhat more likely to continue using the specific term with the same partner than with a new partner, whereas 4 year olds used both specific and general expressions at the same rate regardless of partner.
+
+Expectations of consistency are entwined with being able to take the perspective of an interlocuter.
 Young children are notoriously ego-centric [@epley2004; TODO there are probably better things to cite here instead], but even preschoolers are able to take other's perspectives, especially when the are not under too much cognitive load [@sanjuan2015].
@@
 Older preschoolers are more likely to use more given forms such as pronouns when their interlocuter has more context and indefinite forms when their interlocuter has less context.
-4 to 6 year old children show sensitivity to what their adult partner can or cannot see when picking out objects where there is a size competitor [@nadig2002, @nilsen2009].
-Children are more likely to use a more specific reference expression with a size modifier when their partner can see the size competitor than when their partner cannot see the competitor, but children produce the modifier inconsistently even when it is necessary to distinguish the target.
-
-Another skill children need to tailoring the specificity of their utterances to the context.
-Children struggle with this, with 4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency [@matthews2012]  Children ages 4-8 often produce the same label for a target object regardless of whether the context is close or far, leading to descriptions that are underinformative for close contexts [@leung2024].
+4-6 year old children are more likely to use a more specific reference expression with a size modifier when their partner can see the size competitor than when their partner cannot see the competitor, but children produce the modifier inconsistently even when it is necessary to distinguish the target [@nadig2002; @nilsen2009].
+
+The skill children seem to struggle the most with is appropriately tailoring the specificity of their utterances to the visual context.
+4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency [@matthews2012].
+Children ages 4-8 often produce the same label for a target object regardless of whether the context is close or far, leading to descriptions that are underinformative for close contexts [@leung2024].
 When children must pick out one of two very similar images, 4 and 5 year olds aren't good at mentioning the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
-This suggests that understanding the task and having the task be more similar to children's everyday experiences may be important.
 By age 5, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances, although children still do not ask for clarification all the time [@morisseau2013].
-Their receptive sensitivity to ambiguity is sometimes mirrored in production.
-
-In addition to testing these 3 skills mentioned above, iterated reference games also pose a couple additional challenges.
-Producing descriptions is harder since there are not canonical names for the target objects or their features.
-Additionally, depending on the set up of the game and the number of options, children's limited cognitive resources may impair their performance.
+
 A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e.
@@
 Some of the skills required in iterated reference games are communicative rather than purely linguistic.
-In @bohn2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link (no audio).
-Children's comprehension was good and increased over repetitions, and children showed signs of agreeing on conventionalized gestures with the partner tending to use the same gestures back when roles switched.
-Conventionalization was stronger for 6 year olds than 4 year olds, based on the rating of the similarity of gestures within a dyad versus gestures between dyads.
-
-A recent study on younger children examines children's abilities to play an iterated reference game with a parent.
-@leung2024 had 63 children ages 4-8 year olds play a matching game with their parents.
-This experiment was implemented on tablets, so on each trial, participants saw two imiages on the tablet.
-The describer's tablet had a box around one image and their task was to verbally describe the image so their partner could select the corresponding image on their tablet.
+In @bohn2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link without audio.
+Children's comprehension was good and increased over repetitions, and children, especially the 6 year olds,  showed signs of agreeing on conventionalized gestures with their partner.
+
+A recent study on younger children examined children's abilities to play an iterated reference game with a parent.
+@leung2024 had 63 children ages 4-8 year olds play a matching game on tablets, where, on each trial, participants saw two images.
+The describer's tablet had a box around one image and their task was to verbally describe the image so their partner could select the corresponding image.
 The target images were tangrams, with a total of 10 target images, which repeated across 4 blocks for a total of 40 trials per pair.
@@
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load that can hide early abilities [CITATION WANTED], so we used a simple paradigm where an experimenter could scaffold the children's iteraction as needed.
-We adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
+Young children can be very sensitive to task demands and cognitive load that can hide early abilities [CITATION WANTED], so we adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
 
@@
 On each trial, one child saw two images, one of them in a black box, and was asked to "say what they saw" in the black box so their partner (and Smurfy) could tap the corresponding image.
-The guesser saw the same two images (in a randomized order), but with neither boxed.
+The guesser saw the same two images (in a randomized order).
 Upon tapping an image, both children received feedback in form of a smiley or frowny face and an audible sound.
@@
 
-## Statistical analysis
-
-Statistical analyses were run in brms with weakly informative priors. We present the estimate and 95% credible intervals.
+Statistical analyses were run in TODO CITE brms with weakly informative priors. We present the estimate and 95% credible intervals.
 
@@
 
-Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully (CITATION).
+Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
 As shown in Figure \ref{acc}, children were above chance in their selections.
@@
 
-As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
+[less important, could drop] As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
 We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
@@
 
-In iterated reference games in adults, a canonical finding is that the length of descriptions goes from long to short over repeated references to the initially hard to describe shapes.
+In iterated reference games in adults, a canonical finding is that the length of descriptions goes from long to short over repeated references to the initially hard to describe shapes [@clark1986; @hawkins2020b; @boyce2024].
 We looked at how long the descriptions the children used were to see if the same trend occurred.
@@
 While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length [@boyce2024].
-With only 3 repetitions for each target, we do not have enough data to look at convergence over time.
-However, we can still analalyze the semantics of the descriptions to see if children are influenced by their partner's descriptions or not.
-
 Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions.
-We compare the semantic similarities between descriptions of the same target based on who produced the description.
-Our question is whether the two children within the same game produce descriptions that are more similar to each other than two children in different games.
-As shown in Figure \ref{sim}, different children in the same game do produce more similar descriptions than children in different games, although the descriptions are less similar than descriptions from the same child in different rounds.
+We compared the semantic similarities between descriptions of the same target based on who produced the description.
+Our question was whether the two children within the same game produced descriptions that are more similar to each other than two children in different games.
+
+As shown in Figure \ref{sim}, different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds.
 We modeled this as `r sim_across_between_form |> form()`.
 Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
-The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partner's than to another child's.
-This provides weak evidence for some coordination between chidren.
+The big differences in descriptions between games compared to within games is a measure of partner sensitivity.
 
@@
 In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter.
-As our goal here was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures, children recieved a lot of scaffolding around the experimental interaction.
-Sometimes, this included experimenter echoing children's descriptions "Child A says they see a ...
-Do you see a ...?" which could potentially influence children's responses.
-In Experiment 2, we repeat the same paradigm, with tighter experimenter controls and a larger sample size.
+Our goal was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures.
+so children recieved a lot of scaffolding around the experimental interaction.
+Sometimes, this included experimenter echoing children's descriptions which could potentially influence children's responses.
+In Experiment 2, we repeated the same paradigm, with tighter experimenter controls and a larger sample size.
 
@@
 
-In experiment 1, children were above chance accuracy in their selections.
-In experiment 2, we aimed to repeat the same paradigm with a tighter experimental script to reduce possible influence of experimenters.
-Additionally, we aimed to fix sources of confusion and frustration in experiment 1.
-As most pairs in experiment 1 completed the game fairly quickly, we added a 4th experimental block to allow for more analyses of change over time.
-
-As Experiment 2 was very similar to Experiment 1, here we note the differences. Experiment 2 was pre-registered at https://osf.io/y2dax.
+As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at https://osf.io/y2dax.
 
@@
 In response to some children struggling with the abrupt switch from nameable to non-nameable shapes, we introduced more practice trials for experiment 2.
-We used a total of 4 practice trials, designed to transition from easily recognizable shapes to slightly blockier black and wide shapes to smooth the transition to the critical trials.
+We used a total of 4 practice trials to provide a gradient from more recognizable shapes to less recognizable, blockier shapes (Figure \ref{game}D).
 
@@
 We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in the first experiment).
-The smurfy puppet was swapped out for a more attractive smurfy stuffed animal.
 The experimental script was fully written out so children all received the same instructions.
-To prevent experimenter's influencing children's descriptions or understanding of descriptions, we wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
-TODO link to materials for where the script is.
-The idea was that the experimenter would help with understanding of game mechanics such as who's turn it was to tell and who should press the screen to move the game along if children stalled, but would not provide or repeat any content about the images, what to ask, or whether a description was adequate.
+We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
+Experimenters would help with game mechanics such as who's turn it was to tell and who should press the screen to move the game along, but would not provide or repeat any content about the images, what to ask, or whether a description was adequate.
 
@@
 This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
-
-TODO could include examples of cute things kiddos said here
+Children used a wide range of descriptions successfully, and some example descriptions are shown in Table \ref{example}.
+Often children would use the same description as their partner, although there were some transitions such as from "a person holding a flute" to "a person holding a trumpet".
+
+(not sure how much to say or where to put)
+Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game. 
+In one game, the experimenter provided no input after the practice trials because the children rapidly took turns producing referring expressions and selecting the targets.
+Their advantage could partially be due to being best friends with each other, who may have been more comfortable with each other and had more common ground to start with.
+All children were paired with another child from the same classroom they were willing to play with, but friendship levels varied.
+
+Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
+In one game, child A described the figure as "A human", prompting child B to note "There's two humans." Later in the game, child B used the description "A human", leading child A to ask "Which one?", which child B clarified with"The one that is walking".
+This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+
+```{=latex}
+
+\begin{table}
+	\caption{Example descriptions children used for different tangrams. TODO may want to shorter or pair with the tangram images}
+		\label{example}
+	\begin{tabular}[H]{p{22em}}
+		Descriptions \\
+		\hline
+		%hold
+		"Hold":\\
+		$\bullet$ person\\
+		$\bullet$ a person holding a sandwich\\
+		$\bullet$ a people carrying a box of dirt\\
+		$\bullet$ a monster\\
+		$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon\\
+		\hline
+	"Jump":\\
+		$\bullet$ vampire\\
+		$\bullet$ hopping\\
+		$\bullet$ a person flying\\
+		$\bullet$ somebody skydiving, not in the airplane\\
+		$\bullet$ a person\\
+		$\bullet$ a kite\\
+		$\bullet$ a triangle with a head on it with feet\\
+		\hline
+		"Swim":\\
+		$\bullet$ racecar\\
+		$\bullet$ airplane\\
+		$\bullet$ alligator\\
+		$\bullet$ a person fell down\\
+		$\bullet$ a boat\\
+		$\bullet$ a person that's in a race car that has one triangle and two triangles\\
+		\hline
+		"Walk":\\
+		$\bullet$ person\\
+		$\bullet$ a person walking\\
+		$\bullet$ a person looking down\\
+		$\bullet$ a people, but it doesn't have any arms\\
+		\hline
+	\end{tabular}
+\end{table}
+
+```
 
@@
 
-As a coarse comparison, we repeated the analysis from Experiment 1, comparing the similarity of descriptions to the same target for the same-speaker, same-game, or different-game.
-We modeled this as `r sim_across_between_form |> form()`.
+As a coarse comparison, we repeated the analysis from Experiment 1.
 Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
-The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partner's than to another child's.
 
@@
 
-## TODO Meta-analytic placeholder 
-TODO question for the group: do we want to include any meta analysis across the two expts?
+## Meta-analytic placeholder
+
+Might want to split out and incorporate in along with expt 2 analyses?
 
@@
 
-## using accuracy to predict similarities
+## Accuracy is predictive of similarity to next
 
@@
 As a post-hoc measure, we look at whether descriptions that elicit correct selections have more staying power than descriptions that do not elicit correct selections.
-Here we operationalize "staying power" as similarity to the next block utterance, and we model with `r form(sim_acc_to_next_form)`.
+Here we operationalized "staying power" as similarity to the next block utterance, and we model with `r form(sim_acc_to_next_form)`.
 Correct utterances have an increase in similarity with the next block by `r stats_text(sim_acc_to_next_mod_meta, 2, 3)` with no substantial interaction with block number or same speaker.
@@
 
-## Not sure where to put, cute things kids said
-
-Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
-game 62 is example of played well and fast (also kids bffs) Children sometimes asked clarifying questions of their partner (for instance, when the partner described the shape as "human" or "person", children sometimes sought clarification and arrived at descriptions such as "...
-see game 81, 79?)
-
-game 66 triangle feet
-
-game 61 generally good
-
-For utterances just look through for examples TODO
-
 # General Discussion
@@
 Across 2 experiments and 51 pairs of 4 and 5 year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
-Children varied substantially in how easily they understood the task and what sorts of descriptions they produced, but overall accuracy was high (NUMBER), indicating that children were able to produce adequate descriptions.
-Additionally, children's utterances were more similar to the descriptions their partner used than to other children, and in experiment 2, children's utterances showed signs of converging toward conceptual pacts.
-This provides evidence that at least some preschoolers, in an environment with low task demands and support for understanding the structure of the task, can succeed at iterated referential communication.
-While this task is substantially scaled down relative to measures used for adult competence, and children do not display all the typical trends (i.e no sign of shortening of referential expression), it is a sign that the relevant skills are continuously improving across childhood, rather than only appearing later on.
-
-Limitations.
-The population of children at university nursery schools is non-representative, and the set of materials we used was also not that varied.
+Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (NUMBER), indicating that children were generally able to produce adequate descriptions.
+Additionally, children's utterances showed signs of converging toward conceptual pacts.
+While this task is substantially scaled down relative to measures used for adult competence, and children do not display all the typical trends (i.e no sign of shortening of referential expression), it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+
+At least some preschoolers, in an environment with low task demands and support for understanding the structure of the task, can succeed at iterated referential communication.
+Our task is limited by the non-representative population of children at university nursery schools who participated, and by a limited set of target images.
 This set of tangram images may be easier to distinguish and refer to than some sets used with adults, leading to overall shorter utterances.
-Probably shouldn't say based on this that children can "do reference" at age 4, but it is evidence that under supportive circumstances, a number of children at this age are able to.
-
-We also specifically target the construction of referring expressions that can be jointly understood.
-There are other parts of the coordination where help was provided in children seemed stuck or confused, such as when to make a choice or ask for more information.
-We think this is reasonable, because our goal is on coordination of referring expressions, not larger coordination on tasks or joint instruction following.
-
-Broader implications
-
-This work (along with other work on children's referential communicaiton) suggests that there's a more gradual development of children's communicative and linguistic skills.
-Rather than an abrupt acquisition of a skill, children seem to be refining their abilities across childhood, leading to to the presence of skills early on that become more consistently expressed later.
-Has implications for how we think about children's language development.
-There's debate over how that is ordered and whether communication/pragmatics is a final stage, or how all the stages are boostrapped (CITATIONS WANTED).
-This early ability to use language for communicative purposes is more consistent with the early pragmatics viewpoint.
-Suggestive of a gradual development where children's capabilities are increasing for a wide amount of childhood, as their working memory capacity and executive function improve and they are able to better track other's states of knowledge and keep track of wider arrays of images.
+We specifically targeted children's abilities to construct of referring expressions that can be jointly understood.
+Children were provided scaffolding around the larger coordination problems of taking turns nad talking to their partner.
+
+In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed much later, but pragmatic and communicative competency (CITATIONS), or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning (CITATIONS).
+The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cogntive capacities increase.
+At ages 4-5 years old, children are able to use their linguistic skills communicatively, even as their utterances are not fully adult-like in form.
 

### 2025-01-11 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
updated text

@@
 author-information: > 
-    \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, \\Stanford University \And {\large \bf Bobby Sparks (TODO email) } Department of Psychology, \\Stanford University
+    \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, \\Stanford University \And {\large \bf Robert Z.
+Sparks (bsparks@stanford.edu) } Department of Psychology, \\Stanford University
     \AND {\large \bf Yannick Mofor (TODO email)} \\ TODO affiliation, Stanford University \And {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, \\ Stanford University}
@@
 abstract: >
-  Learning language requires not just learning the facts of a language, but also learning how to use that language to communicate.
-Iterated reference games require rich communicative skills as participants jointly converge on mutually understandable names for initially novel objects.
-Some classical experiments with young children are interpreted as showing that 4-5 year old children were incapable of succeeding at iterated references games with each other.
-More recent work on young children and their parents using simpler paradigms reveal greater abilities even in young children.
-Here, we revisit the question of young children's referential communicative abilites with iterated reference games played between pairs of 4-5 year old preschoolers.
-Across 51 pairs of children (TODO double check N), we find that children were Y% accurate, and that they often used descriptions similar to their partner's.
-Preschoolers' capacity to construct effective referring expressions in novel contexts is consistent with a larger picture of early emerging ability to use language in pragmatic and communicative ways.
+   Learning language requires learning not only the content of language, but also skills that allow us to use language to communicate.
+Iterated reference games provide a window into such skills, requiring rich communication as participants jointly converge on mutually understandable names for initially novel objects.
+Some classical experiments with young children are interpreted as showing that 4-5-year old children are incapable of succeeding at iterated reference games.
+Here, we revisit the question of young children's referential communicative abilities using a simpler, child-friendly paradigm.
+Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in an iterated reference game.
+Children were Y% accurate, and they often used descriptions similar to their partner's.
+These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with a larger picture of the early emerging ability to pragmatically use language in communication.
 TODO fix wrap up sentence!
@@
 
-[way to long, please chop down]
-Learning language requires not just learning the facts of a language, but also learning how to use that language to communicate.
-One common test environment for language use is referential communication, the ability to give a description so that an interlocuter can pick out a target referent from a set of possibilities.
-Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the type of description and the amount of information they provide to their beliefs about the interlocuters knowledge state.
-
-One test of referential communication skill is *iterated reference games* where a describer identifies abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
+Learning a language requires learning not only the content of that language, but also how to use the language to communicate.
+One test environment for language use is referential communication, the ability to give a description so that an interlocutor can pick out a target referent from a set of possibilities.
+Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the type of description and the amount of information they provide to their beliefs about the interlocutor's knowledge state.
+
+One test of referential communication skill is *iterated reference games* where one describes abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
 Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
-Success at this task requires mastery of a number of linguistic and communicative skills including producing adequate initial descriptions, monitoring for comprehension and asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
-
-All of these are skills that children must learn as part of their acquisition of language, but there is debate on the acquisition trajectories of different linguistic and communicative skills.
-Studying how children play iterated reference games can provide insight into the developmental trajectories of the ability to produce expressions to achieve joint understanding.
-
-A study on children's referential communication declared 4-5 year old preschoolers incapable of the task of child-child referential communication [@glucksberg1966].
+Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
+
+Indeed, even young children must make use of these skills in order to achieve success in the task.
+However, there is debate on the developmental trajectories of these communicative skills.
+Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
+
+A study on children's referential communication suggests that 4-5-year old preschoolers struggle to succeed at child-child referential communication [@glucksberg1966].
 In their paradigm, one child was given a set of 6 blocks in a specific order.
-Their task was to describe the images on the block so their partner could pick out their corresponding block.
+Their task was to describe the images on the block so their partner could pick out the corresponding block.
 As children described and selected blocks, they stacked them on pegs, so that the order of blocks on the peg was a measure of correct matching.
-4-5 year old children were successful on practice trials where the blocks had drawings of farm animals and they had visual access to each others blocks; however, younger children (3-4 year olds) were not even able to understand the instructions and complete the practice trials.
-On the critcal trials, the 4-5 year old children were tested with abstract shapes on the blocks and no visual access to the other child's blocks.
-In this condition, even after multiple rounds with the same images, children were not able to produce referential descriptions that allowed for success at the matching task.
-
-Similar experiments with older children indicated a gradual improvement across age (up to the teen years) both at their initial accuracy and on their improvement over repetitions, although even the 9th grade sample was noticeably worse than their adult college student sample [@glucksberg1967; @krauss1969].
-@glucksberg1966 attributed the young children's failures to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
-However, the complexity of the stacking paradigm and the large number of potential targets meant that this was not a clean test of children's ability to produce adequate referent expressions as the experiment had substantial task demands that could mask children's abilities.
-
-Since then, a number of studies have targeted specific skills needed for referential communication, including expectations around consistent labeling ("conversational pacts"), awareness of other's perspectives, and sensitivity to descriptive adequacy in varying contexts.
-A number of these studies have found roots of understanding emerging during hte preschool years, albeit with inconsistent expression.
-
-Young children show some sensitivity to the expectation that during cooperative communicate a person will likely re-use their own descriptions.
-As listeners, preschoolers (3-5 year olds) are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010].
-However, this preference for familiar descriptions is applied regardless of whether they are said by the same person or a new person, and they sometimes verbally protesting the use of a new term ("That's a horse, not a pony!") [@matthews2010].
-As the producers of descriptions, children show some partner-specificity, but are far from consistent.
-@koymen2014 had 4 and 6 year old children do a version of the paradigm from @brennan1996.
-Children used specific referring expressions ("woman's shoe") with one partner in a visual context of different types of shes.
-In later trials when only one shoe was present and thus a general expression like "shoe" was adequate, 6 year olds were somewhat more likely to continue using the specific term with the same partner than with a new partner, whereas 4 year olds used both specific and general expressions at the same rate regardless of partner.
-
-Expectations of consistency are entwined with being able to take the perspective of an interlocuter.
-Young children are notoriously ego-centric [@epley2004; TODO there are probably better things to cite here instead], but even preschoolers are able to take other's perspectives, especially when the are not under too much cognitive load [@sanjuan2015].
-Across the 2-4 year olds range, children are developing sensitivity to what is given their interlocuter based on the discourse history and visual context available to their interlocuter [@matthews2006].
-Older preschoolers are more likely to use more given forms such as pronouns when their interlocuter has more context and indefinite forms when their interlocuter has less context.
-4-6 year old children are more likely to use a more specific reference expression with a size modifier when their partner can see the size competitor than when their partner cannot see the competitor, but children produce the modifier inconsistently even when it is necessary to distinguish the target [@nadig2002; @nilsen2009].
-
-The skill children seem to struggle the most with is appropriately tailoring the specificity of their utterances to the visual context.
-4 year olds showing some improvement after getting specific feedback that models appropriate descriptions, but still a lot of inconsistency [@matthews2012].
-Children ages 4-8 often produce the same label for a target object regardless of whether the context is close or far, leading to descriptions that are underinformative for close contexts [@leung2024].
-When children must pick out one of two very similar images, 4 and 5 year olds aren't good at mentioning the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
-By age 5, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances, although children still do not ask for clarification all the time [@morisseau2013].
+While 4-5-year old children were successful on practice trials where the blocks had drawings of farm animals and they had visual access to each other's blocks; children failed on critical trials where the blocks had drawings of abstract shapes and there was no visual access.
+Even after multiple rounds with the same images, children were not able to produce referential descriptions that allowed for success at the matching task.
+@glucksberg1966 attributed young children's failure to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
+
+Similar experiments with older children indicated a gradual improvement through adolescence both at their initial accuracy and on their improvement over repetitions.
+Still, even the 9th grade sample was noticeably worse than the sample of adult college students [@glucksberg1967; @krauss1969].
+That the task is difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets presents significant task demands that might be masking the referential communication abilities of even the youngest children.
+
+Since then, a number of studies have revealed early emerging skills in the preschool years that support the use of referential communication.
+Studies show that preschool-aged children (3-5-years old) are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and by 6-years more consistently produce partner-specific referential expressions [@koymen2014], suggesting that young children are sensitive to consistent use of descriptions in cooperative communication.
+Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002; @nilsen2009], showing that children at this age can reason about others’ perspectives to more efficiently communication.
+
+While children are sensitive to others’ perspectives, they still seem to struggle with appropriately tailoring the specificity of their utterances to the visual context, resulting in utterances that are under-informative [@matthews2012; @leung2024].
+When children must describe one of two very similar images, 4 and 5-year olds struggle to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
+By 5-years old, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].
 
@@
 fewer distractor referents [@abbot-smith2016].
-This suggests that tailored referential expression production is possible in children, but it is likely to be masked in situations that involve cognitive load.
-
-Since @glucksberg1966, only a few studies have revisited the question of how well children can put together different skills to communicate in an iterated reference game.
-@branigan2016 tested 72 8-10 year olds on iterated referential communication using a paradigm based on @wilkes-gibbs1992.
-In the training period, pairs of children matched tangrams in order from a pool of 8 images.
-Qualitatively, children on average showed the classic patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions.
-However, children's level of accuracy was far below what is typical for adults and varied dramatically from pair to pair.
-
-Some of the skills required in iterated reference games are communicative rather than purely linguistic.
-In @bohn2019, 4 and 6 year olds communicated target images from a pool of 5 images via a video link without audio.
-Children's comprehension was good and increased over repetitions, and children, especially the 6 year olds,  showed signs of agreeing on conventionalized gestures with their partner.
-
-A recent study on younger children examined children's abilities to play an iterated reference game with a parent.
-@leung2024 had 63 children ages 4-8 year olds play a matching game on tablets, where, on each trial, participants saw two images.
-The describer's tablet had a box around one image and their task was to verbally describe the image so their partner could select the corresponding image.
-The target images were tangrams, with a total of 10 target images, which repeated across 4 blocks for a total of 40 trials per pair.
-In this task, even 4 year olds had an initial accuracy above 80% which rose to above 90% on the later repetitions.
-
-Given the potential for task demands in @glucksberg1966 and the level of success young children had with their parents at a simplified, less demanding paradigm, we revisit the question of children's ability to converge on appropriate referential expressions with each other.
-In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified paradigm to reduce cognitive demands.
+This paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
+
+Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game.
+Recent work shows that 8-10-year olds exhibit classic patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions in an integrated reference game,  but children’s accuracy was still far below adult performance and varied dramatically between dyads [@branigan2016].
+4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an integrated reference game where children could only use gestures rather than speech to communicate [@bohn2019].
+A recent study with younger children from 4-8-years old examined children's abilities to play an iterated reference game with a parent using a simple, child-friendly tablet-based task.
+In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
+Together, these findings provide evidence of young children’s ability to successfully communicate in an iterated reference game under the right conditions.
+
+Given the task demands in @glucksberg1966 and work showing that children can succeed in a simplified, less demanding paradigm, we revisit the question of preschool-aged children's ability to converge on appropriate referential expressions with each other.
+In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
+Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in an iterated reference game, suggesting that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought.
 
@@
 Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
-Panel C shows the procedure for Experiment 1; within in critical block targets were ordered randomly.
+Panel C shows the procedure for Experiment 1; within critical block targets were ordered randomly.
 Panel D shows the procedure for Experiment 2.
@@
 ## Participants
-4 and 5 year old children were recruited from a university nursery school during the school day.
+4 and 5-year old children were recruited from a university nursery school during the school day.
 Children played with another child from the same class.
@@
 Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials.
-We had 19 complete games and 1 incomplete, but included game.
+We had 19 complete games and 1 incomplete, but included the game.
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
@@
 We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
-The first trial critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
+The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
 
@@
 Our goal was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures.
-so children recieved a lot of scaffolding around the experimental interaction.
-Sometimes, this included experimenter echoing children's descriptions which could potentially influence children's responses.
+so children received a lot of scaffolding around the experimental interaction.
+Sometimes, this included experimenters echoing children's descriptions which could potentially influence children's responses.
 In Experiment 2, we repeated the same paradigm, with tighter experimenter controls and a larger sample size.
@@
 Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
-Since the two expeirments had different length games, we do not want to pool them in comparing to the last block, but we can pool them comparing to the next block.
+Since the two experiments had different length games, we do not want to pool them in comparing to the last block, but we can pool them comparing to the next block.
 
@@
 
-Across 2 experiments and 51 pairs of 4 and 5 year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
+Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
 Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (NUMBER), indicating that children were generally able to produce adequate descriptions.
@@
 This set of tangram images may be easier to distinguish and refer to than some sets used with adults, leading to overall shorter utterances.
-We specifically targeted children's abilities to construct of referring expressions that can be jointly understood.
-Children were provided scaffolding around the larger coordination problems of taking turns nad talking to their partner.
-
-In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed much later, but pragmatic and communicative competency (CITATIONS), or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning (CITATIONS).
-The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cogntive capacities increase.
-At ages 4-5 years old, children are able to use their linguistic skills communicatively, even as their utterances are not fully adult-like in form.
+We specifically targeted children's abilities to construct referring expressions that can be jointly understood.
+Children were provided scaffolding around the larger coordination problems of taking turns and talking to their partner.
+
+In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed much later, by pragmatic and communicative competency (CITATIONS), or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning (CITATIONS).
+The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.
+At 4-5-years old, children are able to use their linguistic skills communicatively, even as their utterances are not fully adult-like in form.
 

### 2025-01-14 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
start of V edit pass

@@
 ---
-title: "TODO Preschoolers form conventional pacts with each other to communicate about novel shapes"
+title: "Preschoolers form conventional pacts with each other to communicate about novel shapes"
 bibliography: preschool-tangrams.bib
@@
     \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, \\Stanford University \And {\large \bf Robert Z.
-Sparks (bsparks@stanford.edu) } Department of Psychology, \\Stanford University
-    \AND {\large \bf Yannick Mofor (TODO email)} \\ TODO affiliation, Stanford University \And {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, \\ Stanford University}
+Sparks (bsparks@stanford.edu) }\\ Department of Psychology, \\Stanford University
+    \AND {\large \bf Yannick Mofor (yannickm@stanford.edu)} \\ Department of Computer Science, \\ Stanford University \And {\large \bf Michael C.
+Frank (mcfrank@stanford.edu)} \\ Department of Psychology, \\ Stanford University}
 
@@
 Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in an iterated reference game.
-Children were Y% accurate, and they often used descriptions similar to their partner's.
+Children were 85% accurate, and they often used descriptions similar to their partner's.
 These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with a larger picture of the early emerging ability to pragmatically use language in communication.
-TODO fix wrap up sentence!
-TODO possibly shorten!
 
 keywords: >
-    development; iterated reference game; TODO keywords
+    development; iterated reference game; pragmatics; preschoolers
 
 output: cogsci2024::cogsci_paper
-final-submission: \cogscifinalcopy
+header-includes: \cogscifinalcopy
 ---
@@
 
-(Note: we should talk about whether it makes more sense to do all the methods and then all the results versus the current expt 1 than expt 2 layout.)
+<!--(Note: we should talk about whether it makes more sense to do all the methods and then all the results versus the current expt 1 than expt 2 layout.) -->
 
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load that can hide early abilities [CITATION WANTED], so we adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
-
+Young children can be very sensitive to task demands and cognitive load that can hide early abilities (CITATIONS), so we adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
 This experiment was pre-registered at https://osf.io/kcv8j.
@@
 
-Statistical analyses were run in TODO CITE brms with weakly informative priors. We present the estimate and 95% credible intervals.
+Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We present the estimate and 95% credible intervals.
 
@@
 
-[less important, could drop] As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
+<!--[less important, could drop] --> As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
 We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
@@
 We ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
-The initial length was `r stats_text(description_mod_1, 1)` and description length was relatively stable over time (`r stats_text(description_mod_1, 2)`, shown in Figure \ref{length}).
+The initial length was `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` per trial, shown in Figure \ref{length}).
 How long of a description is initially warranted depends on how iconic or easily describable the shape is, as well as how large and close the contrast set is.
-In this case, the low number of options to distinguish and relatively easy to describe shapes may mean long initial descriptions are less necessary.
-(The adult controls in @leung2024 use shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.) However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
+In this case, the low number of options to distinguish and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
+The adult controls in @leung2024 use shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.
+However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
 
@@
 While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length [@boyce2024].
-Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions.
+Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions [@reimers2019].
 We compared the semantic similarities between descriptions of the same target based on who produced the description.
@@
 In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter.
-Our goal was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures.
-so children received a lot of scaffolding around the experimental interaction.
+Our goal was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures, so children received a lot of scaffolding around the experimental interaction.
 Sometimes, this included experimenters echoing children's descriptions which could potentially influence children's responses.
@@
 ```{=latex}
-
 \begin{table}
-	\caption{Example descriptions children used for different tangrams. TODO may want to shorter or pair with the tangram images}
+	\caption{Example descriptions children used for different target tangram images.}
 	\label{example}
-	\begin{tabular}[H]{p{22em}}
-		Descriptions \\
+	\begin{tabular}[H]{p{16em}p{4em}}
 		\hline
 		%hold
-		"Hold":\\
-		$\bullet$ person\\
-		$\bullet$ a person holding a sandwich\\
-		$\bullet$ a people carrying a box of dirt\\
-		$\bullet$ a monster\\
-		$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon\\
+		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\
+
+		$\bullet$ a person holding a sandwich &\\
+		$\bullet$ a people carrying a box of dirt &\\
+		$\bullet$ a monster &\\
+		$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon &\\
 		\hline
-	"Jump":\\
-		$\bullet$ vampire\\
-		$\bullet$ hopping\\
-		$\bullet$ a person flying\\
-		$\bullet$ somebody skydiving, not in the airplane\\
-		$\bullet$ a person\\
-		$\bullet$ a kite\\
-		$\bullet$ a triangle with a head on it with feet\\
+		$\bullet$ vampire & \multirow{7}{*}{\includegraphics[width=.5in]{E1.jpg}}\\
+		$\bullet$ hopping & \\
+		$\bullet$ a person flying &\\
+		$\bullet$ a person &\\
+		$\bullet$ a kite &\\
+		$\bullet$ a triangle with a head on it with feet &\\
+				\multicolumn{2}{l}{$\bullet$  somebody skydiving, not in the airplane}\\
 		\hline
-		"Swim":\\
-		$\bullet$ racecar\\
-		$\bullet$ airplane\\
-		$\bullet$ alligator\\
-		$\bullet$ a person fell down\\
-		$\bullet$ a boat\\
-		$\bullet$ a person that's in a race car that has one triangle and two triangles\\
+		$\bullet$ racecar & \multirow{7}{*}{\includegraphics[width=.5in]{D1.jpg}}\\
+		$\bullet$ airplane &\\
+		$\bullet$ alligator &\\
+		$\bullet$ a person fell down &\\
+		$\bullet$ a boat &\\
+		$\bullet$ a person that's in a race car that has one triangle and two triangles &\\
 		\hline
-		"Walk":\\
-		$\bullet$ person\\
-		$\bullet$ a person walking\\
-		$\bullet$ a person looking down\\
-		$\bullet$ a people, but it doesn't have any arms\\
+
+		$\bullet$ person & \multirow{4}{*}{\includegraphics[width=.5in]{B1.jpg}}\\
+		$\bullet$ a person walking &\\
+		$\bullet$ a person looking down &\\
+		$\bullet$ a people, but it doesn't have any arms& \\
 		\hline
@@
 
-# Post hoc results
-
-## Meta-analytic placeholder
-
-Might want to split out and incorporate in along with expt 2 analyses?
+# Additional results
+
+## Combining across experiments
+
+<!-- Might want to split out and incorporate in along with expt 2 analyses? -->
 
@@
   filter(type != "practice"))
+
+# acc_mod_data_meta |> summarize(m=mean(correct.num))
 
@@
 
-Query -- I did not include experiment # as a random effect here, and maybe should have?
+<!-- Query -- I did not include experiment # as a random effect here, and maybe should have?
 I can rerun if we decide we want this analysis and think expt # as a random effect with game nested within it would be useful.
-
-## Accuracy is predictive of similarity to next
+-->
+
+## Relation between accuracy and convergence
 
@@
 Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
-Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (NUMBER), indicating that children were generally able to produce adequate descriptions.
+Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions.
 Additionally, children's utterances showed signs of converging toward conceptual pacts.

### 2025-01-15 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
edited under the page limit

@@
 Learning a language requires learning not only the content of that language, but also how to use the language to communicate.
-One test environment for language use is referential communication, the ability to give a description so that an interlocutor can pick out a target referent from a set of possibilities.
-Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the type of description and the amount of information they provide to their beliefs about the interlocutor's knowledge state.
+One test environment for language use is referential communication, the ability to describe a target so an interlocutor can pick it out from a set of possibilities.
+Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the description they provide to their beliefs about the interlocutor's knowledge state.
 
@@
 Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
-
-Indeed, even young children must make use of these skills in order to achieve success in the task.
+<!--Indeed, even young children must make use of these skills in order to achieve success in the task.
 However, there is debate on the developmental trajectories of these communicative skills.
-Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
-
-A study on children's referential communication suggests that 4-5-year old preschoolers struggle to succeed at child-child referential communication [@glucksberg1966].
+--> Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
+
+A classic study suggests that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966].
 In their paradigm, one child was given a set of 6 blocks in a specific order.
 Their task was to describe the images on the block so their partner could pick out the corresponding block.
-As children described and selected blocks, they stacked them on pegs, so that the order of blocks on the peg was a measure of correct matching.
-While 4-5-year old children were successful on practice trials where the blocks had drawings of farm animals and they had visual access to each other's blocks; children failed on critical trials where the blocks had drawings of abstract shapes and there was no visual access.
-Even after multiple rounds with the same images, children were not able to produce referential descriptions that allowed for success at the matching task.
-@glucksberg1966 attributed young children's failure to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.
-
+As children described and selected blocks, they stacked them on pegs.
+While 4-5-year old children were successful on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had drawings of abstract shapes and there was no visual access.
+Even after multiple rounds with the same images, children were not able to correctly order the blocks.
+<!--@glucksberg1966 attributed young children's failure to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.-->
 Similar experiments with older children indicated a gradual improvement through adolescence both at their initial accuracy and on their improvement over repetitions.
 Still, even the 9th grade sample was noticeably worse than the sample of adult college students [@glucksberg1967; @krauss1969].
-That the task is difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets presents significant task demands that might be masking the referential communication abilities of even the youngest children.
-
-Since then, a number of studies have revealed early emerging skills in the preschool years that support the use of referential communication.
-Studies show that preschool-aged children (3-5-years old) are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and by 6-years more consistently produce partner-specific referential expressions [@koymen2014], suggesting that young children are sensitive to consistent use of descriptions in cooperative communication.
-Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002; @nilsen2009], showing that children at this age can reason about others’ perspectives to more efficiently communication.
+That the task is difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets present significant task demands that might prevent accurate measurement of children's abilities.
+
+```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental Setup and Procedure.
+Panel A shows the experimental setup with the teller and guesser across the table from each other.
+Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
+Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly.
+Panel D shows the procedure for Experiment 2.
+\\label{game}", cache=FALSE}
+
+knitr::include_graphics("diagram.pdf")
+
+# edit diagram at 
+# https://docs.google.com/drawings/d/1gio1CNSedECJGrkrNDUi_AydPZcYzESfl4WhZD2l9Bc/edit?usp=sharing
+```
+
+More recently, a number of studies have revealed early emerging skills in the preschool years that support the use of referential communication.
+Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
+Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009], showing that children at this age can reason about others’ perspectives to more efficiently communication.
 
@@
 A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e.
-fewer distractor referents [@abbot-smith2016].
+when there are fewer possible referents to consider [@abbot-smith2016].
 This paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
@@
 Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game.
-Recent work shows that 8-10-year olds exhibit classic patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions in an integrated reference game,  but children’s accuracy was still far below adult performance and varied dramatically between dyads [@branigan2016].
-4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an integrated reference game where children could only use gestures rather than speech to communicate [@bohn2019].
-A recent study with younger children from 4-8-years old examined children's abilities to play an iterated reference game with a parent using a simple, child-friendly tablet-based task.
+Recent work shows that 8-10-year olds exhibit adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016].
+4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an iterated reference game where children could only use gestures to communicate [@bohn2019].
+
+A recent study with younger children examined 4-8 year olds' abilities to play an iterated reference game with a parent using a simple, child-friendly tablet-based task.
 In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
-Together, these findings provide evidence of young children’s ability to successfully communicate in an iterated reference game under the right conditions.
-
-Given the task demands in @glucksberg1966 and work showing that children can succeed in a simplified, less demanding paradigm, we revisit the question of preschool-aged children's ability to converge on appropriate referential expressions with each other.
+Together, these findings provide evidence of young children’s ability to successfully communicate about novel referents under the right conditions.
+
+Given the task demands in @glucksberg1966 and work showing that children can succeed in a simplified, less demanding paradigm, we revisit the question of child-child referential communication.
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in an iterated reference game, suggesting that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought.
-
-```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental Setup and Procedure.
-Panel A shows the experimental setup with the teller and guesser across the table from each other.
-Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
-Panel C shows the procedure for Experiment 1; within critical block targets were ordered randomly.
-Panel D shows the procedure for Experiment 2.
-\\label{game}", cache=FALSE}
-
-knitr::include_graphics("diagram.pdf")
-
-# edit diagram at 
-# https://docs.google.com/drawings/d/1gio1CNSedECJGrkrNDUi_AydPZcYzESfl4WhZD2l9Bc/edit?usp=sharing
-```
+Across 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load that can hide early abilities (CITATIONS), so we adapted the experimental framework from @leung2024, but further simplified it by reducing the total pool of targets and the number of trials children completed.
+Young children can be very sensitive to task demands and cognitive load (CITATIONS), so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
 This experiment was pre-registered at https://osf.io/kcv8j.
@@
 ## Participants
-4 and 5-year old children were recruited from a university nursery school during the school day.
+4 and 5-year old children were recruited from a university preschool during the school day.
 Children played with another child from the same class.
@@
 Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials.
-We had 19 complete games and 1 incomplete, but included the game.
+We had 19 complete games and 1 included incomplete game.
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
@@
 We coded the matching game using Empirica and hosted it on a lab server [@almaatouq2020].
-We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate to other websites or applications during the game.
+We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate away from the game.
 
 ## Procedure
-Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room where the game was explained to them.
+Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room.
 Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game.
 Children sat across a table from each other, each with a tablet in front of them (Figure \ref{game}A).
-On each trial, one child saw two images, one of them in a black box, and was asked to "say what they saw" in the black box so their partner (and Smurfy) could tap the corresponding image.
-The guesser saw the same two images (in a randomized order).
-Upon tapping an image, both children received feedback in form of a smiley or frowny face and an audible sound.
-After each trial, children's roles switched.
+On each trial, one child was the "teller" and saw a black box around one of two images on their screen and was asked to "say what they saw" in the black box.
+The "guesser" saw the same two images in a randomized order and tried to select the described image to help Smurfy make a match.
+When the guesser selected an image, both children received feedback in form of a smiley or frowny face and an audible sound.
+After each trial, children switched roles.
 Children passed Smurfy back and forth to help them keep track of whether they were the "guesser" or "teller" on a given trial.
 
-Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 targets (Figure \ref{game}C).
+Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 target images (Figure \ref{game}C).
 Targets were randomly paired with another of the critical images as the foil.
 
-The experimenters running the game did not volunteer descriptions, but did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements.
+The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements.
 The entire interaction was video-recorded.
@@
 
-Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
-As shown in Figure \ref{acc}, children were above chance in their selections.
-To confirm this and test for any changes in accuracy over time, we fit a mixed effects model of accuracy (`r acc_form |> form()`).
-Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
+Our primary measure of interest was whether children could accurately communicate the intended target.
+<!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
+-->  As shown in Figure \ref{acc}, children were above chance in their selections.
+To test for changes in accuracy over time, we fit a mixed effects model of accuracy (`r acc_form |> form()`).
+Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
 
@@
 <!--[less important, could drop] --> As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
-We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`.
-The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)`).
+We ran a Bayesian mixed effects model: `r speed_form |> form()`.
+The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial ).
 
@@
 
-In iterated reference games in adults, a canonical finding is that the length of descriptions goes from long to short over repeated references to the initially hard to describe shapes [@clark1986; @hawkins2020b; @boyce2024].
-We looked at how long the descriptions the children used were to see if the same trend occurred.
-We ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
-The initial length was `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` per trial, shown in Figure \ref{length}).
-How long of a description is initially warranted depends on how iconic or easily describable the shape is, as well as how large and close the contrast set is.
-In this case, the low number of options to distinguish and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
+In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024].
+We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
+On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
+
+How long of a description is initially warranted depends on how easily describable the target is, as well as how large and close the contrast set is.
+In this case, the low number of options and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
 The adult controls in @leung2024 use shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.
 However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
+<!-- some of this is kinda discussion-y, possibly move -->
 
@@
 Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions [@reimers2019].
+
 We compared the semantic similarities between descriptions of the same target based on who produced the description.
@@
 
-As shown in Figure \ref{sim}, different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds.
+Different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds (Figure \ref{sim}).
 We modeled this as `r sim_across_between_form |> form()`.
@@
 In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter.
-Our goal was to see if young children were at all able to provide descriptions to each other adequate to identify abstract figures, so children received a lot of scaffolding around the experimental interaction.
+Our goal was to see if young children were at all able to provide adequate descriptions, so children received a lot of scaffolding around the experimental interaction.
 Sometimes, this included experimenters echoing children's descriptions which could potentially influence children's responses.
-In Experiment 2, we repeated the same paradigm, with tighter experimenter controls and a larger sample size.
+In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.
 
@@
 The same 4 critical images were used as in experiment 1, although this time, children saw these images 4 times.
-In response to some children struggling with the abrupt switch from nameable to non-nameable shapes, we introduced more practice trials for experiment 2.
-We used a total of 4 practice trials to provide a gradient from more recognizable shapes to less recognizable, blockier shapes (Figure \ref{game}D).
+In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
+We used a total of 4 practice trials to provide a gradient from familiar shapes to less recognizable, blockier shapes (Figure \ref{game}D).
 
 ## Procedure
-The procedure was much the same (Figure \ref{game}D).
-We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in the first experiment).
-The experimental script was fully written out so children all received the same instructions.
+The procedure was much the same as Experiment 1(Figure \ref{game}D).
+We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in Experiment 1).
+The experimental script was fully written out and memorized by experimenters so children all received the same instructions.
 We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
-Experimenters would help with game mechanics such as who's turn it was to tell and who should press the screen to move the game along, but would not provide or repeat any content about the images, what to ask, or whether a description was adequate.
-
-## Data processing
-Data was processed in the same way as experiment 1.
-After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
-
-# Experiment 2 Results
-
-We report the same set of analyses and model results as in Experiment 1 as well as additional analyses of how the semantic content of children's descriptions changes over time.
-
-## Accuracy
-
-```{r, eval=F}
-acc_priors <- c(
-  set_prior("normal(0,1)", class = "b"),
-  set_prior("normal(0,1)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-acc_mod_data <- clean_2 |>
-  mutate(
-    correct.num = ifelse(correct == TRUE, 1, 0),
-    trial.num = trial - 4
-  ) |>
-  filter(type != "practice")
-acc_mod <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
-  family = bernoulli(link = "logit"),
-  data = acc_mod_data,
-  file = here(mod_loc, "acc_2.rds"),
-  prior = acc_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r}
-acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
-```
-
-As in Experiment 1, in Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
-This confirms that children are able to communicate with each other about these abstract shapes.
-
-## Speed
-
-```{r}
-speed_priors <- c(
-  set_prior("normal(60,100)", class = "Intercept"),
-  set_prior("normal(0,20)", class = "b"),
-  set_prior("normal(0,20)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-speed_mod_data_2 <- clean_2 |>
-  mutate(
-    time.sec = time / 1000,
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num | gameId) + (1 | target),
-  data = speed_mod_data_2,
-  file = here(mod_loc, "speed_2.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r}
-speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))
-```
-
-In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)`).
-Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
-
-## Description length
-
-```{r, eval=F}
-description_priors <- c(
-  set_prior("normal(5,10)", class = "Intercept"),
-  set_prior("normal(0,5)", class = "b"),
-  set_prior("normal(0,5)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-description_mod_data_2 <- clean_2 |>
-  mutate(
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-description_mod_2 <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
-  data = description_mod_data_2,
-  file = here(mod_loc, "description_2.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r}
-description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))
-```
-
-The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (`r stats_text(description_mod_2, 2)`, Figure \ref{length}).
-This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
-Children used a wide range of descriptions successfully, and some example descriptions are shown in Table \ref{example}.
-Often children would use the same description as their partner, although there were some transitions such as from "a person holding a flute" to "a person holding a trumpet".
-
-(not sure how much to say or where to put)
-Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game. 
-In one game, the experimenter provided no input after the practice trials because the children rapidly took turns producing referring expressions and selecting the targets.
-Their advantage could partially be due to being best friends with each other, who may have been more comfortable with each other and had more common ground to start with.
-All children were paired with another child from the same classroom they were willing to play with, but friendship levels varied.
-
-Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
-In one game, child A described the figure as "A human", prompting child B to note "There's two humans." Later in the game, child B used the description "A human", leading child A to ask "Which one?", which child B clarified with"The one that is walking".
-This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen to move the game along, but avoided contributing or repeating any content about the images or the descriptions.
 
@@
 \begin{table}
-	\caption{Example descriptions children used for different target tangram images.}
+	\caption{Example descriptions children successfully used used to identify different target tangram images in Experiment 2.}
 	\label{example}
-	\begin{tabular}[H]{p{16em}p{4em}}
+	\begin{tabular}[t]{p{16em}p{4em}}
 		\hline
@@
 ```
+## Data processing
+Data was processed in the same way as experiment 1.
+After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
+
+# Experiment 2 Results
+
+We report the same set of analyses and model results as in Experiment 1 as well as additional analyses of how the semantic content of children's descriptions changes over time.
+
+## Accuracy
+
+```{r, eval=F}
+acc_priors <- c(
+  set_prior("normal(0,1)", class = "b"),
+  set_prior("normal(0,1)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+acc_mod_data <- clean_2 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 4
+  ) |>
+  filter(type != "practice")
+acc_mod <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
+  family = bernoulli(link = "logit"),
+  data = acc_mod_data,
+  file = here(mod_loc, "acc_2.rds"),
+  prior = acc_priors,
+  control = list(adapt_delta = .95)
+)
+```
+
+```{r}
+acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
+```
+
+As in Experiment 1, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
+This confirms that children are able to communicate with each other about these abstract shapes.
+
+## Speed
+
+```{r}
+speed_priors <- c(
+  set_prior("normal(60,100)", class = "Intercept"),
+  set_prior("normal(0,20)", class = "b"),
+  set_prior("normal(0,20)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+speed_mod_data_2 <- clean_2 |>
+  mutate(
+    time.sec = time / 1000,
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
+
+speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num | gameId) + (1 | target),
+  data = speed_mod_data_2,
+  file = here(mod_loc, "speed_2.rds"),
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
+```
+
+```{r}
+speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))
+```
+
+In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)`seconds / trial).
+Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
+
+## Description length
+
+```{r, eval=F}
+description_priors <- c(
+  set_prior("normal(5,10)", class = "Intercept"),
+  set_prior("normal(0,5)", class = "b"),
+  set_prior("normal(0,5)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+description_mod_data_2 <- clean_2 |>
+  mutate(
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
+
+description_mod_2 <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
+  data = description_mod_data_2,
+  file = here(mod_loc, "description_2.rds"),
+  prior = speed_priors,
+  control = list(adapt_delta = .95)
+)
+```
+
+```{r}
+description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))
+```
+
+The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial , Figure \ref{length}).
+This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
+
+Children used a wide range of descriptions successfully, and some example descriptions from Experiment 2 are shown in Table \ref{example}.
+Often children would use the same description as their partner, although there were some transitions such as from "a person holding a flute" to "a person holding a trumpet".
 
@@
 
-To look at semantic distance between utterances, we again operationalize similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019].
+<!--To look at semantic distance between utterances, we again operationalized similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019]. -->
 
@@
 
-The greater number of trials in Experiment 2 makes it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
+The greater number of trials in Experiment 2 made it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
 
 To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`.
-Over the first three blocks, descriptions become increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions are more similar to the last block if they come from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
-
-Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
-Although over time descriptions do get more similar to the last block utterance, the distance between adjacent block utterances is relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
-
-As each partnership converges to their shared nicknames, partnerships often diverge from one another as groups focus on distinct aspects of the image.
+Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+Descriptions were more similar to the last block if they came from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
+<!--Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
+--> Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
+
+As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
 We tested whether descriptions in different games to the same target diverged over time: `r sim_across_form |> form()`.
-As the games progress, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
+As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
 
@@
 
-## Combining across experiments
+## Interaction structure
+Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
+In one game, the experimenter provided no input after the practice trials because the children rapidly took turns producing referring expressions and selecting the targets.
+Their advantage could partially be due to being best friends with each other; all children were paired with another child from the same classroom who they were willing to play with, but friendship levels varied.
+
+Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
+In one game, child A described the figure as "A human", prompting child B to note "There's two humans." Later in the game, child B used the description "A human", leading child A to ask "Which one?", which child B clarified with "The one that is walking".
+This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+
+## Meta-analytic models
+
+<!--Combining across experiments
 
@@
 
-As the two experiments were similar to one another, we re-run models pooling across the two experiments. 
+As the two experiments were similar to one another, we re-ran models pooling across the two experiments. 
 Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`.
-
-The average length of descriptions on the first trial was `r stats_text(description_mod_meta, 1)` words and description length was relatively stable over time (`r stats_text(description_mod_meta, 2)`, Figure \ref{length}).
-
-Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
-Since the two experiments had different length games, we do not want to pool them in comparing to the last block, but we can pool them comparing to the next block.
-
+The average length of descriptions on the first trial was `r stats_text(description_mod_meta, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
+
+Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
+<!--Since the two experiments had different length games, we do not pool them in comparisons to the last block, but we can pool them comparing to the next block.
 The distance between adjacent block utterances is relatively constant: `r sim_to_next_mod_meta |> stats_text(2,3)`.
+-->
 
@@
 
-As a post-hoc measure, we look at whether descriptions that elicit correct selections have more staying power than descriptions that do not elicit correct selections.
-Here we operationalized "staying power" as similarity to the next block utterance, and we model with `r form(sim_acc_to_next_form)`.
-Correct utterances have an increase in similarity with the next block by `r stats_text(sim_acc_to_next_mod_meta, 2, 3)` with no substantial interaction with block number or same speaker.
-Numerically, the size of the boost for being correct is similar to the size of the boost from coming form the same speaker `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
+As a post-hoc measure, we looked at whether descriptions that elicited correct selections had more staying power than descriptions that did not elicit correct selections.
+Here we operationalized "staying power" as similarity to the next block utterance, and we modeled it with `r form(sim_acc_to_next_form)`.
+Correct utterances had an increase in similarity with the next block  (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or same speaker.
+Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same speaker `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
 
@@
 
-Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children were at producing referential expressions to allow their partner to find a matching abstract shape.
+Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
 Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions.
 Additionally, children's utterances showed signs of converging toward conceptual pacts.
-While this task is substantially scaled down relative to measures used for adult competence, and children do not display all the typical trends (i.e no sign of shortening of referential expression), it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
-
-At least some preschoolers, in an environment with low task demands and support for understanding the structure of the task, can succeed at iterated referential communication.
-Our task is limited by the non-representative population of children at university nursery schools who participated, and by a limited set of target images.
+Children did not display a reduction in the length of referential expressions over the course of the game.
+
+While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+
+Our task is limited by the non-representative population of children at university nursery schools who participated, and by the limited set of target images.
 This set of tangram images may be easier to distinguish and refer to than some sets used with adults, leading to overall shorter utterances.

### 2025-01-16 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
start to address Alvin Mike comments

@@
 Here, we revisit the question of young children's referential communicative abilities using a simpler, child-friendly paradigm.
-Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in an iterated reference game.
+Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in establishing reference.
 Children were 85% accurate, and they often used descriptions similar to their partner's.
-These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with a larger picture of the early emerging ability to pragmatically use language in communication.
+These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
 
@@
 output: cogsci2024::cogsci_paper
-header-includes: \cogscifinalcopy
+header-includes: \raggedbottom
+#header-includes: \cogscifinalcopy
 ---
@@
     mutate(
-      Estimate = round(Estimate, digits = decimal),
-      Lower = round(lower, digits = decimal),
-      Upper = round(upper, digits = decimal),
+      Estimate = round(Estimate, digits = decimal) |> formatC(format='f', digits=decimal ),
+      Lower = round(lower, digits = decimal) |> formatC(format='f', digits=decimal ),
+      Upper = round(upper, digits = decimal) |> formatC(format='f', digits=decimal ),
       `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
@@
 <!--@glucksberg1966 attributed young children's failure to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.-->
-Similar experiments with older children indicated a gradual improvement through adolescence both at their initial accuracy and on their improvement over repetitions.
+Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions.
 Still, even the 9th grade sample was noticeably worse than the sample of adult college students [@glucksberg1967; @krauss1969].
-That the task is difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets present significant task demands that might prevent accurate measurement of children's abilities.
+That the task was difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets present significant task demands that might prevent accurate measurement of children's abilities.
 
@@
 
-While children are sensitive to others’ perspectives, they still seem to struggle with appropriately tailoring the specificity of their utterances to the visual context, resulting in utterances that are under-informative [@matthews2012; @leung2024].
+While children are sensitive to others’ perspectives, they still seem to struggle with appropriately tailoring the specificity of their utterances to the visual context, often resulting in utterances that are under-informative [@matthews2012; @leung2024].
 When children must describe one of two very similar images, 4 and 5-year olds struggle to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
@@
 4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an iterated reference game where children could only use gestures to communicate [@bohn2019].
-
-A recent study with younger children examined 4-8 year olds' abilities to play an iterated reference game with a parent using a simple, child-friendly tablet-based task.
+Recent work has established that young children can succeed at iterated reference games when playing with their parents [@leung2024].
+4-8 year olds' played an iterated reference game with a parent using a simple, child-friendly tablet-based task.
 In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
@@
 
-Given the task demands in @glucksberg1966 and work showing that children can succeed in a simplified, less demanding paradigm, we revisit the question of child-child referential communication.
+Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, we revisit the question of child-child referential communication.
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
@@
 
-# Experiment 1 Methods
+# Experiment 1
+
+## Methods
 
@@
 
-## Participants
+### Participants
 4 and 5-year old children were recruited from a university preschool during the school day.
@@
 
-## Materials
+### Materials
 For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B).
@@
 
-## Procedure
+### Procedure
 Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room.
@@
 
-## Data processing
+### Data processing
 
@@
 
-# Experiment 1 Results
-
-## Accuracy
+## Results
+
+### Accuracy
 
@@
   ggplot(aes(x = trial, y = correct)) +
-  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
-  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
-  scale_fill_manual(values = time_palette) +
+    geom_smooth(method = "lm", color = "black") +
+  stat_summary(aes(color = type), fun.data = "mean_cl_boot") +
+  #stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
+  scale_color_manual(values = time_palette) +
   # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
-  geom_smooth(method = "lm", color = "black") +
   # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
@@
   ylab("Accuracy") +
-  scale_x_continuous(breaks = seq(1, 16, 4), expand = c(0, 0)) +
-  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0)) +
+  geom_vline(xintercept = 4.5, lty="dashed")+
+    geom_vline(xintercept = 8.5, lty="dashed")+
+    geom_vline(xintercept = 12.5, lty="dashed")+
+  scale_x_continuous(breaks = seq(1, 16, 4), expand = c(0.02, 0.02)) +
+  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
+  coord_cartesian(ylim = c(0.3,1))+
   geom_hline(yintercept = .5) +
@@
     legend.position = "none",
+    axis.ticks.x=element_blank(),
+    axis.text.x=element_blank(),
     panel.grid = element_blank(),
@@
 -->  As shown in Figure \ref{acc}, children were above chance in their selections.
-To test for changes in accuracy over time, we fit a mixed effects model of accuracy (`r acc_form |> form()`).
+To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy (`r acc_form |> form()`).
 Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
 
-## Speed
+### Speed
 
@@
 <!--[less important, could drop] --> As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
-We ran a Bayesian mixed effects model: `r speed_form |> form()`.
-The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial ).
-
-## Description length
+We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds: `r speed_form |> form()`.
+The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial).
+
+### Description length
 
@@
 In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024].
-We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`.
+We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced: `r description_form |> form()`.
 On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
 
-How long of a description is initially warranted depends on how easily describable the target is, as well as how large and close the contrast set is.
-In this case, the low number of options and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
-The adult controls in @leung2024 use shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.
-However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
-<!-- some of this is kinda discussion-y, possibly move -->
-
-## Convergence
-
-```{r}
+### Convergence
+
+```{r, eval=F}
 conv_priors <- c(
@@
 Different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds (Figure \ref{sim}).
-We modeled this as `r sim_across_between_form |> form()`.
+We modeled this as a Bayesian mixed effects linear regression predicting similarity `r sim_across_between_form |> form()`.
 Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
@@
 
-# Experiment 1 Discussion
+## Discussion
 
@@
 In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.
-
-# Experiment 2 Methods
-
-As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at https://osf.io/y2dax.
-
-## Participants
-Experiment 2 was run between March and August of 2024, at the same university preschool as experiment 1.
-No children participated in both experiments.
-30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
-Our target age range was 4 and 5 year olds, but one almost 4-year-old was unintentionally included.
-Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
-
-## Materials
-The same 4 critical images were used as in experiment 1, although this time, children saw these images 4 times.
-In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
-We used a total of 4 practice trials to provide a gradient from familiar shapes to less recognizable, blockier shapes (Figure \ref{game}D).
-
-## Procedure
-The procedure was much the same as Experiment 1(Figure \ref{game}D).
-We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in Experiment 1).
-The experimental script was fully written out and memorized by experimenters so children all received the same instructions.
-We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
-Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen to move the game along, but avoided contributing or repeating any content about the images or the descriptions.
 
@@
 ```
-## Data processing
+
+# Experiment 2
+
+## Methods
+
+As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at https://osf.io/y2dax.
+
+### Participants
+Experiment 2 was run between March and August of 2024, at the same university preschool as experiment 1.
+No children participated in both experiments.
+30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
+Our target age range was 4 and 5 year olds, but one almost 4-year-old was unintentionally included.
+Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
+
+### Materials
+The same 4 critical images were used as in experiment 1, although this time, children saw these images 4 times.
+In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
+We used a total of 4 practice trials to provide a gradient from familiar shapes to less recognizable, blockier shapes (Figure \ref{game}D).
+
+### Procedure
+The procedure was much the same as Experiment 1(Figure \ref{game}D).
+We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in Experiment 1).
+The experimental script was fully written out and memorized by experimenters so children all received the same instructions.
+We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
+Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen to move the game along, but avoided contributing or repeating any content about the images or the descriptions.
+
+### Data processing
 Data was processed in the same way as experiment 1.
@@
 
-# Experiment 2 Results
+## Results
 
 We report the same set of analyses and model results as in Experiment 1 as well as additional analyses of how the semantic content of children's descriptions changes over time.
-
-## Accuracy
+Children used a wide range of descriptions successfully, and some example descriptions from Experiment 2 are shown in Table \ref{example}.
+
+### Accuracy
 
@@
 
-As in Experiment 1, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
+In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy was relatively stable over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
 This confirms that children are able to communicate with each other about these abstract shapes.
 
-## Speed
+### Speed
 
@@
 
-In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)`seconds / trial).
+In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial).
 Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
 
-## Description length
+### Description length
 
@@
 
-Children used a wide range of descriptions successfully, and some example descriptions from Experiment 2 are shown in Table \ref{example}.
-Often children would use the same description as their partner, although there were some transitions such as from "a person holding a flute" to "a person holding a trumpet".
-
-## Convergence
+### Convergence
 
@@
 ```
+
+```{r, include=F}
+sims_2 |>
+  filter(game1 == game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier + 1) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .1, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  facet_wrap(~same_speaker) +
+  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
+```
+
+```{r, include=F}
+sims_2 |>
+  filter(game1 != game2) |>
+  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
+  filter(later == earlier) |>
+  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  geom_point(alpha = .01, position = position_dodge(width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+sim_across_between_2_results <- read_rds(here(mod_results, "sim_across_between_2.rds"))
+
+sim_across_between_form <- read_rds(here(mod_form, "sim_across_between_2.rds"))
+
+sim_across_2_results <- read_rds(here(mod_results, "sim_across_2.rds"))
+
+sim_across_form <- read_rds(here(mod_form, "sim_across_2.rds"))
+
+sim_to_next_2_results <- read_rds(here(mod_results, "sim_to_next_2.rds"))
+
+sim_to_next_form <- read_rds(here(mod_form, "sim_to_next_2.rds"))
+
+sim_to_last_2_results <- read_rds(here(mod_results, "sim_to_last_2.rds"))
+
+sim_to_last_form <- read_rds(here(mod_form, "sim_to_last_2.rds"))
+```
+
+<!--To look at semantic distance between utterances, we again operationalized similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019]. -->
+
+As a coarse comparison, we repeated the analysis from Experiment 1.
+Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
+
+The greater number of trials in Experiment 2 made it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
+
+To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity: `r sim_to_last_form |> form()`.
+Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+Descriptions were more similar to the last block if they came from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
+<!--Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
+--> Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
+
+As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
+We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression: `r sim_across_form |> form()`.
+As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
 
@@
 ```
-
-```{r, include=F}
-sims_2 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier + 1) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .1, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  facet_wrap(~same_speaker) +
-  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
-
-```{r, include=F}
-sims_2 |>
-  filter(game1 != game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .01, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
-
-```{r}
-sim_across_between_2_results <- read_rds(here(mod_results, "sim_across_between_2.rds"))
-
-sim_across_between_form <- read_rds(here(mod_form, "sim_across_between_2.rds"))
-
-sim_across_2_results <- read_rds(here(mod_results, "sim_across_2.rds"))
-
-sim_across_form <- read_rds(here(mod_form, "sim_across_2.rds"))
-
-sim_to_next_2_results <- read_rds(here(mod_results, "sim_to_next_2.rds"))
-
-sim_to_next_form <- read_rds(here(mod_form, "sim_to_next_2.rds"))
-
-sim_to_last_2_results <- read_rds(here(mod_results, "sim_to_last_2.rds"))
-
-sim_to_last_form <- read_rds(here(mod_form, "sim_to_last_2.rds"))
-```
-
-<!--To look at semantic distance between utterances, we again operationalized similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019]. -->
-
-As a coarse comparison, we repeated the analysis from Experiment 1.
-Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
-
-The greater number of trials in Experiment 2 made it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
-
-To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`.
-Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions were more similar to the last block if they came from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
-<!--Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
---> Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
-
-As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
-We tested whether descriptions in different games to the same target diverged over time: `r sim_across_form |> form()`.
-As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
-
 # Additional results
 
-## Interaction structure
+## Qualitative examples of interaction structure
 Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
@@
 
-## Meta-analytic models
-
-<!--Combining across experiments
-
-<!-- Might want to split out and incorporate in along with expt 2 analyses? -->
+## Pooling across experiments
 
@@
 
-As the two experiments were similar to one another, we re-ran models pooling across the two experiments. 
-Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`.
+As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments. 
+Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
 The average length of descriptions on the first trial was `r stats_text(description_mod_meta, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
@@
 
+It is unclear to what extent the uniformly short descriptions are a product of the simplified task or children's behavioral differences from adults.
+The appropriate length for an initial description depends on how easily describable the target is, as well as how large and close the contrast set is.
+In this case, the low number of options and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
+The adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.
+However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
+Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
+
 While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
@@
 The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.
-At 4-5-years old, children are able to use their linguistic skills communicatively, even as their utterances are not fully adult-like in form.
+At 4–5 years old, even though their utterances are not fully adult-like in form, children are nonetheless able to use their linguistic skills communicatively.
 

### 2025-01-21 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
refactor ms to pull out prep and model code into helpers

@@
 
-# experiment 1 imports
-transcript_1 <- read_csv(here(expt_1, "timed_transcript.csv"))
-dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial = trialNum)
-exclusion_1 <- transcript_1 %>%
-  filter(!is.na(description)) %>%
-  filter(role == "S") %>%
-  select(game, trial) %>%
-  unique() %>%
-  mutate("S_talked" = 1)
-
-words_1 <- transcript_1 |>
-  filter(!is.na(description)) |>
-  mutate(words = description |> str_count("\\S+")) |>
-  filter(role == "S") |>
-  group_by(game, trial) |>
-  summarize(words = sum(words))
-
-clean_1 <- dat_1 %>%
-  left_join(exclusion_1) %>%
-  filter(S_talked == 1) |>
-  left_join(words_1) |>
-  mutate(type = case_when(
-    trial < 2 ~ "practice",
-    trial < 6 ~ "block 1",
-    trial < 10 ~ "block 2",
-    trial < 14 ~ "block 3"
-  ))
-
-sims_1 <- read_rds(here(expt_1, "similarities.rds")) |>
-  as_tibble() |>
-  inner_join(clean_1 |> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
-  inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
-  filter(target1 == target2) |>
-  filter(trial1 > 1) |>
-  filter(trial2 > 1) |>
-  mutate(
-    trial1 = trial1 + 2,
-    trial2 = trial2 + 2
-  ) |>
-  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
-  mutate(
-    same_speaker = (speaker1 == speaker2) |> as.numeric(),
-    same_game = (game1 == game2) |> as.numeric(),
-    later = ifelse(block1 > block2, block1, block2),
-    earlier = ifelse(block1 > block2, block2, block1)
-  ) |>
-  mutate(target = target1)
-
-# expt 2 imports
-transcript_2 <- read_csv(here(expt_2, "transcripts.csv"))
-link <- read_csv(here(expt_2, "link_transcripts.csv"))
-
-exclude_2 <- transcript_2 |>
-  filter(!is.na(echo)) |>
-  select(gameId, trial)
-
-data_2 <- read_csv(here(expt_2, "clean_data.csv")) |>
-  rename(trial = trialNum) |>
-  anti_join(exclude_2)
-
-transcript_2 <- transcript_2 |>
-  anti_join(exclude_2) |>
-  filter(!is.na(description))
-
-words_2 <- transcript_2 |>
-  mutate(words = description |> str_count("\\S+")) |>
-  filter(role == "S") |>
-  group_by(gameId, gameConfig, trial) |>
-  summarize(words = sum(words))
-
-clean_2 <- data_2 |>
-  left_join(words_2) |>
-  filter(!is.na(words)) |>
-  mutate(type = case_when(
-    trial < 4 ~ "practice",
-    trial < 8 ~ "block 1",
-    trial < 12 ~ "block 2",
-    trial < 16 ~ "block 3",
-    trial < 20 ~ "block 4"
-  )) |>
-  filter(!is.na(response)) |>
-  mutate(correct = as.numeric(correct))
-
-sims_2 <- read_rds(here(expt_2, "similarities.rds")) |>
-  as_tibble() |>
-  inner_join(clean_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
-  inner_join(clean_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
-  filter(target1 == target2) |>
-  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
-  mutate(
-    same_speaker = (speaker1 == speaker2) |> as.numeric(),
-    same_game = (game1 == game2) |> as.numeric(),
-    later = ifelse(block1 > block2, block1, block2),
-    earlier = ifelse(block1 > block2, block2, block1)
-  ) |>
-  mutate(target = target1)
-```
-
-```{r, eval=F}
-library(tidybayes)
-
-save_summary <- function(model) {
-  intervals <- gather_draws(model, `b_.*`, regex = T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex = T) %>%
-    mutate(above_0 = ifelse(.value > 0, 1, 0)) %>%
-    group_by(.variable) %>%
-    summarize(pct_above_0 = mean(above_0)) %>%
-    left_join(intervals, by = ".variable") %>%
-    mutate(
-      lower = .lower,
-      upper = .upper,
-      Term = str_sub(.variable, 3, -1),
-      Estimate = .value
-    ) %>%
-    select(Term, Estimate, lower, upper)
-
-  stats
-}
-
-do_model <- function(path) {
-  model <- read_rds(here(mod_loc, path))
-  save_summary(model) |> write_rds(here(mod_loc, "summary", path))
-  model$formula |> write_rds(here(mod_loc, "formulae", path))
-  print(summary(model))
-}
-
-mods <- list.files(path = here(mod_loc), pattern = ".*rds") |> walk(~ do_model(.))
-```
-
-```{r}
-stats <- function(model, row, decimal = 2) {
-  model <- model |>
-    mutate(
-      Estimate = round(Estimate, digits = decimal),
-      Lower = round(lower, digits = decimal),
-      Upper = round(upper, digits = decimal),
-      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
-    ) |>
-    select(Term, Estimate, `Credible Interval`)
-  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
-}
-
-stats_text <- function(model, row, decimal = 2) {
-  model <- model |>
-    mutate(
-      Estimate = round(Estimate, digits = decimal) |> formatC(format='f', digits=decimal ),
-      Lower = round(lower, digits = decimal) |> formatC(format='f', digits=decimal ),
-      Upper = round(upper, digits = decimal) |> formatC(format='f', digits=decimal ),
-      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
-    ) |>
-    select(Term, Estimate, `Credible Interval`)
-  str_c(model[row, 2], "  ", model[row, 3])
-}
-
-form <- function(model_form) {
-  dep <- as.character(model_form$formula[2])
-  ind <- as.character(model_form$formula[3])
-
-  str_c(dep, " ~ ", ind) |>
-    str_replace_all(" ", "") |>
-    str_replace_all("\\*", " $\\\\times$ ") |>
-    str_replace_all("\\+", "&nbsp;+ ") |>
-    str_replace_all("~", "$\\\\sim$ ")
-}
+source(here("code/paper_helper.R"))
 ```
@@
 # Introduction
-
-<!-- intro -->
 
@@
 Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
-<!--Indeed, even young children must make use of these skills in order to achieve success in the task.
-However, there is debate on the developmental trajectories of these communicative skills.
---> Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
+Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
 
@@
 Even after multiple rounds with the same images, children were not able to correctly order the blocks.
-<!--@glucksberg1966 attributed young children's failure to the children’s use of idiosyncratic referring expressions unique to their own experiences, making it difficult for the pair to converge on shared descriptions.-->
 Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions.
@@
 
-<!--(Note: we should talk about whether it makes more sense to do all the methods and then all the results versus the current expt 1 than expt 2 layout.) -->
-
 Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
-Young children can be very sensitive to task demands and cognitive load (CITATIONS), so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
+Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
 This experiment was pre-registered at https://osf.io/kcv8j.
@@
 ### Participants
+
 4 and 5-year old children were recruited from a university preschool during the school day.
@@
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
-
-```{r, eval=F}
-# demo <- read_csv(here("demographics.csv")) |>
-#   filter(included == "x") |>
-#   mutate(date_test = mdy(date_test), date_birth = mdy(date_birth)) |>
-#   select(date_test, date_birth, gender) |>
-#   mutate(
-#     age = date_test - date_birth,
-#     age = as.numeric(age) / 365.25 * 12
-#   ) |> 
-#   mutate(expt=ifelse(date_test>ymd("2024-01-01"), "expt2", "expt1")) |> 
-#   group_by(expt) |> 
-#   summarize(female=sum(gender=="female"),
-#             total=n(),
-#             median=median(age),
-#             min=min(age),
-#             max=max(age))
-```
 
@@
 
-The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements.
+The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements (especially when utterances were inaudible or the child did not respond immediately; this aspect of the procedure was modified in Experiment 2).
 The entire interaction was video-recorded.
@@
 Children's selections and the time to selection were recorded from the experiment software.
-Children's descriptions were transcribed from the video-recording, using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters.
+Children's descriptions were transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters.
 Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used.
@@
 
-### Accuracy
+### Accuracy and Speed
 
@@
     geom_vline(xintercept = 12.5, lty="dashed")+
-  scale_x_continuous(breaks = seq(1, 16, 4), expand = c(0.02, 0.02)) +
+  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
   scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
@@
     legend.position = "none",
-    axis.ticks.x=element_blank(),
-    axis.text.x=element_blank(),
     panel.grid = element_blank(),
     strip.background = element_blank()
-  )
-```
-
-```{r, eval=F}
-acc_priors <- c(
-  set_prior("normal(0,1)", class = "b"),
-  set_prior("normal(0,1)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-acc_mod_data_1 <- clean_1 |>
-  mutate(
-    correct.num = ifelse(correct == TRUE, 1, 0),
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-acc_mod_1 <- brm(correct.num ~ trial.num + (trial.num | game) + (1 | target),
-  family = bernoulli(link = "logit"),
-  data = acc_mod_data_1,
-  file = here(mod_loc, "acc_1.rds"),
-  prior = acc_priors,
-  control = list(adapt_delta = .95)
   )
@@
 <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
--->  As shown in Figure \ref{acc}, children were above chance in their selections.
-To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy (`r acc_form |> form()`).
-Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
-
-### Speed
-
-```{r, eval=F}
-speed_priors <- c(
-  set_prior("normal(60,100)", class = "Intercept"),
-  set_prior("normal(0,20)", class = "b"),
-  set_prior("normal(0,20)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-speed_mod_data_1 <- clean_1 |>
-  mutate(
-    time.sec = time / 1000,
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-speed_mod_1 <- brm(time.sec ~ trial.num + (trial.num | game) + (1 | target),
-  data = speed_mod_data_1,
-  file = here(mod_loc, "speed_1.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
+-->  As shown in Figure \ref{acc}, children generally accurate in their selections.
+To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
+This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.
 
@@
 
-<!--[less important, could drop] --> As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
-We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds: `r speed_form |> form()`.
-The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial).
+As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
+We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds.^[`r speed_form |> form()`] The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial).
+This is a measure of efficiency as children were able to achieve the same accuracy levels in less time.
 
@@
 
-```{r, eval=F}
-description_priors <- c(
-  set_prior("normal(5,10)", class = "Intercept"),
-  set_prior("normal(0,5)", class = "b"),
-  set_prior("normal(0,5)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-description_mod_data_1 <- clean_1 |>
-  mutate(
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-description_mod_1 <- brm(words ~ trial.num + (trial.num | game) + (1 | target),
-  data = description_mod_data_1,
-  file = here(mod_loc, "description_1.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of referential expression (in words) produced by the speaker each trial.
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of referential expression (in words) produced by the teller each trial.
 Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs.
@@
   scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
+    geom_vline(xintercept = 4.5, lty="dashed")+
+    geom_vline(xintercept = 8.5, lty="dashed")+
+    geom_vline(xintercept = 12.5, lty="dashed")+
   scale_color_manual(values = time_palette) +
@@
 In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024].
-We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced: `r description_form |> form()`.
-On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
+We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
+Thus, while children were taking less time per trial, their speed-up was not from shorter utterances, instead they were understanding the game better and taking less time to initiate their descriptions and make their selections.
+
+Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example}; these specific examples are from Experiment 2, but are also representative of the types of descriptions produced in Experiment 1.
 
@@
 
-```{r, eval=F}
-conv_priors <- c(
-  set_prior("normal(.5,.2)", class = "Intercept"),
-  set_prior("normal(0,.1)", class = "b"),
-  set_prior("normal(0,.05)", class = "sd")
-)
-
-sim_across_between_data_1 <- sims_1 |>
-  mutate(same_game = case_when(
-    game1 == game2 ~ 1,
-    T ~ 0
-  )) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
-
-sim_across_between_mod_1 <- brm(sim ~ same_game + same_speaker + (1 | target),
-  data = sim_across_between_data_1,
-  file = here(mod_loc, "sim_across_between_1.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between referential descriptions from the speaker to the same target under different circumstances.
-Different games refers to similarities across two speakers from different games, same game to similarities across the two participants in one game, and same speaker to descriptions from the same person in different blocks.
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between referential descriptions from the teller to the same target under different circumstances.
+Different games refers to similarities across two children from different games, same game to similarities across the two participants in one game, and same teller to descriptions from the same person in different blocks.
 Dots are means and lines are bootstrapped 95% CIs.
@@
   mutate(source = case_when(
-    speaker1 == speaker2 ~ "same speaker",
+    speaker1 == speaker2 ~ "same teller",
     game1 == game2 ~ "same game",
@@
     mutate(source = case_when(
-      speaker1 == speaker2 ~ "same speaker",
+      speaker1 == speaker2 ~ "same teller",
       game1 == game2 ~ "same game",
@@
   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
+  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "Target") +
   facet_grid(expt ~ .) +
@@
 
-```{r, include=F}
-sims_1 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == 3) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .1, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  facet_wrap(~same_speaker) +
-  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
+While description length is used as a proxy for measuring convention formation, semantic overlap between utterances does not always track description length.
+@boyce2024 introduced a more sensitive measure of semantic convergence.
+By comparing the content of utterances using word embeddings, @boyce2024 traced when how similarities between utterances in different rounds and games
 
-```{r include=F}
-sims_1 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier + 1) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .1, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  facet_wrap(~same_speaker) +
-  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
-
-```{r, include=F}
-sims_1 |>
-  filter(game1 != game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .01, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
-
-While description length is used as a proxy for measuring convention formation, better measurements for convergence look at the actual content of the utterances, which does not always track description length [@boyce2024].
 Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions [@reimers2019].
@@
 We compared the semantic similarities between descriptions of the same target based on who produced the description.
-Our question was whether the two children within the same game produced descriptions that are more similar to each other than two children in different games.
-
 Different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds (Figure \ref{sim}).
-We modeled this as a Bayesian mixed effects linear regression predicting similarity `r sim_across_between_form |> form()`.
-Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
+We modeled this as a Bayesian mixed effects linear regression predicting similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
 The big differences in descriptions between games compared to within games is a measure of partner sensitivity.
@@
 
+The biggest change between the experiments was increasing the number of repetitions of target stimuli from 3 to 4 (from 12 to 16 trials).
+The greater number of trials in Experiment 2 made it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
+
 ### Participants
@@
 30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
-Our target age range was 4 and 5 year olds, but one almost 4-year-old was unintentionally included.
+Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included.
 Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
@@
 ### Materials
-The same 4 critical images were used as in experiment 1, although this time, children saw these images 4 times.
-In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
+The same 4 critical images were used as in experiment 1.In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
 We used a total of 4 practice trials to provide a gradient from familiar shapes to less recognizable, blockier shapes (Figure \ref{game}D).
@@
 ### Data processing
-Data was processed in the same way as experiment 1.
+Data were processed in the same way as experiment 1.
 After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
@@
 
-We report the same set of analyses and model results as in Experiment 1 as well as additional analyses of how the semantic content of children's descriptions changes over time.
-Children used a wide range of descriptions successfully, and some example descriptions from Experiment 2 are shown in Table \ref{example}.
-
 ### Accuracy
-
-```{r, eval=F}
-acc_priors <- c(
-  set_prior("normal(0,1)", class = "b"),
-  set_prior("normal(0,1)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-acc_mod_data <- clean_2 |>
-  mutate(
-    correct.num = ifelse(correct == TRUE, 1, 0),
-    trial.num = trial - 4
-  ) |>
-  filter(type != "practice")
-acc_mod <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
-  family = bernoulli(link = "logit"),
-  data = acc_mod_data,
-  file = here(mod_loc, "acc_2.rds"),
-  prior = acc_priors,
-  control = list(adapt_delta = .95)
-)
-```
 
@@
 ```{r}
-speed_priors <- c(
-  set_prior("normal(60,100)", class = "Intercept"),
-  set_prior("normal(0,20)", class = "b"),
-  set_prior("normal(0,20)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-speed_mod_data_2 <- clean_2 |>
-  mutate(
-    time.sec = time / 1000,
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-speed_mod_2 <- brm(time.sec ~ trial.num + (trial.num | gameId) + (1 | target),
-  data = speed_mod_data_2,
-  file = here(mod_loc, "speed_2.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r}
 speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))
@@
 
-```{r, eval=F}
-description_priors <- c(
-  set_prior("normal(5,10)", class = "Intercept"),
-  set_prior("normal(0,5)", class = "b"),
-  set_prior("normal(0,5)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-description_mod_data_2 <- clean_2 |>
-  mutate(
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")
-
-description_mod_2 <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
-  data = description_mod_data_2,
-  file = here(mod_loc, "description_2.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
 ```{r}
@@
 This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
+As with Experiment 1, children used a variety of descriptions successfully (Table \ref{example}).
 
 ### Convergence
-
-```{r}
-conv_priors <- c(
-  set_prior("normal(.5,.2)", class = "Intercept"),
-  set_prior("normal(0,.1)", class = "b"),
-  set_prior("normal(0,.05)", class = "sd")
-)
-
-sim_across_between_data_2 <- sims_2 |>
-  mutate(same_game = case_when(
-    game1 == game2 ~ 1,
-    T ~ 0
-  )) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
-
-sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker + (1 | target),
-  data = sim_across_between_data_2,
-  file = here(mod_loc, "sim_across_between_2.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-sim_to_last_data_2 <- sims_2 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == 4) |>
-  mutate(
-    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-    earlier_block.num = earlier - 1
-  )
-
-sim_to_last_mod_2 <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
-  data = sim_to_last_data_2,
-  file = here(mod_loc, "sim_to_last_2.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-sim_to_next_data_2 <- sims_2 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier + 1) |>
-  mutate(
-    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-    earlier_block.num = earlier - 1
-  )
-
-sim_to_next_mod_2 <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
-  data = sim_to_next_data_2,
-  file = here(mod_loc, "sim_to_next_2.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-sim_across_data_2 <- sims_2 |>
-  filter(game1 != game2) |>
-  filter(block1 == block2) |>
-  mutate(block.num = block1 - 1)
-
-sim_across_mod_2 <- brm(sim ~ block.num + (1 | target),
-  data = sim_across_data_2,
-  file = here(mod_loc, "sim_across_2.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r, include=F}
-sims_2 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier + 1) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .1, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  facet_wrap(~same_speaker) +
-  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
-
-```{r, include=F}
-sims_2 |>
-  filter(game1 != game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
-  geom_point(alpha = .01, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
-```
 
@@
 
-<!--To look at semantic distance between utterances, we again operationalized similarity between pairs of utterances as the cosine similarity between their Sentence-BERT embeddings [@reimers2019]. -->
-
 As a coarse comparison, we repeated the analysis from Experiment 1.
@@
 
-The greater number of trials in Experiment 2 made it possible to look for changes over time that could be indicative of convergence to shared descriptions within a game and divergence between games.
-
-To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity: `r sim_to_last_form |> form()`.
-Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
 Descriptions were more similar to the last block if they came from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
-<!--Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`.
---> Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
+Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
 
 As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
-We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression: `r sim_across_form |> form()`.
-As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
+We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
 
 ```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between earlier blocks (1-3) with last block (4) for descriptions to the same image within the same group.
-SImilarity measured as cosine similarity between S-BERT embeddings of referential descriptions.
+Similarity measured as cosine similarity between S-BERT embeddings of referential descriptions.
 Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
@@
   filter(later == 4) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
   ggplot(aes(as.character(earlier), sim, color = target1)) +
@@
   scale_color_solarized() +
-  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
-  theme(legend.position = "bottom")
+  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
+  theme(legend.position = "bottom",
+            strip.background = element_blank(),
+        panel.grid=element_blank()
+)
 ```
-# Additional results
 
-## Qualitative examples of interaction structure
+### Qualitative variation in interaction structure
+
 Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
@@
 
-## Pooling across experiments
-
-```{r, eval=F}
-
-acc_priors <- c(
-  set_prior("normal(0,1)", class = "b"),
-  set_prior("normal(0,1)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-acc_mod_data_meta <-  clean_1 |>
-  mutate(
-    correct.num = ifelse(correct == TRUE, 1, 0),
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice") |> mutate(gameId=game) |> bind_rows(clean_2 |>
-  mutate(
-    correct.num = ifelse(correct == TRUE, 1, 0),
-    trial.num = trial - 4
-  ) |>
-  filter(type != "practice"))
-
-# acc_mod_data_meta |> summarize(m=mean(correct.num))
-
-acc_mod_both <- brm(correct.num ~ trial.num + (trial.num | gameId) + (1 | target),
-  family = bernoulli(link = "logit"),
-  data = acc_mod_data_meta,
-  file = here(mod_loc, "acc_meta.rds"),
-  prior = acc_priors,
-  control = list(adapt_delta = .95)
-)
-
-```
-
-```{r, eval=F}
-description_priors <- c(
-  set_prior("normal(5,10)", class = "Intercept"),
-  set_prior("normal(0,5)", class = "b"),
-  set_prior("normal(0,5)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-description_mod_data_meta <- clean_1 |>
-  mutate(
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice")|> mutate(gameId=game) |> bind_rows(
- clean_2 |>
-  mutate(
-    trial.num = trial - 2
-  ) |>
-  filter(type != "practice"))
-
-description_mod_meta <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
-  data = description_mod_data_meta,
-  file = here(mod_loc, "description_meta.rds"),
-  prior = speed_priors,
-  control = list(adapt_delta = .95)
-)
-
-```
-
-```{r, eval=F}
-conv_priors <- c(
-  set_prior("normal(.5,.2)", class = "Intercept"),
-  set_prior("normal(0,.1)", class = "b"),
-  set_prior("normal(0,.05)", class = "sd")
-)
-
-sim_across_between_data_2 <- sims_2 |>
-  mutate(same_game = case_when(
-    game1 == game2 ~ 1,
-    T ~ 0
-  )) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
-
-sim_across_between_data_1 <- sims_1 |>
-  mutate(same_game = case_when(
-    game1 == game2 ~ 1,
-    T ~ 0
-  )) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))
-
-sim_across_between_data_meta <- sim_across_between_data_2 |> bind_rows(sim_across_between_data_1)
-
-sim_across_between_mod_meta <- brm(sim ~ same_game + same_speaker + (1 | target),
-  data = sim_across_between_data_meta,
-  file = here(mod_loc, "sim_across_between_meta.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-sim_to_next_data_meta <- sims_2 |> bind_rows(sims_1) |> 
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == earlier + 1) |>
-  mutate(
-    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-    earlier_block.num = earlier - 1
-  )
-
-sim_to_next_mod_meta <- brm(sim ~ earlier_block.num + same_speaker + (1 | game1) + (1 | target),
-  data = sim_to_next_data_meta,
-  file = here(mod_loc, "sim_to_next_meta.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-```
+# Joint analysis
 
@@
 
-As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments. 
+As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect.
 Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
@@
 Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
-<!--Since the two experiments had different length games, we do not pool them in comparisons to the last block, but we can pool them comparing to the next block.
-The distance between adjacent block utterances is relatively constant: `r sim_to_next_mod_meta |> stats_text(2,3)`.
--->
-
-<!-- Query -- I did not include experiment # as a random effect here, and maybe should have?
-I can rerun if we decide we want this analysis and think expt # as a random effect with game nested within it would be useful.
--->
-
-## Relation between accuracy and convergence
-
-```{r}
-# try visual first
-
-acc_sim_1 <- clean_1 |>
-  filter(type != "practice") |> mutate(game1=game) |> 
-  mutate(correct=ifelse(correct, 1,0),
-         earlier_trial=trial+2) |> 
-  select(earlier_trial, target, correct, game1)
-
-acc_sim_2 <- clean_2 |>
-  filter(type != "practice") |> select(earlier_trial=trial, target, correct, game1=gameId)
-
-sim_acc_2 <- sims_2 |> 
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
-  filter(later == earlier + 1) |>
-  mutate(
-    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-    earlier_block.num = earlier - 1
-  ) |> 
-  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker) |> 
-  left_join(acc_sim_2)
-
-sim_acc_1 <- sims_1 |> 
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
-  filter(later == earlier + 1) |>
-  mutate(
-    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
-    earlier_block.num = earlier - 1
-  ) |> 
-  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker) |> 
-  left_join(acc_sim_1)
-
-#ggplot(sim_acc_1, aes(x=earlier, y=sim, color=as.factor(correct)))+geom_point(alpha=.4, position=position_dodge(.4))+stat_summary()
-
-#ggplot(sim_acc_2, aes(x=earlier, y=sim, color=as.factor(correct)))+geom_point(alpha=.4, position=position_dodge(.4))+stat_summary()
-
-```
-
-```{r, eval=F}
-conv_priors <- c(
-  set_prior("normal(.5,.2)", class = "Intercept"),
-  set_prior("normal(0,.1)", class = "b"),
-  set_prior("normal(0,.05)", class = "sd")
-)
-sim_acc_to_next_data <- sim_acc_1 |> bind_rows(sim_acc_2)
-sim_acc_to_next_mod_meta <- brm(sim ~ earlier_block.num*correct + same_speaker*correct + (1 | game1) + (1 | target),
-  data = sim_acc_to_next_data,
-  file = here(mod_loc, "sim_acc_to_next_meta.rds"),
-  prior = conv_priors,
-  control = list(adapt_delta = .95)
-)
-
-# more stickiness if correct, no clear interactions 
-# being correct has a similar size effect to same speaker
-```
 
@@
 Here we operationalized "staying power" as similarity to the next block utterance, and we modeled it with `r form(sim_acc_to_next_form)`.
-Correct utterances had an increase in similarity with the next block  (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or same speaker.
-Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same speaker `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
+Correct utterances had an increase in similarity with the next block  (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or same teller.
+Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
 
@@
 
-In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed much later, by pragmatic and communicative competency (CITATIONS), or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning (CITATIONS).
+In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed later by pragmatic competency [for instance, semantics before pragmatic implicatures; @huang2009, @noveck2001], or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning [@bates1974; @bohn2019a; @tomasello2008].
 The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.

### 2025-01-21 code/mods_for_cogsci_paper.R

@@ -1,0 +1,441 @@
+
+library(tidyverse)
+library(here)
+library(brms)
+library(rstan)
+library(rstanarm)
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+expt_1 <- "data/expt_1"
+expt_2 <- "data/expt_2"
+mod_loc <- "code/models"
+mod_results <- "code/models/summary"
+mod_form <- "code/models/formulae"
+
+source(here("code/paper_helper.R"))
+
+### accuracy expt 1
+
+acc_priors <- c(
+  set_prior("normal(0,1)", class = "b"),
+  set_prior("normal(0,1)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+acc_mod_data_1 <- clean_1 |>
+  mutate(
+    correct.num = ifelse(correct == TRUE, 1, 0),
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
+acc_mod_1 <- brm(correct.num ~ trial.num + (trial.num | game) + (1 | target),
+                 family = bernoulli(link = "logit"),
+                 data = acc_mod_data_1,
+                 file = here(mod_loc, "acc_1.rds"),
+                 prior = acc_priors,
+                 control = list(adapt_delta = .95)
+)
+
+### speed expt 1
+
+speed_priors <- c(
+  set_prior("normal(60,100)", class = "Intercept"),
+  set_prior("normal(0,20)", class = "b"),
+  set_prior("normal(0,20)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+speed_mod_data_1 <- clean_1 |>
+  mutate(
+    time.sec = time / 1000,
+    trial.num = trial - 2
+  ) |>
+  filter(type != "practice")
+
+speed_mod_1 <- brm(time.sec ~ trial.num + (trial.num | game) + (1 | target),
+                   data = speed_mod_data_1,
+                   file = here(mod_loc, "speed_1.rds"),
+                   prior = speed_priors,
+                   control = list(adapt_delta = .95)
+)
+
+#### description expt 1
+description_priors <- c(
+  set_prior("normal(5,10)", class = "Intercept"),
+  set_prior("normal(0,5)", class = "b"),
+  set_prior("normal(0,5)", class = "sd"),
+  set_prior("lkj(1)", class = "cor")
+)
+
+description_mod_data_1 <- clean_1 |>
[…]

### 2025-01-21 code/paper_helper.R

@@ -1,0 +1,156 @@
+
+## read in all the data stuff
+# experiment 1 imports
+transcript_1 <- read_csv(here(expt_1, "timed_transcript.csv"))
+dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial = trialNum)
+exclusion_1 <- transcript_1 %>%
+  filter(!is.na(description)) %>%
+  filter(role == "S") %>%
+  select(game, trial) %>%
+  unique() %>%
+  mutate("S_talked" = 1)
+
+words_1 <- transcript_1 |>
+  filter(!is.na(description)) |>
+  mutate(words = description |> str_count("\\S+")) |>
+  filter(role == "S") |>
+  group_by(game, trial) |>
+  summarize(words = sum(words))
+
+clean_1 <- dat_1 %>%
+  left_join(exclusion_1) %>%
+  filter(S_talked == 1) |>
+  left_join(words_1) |>
+  mutate(type = case_when(
+    trial < 2 ~ "practice",
+    trial < 6 ~ "block 1",
+    trial < 10 ~ "block 2",
+    trial < 14 ~ "block 3"
+  ))
+
+
+sims_1 <- read_rds(here(expt_1, "similarities.rds")) |>
+  as_tibble() |>
+  inner_join(clean_1 |> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
+  inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
+  filter(target1 == target2) |>
+  filter(trial1 > 1) |>
+  filter(trial2 > 1) |>
+  mutate(
+    trial1 = trial1 + 2,
+    trial2 = trial2 + 2
+  ) |>
+  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
+  mutate(
+    same_speaker = (speaker1 == speaker2) |> as.numeric(),
+    same_game = (game1 == game2) |> as.numeric(),
+    later = ifelse(block1 > block2, block1, block2),
+    earlier = ifelse(block1 > block2, block2, block1)
+  ) |>
+  mutate(target = target1)
+
+
+# expt 2 imports
+transcript_2 <- read_csv(here(expt_2, "transcripts.csv"))
+link <- read_csv(here(expt_2, "link_transcripts.csv"))
+
+exclude_2 <- transcript_2 |>
[…]

### 2025-01-21 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
pass incorporating mike and alvin comments

@@
 
-```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental Setup and Procedure.
+```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure.
 Panel A shows the experimental setup with the teller and guesser across the table from each other.
@@
 
-Our goal for experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
+Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
 Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
@@
 Targets were randomly paired with another of the critical images as the foil.
-
 The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements (especially when utterances were inaudible or the child did not respond immediately; this aspect of the procedure was modified in Experiment 2).
@@
 Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used.
-
 We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe.
@@
 
-Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We present the estimate and 95% credible intervals.
+Statistical analyses were run in brms [@burkner2018] with weakly informative priors.
+We present the estimate and 95% credible intervals.
+The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at VERONICA SHOULD MAKE AN OSF LINK.
 
@@
 
-### Accuracy and Speed
+### Accuracy and speed
 
 ```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time.
-Experiment 1 had 3 blocks (12 total critical trials) and experiment 2 had 4 blocks (16 critical trials).
 Error bars are bootstrapped 95% CIs with a linear trend line overlaid.
@@
 
-```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of referential expression (in words) produced by the teller each trial.
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of description produced by the teller each trial.
 Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs.
@@
 In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024].
-We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
-Thus, while children were taking less time per trial, their speed-up was not from shorter utterances, instead they were understanding the game better and taking less time to initiate their descriptions and make their selections.
+We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words, and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
+Thus, children's increasing speed was not from shorter utterances, but instead from some combination of improved task understanding, faster utterance planning, and faster decisions of what to select.
 
-Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example}; these specific examples are from Experiment 2, but are also representative of the types of descriptions produced in Experiment 1.
+Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
 
@@
 
-```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between referential descriptions from the teller to the same target under different circumstances.
-Different games refers to similarities across two children from different games, same game to similarities across the two participants in one game, and same teller to descriptions from the same person in different blocks.
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources.
 Dots are means and lines are bootstrapped 95% CIs.
@@
 @boyce2024 introduced a more sensitive measure of semantic convergence.
-By comparing the content of utterances using word embeddings, @boyce2024 traced when how similarities between utterances in different rounds and games
+By comparing the content of utterances using word embeddings, @boyce2024 traced how similarities within and across games changed over time.
+<!--To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game.
+-->
 
-Following @boyce2024, we use Sentence-BERT to embed the descriptions in a semantic vector space, and then use the cosine between embeddings as a measure of semantic similarity between descriptions [@reimers2019].
+With only 3 blocks of descriptions, we do not believe it is meaningful to look at whether the 2nd and 3rd blocks are more similar than the 1st and 3rd blocks.
+However, we can test for a more coarse measure of sensitivity to partner: whether children's utterances are more like their partner's than children in other games'.
+If children are fully ego-centric [as suggested by @glucksberg1966], their choices of descriptions would be independent from their partners.
 
-We compared the semantic similarities between descriptions of the same target based on who produced the description.
-Different children in the same game produced more similar descriptions than children in different games, although the descriptions were less similar than descriptions from the same child in different rounds (Figure \ref{sim}).
-We modeled this as a Bayesian mixed effects linear regression predicting similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_1_results |> stats_text(3,3)`).
-The big differences in descriptions between games compared to within games is a measure of partner sensitivity.
+Following the methods of @boyce2024, we embedded each description in a semantic vector space using S-BERT [@reimers2019], and then used the cosine between embeddings as a measure of semantic similarity.
+
+We compared the semantic similarities between descriptions of the same target based on who produced the description (Figure \ref{sim}).
+We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`).
+Utterances were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves.
+However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating some sensitivity to their partner's expressions.
 
@@
 \begin{table}
-	\caption{Example descriptions children successfully used used to identify different target tangram images in Experiment 2.}
+	\caption{Example descriptions children successfully used used to identify different target images in Experiment 2.}
 	\label{example}
@@
 ### Participants
-Experiment 2 was run between March and August of 2024, at the same university preschool as experiment 1.
+Experiment 2 was run between March and August of 2024, at the same university preschool as Experiment 1.
 No children participated in both experiments.
@@
 ### Materials
-The same 4 critical images were used as in experiment 1.In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for experiment 2.
+The same 4 critical images were used as in Experiment 1.
+In response to some children struggling with the abrupt switch from familiar to non-nameable shapes, we introduced more practice trials for Experiment 2.
 We used a total of 4 practice trials to provide a gradient from familiar shapes to less recognizable, blockier shapes (Figure \ref{game}D).
@@
 ### Data processing
-Data were processed in the same way as experiment 1.
+Data were processed in the same way as Experiment 1.
 After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
@@
 
-### Accuracy
+### Accuracy and speed
 
@@
 
-In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy was relatively stable over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
-This confirms that children are able to communicate with each other about these abstract shapes.
-
-### Speed
-
 ```{r}
@@
 
-In Experiment 2, the first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial).
+In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy was relatively stable over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
+The first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial).
 Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
+Taken together, we find more evidence that children can successfully communicate with each other about these abstract shapes, and do so with increasing efficiency.
 
@@
 
-The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial , Figure \ref{length}).
+The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial, Figure \ref{length}).
 This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
-As with Experiment 1, children used a variety of descriptions successfully (Table \ref{example}).
 
@@
 
-As a coarse comparison, we repeated the analysis from Experiment 1.
+As a course measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1.
 Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
 
-To look for convergence to shared descriptions within games, we compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions were more similar to the last block if they came from the same child who gave the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).
-Although over time descriptions do get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
+As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description.
+We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
+Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
+Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
 
 As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
-We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
+We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions.
+These patterns of increasing similarity within games and increasing divergence between games match the patterns found for adults, albeit weakly [@boyce2024].
 
-```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between earlier blocks (1-3) with last block (4) for descriptions to the same image within the same group.
-Similarity measured as cosine similarity between S-BERT embeddings of referential descriptions.
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block.
 Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
@@
 
-### Qualitative variation in interaction structure
+<!--### Qualitative variation in interaction structure
 
@@
 This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+
+-->
 
@@
 As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect.
-Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
-The average length of descriptions on the first trial was `r stats_text(description_mod_meta, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
-
-Utterances were more similar if they came from the same partnership (`r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
+Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly, but unreliably, increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
+Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
 
@@
 
-As a post-hoc measure, we looked at whether descriptions that elicited correct selections had more staying power than descriptions that did not elicit correct selections.
-Here we operationalized "staying power" as similarity to the next block utterance, and we modeled it with `r form(sim_acc_to_next_form)`.
-Correct utterances had an increase in similarity with the next block  (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or same teller.
-Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
+We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds.
+Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt.
+To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar with the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
+This pattern of results is consistent with children sticking to their own description more when it was previously successful last time, and children repeat the other child's description more when they understood it.
+<!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
+-->
 
-# General Discussion
+# General discussion
+
+Prominent classical findings have been interpreted as evidence that young children are cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+However, an emerging pattern of research finds that  young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
+<!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
+-->We re-addressed the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
 
@@
 Children did not display a reduction in the length of referential expressions over the course of the game.
+While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
 
 It is unclear to what extent the uniformly short descriptions are a product of the simplified task or children's behavioral differences from adults.
-The appropriate length for an initial description depends on how easily describable the target is, as well as how large and close the contrast set is.
-In this case, the low number of options and relatively easy to describe shapes may have meant long initial descriptions were less necessary.
-The adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images.
-However, young children may also choose to provide shorter descriptions than adults would, and young children may accept shorter descriptions as adequate whereas adults may ask for more information to increase their confidence level before making a guess.
+In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions.
+Indeed, adults controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024].
+However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification.
 Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
 
-While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+The generalizability of our results is limited by the target population, the target images, and the task structure.
+We sampled a convenience poplaution of children at a university nursery school, who are far from representative of all 4 and 5 year olds.
+The set of tangram images may be easier to distinguish have higher codability than other target images used in adult reference games.
+We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around the larger coordination problems of taking turns and talking to their partner.
+Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children demonstrate these abilities.
 
-Our task is limited by the non-representative population of children at university nursery schools who participated, and by the limited set of target images.
-This set of tangram images may be easier to distinguish and refer to than some sets used with adults, leading to overall shorter utterances.
-We specifically targeted children's abilities to construct referring expressions that can be jointly understood.
-Children were provided scaffolding around the larger coordination problems of taking turns and talking to their partner.
-
-In the broader picture of language acquisition, there's debate over whether children learn the facts of their language first, followed later by pragmatic competency [for instance, semantics before pragmatic implicatures; @huang2009, @noveck2001], or whether sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning [@bates1974; @bohn2019a; @tomasello2008].
+In the broader picture of language acquisition, there's debate over the timing of emergence of communicative and pragmatic abilities relative to the acquisition of grammar and meaning.
+On one side, children seem to learn literal semantics far before they display an understanding of some pragmatic implicatures [@huang2009, @noveck2001]; on the other, sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning [@bates1974; @bohn2019a; @tomasello2008].
 The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.

### 2025-01-21 experiments/expt_1/client/main.js
add old expts to expts folder and rename

@@ -1,0 +1,58 @@
+import { render } from "react-dom";
+
+import "@blueprintjs/icons/lib/css/blueprint-icons.css";
+import "@blueprintjs/core/lib/css/blueprint.css";
+
+
+
+import Empirica from "meteor/empirica:core";
+
+
+import Overview from "./intro/Overview.jsx";
+
+import Round from "./game/Round.jsx";
+import Thanks from "./exit/Thanks.jsx";
+
+
+import customBreadcrumb from "./game/Breadcrumb.jsx";
+
+
+ Empirica.header(() => null);
+
+// Set the Consent Component you want to present players (optional).
+//Empirica.consent(Consent);
+
+// Introduction pages to show before they play the game (optional).
+// At this point they have been assigned a treatment. You can return
+// different instruction steps depending on the assigned treatment.
+Empirica.introSteps((game, treatment) => {
+  return [Overview];
+});
+
+// The Round component containing the game UI logic.
+// This is where you will be doing the most development.
+// See client/game/Round.jsx to learn more.
+Empirica.round(Round);
+
+// End of Game pages. These may vary depending on player or game information.
+// For example we can show the score of the user, or we can show them a
+// different message if they actually could not participate the game (timed
+// out), etc.
+// The last step will be the last page shown to user and will be shown to the
+// user if they come back to the website.
+// If you don't return anything, or do not define this function, a default
+// exit screen will be shown.
+Empirica.exitSteps((game, player) => {
+    return [Thanks];
+});
+
+// Empirica.breadcrumb would probably go here
+Empirica.breadcrumb(customBreadcrumb);
+
+// Start the app render tree.
+// NB: This must be called after any other Empirica calls (Empirica.round(),
+// Empirica.introSteps(), ...).
+// It is required and usually does not need changing.
+Meteor.startup(() => {
+  render(Empirica.routes(), document.getElementById("app"));
+});

### 2025-01-21 experiments/expt_1/server/constants.js

@@ -1,0 +1,57 @@
+import Empirica from "meteor/empirica:core";
+
+export const targetSets = {
+  'setA' : [
+    "/experiment/B1.jpg",
+    "/experiment/D1.jpg",
+    "/experiment/E1.jpg",
+    "/experiment/I1.jpg"
+  ]
+};
+
+export const practiceSet = [
+  "/experiment/apple.png",
+  "/experiment/banana.png",
+  "/experiment/dog.png",
+  "/experiment/bird.jpeg"
+]
+
+export const names = [
+    "Repi",
+    "Minu",
+    "Laju",
+    "Hera",
+    "Zuda",
+    "Bavi",
+    "Lika",
+    "Felu",
+    "Wepi",
+    "Dace",
+  ]; // for the players names to match avatar color
+
+export const avatarNames = [
+    "Samuel",
+    "Alaina",
+    "Katherine",
+    "Riley",
+    "Mason",
+    "Kaylee",
+    "Matthew",
+    "Leah",
+    "Ethan",
+    "Gabriel",
+  ]// to do more go to https://jdenticon.com/#icon-D3
+
+export const nameColors = [
+    "#8138A8",
+    "#7ACC66",
+    "#3D45B7",
+    "#997E32",
+    "#BD75D1",
+    "#A83F38",
+    "#2E808C",
+    "#59C7B1",
+    "#A8385F",
+    "#996832"
+  ] //these match the avatarNames colors -- need to code by hand with html color picker type thing
+

### 2025-01-21 experiments/expt_1/server/main.js

@@ -1,0 +1,147 @@
+import Empirica from "meteor/empirica:core";
+
+import "./callbacks.js";
+import { practiceSet, targetSets } from "./constants";
+import _ from "lodash";
+
+
+function createRoles(players, info) {
+  const l = _.shuffle(players);
+  let order=[]
+  if (info.rotate===false){
+    order=_.times(info.numBlocks, _.constant(0))
+  }else if (info.rotate===true) {
+    order=_.range(info.numBlocks).map(p => p % info.numPlayers)  }
+  const speaker=_.times(info.numTrialsPerBlock, _.constant("speaker"))
+  const listener=_.times(info.numTrialsPerBlock, _.constant("listener"))
+
+  let role_list=[]
+  _.times(info.numPlayers, player_num => {
+    let current=[]
+    _.times(info.numBlocks, block =>{
+      order[block]==player_num ?
+        current.push(...speaker):
+        current.push(...listener)
+    })
+    role_list.push(current)
+  })
+  const roles=_.zipObject(l,role_list);
+  console.log(roles);
+  return roles;
+}
+
+
+// gameInit is where the structure of a game is defined.  Just before
+// every game starts, once all the players needed are ready, this
+// function is called with the treatment and the list of players.  You
+// must then add rounds and stages to the game, depending on the
+// treatment and the players. You can also get/set initial values on
+// your game, players, rounds and stages (with get/set methods), that
+// will be able to use later in the game.
+Empirica.gameInit((game, treatment) => {
+  console.log(
+    "Game with a treatment: ",
+    treatment,
+    " will start, with workers",
+    _.map(game.players, "id")
+  );
+
+
+  game.set("targetSet", 'setA'); 
+  game.set('context', targetSets['setA']);
+  game.set("feedback", treatment.feedback)
+  game.set("chat", treatment.chat)
+  game.set("rotateInterval", treatment.rotateInterval ? treatment.rotateInterval : 12)
+  //console.log(game.get("rotateInterval"))
+  const targets = game.get('context');
+  const reps = treatment.rounds;
+  const numTargets = targets.length;
+  const info = {
+    numTrialsPerBlock : numTargets,
+    numBlocks : reps,
+    numTotalTrials: reps * numTargets,
+    numPlayers: game.players.length,
+    rotate: treatment.rotateSpeaker,
+  };
+  
+  // I use this to play the sound on the UI when the game starts
+  //game.set("justStarted", true);
+
+  // Make role list
[…]

### 2025-01-21 experiments/expt_2/client/main.js

@@ -1,0 +1,60 @@
+import { render } from "react-dom";
+
+import "@blueprintjs/icons/lib/css/blueprint-icons.css";
+import "@blueprintjs/core/lib/css/blueprint.css";
+
+
+
+import Empirica from "meteor/empirica:core";
+
+
+import Overview from "./intro/Overview.jsx";
+import Bubbles from "./intro/bubbles.jsx";
+import Smileys from "./intro/smileys.jsx";
+import Round from "./game/Round.jsx";
+import Thanks from "./exit/Thanks.jsx";
+import Pause from "./intro/Pause.jsx";
+
+import customBreadcrumb from "./game/Breadcrumb.jsx";
+
+
+ Empirica.header(() => null);
+
+// Set the Consent Component you want to present players (optional).
+//Empirica.consent(Consent);
+
+// Introduction pages to show before they play the game (optional).
+// At this point they have been assigned a treatment. You can return
+// different instruction steps depending on the assigned treatment.
+Empirica.introSteps((game, treatment) => {
+  return [Overview, Bubbles, Smileys, Pause, Overview];
+  //return[];
+});
+
+// The Round component containing the game UI logic.
+// This is where you will be doing the most development.
+// See client/game/Round.jsx to learn more.
+Empirica.round(Round);
+
+// End of Game pages. These may vary depending on player or game information.
+// For example we can show the score of the user, or we can show them a
+// different message if they actually could not participate the game (timed
+// out), etc.
+// The last step will be the last page shown to user and will be shown to the
+// user if they come back to the website.
+// If you don't return anything, or do not define this function, a default
+// exit screen will be shown.
+Empirica.exitSteps((game, player) => {
+    return [Thanks];
+});
+
+// Empirica.breadcrumb would probably go here
+Empirica.breadcrumb(customBreadcrumb);
+
+// Start the app render tree.
+// NB: This must be called after any other Empirica calls (Empirica.round(),
+// Empirica.introSteps(), ...).
+// It is required and usually does not need changing.
+Meteor.startup(() => {
+  render(Empirica.routes(), document.getElementById("app"));
+});

### 2025-01-21 experiments/expt_2/server/constants.js

@@ -1,0 +1,58 @@
+import Empirica from "meteor/empirica:core";
+
+export const targetSets = {
+  'setA' : [
+    "/experiment/B1.jpg",
+    "/experiment/D1.jpg",
+    "/experiment/E1.jpg",
+    "/experiment/I1.jpg"
+  ]
+};
+
+export const practiceSet = [
+  ["/experiment/apple.png", "/experiment/banana.png"],
+  ["/experiment/dog.png","/experiment/bird.jpeg"],
+  ["/experiment/boat.jpg", "/experiment/house.jpg"],
+  ["/experiment/tree.jpg", "/experiment/swan.jpg"],
+  //["/experiment/horse.jpg", "/experiment/cat.jpg"]
+]
+
+export const names = [
+    "Repi",
+    "Minu",
+    "Laju",
+    "Hera",
+    "Zuda",
+    "Bavi",
+    "Lika",
+    "Felu",
+    "Wepi",
+    "Dace",
+  ]; // for the players names to match avatar color
+
+export const avatarNames = [
+    "Samuel",
+    "Alaina",
+    "Katherine",
+    "Riley",
+    "Mason",
+    "Kaylee",
+    "Matthew",
+    "Leah",
+    "Ethan",
+    "Gabriel",
+  ]// to do more go to https://jdenticon.com/#icon-D3
+
+export const nameColors = [
+    "#8138A8",
+    "#7ACC66",
+    "#3D45B7",
+    "#997E32",
+    "#BD75D1",
+    "#A83F38",
+    "#2E808C",
+    "#59C7B1",
+    "#A8385F",
+    "#996832"
+  ] //these match the avatarNames colors -- need to code by hand with html color picker type thing
+

### 2025-01-21 experiments/expt_3/server/callbacks.js

@@ -1,0 +1,229 @@
+import Empirica from "meteor/empirica:core";
+import {names, avatarNames, nameColors} from './constants.js';
+import _ from "lodash";
+
+
+
+// onGameStart is triggered opnce per game before the game starts, and before
+// the first onRoundStart. It receives the game and list of all the players in
+// the game.
+Empirica.onGameStart((game) => {
+  const players = game.players;
+  console.debug("game ", game._id, " started");
+
+  const roleList = game.get('roleList');
+  game.set("currentSpeaker", null)
+  game.set("justStarted", true);
+  game.set("activePlayerCount", game.players.length)
+
+})
+
+// onRoundStart is triggered before each round starts, and before onStageStart.
+// It receives the same options as onGameStart, and the round that is starting.
+Empirica.onRoundStart((game, round) => {
+  
+  const players = game.players;
+  round.set("countCorrect",0);
+  round.set('speaker', "")
+  round.set('submitted', false);
+  round.set("activePlayerCount", game.get("activePlayerCount"));
+  const activePlayers=_.reject(game.players, p => p.get("exited"))
+
+  let queue=game.get("speakerQueue")
+
+  if (game.treatment.rotateSpeaker){
+    console.log(round.get("trialNum") % game.get("rotateInterval"));
+    //console.log(game.get("rotateInterval"));
+    console.log(round.get("trialNum") % game.get("rotateInterval") ==0 );
+    if (round.get("trialNum") % game.get("rotateInterval") ==0 ) {// time for a new speaker
+      game.set("currentSpeaker", queue.shift())
+      console.log(game.get("currentSpeaker"));
+      queue.push(game.get("currentSpeaker"));
+      //console.log(queue);
+      game.set("speakerQueue", queue) // move them to the end of the queue
+      console.log(game.get("speakerQueue"))
+    }
+  } 
+  if (!game.get("currentSpeaker")){ // it's null either from start of game or from speaker quitting
+    //console.log("here")
+    //console.log(game.get("speakerQueue"))
+
+    game.set("currentSpeaker", queue.shift())
+    queue.push(game.get("currentSpeaker"));
+    console.log(queue);
+    game.set("speakerQueue", queue) // move them to the end of the queue
+  }
+  activePlayers.forEach(player => {
+    let tangrams=round.get("tangramURLs")
+    player.set("tangramURLs", _.shuffle(tangrams))
+    if (player._id==game.get("currentSpeaker")){
[…]

### 2025-01-21 experiments/expt_3/server/main.js

@@ -1,0 +1,136 @@
+import Empirica from "meteor/empirica:core";
+
+import "./callbacks.js";
+import { practiceSet, targetSets } from "./constants";
+import _ from "lodash";
+
+
+function createRoles(players, info) {
+  const l = _.shuffle(players);
+  let order=[]
+  if (info.rotate===false){
+    order=_.times(info.numBlocks, _.constant(0))
+  }else if (info.rotate===true) {
+    order=_.range(info.numBlocks).map(p => p % info.numPlayers)  }
+  const speaker=_.times(info.numTrialsPerBlock, _.constant("speaker"))
+  const listener=_.times(info.numTrialsPerBlock, _.constant("listener"))
+
+  let role_list=[]
+  _.times(info.numPlayers, player_num => {
+    let current=[]
+    _.times(info.numBlocks, block =>{
+      order[block]==player_num ?
+        current.push(...speaker):
+        current.push(...listener)
+    })
+    role_list.push(current)
+  })
+  const roles=_.zipObject(l,role_list);
+  console.log(roles);
+  return roles;
+}
+
+
+// gameInit is where the structure of a game is defined.  Just before
+// every game starts, once all the players needed are ready, this
+// function is called with the treatment and the list of players.  You
+// must then add rounds and stages to the game, depending on the
+// treatment and the players. You can also get/set initial values on
+// your game, players, rounds and stages (with get/set methods), that
+// will be able to use later in the game.
+Empirica.gameInit((game, treatment) => {
+  console.log(
+    "Game with a treatment: ",
+    treatment,
+    " will start, with workers",
+    _.map(game.players, "id")
+  );
+
+
+  game.set("targetSet", 'setA'); 
+  game.set('context', targetSets['setA']);
+  game.set("feedback", treatment.feedback)
+  game.set("chat", treatment.chat)
+  game.set("rotateInterval", treatment.rotateInterval ? treatment.rotateInterval : 12)
+  //console.log(game.get("rotateInterval"))
+  const targets = game.get('context');
+  const reps = treatment.rounds;
+  const numTargets = targets.length;
+  const info = {
+    numTrialsPerBlock : numTargets,
+    numBlocks : reps,
+    numTotalTrials: reps * numTargets,
+    numPlayers: game.players.length,
+    rotate: treatment.rotateSpeaker,
+  };
+  
+  // I use this to play the sound on the UI when the game starts
+  //game.set("justStarted", true);
+
+  // Make role list
[…]

### 2025-01-27 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
Feedback + Edits

@@
 Children sat across a table from each other, each with a tablet in front of them (Figure \ref{game}A).
-On each trial, one child was the "teller" and saw a black box around one of two images on their screen and was asked to "say what they saw" in the black box.
+On each trial, one child was the "teller" and saw a black box around one of two images on their screen and was asked to "tell Smurfy what they see" in the black box.
 The "guesser" saw the same two images in a randomized order and tried to select the described image to help Smurfy make a match.
@@
 <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
--->  As shown in Figure \ref{acc}, children generally accurate in their selections.
+-->  As shown in Figure \ref{acc}, children were generally accurate in their selections.
 To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
@@
 
-While description length is used as a proxy for measuring convention formation, semantic overlap between utterances does not always track description length.
-@boyce2024 introduced a more sensitive measure of semantic convergence.
-By comparing the content of utterances using word embeddings, @boyce2024 traced how similarities within and across games changed over time.
+While description length is often used as a proxy for measuring convention formation, it does not capture semantic overlap between utterances.
+@boyce2024 introduced a more sensitive measure of semantic convergence that compares the content of utterances using word embeddings to trace how similarities within and across games change over time.
 <!--To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game.
--->
-
-With only 3 blocks of descriptions, we do not believe it is meaningful to look at whether the 2nd and 3rd blocks are more similar than the 1st and 3rd blocks.
+--> With only 3 blocks of descriptions, we do not expect semantic similarity for descriptions of a given target to show any meaningful change over time.
 However, we can test for a more coarse measure of sensitivity to partner: whether children's utterances are more like their partner's than children in other games'.
@@
 Utterances were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves.
-However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating some sensitivity to their partner's expressions.
+However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.
 
@@
 As a course measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1.
-Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
+Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
 
@@
 To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar with the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
-This pattern of results is consistent with children sticking to their own description more when it was previously successful last time, and children repeat the other child's description more when they understood it.
+This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful.
 <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`.
@@
 
-Prominent classical findings have been interpreted as evidence that young children are cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+Prominent classical findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
 However, an emerging pattern of research finds that  young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
@@
 Additionally, children's utterances showed signs of converging toward conceptual pacts.
-Children did not display a reduction in the length of referential expressions over the course of the game.
 While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+
+Unlike adults, Children did not display an increase in accuracy nor a reduction in the length of referential expressions over the course of the game.
+This might be taken to mean that children are not adapting their descriptions over time.
+Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that childen are adapting their descriptions as the game unfolds.
+These null findings are likely a result of initially high accuracy in the first block and initially short utterances that have less possibility for reduction when compared to adult utterances.
 
@@
 In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions.
-Indeed, adults controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024].
+Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024].
 However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification.
@@
 The generalizability of our results is limited by the target population, the target images, and the task structure.
-We sampled a convenience poplaution of children at a university nursery school, who are far from representative of all 4 and 5 year olds.
-The set of tangram images may be easier to distinguish have higher codability than other target images used in adult reference games.
+We sampled a convenience poplaution of children at a university nursery school, who are far from representative of all 4 and 5-year olds.
+The set of tangram images may be easier to distinguish and have higher codability than other target images used in adult reference games.
 We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around the larger coordination problems of taking turns and talking to their partner.
-Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children demonstrate these abilities.
+Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.
 
@@
 The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.
-At 4–5 years old, even though their utterances are not fully adult-like in form, children are nonetheless able to use their linguistic skills communicatively.
+At 4–5-years old, even though their utterances are not fully adult-like in form, children are nonetheless able to use their linguistic skills communicatively.
 

### 2025-01-27 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
add osf link

@@
 We present the estimate and 95% credible intervals.
-The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at VERONICA SHOULD MAKE AN OSF LINK.
+The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at \url{https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2}.
 

### 2025-01-29 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
final pass of edits from Yannick and mike, and length futzing

@@
 ---
-title: "Preschoolers form conventional pacts with each other to communicate about novel shapes"
+title: >
+  Preschoolers can form conventional pacts with each other 
+  to communicate about novel referents
 bibliography: preschool-tangrams.bib
@@
 abstract: > 
-   Learning language requires learning not only the content of language, but also skills that allow us to use language to communicate.
-Iterated reference games provide a window into such skills, requiring rich communication as participants jointly converge on mutually understandable names for initially novel objects.
-Some classical experiments with young children are interpreted as showing that 4-5-year old children are incapable of succeeding at iterated reference games.
-Here, we revisit the question of young children's referential communicative abilities using a simpler, child-friendly paradigm.
-Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in establishing reference.
+  Learning language requires learning not only the content of language, but also how to use language to communicate.
+Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents.
+Some early experiments are interpreted as evidence that 4-5-year old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game.
+Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm.
+Across 51 pairs of children, we found that 4-5 year olds successfully established reference with each other.
 Children were 85% accurate, and they often used descriptions similar to their partner's.
@@
 Learning a language requires learning not only the content of that language, but also how to use the language to communicate.
-One test environment for language use is referential communication, the ability to describe a target so an interlocutor can pick it out from a set of possibilities.
+One case study for language use is referential communication, the ability to describe a target so an interlocutor can pick it out from a set of possibilities.
 Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the description they provide to their beliefs about the interlocutor's knowledge state.
 
-One test of referential communication skill is *iterated reference games* where one describes abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
-Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
-Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds.
+Iterated reference games provide an important paradigm for studying referential communication.
+In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
+Over repetition, features of the initial descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
+Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform later referring expressions.
 Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.
 
-A classic study suggests that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966].
+One influential early study suggested that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966].
 In their paradigm, one child was given a set of 6 blocks in a specific order.
@@
 As children described and selected blocks, they stacked them on pegs.
-While 4-5-year old children were successful on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had drawings of abstract shapes and there was no visual access.
+While 4-5-year old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access.
 Even after multiple rounds with the same images, children were not able to correctly order the blocks.
+@glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.
+
 Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions.
-Still, even the 9th grade sample was noticeably worse than the sample of adult college students [@glucksberg1967; @krauss1969].
-That the task was difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets present significant task demands that might prevent accurate measurement of children's abilities.
+Still, even the 9th grade sample was noticeably worse than the adult college student sample [@glucksberg1967; @krauss1969].
+Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands preventing accurate measurement of children's abilities.
 
 ```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure.
-Panel A shows the experimental setup with the teller and guesser across the table from each other.
+Panel A shows the experimental setup.
 Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
@@
 
-More recently, a number of studies have revealed early emerging skills in the preschool years that support the use of referential communication.
+More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
 Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
-Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009], showing that children at this age can reason about others’ perspectives to more efficiently communication.
+Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009].
+By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015].
+Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.
 
-While children are sensitive to others’ perspectives, they still seem to struggle with appropriately tailoring the specificity of their utterances to the visual context, often resulting in utterances that are under-informative [@matthews2012; @leung2024].
-When children must describe one of two very similar images, 4 and 5-year olds struggle to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
-By 5-years old, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].
+While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting under-informative utterances [@matthews2012; @leung2024].
+When children must describe one of two very similar images, 4 and 5-year olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
+By 5-years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].
 
-A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e.
-when there are fewer possible referents to consider [@abbot-smith2016].
-This paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
+A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e.
+when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].
+These findings paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
 
 Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game.
-Recent work shows that 8-10-year olds exhibit adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016].
-4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an iterated reference game where children could only use gestures to communicate [@bohn2019].
-Recent work has established that young children can succeed at iterated reference games when playing with their parents [@leung2024].
-4-8 year olds' played an iterated reference game with a parent using a simple, child-friendly tablet-based task.
+In one study, 8-10-year olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016].
+4-6-year olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019].
+4-8 year old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024].
 In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
-Together, these findings provide evidence of young children’s ability to successfully communicate about novel referents under the right conditions.
+Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.
 
-Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, we revisit the question of child-child referential communication.
+Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication.
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
+Across an initial study and a recplication including a total of 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials.
-We had 19 complete games and 1 included incomplete game.
+We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials.
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
@@
 For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B).
-We coded the matching game using Empirica and hosted it on a lab server [@almaatouq2020].
-We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate away from the game.
+We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on ablets that were locked in a kiosk mode so children could not navigate away from the game.
 
@@
 The "guesser" saw the same two images in a randomized order and tried to select the described image to help Smurfy make a match.
-When the guesser selected an image, both children received feedback in form of a smiley or frowny face and an audible sound.
+When the guesser selected an image, both children received feedback in form of a smiley or frowny face and an excited or disappointed sound.
 After each trial, children switched roles.
-Children passed Smurfy back and forth to help them keep track of whether they were the "guesser" or "teller" on a given trial.
+Children passed Smurfy back and forth to keep track of whether they were the "guesser" or "teller" on a given trial.
 
@@
 Children's selections and the time to selection were recorded from the experiment software.
-Children's descriptions were transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters.
+Children's descriptions were automatically transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters.
 Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used.
 We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe.
-After exclusions, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials remaining.
+After exclusions, we had `r clean_1 |> filter(!is.na(repNum)) |> nrow()` trials remaining.
 
@@
 ## Results
-
 ### Accuracy and speed
@@
 <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
--->  As shown in Figure \ref{acc}, children were generally accurate in their selections.
-To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).
+--> To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}).
 This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.
@@
 
-As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time.
+As another measure of children's performance, we looked at how long children spent on each trial.
 We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds.^[`r speed_form |> form()`] The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial).
-This is a measure of efficiency as children were able to achieve the same accuracy levels in less time.
+Children were able to achieve the same accuracy in less time, suggesting that they were becoming more efficient at completing the task.
 
@@
 We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`).
-Utterances were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves.
+Utterances were slightly more similar still if they came from the same person (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves.
 However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.
@@
 
-In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter.
+In Experiment 1, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set-up and making it shorter.
 Our goal was to see if young children were at all able to provide adequate descriptions, so children received a lot of scaffolding around the experimental interaction.
-Sometimes, this included experimenters echoing children's descriptions which could potentially influence children's responses.
+Sometimes, this scaffolding included experimenters echoing children's descriptions, which could potentially influence children's responses.
 In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.
@@
 ### Participants
-Experiment 2 was run between March and August of 2024, at the same university preschool as Experiment 1.
+Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1.
 No children participated in both experiments.
-30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials.
+30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials.
 Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included.
@@
 The procedure was much the same as Experiment 1(Figure \ref{game}D).
-We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in Experiment 1).
+We added an initial "bubble popping" exercise to give children practice tapping the tablet appropriately (this was an issue for some children in Experiment 1).
 The experimental script was fully written out and memorized by experimenters so children all received the same instructions.
 We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections.
-Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen to move the game along, but avoided contributing or repeating any content about the images or the descriptions.
+Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen, but avoided contributing or repeating any content about the images or the descriptions.
 
@@
 
-In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy was relatively stable over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
+In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and relatively stable over time (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}).
 The first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial).
@@
 The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial, Figure \ref{length}).
-This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
+This finding is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.
 
@@
 
-As a course measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1.
-Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).
+As a coarse measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1.
+Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar if they came from the same person (`r sim_across_between_2_results |> stats_text(3,3)`).
 
@@
 Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
-Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.
+Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).
 
-As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image.
+Partnerships often diverged from one another as groups focused on distinct aspects of the image.
 We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions.
-These patterns of increasing similarity within games and increasing divergence between games match the patterns found for adults, albeit weakly [@boyce2024].
+These patterns of increasing similarity within games and increasing divergence between games qualitatively match the patterns found for adults [@boyce2024].
 
@@
   mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
-  ggplot(aes(as.character(earlier), sim, color = target1)) +
+  ggplot(aes(earlier, sim, color = target1)) +
+    geom_smooth(color="black", method="lm")+
   geom_point(alpha = .1, position = position_dodge(width = .2)) +
   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
+  #  geom_smooth( method="lm")+
   facet_wrap(~same_speaker) +
   scale_color_solarized() +
+  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
   labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
@@
             strip.background = element_blank(),
-        panel.grid=element_blank()
+        #panel.grid=element_blank()
 )
+
 ```
@@
 As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect.
-Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly, but unreliably, increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+Pooling the two experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
 Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
@@
 Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt.
-To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar with the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
+To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
 This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful.
@@
 
-Prominent classical findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
-However, an emerging pattern of research finds that  young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
+Prominent early findings been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+However, more recent developmental have found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
 <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
--->We re-addressed the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
+-->We revisited the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
 
@@
 
-Unlike adults, Children did not display an increase in accuracy nor a reduction in the length of referential expressions over the course of the game.
-This might be taken to mean that children are not adapting their descriptions over time.
-Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that childen are adapting their descriptions as the game unfolds.
-These null findings are likely a result of initially high accuracy in the first block and initially short utterances that have less possibility for reduction when compared to adult utterances.
+Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game.
+Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds.
+These null findings are likely a result of initially high accuracy in the first block and initially short utterances that leave little room for reduction.
 
-It is unclear to what extent the uniformly short descriptions are a product of the simplified task or children's behavioral differences from adults.
+It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults.
 In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions.
@@
 The generalizability of our results is limited by the target population, the target images, and the task structure.
-We sampled a convenience poplaution of children at a university nursery school, who are far from representative of all 4 and 5-year olds.
+We sampled a convenience population of children at a university nursery school.
 The set of tangram images may be easier to distinguish and have higher codability than other target images used in adult reference games.
-We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around the larger coordination problems of taking turns and talking to their partner.
+We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner.
 Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.
 
-In the broader picture of language acquisition, there's debate over the timing of emergence of communicative and pragmatic abilities relative to the acquisition of grammar and meaning.
-On one side, children seem to learn literal semantics far before they display an understanding of some pragmatic implicatures [@huang2009, @noveck2001]; on the other, sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning [@bates1974; @bohn2019a; @tomasello2008].
-The findings in the current work are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.
-At 4–5-years old, even though their utterances are not fully adult-like in form, children are nonetheless able to use their linguistic skills communicatively.
+In the broader picture of language acquisition, there is debate over the timing of the emergence of communicative and pragmatic abilities relative to the acquisition of grammar and meaning.
+On one side, children seem to learn literal semantics far before they display an understanding of some pragmatic implicatures [@huang2009; @noveck2001]; on the other, sensitivity to communicative intent is an early emerging skill that develops in parallel with linguistic knowledge and may bootstrap language learning [@bates1974; @bohn2019a; @tomasello2008].
+Our current findings are most consistent with a gradual development of children's communicative and linguistic skills, where the skills emerge early and then are refined over time, as children's cognitive capacities increase.
+At 4-5 years old, children are already able to establish novel referential conventions with one another as part of their broader ability to communicate and coordinate.
 

### 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
cleaning up

@@
 Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents.
-Some early experiments are interpreted as evidence that 4-5-year old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game.
+Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game.
 Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm.
-Across 51 pairs of children, we found that 4-5 year olds successfully established reference with each other.
+Across 51 pairs of children, we found that 4-5-year-olds successfully established reference with each other.
 Children were 85% accurate, and they often used descriptions similar to their partner's.
@@
 
-One influential early study suggested that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966].
+One influential early study suggested that 4-5-year-old preschoolers struggle with child-child referential communication [@glucksberg1966].
 In their paradigm, one child was given a set of 6 blocks in a specific order.
@@
 As children described and selected blocks, they stacked them on pegs.
-While 4-5-year old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access.
+While 4-5-year-old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access.
 Even after multiple rounds with the same images, children were not able to correctly order the blocks.
@@
 More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
-Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
+Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
 Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009].
@@
 
-While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting under-informative utterances [@matthews2012; @leung2024].
-When children must describe one of two very similar images, 4 and 5-year olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
-By 5-years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].
+While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting in under-informative utterances [@matthews2012; @leung2024].
+When children must describe one of two very similar images, 4 and 5-year-olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019].
+By 5 years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].
 
@@
 when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].
-These findings paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
+These findings paired with evidence of emerging communicative skills in young children suggest that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
 
 Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game.
-In one study, 8-10-year olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016].
-4-6-year olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019].
-4-8 year old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024].
-In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
+In one study, 8-10-year-olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance and highly variable between dyads [@branigan2016].
+4-6-year-olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019].
+4-8-year-old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024].
+In this task, even 4-year-olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
 Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.
@@
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across an initial study and a recplication including a total of 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
+Across an initial study and a replication including a total of 51 pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 
-Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand.
+Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions of abstract shapes that their partner could understand.
 Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
@@
 
-4 and 5-year old children were recruited from a university preschool during the school day.
+4 and 5-year-old children were recruited from a university preschool during the school day.
 Children played with another child from the same class.
@@
 For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B).
-We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on ablets that were locked in a kiosk mode so children could not navigate away from the game.
+We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.
 
@@
 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials.
-Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included.
+Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included.
 Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
@@
 
-Prominent early findings been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
-However, more recent developmental have found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
+Prominent early findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
 <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
--->We revisited the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
+-->We revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
 
-Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
+Across 2 experiments and 51 pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
 Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions.

### 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
anonymize pre-regs & rmd

@@
 author-information: > 
-    \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, \\Stanford University \And {\large \bf Robert Z.
-Sparks (bsparks@stanford.edu) }\\ Department of Psychology, \\Stanford University
-    \AND {\large \bf Yannick Mofor (yannickm@stanford.edu)} \\ Department of Computer Science, \\ Stanford University \And {\large \bf Michael C.
-Frank (mcfrank@stanford.edu)} \\ Department of Psychology, \\ Stanford University}
+    \author{Anonymized}
 
@@
 Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
-This experiment was pre-registered at https://osf.io/kcv8j.
+This experiment was pre-registered at https://osf.io/kcv8j?view_only=fd55733671194962b868a4ba19377631.
 
@@
 
-As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at https://osf.io/y2dax.
+As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1.
+Experiment 2 was pre-registered at https://osf.io/y2dax/?view_only=fd55733671194962b868a4ba19377631.
 

### 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
rerun models because of wrong priors, proofreading

@@
 ---
-title: >
-  Preschoolers can form conventional pacts with each other 
-  to communicate about novel referents
+title: |
+  | Preschoolers can form conventional pacts with each other 
+  | to communicate about novel referents
 bibliography: preschool-tangrams.bib
@@
 In their paradigm, one child was given a set of 6 blocks in a specific order.
-Their task was to describe the images on the block so their partner could pick out the corresponding block.
+Their task was to describe the image on each block so their partner could pick out their corresponding block.
 As children described and selected blocks, they stacked them on pegs.
@@
 Still, even the 9th grade sample was noticeably worse than the adult college student sample [@glucksberg1967; @krauss1969].
-Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands preventing accurate measurement of children's abilities.
+Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands that prevented accurate measurement of children's abilities.
 
@@
 
-More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
+More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
 Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
@@
 
-Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions of abstract shapes that their partner could understand.
+Our goal for Experiment 1 was to test young children's ability to coordinate to produce descriptions of abstract shapes that their partner could understand.
 Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials.
-This experiment was pre-registered at https://osf.io/kcv8j?view_only=fd55733671194962b868a4ba19377631.
+This experiment was pre-registered at [anonymized link](https://osf.io/kcv8j?view_only=fd55733671194962b868a4ba19377631).
 
@@
 Targets were randomly paired with another of the critical images as the foil.
+
 The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements (especially when utterances were inaudible or the child did not respond immediately; this aspect of the procedure was modified in Experiment 2).
@@
 Statistical analyses were run in brms [@burkner2018] with weakly informative priors.
-We present the estimate and 95% credible intervals.
-The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at \url{https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2}.
+We report estimates and 95% credible intervals.
+The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).
 
@@
 \begin{table}
-	\caption{Example descriptions children successfully used used to identify different target images in Experiment 2.}
+	\caption{Example descriptions children successfully used to identify different target images in Experiment 2.}
 	\label{example}
@@
 As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1.
-Experiment 2 was pre-registered at https://osf.io/y2dax/?view_only=fd55733671194962b868a4ba19377631.
+Experiment 2 was pre-registered at [anonymized link](https://osf.io/y2dax/?view_only=fd55733671194962b868a4ba19377631).
 
@@
 We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
+Descriptions were more similar if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
 Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).
@@
 Partnerships often diverged from one another as groups focused on distinct aspects of the image.
-We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions.
+We tested whether descriptions in different games diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions to the same target from different games became slightly less similar (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions.
 These patterns of increasing similarity within games and increasing divergence between games qualitatively match the patterns found for adults [@boyce2024].
 
-```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block.
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiment 2.
 Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
@@
 Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt.
-To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
+To test whether accuracy is predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
 This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful.
@@
 
-Prominent early findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
 However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
 <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
--->We revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
+-->Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
 
@@
 It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults.
-In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions.
+In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions.
 Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024].

### 2025-01-31 code/mods_for_cogsci_paper.R

@@ -77,7 +77,7 @@ description_mod_data_1 <- clean_1 |>
 description_mod_1 <- brm(words ~ trial.num + (trial.num | game) + (1 | target),
                          data = description_mod_data_1,
                          file = here(mod_loc, "description_1.rds"),
-                         prior = speed_priors,
+                         prior = description_priors,
                          control = list(adapt_delta = .95)
 )
 
@@ -169,7 +169,7 @@ description_mod_data_2 <- clean_2 |>
 description_mod_2 <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target),
                          data = description_mod_data_2,
                          file = here(mod_loc, "description_2.rds"),
-                         prior = speed_priors,
+                         prior = description_priors,
                          control = list(adapt_delta = .95)
 )
 
@@ -297,7 +297,7 @@ description_mod_data_meta <- clean_1 |>
 description_mod_meta <- brm(words ~ trial.num + (trial.num | gameId) + (1 | target) + (1|expt),
                             data = description_mod_data_meta,
                             file = here(mod_loc, "description_meta.rds"),
-                            prior = speed_priors,
+                            prior = description_priors,
                             control = list(adapt_delta = .95)
 )
 

### 2025-08-07 code/demog.qmd
add demog (age in mo)

---
title: "demog"
format: html
---

```{r}
library(tidyverse)
library(here)
raw_demog <- read_csv(here("demographics.csv"))
```

```{r}
expt_1 <- read_csv(here("data/expt_1/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 1) |>
  mutate(sid = id)

id_sub <- tribble(
  ~id, ~sid,
  "apple123", "id74",
  "pear124", "id75",
  "apple127", "id76",
  "pear128", "id77",
  "80", "id78",
  "81", "id79",
  "apple111", "id90",
  "pear222", "id91",
  "apple333", "id92",
  "pear444", "id93",
  "apple255", "id94",
  "pear257", "id95"
)

expt_2 <- read_csv(here("data/expt_2/clean_data.csv")) |>
  select(game = gameId, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 2) |>
  left_join(id_sub) |>
  mutate(sid = case_when(
    !is.na(sid) ~ sid,
    str_detect(id, "id") ~ id,
    T ~ str_c("id", id)
  ))

kids <- expt_1 |> bind_rows(expt_2) # included kids!
```

```{r}
demogs <- raw_demog |>
  mutate(
    date_test = mdy(date_test),
    date_birth = mdy(date_birth)
  ) |>
  select(sid, included, date_test, date_birth, hispanic, ethnicity, gender) |>
  full_join(kids) |>
  filter(!is.na(included)) |>
  mutate(
    age_days = (date_test - date_birth) |> as.numeric(),
    age_month = floor(age_days / 30.43)
  ) |>
  select(-age_days) |>
  select(expt, id, game, sid, gender, age_month, ethnicity, hispanic) |>
  write_csv(here("data/clean_demog_1_2.csv"))
```

### 2025-08-12 code/expt_3/data_prep_3.Rmd
add preliminary expt 3 data

---
title: "Wrangle transcripts"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_source <- "data/expt_3/rounds.csv"

player_source <- "data/expt_3/players.csv"

real_ids <- c(str_c("", c(
  159:162, 167:170,
  173:174, 181:184,
  187:188, 191:196,
  199:200, 203:204,
  207:208, 227:234,
  237:240, 243:252
)))
# kids ids to include
```

# Empirica output prep

```{r}
players <- read_csv(here(player_source)) |>
  filter(id %in% real_ids) %>%
  select(id, playerId = `_id`)

# expect 50 kids (25 games)
```

```{r}
raw_data <- read.csv(here(data_source), header = TRUE) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  # select(-chat) |>
  pivot_longer(
    cols = starts_with("player"), names_to = c("playerId", "info"), names_prefix = "player_", names_sep = "_",
    values_to = "values", values_transform = as.character, values_drop_na = T
  ) %>%
  filter(values != "") %>%
  filter(playerId != speaker) %>%
  pivot_wider(names_from = info, values_from = values) |>
  left_join(players, by = c("playerId")) |>
  select(-playerId) |>
  rename(playerId = id) |>
  filter(!is.na(playerId)) |>
  left_join(players, by = c("speaker" = "playerId")) |>
  select(-speaker) |>
  rename(speaker = id)

raw_data |>
  filter(!is.na(repNum)) |>
  filter(submitted == "true") |>
  group_by(gameId) |>
  tally()

label_image <- function(string) {
  new_string <- case_when(
    string == "/experiment/tangram_E.png" ~ "bunny",
    string == "/experiment/tangram_G.png" ~ "dancer",
    string == "/experiment/tangram_H.png" ~ "baby",
    string == "/experiment/tangram_A.png" ~ "zombie",
    string == "/experiment/banana.png" ~ "banana",
    string == "/experiment/bird.jpeg" ~ "bird",
    string == "/experiment/dog.png" ~ "dog",
    string == "/experiment/apple.png" ~ "apple",
    string == "/experiment/star.png" ~ "star",
    string == "/experiment/crown.png" ~ "crown",
    string == "/experiment/lemon.png" ~ "lemon",
    string == "/experiment/pinwheel.png" ~ "pinwheel"
  )
  return(new_string)
}

less_raw_data <- raw_data %>%
  select(target, distractor, trialNum, targetNum, repNum, response, correct, time, listener = playerId, speaker, gameId) %>%
  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
  write_csv(here("data/expt_2/clean_data.csv"))
```

below here copied from expt 2 and not updated!

# Transcript wrangle - new

```{r}
library(readxl)

wrangle_transcript <- function(game_num) {
  read_xlsx(here("data/expt_2/raw_transcripts", str_c(game_num, ".xlsx"))) |>
    select(start, end, text, speaker, trial, role, echo, description) |>
    fill(trial) %>%
    mutate(start = start / 1000, end = end / 1000) |>
    mutate(trial = as.character(trial)) |>
    nest(data = c(start, end, text, speaker, trial, role, echo, description))
}

transcripts <- read_csv(here("data/expt_2/link_transcripts.csv")) |>
  select(gameId, gameConfig) |>
  mutate(transcript = map_df(gameConfig, wrangle_transcript)) |>
  unnest(transcript) |>
  unnest(data)

write_csv(transcripts, here("data/expt_2/transcripts.csv"))
```

# Checks

how much echoing is there?

```{r}
transcripts |> filter(!is.na(echo))

# not much, yay!
```

# Sbert data prep

```{r}
transcript <- read_csv(here("data/expt_2/transcripts.csv"))

dat <- read_csv(here("data/expt_2/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)
transcript %>%
  filter(!is.na(description)) %>%
  group_by(trial, gameId) %>%
  mutate(trial = as.numeric(trial)) |>
  filter(trial > 3) |>
  summarize(words = str_c(description, sep = " ", collapse = ", ")) %>%
  left_join(dat) %>%
  write_csv(here("data/expt_2/pre_sbert.csv"))
```

```{r}
library(reticulate)
np <- import("numpy")
mat <- np$load(here("data/expt_2/post_sbert.npy"))
saveRDS(mat, here("data/expt_2/post_sbert.RData"))
```

```{r}
sbert <- read_csv(here("data/expt_2/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/expt_2/post_sbert.RData")) %>%
  as_tibble())

### helper funcs
get_sim_matrix <- function(df, F_mat, method = "cosine") {
  feats <- F_mat[df$feature_ind, ]
  if (method == "cor") {
    return(cor(t(feats), method = "pearson"))
  } else if (method == "euclidean") {
    return(as.matrix(dist(feats, method = "euclidean")))
  } else if (method == "cosine") {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0("unknown method", method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(
      dim1 = as.character(dim1),
      dim2 = as.character(dim2)
    )
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(
      get_sim_matrix(., F_mat, method = method),
      as.character(.$combinedId)
    ))
}

### funcs
F_mat <- sbert %>%
  select(starts_with("V")) %>%
  as.matrix() # Features
M_mat <- sbert %>%
  select(-starts_with("V")) %>%
  mutate(feature_ind = row_number())

similarities <- M_mat %>%
  mutate(combinedId = str_c(gameId, trial, sep = "_")) %>%
  make_across_df(F_mat, "cosine") %>%
  separate(dim1, into = c("game1", "trial1"), convert = T, sep = "_") %>%
  separate(dim2, into = c("game2", "trial2"), convert = T, sep = "_") %>%
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  write_rds(here("data/expt_2/similarities.rds"))
```

### 2025-10-18 code/expt_3/data_prep_3.Rmd
start transcripts for expt 3

@@
   mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
-  write_csv(here("data/expt_2/clean_data.csv"))
+  write_csv(here("data/expt_3/clean_data.csv"))
 ```

### 2025-10-23 code/demog.qmd
data prep and more annotation

@@
 ```
+
+# expt 3
+
+```{r}
+
+expt_3_demog <- read_csv(here("expt_3_demog.csv"))
+
+expt_3 <- read_csv(here("data/expt_3/clean_data.csv")) |>
+  select(game = gameId, speaker, listener) |>
+  pivot_longer(c("speaker", "listener"), values_to = "sid") |>
+  select(-name) |>
+  unique() |> 
+  mutate(expt=3)
+
+expt_3_demog <- read_csv(here("expt_3_demog.csv")) |> 
+ mutate(
+    date_test = mdy(date_test),
+    date_birth = ymd(date_birth)
+  ) |>
+  select(sid=SID, date_test, date_birth, hispanic, ethnicity, gender) |> 
+  full_join(expt_3) |> 
+  filter(!is.na(game)) |> 
+  mutate(
+    age_days = (date_test - date_birth) |> as.numeric(),
+    age_month = floor(age_days / 30.43)
+  ) |>
+  select(-age_days) |>
+  mutate(id=sid) |> 
+  select(expt, game, id, sid, gender, age_month, ethnicity, hispanic) |> write_csv(here("data/clean_demog_3.csv"))
+
+```

### 2025-10-23 code/expt_3/data_prep_3.Rmd
data prep and more annotation

@@
 
-raw_data |>
-  filter(!is.na(repNum)) |>
-  filter(submitted == "true") |>
-  group_by(gameId) |>
-  tally()
-
 label_image <- function(string) {
@@
 wrangle_transcript <- function(game_num) {
-  read_xlsx(here("data/expt_2/raw_transcripts", str_c(game_num, ".xlsx"))) |>
+  read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv"))) |>
     select(start, end, text, speaker, trial, role, echo, description) |>
@@
     mutate(start = start / 1000, end = end / 1000) |>
-    mutate(trial = as.character(trial)) |>
+    mutate(trial = trial-1) |>
     nest(data = c(start, end, text, speaker, trial, role, echo, description))
@@
 
-transcripts <- read_csv(here("data/expt_2/link_transcripts.csv")) |>
+transcripts <- read_tsv(here("data/expt_3/link_transcripts.csv")) |> 
   select(gameId, gameConfig) |>
@@
 
-write_csv(transcripts, here("data/expt_2/transcripts.csv"))
+write_csv(transcripts, here("data/expt_3/transcripts.csv"))
 ```
@@
 ```{r}
-transcript <- read_csv(here("data/expt_2/transcripts.csv"))
+transcript <- read_csv(here("data/expt_3/transcripts.csv"))
 
-dat <- read_csv(here("data/expt_2/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)
+dat <- read_csv(here("data/expt_3/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)
 transcript %>%
@@
   left_join(dat) %>%
-  write_csv(here("data/expt_2/pre_sbert.csv"))
+  write_csv(here("data/expt_3/pre_sbert.csv"))
 ```
@@
 np <- import("numpy")
-mat <- np$load(here("data/expt_2/post_sbert.npy"))
-saveRDS(mat, here("data/expt_2/post_sbert.RData"))
+mat <- np$load(here("data/expt_3/post_sbert.npy"))
+saveRDS(mat, here("data/expt_3/post_sbert.RData"))
 ```
@@
 ```{r}
-sbert <- read_csv(here("data/expt_2/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/expt_2/post_sbert.RData")) %>%
+sbert <- read_csv(here("data/expt_3/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/expt_3/post_sbert.RData")) %>%
   as_tibble())
@@
   mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
-  write_rds(here("data/expt_2/similarities.rds"))
+  write_rds(here("data/expt_3/similarities.rds"))
 ```

### 2025-10-28 code/demog.qmd
clean up of transcripts for referential, matching ids

--- 2025-08-07 code/demog.qmd
@@
   unique() |>
-  mutate(expt = 1) |>
-  mutate(sid = id)
-
-id_sub <- tribble(
-  ~id, ~sid,
-  "apple123", "id74",
-  "pear124", "id75",
-  "apple127", "id76",
-  "pear128", "id77",
-  "80", "id78",
-  "81", "id79",
-  "apple111", "id90",
-  "pear222", "id91",
-  "apple333", "id92",
-  "pear444", "id93",
-  "apple255", "id94",
-  "pear257", "id95"
-)
+  mutate(expt = 1)
 
 expt_2 <- read_csv(here("data/expt_2/clean_data.csv")) |>
-  select(game = gameId, speaker, listener) |>
+  select(game, speaker, listener) |>
   pivot_longer(c("speaker", "listener"), values_to = "id") |>
@@
   unique() |>
-  mutate(expt = 2) |>
-  left_join(id_sub) |>
-  mutate(sid = case_when(
-    !is.na(sid) ~ sid,
-    str_detect(id, "id") ~ id,
-    T ~ str_c("id", id)
-  ))
+  mutate(expt = 2)
 
@@
   ) |>
-  select(sid, included, date_test, date_birth, hispanic, ethnicity, gender) |>
+  select(id=sid, included, date_test, date_birth, hispanic, ethnicity, gender) |> 
   full_join(kids) |>
@@
   select(-age_days) |>
-  select(expt, id, game, sid, gender, age_month, ethnicity, hispanic) |>
-  write_csv(here("data/clean_demog_1_2.csv"))
+  mutate(gender=str_to_lower(gender)) |> 
+  select(expt, id, game, gender, age_month, ethnicity, hispanic)
+
+demogs |> filter(expt==1) |> write_csv(here("data/expt_1/clean_demog.csv"))
+
+demogs |> filter(expt==2) |> write_csv(here("data/expt_2/clean_demog.csv"))
+
 ```
+
+# expt 3
+
+```{r}
+
+expt_3_demog <- read_csv(here("expt_3_demog.csv"))
+
+expt_3 <- read_csv(here("data/expt_3/clean_data.csv")) |>
+  select(game, speaker, listener) |>
+  pivot_longer(c("speaker", "listener"), values_to = "id") |>
+  select(-name) |>
+  unique() |> 
+  mutate(expt=3)
+
+expt_3_demog <- read_csv(here("expt_3_demog.csv")) |> 
+ mutate(
+    date_test = mdy(date_test),
+    date_birth = ymd(date_birth),
+    id=str_c("id", SID)
+  ) |>
+  select(id, date_test, date_birth, hispanic, ethnicity, gender) |> 
+  full_join(expt_3) |> 
+  filter(!is.na(game)) |> 
+  mutate(
+    age_days = (date_test - date_birth) |> as.numeric(),
+    age_month = floor(age_days / 30.43)
+  ) |>
+  select(-age_days) |>
+  mutate(gender=str_to_lower(gender)) |> 
+  select(expt, game, id, gender, age_month, ethnicity, hispanic) |> write_csv(here("data/expt_3/clean_demog.csv"))
+
+```

### 2025-10-28 code/expt_1/analysis_1.Rmd
clean up of transcripts for referential, matching ids

@@
 theme_set(theme_classic())
+
 transcript <- read_csv(here("data/timed_transcript.csv"))
@@
 ```
+.
 
-Trials were excluded if the speaker did not provide any description before the listener clicked.
+# Exclusions
+
+```{r}
+transcript <- read_csv(here(data_loc, "transcripts.csv"))
+data <- read_csv(here(data_loc, "clean_data.csv"))
+link <- read_tsv(here(data_loc, "link_transcripts.csv"))
+
+data <- read_csv(here(data_loc, "clean_data.csv")) |>
+  rename(trial = trialNum) |>
+  inner_join(link |> filter(is.na(exclude)))
+
+transcript <- transcript |>
+    inner_join(link |> filter(is.na(exclude))) |> 
+  filter(!is.na(description))
+
+words <- transcript |> mutate(words=description |> str_count("\\S+")) |>
+  filter(role=="S") |>
+  group_by(gameId, game, trial) |>
+  summarize(words=sum(words))
+
+all_data <- data |>left_join(words) |> 
+  filter(!is.na(words)) |> 
+   mutate(type = case_when(
+    trial < 4 ~ "practice",
+    trial < 8 ~ "block 1",
+    trial < 12 ~ "block 2",
+    trial < 16 ~ "block 3",
+    trial < 20 ~ "block 4"
+  )) |>
+  filter(!is.na(response)) |> 
+  mutate(correct=as.numeric(correct))
+
+all_data |> select(game, speaker) |> distinct()
+
+```
 

### 2025-10-28 code/expt_1/data_prep.Rmd
clean up of transcripts for referential, matching ids

@@
 
-pilot_start=lubridate::ymd('2023-07-07') 
-pilot_end=lubridate::ymd('2023-07-12')
-
-real_start=lubridate::ymd('2023-07-17')
-real_end=lubridate::ymd('2023-08-09')
-
 data_source <- "data/expt_1/rounds.csv"
@@
 
-pilot_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test
-
-real_ids <-str_c("id", c(13:16, 19:36, 39:54, 57:58)) # skipping kids who didn't complete enough trials
 ```
@@
 
-players <- read_csv(here(player_source)) |> filter(id %in% real_ids) %>% select(id, playerId=`_id`)
-
-games <- read_csv(here("data/expt_1/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)
+players <- read_csv(here(player_source)) |> select(id, playerId=`_id`)
 
@@
 raw_data <- read.csv(here(data_source), header = TRUE) %>% 
-  filter(createdAt >= real_start) %>%
-  filter(createdAt <= real_end) %>% 
   rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
@@
   select(-speaker) |> 
-  rename(speaker=id) %>% 
-  left_join(games) |> 
-  rename(game=gameConfig) |> 
-  select(-gameId, -offset_sec) |> 
-  filter(!is.na(game)) |> 
-  mutate(time=as.numeric(time))
+  rename(speaker=id)
 
@@
 
-less_raw_data <- raw_data %>% select(game, target, distractor, trialNum, targetNum, repNum, response, correct, time, listener=playerId, speaker) %>% 
-  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>% write_csv(here("data/clean_data.csv"))
-```
-
-```{r}
-#just for pilot, need to fix offset
-#fix <- raw_data |> filter(trialNum==0) |> select(game, time)
-
-#real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)
-
-real_offset <- games
+less_raw_data <- raw_data %>% select(gameId, target, distractor, trialNum, targetNum, repNum, response, correct, time, listener=playerId, speaker) %>% 
+  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
+ inner_join(read_csv(here("data/expt_1/link_transcripts.csv")) |> select(gameId, game)) |> 
+write_csv(here("data/expt_1/clean_data.csv"))
 
@@
 
-# Transcript wrangle - new
+# Transcript wrangle
+```{r}
+wrangle_transcript <- function(game_num) {
+  if(file.exists(here("data/expt_1/raw_transcripts", str_c(game_num, ".csv")))){
+  file = read_csv(here("data/expt_1/raw_transcripts", str_c(game_num, ".csv")))
+  }
+  else{
+    if(file.exists(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))){
+        #file = read_csv(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))
+         file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+    }
+    else{
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+    }
+  }
+  file|>
+    select(start, end, text, speaker, trial, role, echo, description) |>
+    fill(trial) |> 
+    mutate(start = start / 1000, end = end / 1000) |>
+   nest(data = c(start, end, text, speaker, trial, role, echo, description))
 
-```{r}
-
-wrangle_transcript <- function(game_num){
- read_csv(here("data/expt_1/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
-  select(start, end, text, speaker, trial, role, echo, description, backchannel) |> 
-    fill(trial) %>% 
-  mutate(start=start/1000, end=end/1000) |> 
-    nest(data=c(start, end, text, speaker, trial, role, echo, description, backchannel))
 }
 
-#wrangle_transcript("game41") %>% unnest(data) %>% View()
+transcripts <- read_csv(here("data/expt_1/link_transcripts.csv")) |> 
+  select(gameId, game) |>
+  mutate(transcript = map_df(game, wrangle_transcript)) |>
+  unnest(transcript) |>
+  unnest(data)
 
-real_offset$transcripts <- map_df(real_offset$gameConfig, wrangle_transcript)
-
-transcripts <- real_offset |> unnest(transcripts) |> unnest(data)
-
-# this section is janky as hell, but it does work
-
-```
-
-```{r}
-
-stages <- read_csv(here("data/expt_1/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)
-
-gameStarts <- stages |> filter(index==0) |> rename(first=startTimeAt) |> select(-index)
-
-stage_timing <- stages |> left_join(gameStarts) |>
-  mutate(rel_time=startTimeAt-first) |> 
-  mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> 
-   mutate(trialNum=index/2,
-          start=rel_time,
-          end=lead(rel_time)) |> 
-  select(gameId, trialNum,start, end) |>  left_join(games) |>  ungroup() |> 
-  mutate(trialNumStart=trialNum, trialNumEnd=trialNum) |> 
-  select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
-```
-
-```{r}
-timed_transcripts <-  transcripts |>
-  mutate(start=start-offset_sec, end=end-offset_sec) |> 
-  select(-offset_sec, -gameId) |> 
-  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
-  rename(game=gameConfig) |> 
-  bind_rows(stage_timing) |> 
-  group_by(game) |> 
-  arrange(start) |> 
-  fill(trialNumStart) |> 
-  arrange(end) |> 
-  fill(trialNumEnd, .direction="up") |> 
-  arrange(game) |> 
-  filter(!is.na(text))
-
-echoes <- timed_transcripts %>% select(game,trial, echo) %>% filter(echo=="Y") %>% unique()
-
-trial_timing <- stage_timing %>% select(game, trial=trialNumStart, trial_start=start)
-
-transcript_full <- timed_transcripts %>% select(-echo, -exclude) %>% left_join(echoes) %>% 
-  left_join(trial_timing) %>% 
-  rename(start_time=start, end_time=end) %>% 
-  write_csv(here("data/expt_1/timed_transcript.csv"))
-```
-
-# Checks
-```{r}
-# Quick check -- for kid talking, so auto-trial and hand-trial align?
-
-transcript_full %>% filter(!is.na(trialNumStart)) %>% filter(str_detect(speaker, "id")) %>% filter(!is.na(role)) %>% filter(trial!=trialNumStart)
-
-# we'll go with the by-hand method -- these mostly seem to be when there's a wide span of time! possibly due to whisper being silly or us hand filling due to whisper missing things
-
-```
-
-checks on data:
-* anything that has a description should be from a speaker and have Y for backchannel
-* anything that has a Y for backchannel should be from a child
-
-```{r}
-transcript_full %>% filter(!is.na(description)) %>% filter(is.na(backchannel))
-transcript_full %>% filter(!is.na(description)) %>% ungroup() %>% select(role) %>% unique()
-transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% select(role) %>% unique() 
-transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% filter(!str_detect(speaker,"id"))
+write_csv(transcripts, here("data/expt_1/transcripts.csv"))
 ```
@@
 ```{r}
-transcript <- read_csv(here("data/expt_1/timed_transcript.csv"))
+transcript <- read_csv(here("data/expt_1/transcripts.csv"))
 
-dat <- read_csv(here("data/expt_1/clean_data.csv")) %>% select(trial=trialNum, game, target, distractor)
+dat <- read_csv(here("data/expt_1/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)
 
-transcript %>% filter(!is.na(description)) %>% group_by(trial, game) %>% 
-  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% left_join(dat) %>% write_csv(here("data/expt_1/pre_sbert.csv"))
+transcript %>%
+  filter(!is.na(description)) %>%
+  group_by(trial, gameId) %>%
+  mutate(trial = as.numeric(trial)) |>
+  filter(trial > 1) |>
+  summarize(words = str_c(description, sep = " ", collapse = ", ")) %>%
+  left_join(dat) %>%
+  write_csv(here("data/expt_1/pre_sbert.csv"))
 
@@
 similarities <- M_mat %>%
-    mutate(combinedId=str_c(game,trial,sep="_")) %>%
+    mutate(combinedId=str_c(gameId,trial,sep="_")) %>%
     make_across_df(F_mat, 'cosine') %>%

### 2025-10-28 code/expt_3/data_prep_3.Rmd
clean up of transcripts for referential, matching ids

@@
 
-real_ids <- c(str_c("", c(
-  159:162, 167:170,
-  173:174, 181:184,
-  187:188, 191:196,
-  199:200, 203:204,
-  207:208, 227:234,
-  237:240, 243:252
-)))
-# kids ids to include
 ```
@@
 ```{r}
-players <- read_csv(here(player_source)) |>
-  filter(id %in% real_ids) %>%
-  select(id, playerId = `_id`)
 
-# expect 50 kids (25 games)
-```
+id_sub <- function(id){
+  new_id = case_when(
+    id=="pear" ~ "id153",
+    id=="apple" ~ "id154",
+    id=="pear155" ~ "id155",
+    id=="apple156" ~ "id156",
+    id=="pear157" ~ "id157",
+    id=="apple158" ~ "id158",
+    id=="pear123" ~ "id163",
+    id=="apple124" ~ "id164",
+    id=="pear124"~"id165",
+    id=="apple125" ~ "id166",
+    T ~ str_c("id", id)
+  )
+  return(new_id)
+}
 
-```{r}
+players <- read_csv(here(player_source)) |> select(playerId=`_id`, id) |> rowwise() |>  mutate(id=id_sub(id)) |> ungroup()
+
 raw_data <- read.csv(here(data_source), header = TRUE) %>%
   rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
-  # select(-chat) |>
   pivot_longer(
@@
   select(target, distractor, trialNum, targetNum, repNum, response, correct, time, listener = playerId, speaker, gameId) %>%
-  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
+  mutate(across(c(target, distractor, response), ~ label_image(.x))) |> 
+  inner_join(read_tsv(here("data/expt_3/link_transcripts.csv")) |> select(gameId, game)) |> 
   write_csv(here("data/expt_3/clean_data.csv"))
+
 ```
 
-below here copied from expt 2 and not updated!
-
-# Transcript wrangle - new
+# Transcript wrangle
 
 ```{r}
-library(readxl)
 
 wrangle_transcript <- function(game_num) {
-  read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv"))) |>
+  if(file.exists(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))){
+  file = read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))
+  }
+  else{
+    if(file.exists(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))){
+        file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
+    }
+    else{
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+    }
+  }
+  file|>
     select(start, end, text, speaker, trial, role, echo, description) |>
@@
     nest(data = c(start, end, text, speaker, trial, role, echo, description))
+
 }
@@
 transcripts <- read_tsv(here("data/expt_3/link_transcripts.csv")) |> 
-  select(gameId, gameConfig) |>
-  mutate(transcript = map_df(gameConfig, wrangle_transcript)) |>
+  select(gameId, game) |>
+  mutate(transcript = map_df(game, wrangle_transcript)) |>
   unnest(transcript) |>
@@
 write_csv(transcripts, here("data/expt_3/transcripts.csv"))
-```
-
-# Checks
-
-how much echoing is there?
-
-```{r}
-transcripts |> filter(!is.na(echo))
-
-# not much, yay!
 ```
@@
 
+(go run sbert here)
 ```{r}

### 2025-10-28 write-ups/manuscript/kid-tangrams-ms.Rmd
start ms

--- 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
@@
 ---
-title: |
+title             : |
   | Preschoolers can form conventional pacts with each other 
   | to communicate about novel referents
-bibliography: preschool-tangrams.bib
-csl: apa7.csl
-document-params: "10pt, letterpaper"
 
-author-information: > 
-    \author{Anonymized}
+author: 
+  - name          : "Veronica Boyce"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    email         : "vboyce@stanford.edu"
+  - name          : "Robert Z. Sparks"
+    affiliation   : "1"
+  - name          : "Michael C. Frank"
+    affiliation   : "1"
+
+affiliation:
+  - id            : "1"
+    institution   : "Stanford University"
+
+authornote: |
+  TODO credit taxonomy etc
 
@@
 Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm.
-Across 51 pairs of children, we found that 4-5-year-olds successfully established reference with each other.
-Children were 85% accurate, and they often used descriptions similar to their partner's.
+Across TODO pairs of children, we found that 4-5-year-olds successfully established reference with each other.
+Children were TODO% accurate, and they often used descriptions similar to their partner's.
 These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
+  <!-- https://tinyurl.com/ybremelq -->
 
-keywords: >
-    development; iterated reference game; pragmatics; preschoolers
+keywords          : development; iterated reference game; pragmatics; preschoolers
+wordcount         : "X"
 
-output: cogsci2024::cogsci_paper
-header-includes: \raggedbottom
-#header-includes: \cogscifinalcopy
+bibliography      : preschool-tangrams.bib
+
+floatsintext      : yes
+linenumbers       : yes
+draft             : no
+mask              : no
+
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+
+classoption       : "man"
+csl               : apa7.csl
+output            : papaja::apa6_pdf
+header-includes   :
+  - \usepackage{setspace}\singlespacing
 ---
 
-```{r global_options, include=FALSE}
+```{r setup, include = FALSE}
+library("papaja")
+
 knitr::opts_chunk$set(
@@
 expt_2 <- "data/expt_2"
+expt_3 <- "data/expt_3"
 mod_loc <- "code/models"
@@
 
-source(here("code/paper_helper.R"))
+library(ggtext)
+labels <- tribble(~target, ~img,
+                  "hold", "I1.jpg",
+                  "swim", "D1.jpg",
+                  "walk", "B1.jpg",
+                  "jump", "E1.jpg",
+                  "baby","tangram_H.png",
+                  "bunny","tangram_E.png",
+                  "dancer","tangram_G.png",
+                  "zombie", "tangram_A.png"
+                  ) |> 
+  mutate(img_source= str_c("<img src='",img,"' width='20' />"))
+
+#source(here("code/paper_helper.R"))
+joint_data <- read_csv(here("data/processed_for_paper/joint_with_demog.csv"))
+joint_similarities <-   read_csv(here("data/processed_for_paper/sims.csv"))
+
+### functions for printing things nicely!
+stats <- function(model, row, decimal = 2) {
+  model <- model |>
+    mutate(
+      Estimate = round(Estimate, digits = decimal),
+      Lower = round(lower, digits = decimal),
+      Upper = round(upper, digits = decimal),
+      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
+    ) |>
+    select(Term, Estimate, `Credible Interval`)
+  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
+}
+
+stats_text <- function(model, row, decimal = 2) {
+  model <- model |>
+    mutate(
+      Estimate = round(Estimate, digits = decimal) |> formatC(format='f', digits=decimal ),
+      Lower = round(lower, digits = decimal) |> formatC(format='f', digits=decimal ),
+      Upper = round(upper, digits = decimal) |> formatC(format='f', digits=decimal ),
+      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
+    ) |>
+    select(Term, Estimate, `Credible Interval`)
+  str_c(model[row, 2], "  ", model[row, 3])
+}
+
+form <- function(model_form) {
+  dep <- as.character(model_form$formula[2])
+  ind <- as.character(model_form$formula[3])
+
+  str_c(dep, " ~ ", ind) |>
+    str_replace_all(" ", "") |>
+    str_replace_all("\\*", " $\\\\times$ ") |>
+    str_replace_all("\\+", "&nbsp;+ ") |>
+    str_replace_all("~", "$\\\\sim$ ")
+}
 ```
+
+TODO all figures could be smaller!
 
@@
 
-```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure.
+```{r interface, fig.env = "figure", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure.
 Panel A shows the experimental setup.
@@
 
+TODO review the literature for 
+* any other studies we should include
+* any studies we want to say more about than we had room for in cogsci
+
 More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
@@
 Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.
+* [@sanjuan2025] When 4-5 year old children learn that some objects are only visible to them and not their partner, they are faster to rule out these competitors and select a target than when they believe the interlocuter is aware of the competitors as well.
+* [@nilsen2009] take into account partners perspective for both speaker and listening (4-5 yos); 3-4 year olds also on listening only; relate this to inhibitory control.
 
@@
 
+- Bohn, Kachel, Tomasello (2019) Young Children spontaneously recreate core properties of language in a new modality
+
+    - 4, 6, (and 8 in a separate thing) communicating target image via video link (no audio after practice trials)
+
+    - comprehension good and increasing, gestures used back by second partner when roles switch
+
+    - (out of 5 pictures, including “blank”)
+
+    - do see more similaritiy w/i dyads especially for 6 versus 4 yos
+
+May want to set up for change over age?
+And a little more on why 4-5 is a good age -- it's a time of rapid development, we think from prior work that children can comprehend the task.
+Just before big changes in linguistic input from school + reading.
+...
+
 Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication.
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across an initial study and a replication including a total of 51 pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
+Across three studies including a total of ZZ pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
+An additional XX games were excluded for being pilot runs or not completing enough trials.
 
@@
 We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe.
-After exclusions, we had `r clean_1 |> filter(!is.na(repNum)) |> nrow()` trials remaining.
+After exclusions, we had `r joint_data |> filter(expt=="1") |> nrow()` trials remaining.
 
@@
 We report estimates and 95% credible intervals.
-The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).
+The experimental set-up, analysis code, de-identified transcripts, and performance data for all experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).
 
@@
 \\label{acc}" }
-clean_2 |>
-  filter(type != "practice") |>
-  mutate(trial = trial - 3, expt = "Expt 2") |>
-  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
-  ggplot(aes(x = trial, y = correct)) +
+joint_data |>
+  ggplot(aes(x = trial+1, y = correct)) +
     geom_smooth(method = "lm", color = "black") +
@@
   # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
+  labs(color="Block")+
   xlab("Trial") +
@@
   geom_hline(yintercept = .5) +
-  facet_grid(expt ~ .) +
+  facet_grid(str_c("Expt ",expt) ~ .) +
   theme(
-    legend.position = "none",
     panel.grid = element_blank(),
@@
 
-Our primary measure of interest was whether children could accurately communicate the intended target.
-<!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
---> To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}).
+Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966].
+To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}).
 This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.
@@
 \\label{length}" }
-clean_2 |>
-  filter(type != "practice") |>
-  mutate(trial = trial - 3, expt = "Expt 2") |>
-  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
-  ggplot(aes(x = trial, y = words, color = type)) +
+joint_data |> 
+  ggplot(aes(x = trial+1, y = words, color = type)) +
   geom_jitter(alpha = .5, color = "grey") +
@@
   ylab("Words") +
-  facet_grid(expt ~ .) +
+  facet_grid(str_c("Expt ",expt) ~ .) +
   theme(
@@
 Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
+TODO add examples from expt 3 as well??
 
@@
 \\label{sim}" }
-sims_2 |>
+
+joint_similarities|>
   mutate(source = case_when(
@@
     game1 == game2 ~ "same game",
-    T ~ "different games"
-  )) |>
-  mutate(expt = "Expt 2") |>
-  bind_rows(sims_1 |>
-    mutate(source = case_when(
-      speaker1 == speaker2 ~ "same teller",
-      game1 == game2 ~ "same game",
-      T ~ "different games"
-    )) |> mutate(expt = "Expt 1")) |>
-  ggplot(aes(x = source, y = sim, color = target1)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+    T ~ "different games")) |> 
+  left_join(labels) |> 
+  ggplot(aes(x = source, y = sim, color = img_source)) +
+      stat_summary(aes(group=img_source),fun.data = "mean_cl_boot", position = position_dodge(width = .3), geom="line") +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3)) +
   labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "Target") +
-  facet_grid(expt ~ .) +
-  scale_color_solarized() +
+  facet_grid(str_c("Expt ",expt) ~ .) +
+  scale_color_brewer(palette="Dark2") +
   theme(
-    legend.position = "bottom",
+    legend.position = "right",
+    legend.text = element_markdown(color = "black", size = 11),
     panel.grid = element_blank(),
@@
 @boyce2024 introduced a more sensitive measure of semantic convergence that compares the content of utterances using word embeddings to trace how similarities within and across games change over time.
-<!--To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game.
---> With only 3 blocks of descriptions, we do not expect semantic similarity for descriptions of a given target to show any meaningful change over time.
+To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game.
+With only 3 blocks of descriptions, we do not expect semantic similarity for descriptions of a given target to show any meaningful change over time.
 However, we can test for a more coarse measure of sensitivity to partner: whether children's utterances are more like their partner's than children in other games'.
@@
 However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.
+%TODO consider dropping this because it's not that strong an analysis?
 
@@
 In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.
+
+TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
 
@@
 Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
+An additional XX games were excluded for being pilot runs or not completing enough trials.
 
@@
 Data were processed in the same way as Experiment 1.
-After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.
+After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r joint_data |> filter(expt=="2") |> nrow()` trials total.
 
@@
 
-```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiment 2.
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.
 Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
 \\label{tolast}" }
-sims_2 |>
-  filter(game1 == game2) |>
-  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
-  filter(later == 4) |>
-  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
-  ggplot(aes(earlier, sim, color = target1)) +
+joint_similarities |> 
+  filter(game1==game2) |> 
+  filter(later==3) |> 
+  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
+  ggplot(aes(earlier+1, sim, color = img_source)) +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
+    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
     geom_smooth(color="black", method="lm")+
-  geom_point(alpha = .1, position = position_dodge(width = .2)) +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
-    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
-  #  geom_smooth( method="lm")+
-  facet_wrap(~same_speaker) +
-  scale_color_solarized() +
+  facet_grid(str_c("Expt ", expt)~same_speaker) +
   scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
   labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
-  theme(legend.position = "bottom",
-            strip.background = element_blank(),
-        #panel.grid=element_blank()
+scale_color_brewer(palette="Dark2") +
+  theme(
+    legend.position = "right",
+    legend.text = element_markdown(color = "black", size = 11),
+    panel.grid = element_blank(),
+    strip.background = element_blank()
   )
@@
 
-<!--### Qualitative variation in interaction structure
+# Experiment 3
 
-Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
-In one game, the experimenter provided no input after the practice trials because the children rapidly took turns producing referring expressions and selecting the targets.
-Their advantage could partially be due to being best friends with each other; all children were paired with another child from the same classroom who they were willing to play with, but friendship levels varied.
+The first two experiments demonstrated children's ability to communicate referentially, but were limited in their generality by the use of a specific set of 4 fairly humanoid tangram figures.
+To explore the generality of the results to a different set of potentially harder tangram stimuli, we conducted Experiment 3.
 
-Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
-In one game, child A described the figure as "A human", prompting child B to note "There's two humans." Later in the game, child B used the description "A human", leading child A to ask "Which one?", which child B clarified with "The one that is walking".
-This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+## Methods
+As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2. Experiment 3 was pre-registered at TODO [anonymized link].
 
--->
+The procedure for Experiment 3 matched Experiment 2, and the only change was in the stimuli used.
+
+### Participants
+Experiment 3 was run between XX and YY , at the same preschool as Experiments 1 and 2.
+Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and ZZ participants in experiment 2 were also in experiment 3.
+TODO pairs of children completed all 16 critical trials, and TODO.
+Our target age range was 4 and 5-year-olds.
+Of the TODO children, QQ were girls, and the children had a median age of ZZ months, and a range of XX-YY months.
+An additional XX games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.
+
+### Materials
+A different set of critical images were used (see Figure TODO).
+These were chosen from a set of commonly used tangram images (CITATIONS), to be fairly distinct from each other.
+We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having convention names (TODO figure).
+
+### Procedure
+The procedure was the same as for Experiment 2.
+The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop.
+
+### Data processing
+Data were processed in the same way as Experiment 1. After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had TODO trials total.
+
+## Results
+
+### Accuracy and speed
+
+### Description length
+
+### Convergence
 
 # Joint analysis
+
+## Pooled analysis
 
@@
 
+## Sensitivity to prior result
+
 ```{r}
@@
 
+## Age and gender as moderators of referential success
+
+We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners.
+As children of this age are rapidly developing their linguistic skills (CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.
+
+```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair.
+Each dot is a game, and linear fits are overlayed.
+\\label{agewords}" }
+joint_data |>
+  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
+  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
+  labs(x="Mean age of participants (months)", y="Accuracy", color="Number of girls")+
+  geom_jitter(data=joint_data|>
+  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
+  scale_color_manual(values=c("#DC322FFF","#2AA198FF" ,"#268BD2FF"  ))+
+  geom_smooth(method = "lm", show.legend=F)
+```
+
+```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender.
+Each dot is a game, and linear fits are overlayed.
+\\label{ageacc}" }
+joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
+  geom_smooth(method = "lm", show.legend=F)+
+  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
+  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
+  labs(x="Teller Age (months)", y="Words", color="Teller gender")
+```
+
+For accuracy ... TODO
+
+For description length ... TODO
+
+These findings of a slight effect of age and a slight female-advantage are consistent with a wider literature of ...
+
+# Qualitative variation in interaction structure
+
+Pairs of children varied significantly in how well they could scaffold their own interaction versus needed reminders from experimenters about the structure of the game.
+In one game, the experimenter provided no input after the practice trials because the children rapidly took turns producing referring expressions and selecting the targets.
+Their advantage could partially be due to being best friends with each other; all children were paired with another child from the same classroom who they were willing to play with, but friendship levels varied.
+
+Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
+In one game, child A described the figure as "A human", prompting child B to note "There's two humans." Later in the game, child B used the description "A human", leading child A to ask "Which one?", which child B clarified with "The one that is walking".
+This anecdote illustrates how emerging abilities can be inconsistent -- finding a description inadequate as a guesser does not always translate into providing a more informative description as a teller.
+
+* maybe cut this section
+* do want to make the point that there's really wide variability (consistent with older kids and also adults for that matter) -- some pairs one kid does the describing including of options and the other chooses
+
+* may want to be clear that structuring the interaction in an effective way is a skill that not all the kiddos have -- some relied more on the experimenter saying whose turn and needed more prompting with "do you want to give more hints" or "do you want to ask a question".
+whereas other groups rapidly produced descriptions and then either clicked or asked for (and got) clarification (either completely on their own, or after the option was questions offered as a possibility).
+that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways we that followed the "rules" of the game)
+where we see what looks like failures it may be hard to tell what part isn't working (is it knowing what i know, or what other knows, or cognitive control, or producing questions, or sampling)
+
+# (unclear if include)
+could also do some qualitative comparison of the *types* of descriptions that kiddos give versus kid-parent vs adult ?
+
 # General discussion
@@
 However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
-<!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
--->Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
+Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
+Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
 
-Across 2 experiments and 51 pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
-Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions.
+Across 3 experiments and TODO pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
+Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (TODO%), indicating that children were generally able to produce adequate descriptions.
 Additionally, children's utterances showed signs of converging toward conceptual pacts.
@@
 
+Even within this population, we observed some substantial pair to pair differences.
+Pairs varied in what types of descriptions they have, how readily the gave them, as well as how well they understood the "rules" of the game and how much they cooperated with the rules.
+While we did not aim to test 3 year olds, the few 3 year olds who we tested accidentally did fine.
+
+Our qualitative impression is that it was more common for children to hit limits around understanding the game structure (ex.
+that the "teller" doesn't press the ipad),  cooperating with the game (ex.
+not trying to show the ipad to their partner, not pressing buttons before their partner has a change to describe), and understanding that this wasn't a test about the "name" of the shape, moreso than being truely able to describe an unknown referent adequately.
+Pairs of children did very more in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.
+Generally, it's hard to push at the lower limits of some of children's abilities because many skills are changing over the same time frame and variables that may not be of primary interest, such as patience and instruction-following, can influence the outcomes.
+
 In the broader picture of language acquisition, there is debate over the timing of the emergence of communicative and pragmatic abilities relative to the acquisition of grammar and meaning.
@@
 
+\newpage
+
 # References
 
-```{r}
-# References will be generated automatically by Pandoc and included here.
-# The following code is some latex to format the bibliography. Do not remove it.
-```
-
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-\noindent
+::: {#refs custom-style="Bibliography"}
+:::

### 2025-10-29 code/expt_1/data_prep.Rmd
run models, annotations

@@
         #file = read_csv(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))
-         file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+         file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA, on_topic=NA)
     }
     else{
-      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA, on_topic=NA)
     }
@@
   file|>
-    select(start, end, text, speaker, trial, role, echo, description) |>
+    select(start, end, text, speaker, trial, role, echo, description,on_topic) |>
     fill(trial) |> 
     mutate(start = start / 1000, end = end / 1000) |>
-   nest(data = c(start, end, text, speaker, trial, role, echo, description))
+   nest(data = c(start, end, text, speaker, trial, role, echo, description, on_topic))
 

### 2025-10-29 code/expt_3/analysis_3.Rmd
run models, annotations

---
title: "Kid tg 2"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)
library(brms)
library(rstan)
library(rstanarm)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_loc <- "data/expt_3"
mod_loc <- "code/expt_3/models"
```

# Exclusions

```{r}
transcript <- read_csv(here(data_loc, "transcripts.csv"))
data <- read_csv(here(data_loc, "clean_data.csv"))
link <- read_tsv(here(data_loc, "link_transcripts.csv"))

 exclude_echo <- transcript |>
   filter(!is.na(echo)) |>
   select(game, gameId, trial)

data <- read_csv(here(data_loc, "clean_data.csv")) |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |> 
  anti_join(exclude)

transcript <- transcript |>
    inner_join(link |> filter(is.na(exclude))) |> 
  anti_join(exclude) |>
  filter(!is.na(description))

words <- transcript |> mutate(words=description |> str_count("\\S+")) |>
  filter(role=="S") |>
  group_by(gameId, game, trial) |>
  summarize(words=sum(words))

all_data <- data |>left_join(words) |> 
  filter(!is.na(words)) |> 
   mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |> 
  mutate(correct=as.numeric(correct))

all_data |> select(game, speaker) |> distinct()

```

# Accuracy

```{r}

time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

all_data |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot", geom="point", size=2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = all_data |> filter(trial > 3), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  theme(
    aspect.ratio = 4 / 5,
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```
```{r}
all_data |> filter(type!="practice") |> ggplot(aes(x=reorder(gameId, as.numeric(correct)), y=as.numeric(correct)))+stat_summary(fun.data="mean_cl_boot")+theme(axis.text.x=element_blank())
```

```{r}
all_data |>
  filter(type != "practice") |>
  ggplot(aes(x = type, y = as.numeric(correct), color = target)) +
  theme_bw() +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0, 1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  labs(x = "", y = "Fraction correct") +
  theme(legend.position = "bottom")
```

# Description length

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(all_data, aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = all_data|> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  theme(legend.position = "none")
```

# Speed

```{r}

time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(all_data, aes(x = trial + 1, y = time / 1000, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = all_data |> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  theme(legend.position = "none")
```

# Similarities

```{r}
sims <- read_rds(here(data_loc, "similarities.rds")) |>
  as_tibble() |>
  inner_join(data |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(data |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
         same_game=(game1==game2) |> as.numeric(),
        later = ifelse(block1 > block2, block1, block2), 
        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
```

```{r}
sims |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  theme(legend.position = "bottom")
```

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```
```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```
```{r}
sims |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

# Run Analyses

```{r}

acc_priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod <- brm(correct ~ trial + (trial | gameId) + (1 | target),
  family = bernoulli(link = "logit"),
  data = all_data |> filter(!is.na(repNum)),
  file = here(mod_loc, "acc"),
  prior = acc_priors,
  control = list(adapt_delta = .95)
)
```

```{r}

red_priors <- c(
  set_prior("normal(5, 10)", class = "Intercept"),
  set_prior("normal(0,5)", class="b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

red_mod <- brm(words ~ trial + (trial | gameId) + (1 | target),
  data = all_data |> filter(!is.na(repNum)),
  file = here(mod_loc, "red"),
  prior = red_priors,
  control = list(adapt_delta = .95)
)
```

```{r}

speed_priors <- c(
  set_prior("normal(60, 100)", class = "Intercept"),
  set_prior("normal(0,20)", class="b"),
  set_prior("normal(0,20)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

speed_mod <- brm(time_seconds ~ trial + (trial | gameId) + (1 | target),
  data = all_data |> filter(!is.na(repNum)) |> mutate(time_seconds=time/1000),
  file = here(mod_loc, "speed"),
  prior = speed_priors,
  control = list(adapt_delta = .95)
)
```

```{r}
sim_priors <- c(
  set_prior("normal(.5, .2)", class = "Intercept"),
  set_prior("normal(0,.1)", class="b"),
  set_prior("normal(0,.05)", class = "sd"))

sim_no_time <- brm(sim ~ same_game + same_speaker + (1|target),
                   data=sims ,
                   file = here(mod_loc, "sim_no_time"),
  prior = sim_priors,
  control = list(adapt_delta = .95))
```

```{r}

sim_to_last <- brm(sim ~ earlier+same_speaker+(1|game)+(1|target),
                   data=sims |> filter(later==4) |> filter(game1==game2) |> mutate(game=game1),
                   file = here(mod_loc, "sim_to_last"),
  prior = sim_priors,
  control = list(adapt_delta = .95))

sim_to_next <- brm(sim ~ earlier+same_speaker+(1|game)+(1|target),
                   data=sims |> filter(later==earlier+1) |> filter(game1==game2) |> mutate(game=game1),
                   file = here(mod_loc, "sim_to_next"),
  prior = sim_priors,
  control = list(adapt_delta = .95))

sim_across <- brm(sim ~ block+(1|target),
                   data=sims |> filter(later==earlier) |> filter(game1!=game2) |> mutate(block=earlier),
                   file = here(mod_loc, "sim_across"),
  prior = sim_priors,
  control = list(adapt_delta = .95))
```

# Analysis results

[x] Logistic model of accuracy: correct ~ trialNum + (trialNum|game) + (1|target) (prior normal(0,1) for beta and sd, lkj(1) for correlation)

```{r}

summary(acc_mod)

```

no effect of trial on accuracy; accuracy generally high

[x] Linear model of description length: words ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(5,10), beta and sd priors of normal(0,5), and correlation of lkj(1))

```{r}

summary(red_mod)

```

no effect of trial on length; generally pretty short throughout

[x] Speed: time_seconds ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(60,100), beta and sd priors of (0,20) and correlation of lkj(1))

```{r}

summary(speed_mod)

```

kids get faster over time (by about .7 seconds a pop)

We will interpret the fixed effect of trial number as being about changes in referring over time.

## Sbert models
For utterance similarities, we will embed the describer’s descriptions using S-BERT and use cosine similarity as a proxy for utterance similarity.

We are unsure whether 4 blocks will be sufficient to see potential change over time, so we will

[x]* compare similarities of utterances between different games v within one game v within one speaker: sim ~ same_game + same_speaker + (1|target)

```{r}
summary(sim_no_time)
```

utterances from within the same game are more similar (.34 across games v .59 within), smaller effect of same speaker(+.08, so .67)

[x]* compare similarities of utterances to the last utterance within a game: sim_to_last ~ earlier_block + same_speaker + (1|game) + (1|target)

```{r}
summary(sim_to_last)
```

utterances do get (slightly) more similar to last utterance over time, with a probable bump if same speaker

[x]* compare similarities of utterances to the next utterance: sim_to_next ~ earlier_block + same_speaker + (1|game) + (1|target)

```{r}
summary(sim_to_next)
```

utterances are again more similar from same speaker, but similarity to next is not clearly increasing over time

[x]* compare the similarity level of descriptions from the same block in different games: sim ~ block + (1|target)

```{r}
summary(sim_across)
```

distance of descriptions across games is stable (not seeing divergence)

Priors for the above similarity models: normal(.5,.2) for intercept, normal(0,.1) for
beta, and normal(0,.05) for sd.

We will also qualitatively explore children’s turn-taking structure and how they understand each other.

TODO: the qualitative thing

### 2025-10-29 code/expt_3/data_prep_3.Rmd
run models, annotations

@@
     if(file.exists(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))){
-        file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
+        #file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
+        file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA,description=NA, on_topic=NA)
     }
     else{
-      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA,description=NA, on_topic=NA)
     }
@@
   file|>
-    select(start, end, text, speaker, trial, role, echo, description) |>
+    select(start, end, text, speaker, trial, role, echo, description, on_topic) |>
     fill(trial) %>%
@@
     mutate(trial = trial-1) |>
-    nest(data = c(start, end, text, speaker, trial, role, echo, description))
+    nest(data = c(start, end, text, speaker, trial, role, echo, description, on_topic))
 

### 2025-10-29 code/ms_analysis.Rmd
run models, annotations

---
title: "Kid tangrams joint analysis"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)
library(brms)
library(rstan)
library(rstanarm)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_loc_1 <- "data/expt_1"
data_loc_2 <- "data/expt_2"
data_loc_3 <- "data/expt_3"
mod_loc <- "code/mods_for_paper"
```

# Read in data

expt 1

```{r}
transcript <- read_csv(here(data_loc_1, "transcripts.csv"))
data <- read_csv(here(data_loc_1, "clean_data.csv"))
link <- read_csv(here(data_loc_1, "link_transcripts.csv"))

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude)))

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_1 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 2 ~ "practice",
    trial < 6 ~ "block 1",
    trial < 10 ~ "block 2",
    trial < 14 ~ "block 3",
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "1")
```

expt 2

```{r}
transcript <- read_csv(here(data_loc_2, "transcripts.csv"))
data <- read_csv(here(data_loc_2, "clean_data.csv"))
link <- read_csv(here(data_loc_2, "link_transcripts.csv"))

exclude_echo <- transcript |>
  filter(!is.na(echo)) |>
  select(game, gameId, trial)

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo)

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_2 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "2")
```

expt 3

```{r}
transcript <- read_csv(here(data_loc_3, "transcripts.csv"))
data <- read_csv(here(data_loc_3, "clean_data.csv"))
link <- read_tsv(here(data_loc_3, "link_transcripts.csv")) |> select(gameId, game, exclude, exclusion_reason)

exclude_echo <- transcript |>
  filter(!is.na(echo)) |>
  select(game, gameId, trial)

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo)

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_3 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "3")
```

```{r}
demogs <- read_csv(here(data_loc_1, "clean_demog.csv")) |>
  bind_rows(read_csv(here(data_loc_2, "clean_demog.csv"))) |>
  bind_rows(read_csv(here(data_loc_3, "clean_demog.csv"))) |>
  mutate(expt = as.character(expt))
```

```{r}
joint <- all_data_1 |>
  bind_rows(all_data_2) |>
  bind_rows(all_data_3)

speaker_demogs <- demogs |> select(speaker = id, expt, game, speaker_gender = gender, speaker_age = age_month)

listener_demogs <- demogs |> select(listener = id, expt, game, listener_gender = gender, listener_age = age_month)

joint_demog <- joint |>
  left_join(speaker_demogs) |>
  left_join(listener_demogs) |>
  mutate(trial = case_when(
    expt=="1"~ trial-2,
    T ~ trial-4,
  ),     time.sec=time/1000
) |> filter(type!="practice") |> select(-time) |> 
  write_csv(here("data/processed_for_paper/joint_with_demog.csv"))
```

```{r}
sims_1 <- read_rds(here(data_loc_1, "similarities.rds")) |>
  as_tibble() |> 
  inner_join(all_data_1 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_1 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |> 
  filter(target1 == target2) |> 
  mutate(block1 = (trial1-2) %/% 4, block2 = (trial2-2) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "1")

sims_2 <- read_rds(here(data_loc_2, "similarities.rds")) |>
  as_tibble() |>
  inner_join(all_data_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = (trial1-4) %/% 4, block2 = (trial2-4) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "2")

sims_3 <- read_rds(here(data_loc_3, "similarities.rds")) |>
  as_tibble() |>
  inner_join(all_data_3 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_3 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = (trial1-4) %/% 4, block2 = (trial2-4) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "3")

sims <- sims_1 |>
  bind_rows(sims_2) |>
  bind_rows(sims_3) |>
  select(-target1, -target2) |> 
  write_csv(here("data/processed_for_paper/sims.csv"))
```

# QC

TODO TURN INTO ASSERTIONS
```{r, eval=F}
joint_demog |>
  select(speaker, game, expt) |>
  distinct() |>
  group_by(game, expt) |>
  tally() |>
  filter(n != 2)

joint_demog |>
  select(listener, game, expt) |>
  distinct() |>
  group_by(game, expt) |>
  tally() |>
  filter(n != 2)

joint_demog |>
  select(speaker, game, expt, trial) |>
  distinct() |>
  group_by(game, expt, trial) |>
  tally() |>
  filter(n != 1)

joint_demog |>
  select(listener, game, expt, trial) |>
  distinct() |>
  group_by(game, expt, trial) |>
  tally() |>
  filter(n != 1)
# targetNum
# repNum
```

```{r, eval=F}
transcript_1 <- read_csv(here(data_loc_1, "transcripts.csv")) |>
  left_join(read_csv(here(data_loc_1, "link_transcripts.csv"))) |>
  mutate(expt = "1")

transcript_2 <- read_csv(here(data_loc_2, "transcripts.csv")) |>
  left_join(read_csv(here(data_loc_2, "link_transcripts.csv"))) |>
  mutate(expt = "2")

transcript_3 <- read_csv(here(data_loc_3, "transcripts.csv")) |>
  left_join(read_tsv(here(data_loc_3, "link_transcripts.csv"))) |>
  mutate(expt = "3")

transcripts <- transcript_1 |>
  bind_rows(transcript_2) |>
  bind_rows(transcript_3)

# confirm that all S and L tags go with kids

transcripts |>
  filter(role %in% c("S", "L")) |>
  filter(!str_detect(speaker, "id")) |>
  select(expt, game, speaker, role, text)

transcripts |>
  select(trial, expt) |>
  distinct() # should be 0-19 (0-13 for expt 1) + NAs

transcripts |>
  filter(!is.na(trial)) |>
 filter(str_detect(speaker, "id")) |>
  filter(is.na(role)) |>
  select(expt, game, trial, text) |> View() # to check on later!!!

transcripts |> filter(!is.na(trial)) |> 
    filter(!is.na(role)) |>
  select(game, expt, trial, speaker, role) |> 
  distinct()|>group_by(game, expt, trial, speaker) |> tally() |> filter(n!=1)

transcripts |> filter(!is.na(trial)) |> 
    filter(!is.na(role)) |>
  select(game, expt, trial, speaker, role) |> 
  distinct()|>group_by(game, expt, trial, role) |> tally() |> filter(n!=1)

transcripts |>
  filter(!is.na(trial)) |>
  filter(!is.na(role)) |>
  select(game, expt, trial, speaker_transcript = speaker, role, text) |>
  filter(role == "S") |>
  full_join(joint_demog) |>
  filter(speaker != speaker_transcript) |>
  select(expt, game, trial, text, speaker_transcript, speaker, listener)
```

# Graphs
## Accuracy
```{r, fig.width=5, fig.height=5}
time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

joint_demog |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot", geom = "point", size = 2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  geom_hline(yintercept = .5) +
  facet_grid(expt ~ .) +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```

```{r}
joint_demog |>
  filter(type != "practice") |>
  ggplot(aes(x = type, y = as.numeric(correct), color = target)) +
  theme_bw() +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0, 1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  labs(x = "", y = "Fraction correct") +
  theme(legend.position = "bottom")
```

## Speed

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = time.sec, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

## Reduction

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = all_words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

# Models
## Accuracy
```{r}
acc_priors <- c(
  set_prior("normal(0,1)", class = "b"),
  set_prior("normal(0,1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod_1 <- brm(correct ~ trial + (trial | game) + (1 | target),
                 family = bernoulli(link = "logit"),
                 data = joint_demog |> filter(expt=="1"),
                 file = here(mod_loc, "acc_1.rds"),
                 prior = acc_priors,
                 control = list(adapt_delta = .95)
)

acc_mod_2 <- brm(correct ~ trial+ (trial| gameId) + (1 | target),
               family = bernoulli(link = "logit"),
               data = joint_demog |> filter(expt=="2"),
               file = here(mod_loc, "acc_2.rds"),
               prior = acc_priors,
               control = list(adapt_delta = .95)
)

acc_mod_3 <- brm(correct ~ trial + (trial | gameId) + (1 | target),
               family = bernoulli(link = "logit"),
               data = joint_demog |> filter(expt=="3"),
               file = here(mod_loc, "acc_3.rds"),
               prior = acc_priors,
               control = list(adapt_delta = .95)
)

acc_mod_all <- brm(correct ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                    family = bernoulli(link = "logit"),
                    data = joint_demog,
                    file = here(mod_loc, "acc_meta.rds"),
                    prior = acc_priors,
                    control = list(adapt_delta = .95)
)

```

## Speed
```{r}
speed_priors <- c(
  set_prior("normal(60,100)", class = "Intercept"),
  set_prior("normal(0,20)", class = "b"),
  set_prior("normal(0,20)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

speed_mod_1 <- brm(time.sec ~ trial + (trial | game) + (1 | target),
                   data = joint_demog |> filter(expt=="1"),
                   file = here(mod_loc, "speed_1.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_2 <- brm(time.sec ~ trial + (trial | gameId) + (1 | target),
                   data = joint_demog |> filter(expt=="2"),
                   file = here(mod_loc, "speed_2.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_3 <- brm(time.sec ~ trial + (trial | gameId) + (1 | target),
                   data = joint_demog |> filter(expt=="3"),
                   file = here(mod_loc, "speed_3.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_meta <- brm(time.sec ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                   data = joint_demog,
                   file = here(mod_loc, "speed_meta.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)
```

## Description length

```{r}
description_priors <- c(
  set_prior("normal(5,10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0,5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

description_mod_1 <- brm(words ~ trial + (trial | game) + (1 | target),
                         data = joint_demog |> filter(expt=="1"),
                         file = here(mod_loc, "description_1.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_2 <- brm(words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="2"),
                         file = here(mod_loc, "description_2.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_3 <- brm(words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="3"),
                         file = here(mod_loc, "description_3.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_meta <- brm(words ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                            data = joint_demog,
                            file = here(mod_loc, "description_meta.rds"),
                            prior = description_priors,
                            control = list(adapt_delta = .95)
)
```

## Description -- robustness
we can instead of just the cleaned up description, look at all kids on topic words (including listener ?s, self repeitions, saying I don't know...)

this excludes off topic things (I have a stuffy too / comments about smurfy), stuff about the mechanics of the game (do I press it?), and celebration/non-referential commentary on how the game is going (yay!, I'm winning. that was wrong)

TODO 
```{r}
description_priors <- c(
  set_prior("normal(5,10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0,5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

all_description_mod_1 <- brm(all_words ~ trial + (trial | game) + (1 | target),
                         data = joint_demog |> filter(expt=="1"),
                         file = here(mod_loc, "all_description_1.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_2 <- brm(all_words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="2"),
                         file = here(mod_loc, "all_description_2.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_3 <- brm(all_words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="3"),
                         file = here(mod_loc, "all_description_3.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_meta <- brm(all_words ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                            data = joint_demog,
                            file = here(mod_loc, "all_description_meta.rds"),
                            prior = description_priors,
                            control = list(adapt_delta = .95)
)
```

# S-BERT graphs

## Sim to self/partner/others
```{r}
sims |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  facet_wrap(~expt) +
  theme(legend.position = "bottom")
```

## Sim to last round

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_grid(expt ~ same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## Sim to next round

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## Sim across game

```{r}
sims |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## similarity and accuracy

```{r}
sim_acc_to_next <- sims |> 
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
    earlier_block.num = earlier
  ) |> 
  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker, expt) |> left_join(joint_demog |> select(target, earlier_trial=trial, game1=gameId, correct, expt)) |> filter(!is.na(correct))

sim_acc_to_next |> ggplot(aes(as.character(earlier), sim, color = as.character(correct))) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity") +
  theme(legend.position = "bottom")+facet_wrap(~expt)
```

# S-BERT models

# across-between 
```{r}
conv_priors <- c(
  set_prior("normal(.5,.2)", class = "Intercept"),
  set_prior("normal(0,.1)", class = "b"),
  set_prior("normal(0,.05)", class = "sd")
)

sim_across_between_data <- sims |>
  mutate(same_game = case_when(
    game1 == game2 ~ 1,
    T ~ 0
  )) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))

sim_across_between_mod_1 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="1"),
                                file = here(mod_loc, "sim_across_between_1.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="2"),
                                file = here(mod_loc, "sim_across_between_2.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_3 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="3"),
                                file = here(mod_loc, "sim_across_between_3.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_meta <- brm(sim ~ same_game + same_speaker + (1 | target) + (1|expt),
                                   data = sim_across_between_data,
                                   file = here(mod_loc, "sim_across_between_meta.rds"),
                                   prior = conv_priors,
                                   control = list(adapt_delta = .95)
)
```

## to last

```{r}
sim_to_last_data <- sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
  )

sim_to_last_mod_2 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_last_data |> filter(expt=="2"),
                         file = here(mod_loc, "sim_to_last_2.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
sim_to_last_mod_3 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_last_data |> filter(expt=="3"),
                         file = here(mod_loc, "sim_to_last_3.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_last_mod_meta <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target)+ (1|expt),
                         data = sim_to_last_data,
                         file = here(mod_loc, "sim_to_last_meta.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
```

## to next

```{r}
sim_to_next_data <- sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
  )

sim_to_next_mod_2 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_next_data |> filter(expt=="2"),
                         file = here(mod_loc, "sim_to_next_2.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_next_mod_3 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_next_data |> filter(expt=="3"),
                         file = here(mod_loc, "sim_to_next_3.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_next_mod_meta <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target)+ (1|expt),
                         data = sim_to_next_data,
                         file = here(mod_loc, "sim_to_next_meta.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
```

## cross game

```{r}
sim_across_data <- sims |>
  filter(game1 != game2) |>
  filter(block1 == block2)

sim_across_mod_2 <- brm(sim ~ block1 + (1 | target),
                        data = sim_across_data |> filter(expt=="2"),
                        file = here(mod_loc, "sim_across_2.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

sim_across_mod_3 <- brm(sim ~ block1 + (1 | target),
                        data = sim_across_data |> filter(expt=="3"),
                        file = here(mod_loc, "sim_across_3.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

sim_across_mod_meta <- brm(sim ~ block1 + (1 | target)+ (1|expt),
                        data = sim_across_data,
                        file = here(mod_loc, "sim_across_meta.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

```

## similarity & accuracy

may need to double check this one!
```{r}

sim_acc_to_next <- sims |> 
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0)
  ) |> 
  select(target, game1, earlier, later, sim, same_speaker, expt) |> 
left_join(joint_demog |> select(target, earlier=repNum, game1=gameId, correct, expt ))

conv_priors <- c(
  set_prior("normal(.5,.2)", class = "Intercept"),
  set_prior("normal(0,.1)", class = "b"),
  set_prior("normal(0,.05)", class = "sd")
)

sim_acc_to_next_mod_meta <- brm(sim ~ earlier*correct + same_speaker*correct + (1 | game1) + (1 | target) + (1|expt),
                                data = sim_acc_to_next,
                                file = here(mod_loc, "sim_acc_to_next_meta.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

```

# Demog graphs

## Accuracy
```{r}
joint_demog |> ggplot(aes(x = speaker_age, y = correct, color = interaction(speaker_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(correct=mean(correct, na.rm=T)))

joint_demog |> ggplot(aes(x = listener_age, y = correct, color = interaction(listener_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(listener_age, expt, game, listener_gender) |> summarize(correct=mean(correct, na.rm=T)))

joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
  labs(x="Mean age of participants", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
  geom_smooth(method = "lm")
```

## words

```{r}
joint_demog |> ggplot(aes(x = speaker_age, y = words, color = interaction(speaker_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))

joint_demog |> ggplot(aes(x = listener_age, y = words, color = interaction(listener_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(listener_age, expt, game, listener_gender) |> summarize(words=mean(words, na.rm=T)))

joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = words, color = as.character(girlness))) +
  labs(x="Mean age of participants", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(words=mean(words, na.rm=T))) +
  geom_smooth(method = "lm")
```

# Demog models
## Accuracy

```{r}

mean_age <- joint_demog |> distinct(expt, speaker, speaker_age) |> summarize(mean_age=mean(speaker_age)) |> pluck(1,1)

centered_demog <- joint_demog |> 
  mutate(
    centered_speaker_age = speaker_age - mean_age,
    centered_listener_age = listener_age - mean_age,
    speaker_gender_num = ifelse(speaker_gender == "female", .5, -.5),
    listener_gender_num = ifelse(listener_gender == "female", .5, -.5)
  )

acc_priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod_demog <- brm(
  correct ~ trial +
    centered_speaker_age + speaker_gender_num +
    centered_listener_age + listener_gender_num +
    (trial | game) + (1 | target),
  family = bernoulli(link = "logit"),
  data = centered_demog,
  file = here(mod_loc, "acc_demog.rds"),
  prior = acc_priors,
  control = list(adapt_delta = .95)
)

```

## Description length

```{r}
red_priors <- c(
  set_prior("normal(5, 10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

red_mod <- brm(
  words ~ trial*(centered_speaker_age + speaker_gender_num +
    centered_listener_age + listener_gender_num) +
    (trial | game) + (1 | target),
  data = centered_demog,
  file = here(mod_loc, "description_demog.rds"),
  prior = red_priors,
  control = list(adapt_delta = .95)
)

```

# Save model summaries

```{r}
library(tidybayes)

save_summary <- function(model) {
  intervals <- gather_draws(model, `b_.*`, regex = T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex = T) %>%
    mutate(above_0 = ifelse(.value > 0, 1, 0)) %>%
    group_by(.variable) %>%
    summarize(pct_above_0 = mean(above_0)) %>%
    left_join(intervals, by = ".variable") %>%
    mutate(
      lower = .lower,
      upper = .upper,
      Term = str_sub(.variable, 3, -1),
      Estimate = .value
    ) %>%
    select(Term, Estimate, lower, upper)

  stats
}

do_model <- function(path) {
  model <- read_rds(here(mod_loc, path))
  save_summary(model) |> write_rds(here(mod_loc, "summary", path))
  model$formula |> write_rds(here(mod_loc, "formulae", path))
  print(summary(model))
}

mods <- list.files(path = here(mod_loc), pattern = ".*rds") |> walk(~ do_model(.))
```

### 2025-10-29 write-ups/manuscript/kid-tangrams-ms.Rmd
run models, annotations

@@
 expt_3 <- "data/expt_3"
-mod_loc <- "code/models"
-mod_results <- "code/models/summary"
-mod_form <- "code/models/formulae"
+mod_loc <- "code/mods_for_paper"
+mod_results <- "code/mods_for_paper/summary"
+mod_form <- "code/mods_for_paper/formulae"
 time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple")
@@
 Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands that prevented accurate measurement of children's abilities.
+
+TODO update diagram with expt 3 stim.
 
@@
 description_form <- read_rds(here(mod_form, "description_1.rds"))
+
+description_robustness_mod_1 <- read_rds(here(mod_results, "all_description_1.rds"))
 ```
@@
 
-Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
+TODO this description metric is just the length of the description, taking out repetition/rephrasing from the child.
+We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc)
+
+TOOD maybe move this paragraph Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
 TODO add examples from expt 3 as well??
@@
 However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.
-%TODO consider dropping this because it's not that strong an analysis?
+TODO consider dropping this because it's not that strong an analysis?
 
@@
 The first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial).
-Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
-Taken together, we find more evidence that children can successfully communicate with each other about these abstract shapes, and do so with increasing efficiency.
+<!--Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens.
+-->Taken together, we find more evidence that children can successfully communicate with each other about these abstract shapes, and do so with increasing efficiency.
 
@@
 description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))
+
+description_robustness_mod_2 <- read_rds(here(mod_results, "all_description_2.rds"))
+
 ```
@@
 We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions were more similar if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
+Descriptions were marginally more similar if they came from the same child, which is expected as a sign of internal consistency TODO rephrase (`r sim_to_last_2_results |> stats_text(3,3)`).
 Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).
@@
 ## Results
+### Accuracy and speed
 
-### Accuracy and speed
+```{r}
+acc_mod_3 <- read_rds(here(mod_results, "acc_3.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
+
+acc_pct<- joint_data |> group_by(expt) |> summarize(acc=mean(correct, na.rm=T)*100)
+```
+
+```{r}
+speed_mod_3 <- read_rds(here(mod_results, "speed_3.rds"))
+```
+
+In Experiment 3, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_3,1)`) and relatively stable over time (OR of one trial later: `r stats_text(acc_mod_3, 2)`, Figure \ref{acc}).
+Children's overall accuracy was lower in Experiment 3(`r round(acc_pct |> pluck(2,3))`%) than in the prior experiments (`r round(acc_pct |> pluck(2,2))`%), potentially indicating that the stimuli were more difficult.
+
+The first critical trial averaged `r stats_text(speed_mod_3, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_3, 2)` seconds / trial).
+Overall, we find evidence that children's ability to communicate with increasing efficiency generalizes to different abstract images.
 
@@
 
+```{r}
+description_mod_3 <- read_rds(here(mod_results, "description_3.rds"))
+
+description_robustness_mod_3 <- read_rds(here(mod_results, "all_description_3.rds"))
+
+```
+
+The average length of descriptions on the first trial was `r stats_text(description_mod_3, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_3, 2)` words / trial, Figure \ref{length}).
+TODO after we double check data again, but might be overall more words than other expts?
+
 ### Convergence
+
+TODO reverify everything after data review + full sample
+
+```{r}
+sim_across_between_3_results <- read_rds(here(mod_results, "sim_across_between_3.rds"))
+
+sim_across_3_results <- read_rds(here(mod_results, "sim_across_3.rds"))
+
+sim_to_next_3_results <- read_rds(here(mod_results, "sim_to_next_3.rds"))
+
+sim_to_last_3_results <- read_rds(here(mod_results, "sim_to_last_3.rds"))
+
+```
+
+We repeat the semantic analyses from Experiment 2.
+To start with, we again found that utterances were more similar if they came from the same partnership (`r sim_across_between_3_results |> stats_text(2,3)`) and if they came from the same person (`r sim_across_between_3_results |> stats_text(3,3)`).
+
+Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_3_results |> stats_text(2,3)`).
+Over time, the semantic distance between adjacent block utterances increased slightly (`r sim_to_next_2_results |> stats_text(2,3)`).
+
+Unlike in Experiment 2, we did not find that different games became less similar over time (`r sim_across_3_results |> stats_text(2,3)`). 
+TODO interpretation
 
@@
 
+description_robustness_mod_meta <- read_rds(here(mod_results, "all_description_meta.rds"))
+
 sim_across_between_mod_meta <- read_rds(here(mod_results, "sim_across_between_meta.rds"))
 
+sim_to_last_mod_meta <- read_rds(here(mod_results, "sim_to_last_meta.rds"))
+
 sim_to_next_mod_meta <- read_rds(here(mod_results, "sim_to_next_meta.rds"))
+
+sim_across_mod_meta <- read_rds(here(mod_results, "sim_across_meta.rds"))
 ```
 
-As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect.
-Pooling the two experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+As the two experiments were similar to one another, we re-ran models pooling the data across the three experiments, using experiment number as a random effect.
+Pooling the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+
+TODO decide whether to discuss speed here again.
+
 Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
-Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
+Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).
+
+For other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks.
+In the combined model, we found that the similarity the last block utterance increased across blocks (`r stats_text(sim_to_last_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_last_mod_meta, 3,3)`).
+In a combined model, similarity to the next block utterance also increased across blocks (`r stats_text(sim_to_next_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_next_mod_meta, 3,3)`).
+Across the two experiments, utterances from different pairs to the same stimulus tended to decrease in similarity across blocks (`r stats_text(sim_across_mod_meta, 2,3)`).
+
+While some of these effects were small and thus not all effects were robustness detectable in each experiment by itself, the overall picture is consistent in direction with the patterns of convergence and divergence reported in reference games between adults [@boyce2024].
 
@@
 Each dot is a game, and linear fits are overlayed.
-\\label{agewords}" }
+\\label{ageacc}" }
 joint_data |>
@@
 Each dot is a game, and linear fits are overlayed.
-\\label{ageacc}" }
+\\label{agewords}" }
 joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
@@
 
-For accuracy ... TODO
+```{r}
+acc_demog_meta <- read_rds(here(mod_results, "acc_demog.rds")) |> mutate(across(c(`Estimate`, `lower`, `upper`), ~ exp(.)))
 
-For description length ... TODO
+acc_demog_form <- read_rds(here(mod_form, "acc_demog.rds"))
 
-These findings of a slight effect of age and a slight female-advantage are consistent with a wider literature of ...
+description_demog_meta <- read_rds(here(mod_results, "description_demog.rds"))
+description_demog_form <- read_rds(here(mod_form, "description_demog.rds"))
+```
+
+The effects of children's age and gender on their performance are shown in Figure \ref{ageacc}.
+We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child^[`r form(acc_demog_form)`].
+Accuracy was higher for older children (OR for one month older speaker: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older listener: `r stats_text(acc_demog_meta, 2,2)`).
+Accuracy was also higher when the listener was a girl (OR: `r stats_text(acc_demog_meta, 4,2)`), but was not credibly higher when the speaker was a girl (OR: `r stats_text(acc_demog_meta, 5,2)`).
+
+For description length, we tested whether children's age or gender effected how many words they produced in descriptions when in the teller role (Figure \ref{agewords}).
+We also ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors^[`r form(description_demog_form)`].
+Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of listener (`r stats_text(description_demog_meta, 2,2)`).
+For gender, there was no effect of teller gender on description length (girls produced `r stats_text(description_demog_meta, 5,2)` words compared to boys), but tellers produced longer descriptions for boy guessers (`r stats_text(description_demog_meta, 4,2)`).
+There were not substantial effects of trial order or interactions between demographics and trial order.
+
+It is curious that listener gender but not speaker gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic.
+For accuracy, the age and gender effects could be reflective of any of a number of developing capacities, including linguistic, executive functioning, or social/theory of mind.
+These findings of a slight effect of age and a slight female-advantage are consistent with a wider literature of TODO
 
@@
 whereas other groups rapidly produced descriptions and then either clicked or asked for (and got) clarification (either completely on their own, or after the option was questions offered as a possibility).
-that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways we that followed the "rules" of the game)
+that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways that followed the "rules" of the game)
 where we see what looks like failures it may be hard to tell what part isn't working (is it knowing what i know, or what other knows, or cognitive control, or producing questions, or sampling)

### 2025-10-29 code/cogsci_paper_helper.R

@@ -1,0 +1,155 @@
+## read in all the data stuff
+# experiment 1 imports
+transcript_1 <- read_csv(here(expt_1, "timed_transcript.csv"))
+dat_1 <- read_csv(here(expt_1, "clean_data.csv")) |> rename(trial = trialNum)
+exclusion_1 <- transcript_1 %>%
+  filter(!is.na(description)) %>%
+  filter(role == "S") %>%
+  select(game, trial) %>%
+  unique() %>%
+  mutate("S_talked" = 1)
+
+words_1 <- transcript_1 |>
+  filter(!is.na(description)) |>
+  mutate(words = description |> str_count("\\S+")) |>
+  filter(role == "S") |>
+  group_by(game, trial) |>
+  summarize(words = sum(words))
+
+clean_1 <- dat_1 %>%
+  left_join(exclusion_1) %>%
+  filter(S_talked == 1) |>
+  left_join(words_1) |>
+  mutate(type = case_when(
+    trial < 2 ~ "practice",
+    trial < 6 ~ "block 1",
+    trial < 10 ~ "block 2",
+    trial < 14 ~ "block 3"
+  ))
+
+
+sims_1 <- read_rds(here(expt_1, "similarities.rds")) |>
+  as_tibble() |>
+  inner_join(clean_1 |> select(game1 = game, trial1 = trial, target1 = target, speaker1 = speaker)) |>
+  inner_join(clean_1 |> select(game2 = game, trial2 = trial, target2 = target, speaker2 = speaker)) |>
+  filter(target1 == target2) |>
+  filter(trial1 > 1) |>
+  filter(trial2 > 1) |>
+  mutate(
+    trial1 = trial1 + 2,
+    trial2 = trial2 + 2
+  ) |>
+  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |>
+  mutate(
+    same_speaker = (speaker1 == speaker2) |> as.numeric(),
+    same_game = (game1 == game2) |> as.numeric(),
+    later = ifelse(block1 > block2, block1, block2),
+    earlier = ifelse(block1 > block2, block2, block1)
+  ) |>
+  mutate(target = target1)
+
+
+# expt 2 imports
+transcript_2 <- read_csv(here(expt_2, "transcripts.csv"))
+link <- read_csv(here(expt_2, "link_transcripts.csv"))
+
+exclude_2 <- transcript_2 |>
[…]

### 2025-10-29 code/run_sbert.py

@@ -1,0 +1,30 @@
+
+import numpy as np
+import pandas as pd
+from sentence_transformers import SentenceTransformer
+model = SentenceTransformer('paraphrase-MiniLM-L6-v2')
+
+
+
+sample=pd.read_csv('../data/expt_3/pre_sbert.csv')
+sentence_list=sample['words'].tolist()
+
+embeddings = model.encode(sentence_list)
+
+np.save('../data/expt_3/post_sbert.npy', embeddings)
+
+
+sample=pd.read_csv('../data/expt_2/pre_sbert.csv')
+sentence_list=sample['words'].tolist()
+
+embeddings = model.encode(sentence_list)
+
+np.save('../data/expt_2/post_sbert.npy', embeddings)
+
+
+sample=pd.read_csv('../data/expt_1/pre_sbert.csv')
+sentence_list=sample['words'].tolist()
+
+embeddings = model.encode(sentence_list)
+
+np.save('../data/expt_1/post_sbert.npy', embeddings)

### 2025-10-30 code/ms_analysis.Rmd
double checking standardization of description , typo catches

@@
   select(expt, game, trial, text, speaker_transcript, speaker, listener)
+
+transcripts  |> filter(on_topic=="Y") |> filter(role=="S")|> select(expt, game, trial, text, description) |> View()
 ```

### 2025-10-31 write-ups/manuscript/kid-tangrams-ms.Rmd
figs, intro, cites

@@
   | Preschoolers can form conventional pacts with each other 
-  | to communicate about novel referents
-
+  | to communicate about novel referents (TODO?)
+shorttitle: Preschoolers can form conventional pacts (TODO?)
 author: 
@@
 Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents.
-Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game.
+Some early experiments are interpreted as evidence that 4-5-year-old children cannot generate the mutually understandable names needed to succeed in an iterated reference game.
 Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm.
@@
 
-TODO all figures could be smaller!
-
 # Introduction
@@
 Iterated reference games provide an important paradigm for studying referential communication.
-In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020b; @boyce2024].
+In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020dynamics; @boyce2024].
 Over repetition, features of the initial descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image.
@@
 
-TODO update diagram with expt 3 stim.
-
 ```{r interface, fig.env = "figure", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure.
 Panel A shows the experimental setup.
-Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only.
-Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly.
-Panel D shows the procedure for Experiment 2.
-\\label{game}", cache=FALSE}
+Panel B shows the procedures for each experiment;  within critical blocks targets were ordered randomly.\\label{game}", cache=FALSE}
 
@@
 # edit diagram at 
-# https://docs.google.com/drawings/d/1gio1CNSedECJGrkrNDUi_AydPZcYzESfl4WhZD2l9Bc/edit?usp=sharing
+# https://docs.google.com/drawings/d/1kLfIGdPGeEAQgXG--FE7eyLKzX0DjTZxwfQOIhvW19I/edit?usp=sharing
 ```
@@
 More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication.
-Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication.
+Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2013; @matthews2010].
+When a change in visual context renders previous descriptions overinformative, 6-year-olds (but not 4-year-olds) are more likely to use consistent referential expressions when they keep talking to the same partner, but switch to simpler terms with a new partner [@koymen2014] These studies suggest that young children are sensitive to the norm of consistent descriptions in cooperative communication.
+
 Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009].
 By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015].
-Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.
-* [@sanjuan2025] When 4-5 year old children learn that some objects are only visible to them and not their partner, they are faster to rule out these competitors and select a target than when they believe the interlocuter is aware of the competitors as well.
-* [@nilsen2009] take into account partners perspective for both speaker and listening (4-5 yos); 3-4 year olds also on listening only; relate this to inhibitory control.
+4-5 year old children are faster to rule out competitors and select the correct target when they have learned that their partner is not aware of the competitor object.
+Children's ability to take partner's perspectives and knowledge states into account is evolving over the preschool years, but overall, by late preschool, children can reason about others’ perspectives to communicate more effectively.
 
@@
 A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e.
-when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].
+when there are fewer possible referents [@abbot-smith2015; @sanjuan2015].
 These findings paired with evidence of emerging communicative skills in young children suggest that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.
@@
 In one study, 8-10-year-olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance and highly variable between dyads [@branigan2016].
-4-6-year-olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019].
-4-8-year-old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024].
-In this task, even 4-year-olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024].
-Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.
+In an iterated reference game where 4- and 6- year-olds communicated to each other via gestures, children were successful at communicating a repertoire of 5 image options with each other [@bohn2019].
+Children's comprehension increased over repetitions, and when children switched roles, they often used their partner's gestures back at each other.
+The 6-year-olds showed greater within dyad similarity in gestures than the 4-year-olds, indicating that children's abilities to form conventional pacts is increasing in this age range.
 
-- Bohn, Kachel, Tomasello (2019) Young Children spontaneously recreate core properties of language in a new modality
-
-    - 4, 6, (and 8 in a separate thing) communicating target image via video link (no audio after practice trials)
-
-    - comprehension good and increasing, gestures used back by second partner when roles switch
-
-    - (out of 5 pictures, including “blank”)
-
-    - do see more similaritiy w/i dyads especially for 6 versus 4 yos
-
-May want to set up for change over age?
-And a little more on why 4-5 is a good age -- it's a time of rapid development, we think from prior work that children can comprehend the task.
-Just before big changes in linguistic input from school + reading.
-...
+In an iterated reference game using natural language to communicate, 4-8-year-old children succeeded when playing with a parent using a simple, child-friendly tablet-based task [@leung2024].
+In this task, even 4-year-olds had an initial accuracy above 80% on two-way choice, which rose to above 90% in later repetitions [@leung2024].
+Together, the recent work on both iterated and non-iterated reference games with young children provides evidence of young children’s ability to communicate with each other and with adults under the right conditions.
+The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 [TODO maybe more citations, or re-citations].
 
@@
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across three studies including a total of ZZ pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
+Across three studies including a total of ZZ pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B).
-We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.
+We coded the matching game using Empirica [@almaatouq2021], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.
 
@@
 
-```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time.
+```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time.
 Error bars are bootstrapped 95% CIs with a linear trend line overlaid.
@@
 
-```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of description produced by the teller each trial.
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial.
 Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs.
@@
   theme(
-    legend.position = "none",
+    legend.position = "right",
     panel.grid = element_blank(),
@@
 
-In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024].
+In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020dynamics; @boyce2024].
 We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words, and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}).
@@
 
-```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources.
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources.
 Dots are means and lines are bootstrapped 95% CIs.
@@
 
-TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
-
 ```{=latex}
 \begin{table}
-	\caption{Example descriptions children successfully used to identify different target images in Experiment 2.}
+	\caption{Example descriptions children successfully used to identify different target images in Experiment 2. TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
+}
 	\label{example}
@@
 
-```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.
+```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.
 Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
@@
 We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners.
-As children of this age are rapidly developing their linguistic skills (CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.
+As children of this age are rapidly developing their linguistic skills (TODO CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.
 
-```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair.
+```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair.
 Each dot is a game, and linear fits are overlayed.
@@
   mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
+  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
+  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
+  coord_cartesian(ylim = c(0.3,1))+
+  geom_hline(yintercept = .5) +
   scale_color_manual(values=c("#DC322FFF","#2AA198FF" ,"#268BD2FF"  ))+
@@
 
-```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender.
+```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender.
 Each dot is a game, and linear fits are overlayed.
@@
 In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions.
-Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024].
+Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020dynamics; @boyce2024].
 However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification.

### 2025-10-31 write-ups/manuscript/kid-tangrams-ms.Rmd
intro and discussion

@@
 authornote: |
-  TODO credit taxonomy etc
+  TODO credit taxonomy, acknowledgements, etc
 
@@
 These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
-  <!-- https://tinyurl.com/ybremelq -->
 
@@
 ```
-
-TODO review the literature for 
-* any other studies we should include
-* any studies we want to say more about than we had room for in cogsci
 
@@
 TODO this description metric is just the length of the description, taking out repetition/rephrasing from the child.
-We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc)
+We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc), which is also very flat.
 
-TOOD maybe move this paragraph Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
-TODO add examples from expt 3 as well??
+[TODO maybe move this paragraph] Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
 
@@
 However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.
+
 TODO consider dropping this because it's not that strong an analysis?
@@
 ### Participants
-Experiment 3 was run between XX and YY , at the same preschool as Experiments 1 and 2.
+Experiment 3 was run between XX and YY, at the same preschool as Experiments 1 and 2.
 Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and ZZ participants in experiment 2 were also in experiment 3.
-TODO pairs of children completed all 16 critical trials, and TODO.
+TODO pairs of children completed all 16 critical trials, and TODO others completed ...
+.
 Our target age range was 4 and 5-year-olds.
@@
 ### Data processing
-Data were processed in the same way as Experiment 1. After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had TODO trials total.
+Data were processed in the same way as Experiment 1.
+After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r joint_data |> filter(expt=="3") |> nrow()` trials total.
 
@@
 
-TODO reverify everything after data review + full sample
+TODO reverify everything after full sample
 
@@
 
-We repeat the semantic analyses from Experiment 2.
-To start with, we again found that utterances were more similar if they came from the same partnership (`r sim_across_between_3_results |> stats_text(2,3)`) and if they came from the same person (`r sim_across_between_3_results |> stats_text(3,3)`).
+We repeated the semantic analyses from Experiment 2.
+We again found that utterances were more similar if they came from the same partnership (`r sim_across_between_3_results |> stats_text(2,3)`) and if they came from the same person (`r sim_across_between_3_results |> stats_text(3,3)`).
 
 Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_3_results |> stats_text(2,3)`).
-Over time, the semantic distance between adjacent block utterances increased slightly (`r sim_to_next_2_results |> stats_text(2,3)`).
+Over time, the semantic similarity between adjacent block utterances increased slightly (`r sim_to_next_2_results |> stats_text(2,3)`).
 
@@
 
-As the two experiments were similar to one another, we re-ran models pooling the data across the three experiments, using experiment number as a random effect.
-Pooling the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+As the experiments were all similar to one another, we re-ran models pooling the data across the three experiments, with experiment number as a random effect.
+Pooling across the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
 
-TODO decide whether to discuss speed here again.
+TODO decide whether to include speed here again.
 
@@
 
-For other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks.
+For the other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks.
 In the combined model, we found that the similarity the last block utterance increased across blocks (`r stats_text(sim_to_last_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_last_mod_meta, 3,3)`).
@@
 
-While some of these effects were small and thus not all effects were robustness detectable in each experiment by itself, the overall picture is consistent in direction with the patterns of convergence and divergence reported in reference games between adults [@boyce2024].
+While some of these effects were small and thus not all effects were robustly detectable in every experiment, the overall pattern of results is consistent in direction with the patterns of convergence and divergence reported in reference games between adults [@boyce2024].
 
@@
 
-We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners.
-As children of this age are rapidly developing their linguistic skills (TODO CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.
+Children in the late preschooler years are undergoing a time of rapidly developing linguistic and cognitive skills, so children may show age-based differences in how well they complete tasks that use language or executive functioning [citation needed].
+Working memory and other executive capacities are developing across childhood [@diamond2013].
+Additionally, there may be gender differences at this age.
+Young girls have on average an advantage in early linguistic skills, showing faster earlier vocabulary growth that boys [@marchman; @rinaldi2023], and this linguistic advantage continues until ....
+TODO
+
+We did not design these experiments to test for the affect of children's age or gender.
+Children were not matched with their partner based on age and gender, so different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners.
+We consider the effects of participant age and gender as an exploratory measure.
 
@@
 The effects of children's age and gender on their performance are shown in Figure \ref{ageacc}.
-We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child^[`r form(acc_demog_form)`].
+We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child on the accuracy of their selections^[`r form(acc_demog_form)`].
 Accuracy was higher for older children (OR for one month older speaker: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older listener: `r stats_text(acc_demog_meta, 2,2)`).
@@
 
-For description length, we tested whether children's age or gender effected how many words they produced in descriptions when in the teller role (Figure \ref{agewords}).
-We also ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors^[`r form(description_demog_form)`].
+TODO be consistent about teller/guesser versus speaker/listener
+
+We also tested whether children's age or gender affected how many words they produced in their descriptions (Figure \ref{agewords}).
+We ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors of description length^[`r form(description_demog_form)`].
 Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of listener (`r stats_text(description_demog_meta, 2,2)`).
@@
 It is curious that listener gender but not speaker gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic.
-For accuracy, the age and gender effects could be reflective of any of a number of developing capacities, including linguistic, executive functioning, or social/theory of mind.
-These findings of a slight effect of age and a slight female-advantage are consistent with a wider literature of TODO
+At one end, longer descriptions may provide more information than short non-specific labels such as "person" or "diamond".
+On the other, succinct descriptions that reflect distinguishing features ("mountain", "man with backpack") are better than long descriptions about the component shapes ("there's a triangle and then a square and another square...").
+
+For accuracy, the age and gender effects could be reflective of any of a number of developing capacities, including linguistic sophistication, improving executive functioning and impulse control, and better perspective taking and theory of mind.
 
@@
 
-# (unclear if include)
-could also do some qualitative comparison of the *types* of descriptions that kiddos give versus kid-parent vs adult ?
-
 # General discussion
@@
 Additionally, children's utterances showed signs of converging toward conceptual pacts.
-While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+While this task was substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
 
@@
 We sampled a convenience population of children at a university nursery school.
-The set of tangram images may be easier to distinguish and have higher codability than other target images used in adult reference games.
+We used two sets of 4 tangram images, which may be easier to distinguish and have higher codability than other target images used in adult reference games.
 We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner.
@@
 Even within this population, we observed some substantial pair to pair differences.
-Pairs varied in what types of descriptions they have, how readily the gave them, as well as how well they understood the "rules" of the game and how much they cooperated with the rules.
-While we did not aim to test 3 year olds, the few 3 year olds who we tested accidentally did fine.
+Pairs varied in what types of descriptions used, how readily they produced them, as well as how well they understood the "rules" of the game and how much they cooperated with the rules.
 
 Our qualitative impression is that it was more common for children to hit limits around understanding the game structure (ex.
-that the "teller" doesn't press the ipad),  cooperating with the game (ex.
-not trying to show the ipad to their partner, not pressing buttons before their partner has a change to describe), and understanding that this wasn't a test about the "name" of the shape, moreso than being truely able to describe an unknown referent adequately.
-Pairs of children did very more in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.
+that the "teller" doesn't press the iPad),  cooperating with the game (ex.
+not turning their iPad to show their partner, not pressing buttons before their partner has a chance to describe), and understanding that this wasn't a test about the "name" of the shape, than for children to specifically struggle with how to describe an unknown referent adequately.
+Pairs of children did vary more in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.
 Generally, it's hard to push at the lower limits of some of children's abilities because many skills are changing over the same time frame and variables that may not be of primary interest, such as patience and instruction-following, can influence the outcomes.
+[TODO do we want to explicitly comment that we aren't making claims about what the lower limit is]
 

### 2025-11-14 code/ms_analysis.Rmd
examples for expt 3, latex adventures

@@
   theme(legend.position = "none")
+```
+# Successful descriptions
+
+```{r}
+transcript_1 <- read_csv(here(data_loc_1, "transcripts.csv")) |>
+  left_join(read_csv(here(data_loc_1, "link_transcripts.csv"))) |>
+  mutate(expt = "1") |> mutate(trial=trial-2)
+
+transcript_2 <- read_csv(here(data_loc_2, "transcripts.csv")) |>
+  left_join(read_csv(here(data_loc_2, "link_transcripts.csv"))) |>
+  mutate(expt = "2") |> mutate(trial=trial-4)
+
+transcript_3 <- read_csv(here(data_loc_3, "transcripts.csv")) |>
+  left_join(read_tsv(here(data_loc_3, "link_transcripts.csv"))) |>
+  mutate(expt = "3") |> mutate(trial=trial-4)
+
+transcripts <- transcript_1 |>
+  bind_rows(transcript_2) |>
+  bind_rows(transcript_3) |> filter(role=="S") |> filter(!is.na(description)) |> 
+  group_by(trial, speaker, expt, game) |> summarize(description=str_c(description, sep=", ", collapse=", "))
+
+joint_demog |> filter(correct==1) |> left_join(transcripts) |> select(target, expt, game, speaker, trial, description) |> View()
 ```

### 2025-11-14 write-ups/manuscript/kid-tangrams-ms.Rmd
examples for expt 3, latex adventures

@@
 
+```{r}
+# for use in abstract 
+pairs <- joint_data |> distinct(game, expt) |> nrow()
+# eventually expect 81
+
+pct <- joint_data |> summarize(m=round(mean(correct)*100)) |> pluck()
+```
+
 # Introduction
@@
 Together, the recent work on both iterated and non-iterated reference games with young children provides evidence of young children’s ability to communicate with each other and with adults under the right conditions.
-The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 [TODO maybe more citations, or re-citations].
+The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 years old.
 
@@
 In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm.
-Across three studies including a total of ZZ pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
+Across three studies including a total of `r pairs` pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.
 
@@
 Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.
-An additional XX games were excluded for being pilot runs or not completing enough trials.
+An additional 8 games were excluded for being pilot runs or not completing enough trials.
 
@@
 
-```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial.
+```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial, shown on a log scale.
 Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs.
@@
   geom_smooth(method = "lm", color = "black") +
+  scale_y_log10()+
   xlab("Trial") +
   ylab("Words") +
+  labs(color="Block")+
   facet_grid(str_c("Expt ",expt) ~ .) +
@@
 
-TODO this description metric is just the length of the description, taking out repetition/rephrasing from the child.
+<!-- this description metric is just the length of the description, taking out repetition/rephrasing from the child.
 We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc), which is also very flat.
+-->
 
-[TODO maybe move this paragraph] Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).
+Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example}.
 
@@
 
-```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources.
+```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="101%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources.
 Dots are means and lines are bootstrapped 95% CIs.
@@
 
-TODO consider dropping this because it's not that strong an analysis?
-
 ## Discussion
@@
 ```{=latex}
+
 \begin{table}
-	\caption{Example descriptions children successfully used to identify different target images in Experiment 2. TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
+	\caption{Example descriptions children successfully used to identify different target images in Experiments 2 and 3.
 }
 	\label{example}
-	\begin{tabular}[t]{p{16em}p{4em}}
+	\begin{tabular}[t]{p{14em}p{3em}}
 		\hline
 		%hold
-		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\
 
+		$\bullet$ vampire & \multirow{4}{*}{\includegraphics[width=.5in]{E1.jpg}}\\
+		$\bullet$ a person flying &\\
+		$\bullet$ a person &\\
+		$\bullet$ a kite &\\
+		\multicolumn{2}{l}{$\bullet$ a triangle with a head on it with feet}\\
+				\multicolumn{2}{p{17em}}{$\bullet$  somebody skydiving, not in the airplane}\\
+		\hline
+		$\bullet$ airplane & \multirow{4}{*}{\includegraphics[width=.5in]{D1.jpg}}\\
+		$\bullet$ alligator &\\
+		$\bullet$ a person fell down &\\
+		$\bullet$ a boat &\\
+		\multicolumn{2}{p{17em}}{$\bullet$ a person that's in a race car that has one triangle and two triangles}\\
+		\hline
+			$\bullet$ person & 		\multirow{4}{*}{\includegraphics[width=.5in]{I1.jpg}}\\
 		$\bullet$ a person holding a sandwich &\\
@@
 		$\bullet$ a monster &\\
-		$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon &\\
+		\multicolumn{2}{p{17em}}{$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon}\\
 		\hline
-		$\bullet$ vampire & \multirow{7}{*}{\includegraphics[width=.5in]{E1.jpg}}\\
-		$\bullet$ hopping & \\
-		$\bullet$ a person flying &\\
-		$\bullet$ a person &\\
-		$\bullet$ a kite &\\
-		$\bullet$ a triangle with a head on it with feet &\\
-				\multicolumn{2}{l}{$\bullet$  somebody skydiving, not in the airplane}\\
-		\hline
-		$\bullet$ racecar & \multirow{7}{*}{\includegraphics[width=.5in]{D1.jpg}}\\
-		$\bullet$ airplane &\\
-		$\bullet$ alligator &\\
-		$\bullet$ a person fell down &\\
-		$\bullet$ a boat &\\
-		$\bullet$ a person that's in a race car that has one triangle and two triangles &\\
-		\hline
-
-		$\bullet$ person & \multirow{4}{*}{\includegraphics[width=.5in]{B1.jpg}}\\
+		$\bullet$ person & \multirow{3}{*}{\includegraphics[width=.5in]{B1.jpg}}\\
 		$\bullet$ a person walking &\\
 		$\bullet$ a person looking down &\\
-		$\bullet$ a people, but it doesn't have any arms& \\
+		\multicolumn{2}{p{17em}}{$\bullet$ a people, but it doesn't have any arms} \\
+		\hline
+	\end{tabular}
+	\quad
+	\begin{tabular}[t]{p{14em}p{3em}}
+		\hline
+		$\bullet$ a people & 		\multirow{4}{*}{\includegraphics[width=.5in]{tangram_A.png}}\\
+	$\bullet$ a horse & \\
+	$\bullet$ stomping & 	\\
+		\multicolumn{2}{p{17em}}{$\bullet$ a person that’s walking and has a square, like on his shoulder}\\
+	\multicolumn{2}{p{17em}}{$\bullet$ a diamond on the back, this leg is one leg on the ground, one leg up in the air}\\
+		\hline
+		$\bullet$ a bunny & \multirow{4}{*}{\includegraphics[width=.5in]{tangram_E.png}}\\
+$\bullet$ a kangaroo &\\
+$\bullet$ a mountain with some triangles &\\
+$\bullet$ two triangles on the top &\\
+\multicolumn{2}{p{17em}}{$\bullet$ a person looking backwards with a dress and a bow}\\
+		\hline
+		$\bullet$ a shoe &  \multirow{3}{*}{\includegraphics[width=.5in]{tangram_G.png}}\\
+		$\bullet$  a walrus &\\
+$\bullet$ something looking down &\\
+\multicolumn{2}{p{17em}}{$\bullet$ a pterodactyl with a flag on its back }\\
+
+\multicolumn{2}{p{17em}}{$\bullet$  a diamond on the top, and a triangle on the bottom}\\
+		\hline
+		$\bullet$ a bottle & \multirow{4}{*}{\includegraphics[width=.5in]{tangram_H.png}}\\
+		$\bullet$ a diamond on the top &\\
+ $\bullet$ a weird looking banana &\\
+$\bullet$ candy &\\
+
 		\hline
@@
 Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.
-An additional XX games were excluded for being pilot runs or not completing enough trials.
+An additional 6 games were excluded for being pilot runs or not completing enough trials.
 
@@
 We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`).
-Descriptions were marginally more similar if they came from the same child, which is expected as a sign of internal consistency TODO rephrase (`r sim_to_last_2_results |> stats_text(3,3)`).
+Descriptions were marginally more similar if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`).
 Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).
@@
 ```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.
-Heavy dots are means with bootstrapped 95% CIs; light dots are individual values.
+Heavy dots are means with bootstrapped 95% CIs.
 \\label{tolast}" }
@@
 ## Methods
-As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2. Experiment 3 was pre-registered at TODO [anonymized link].
+As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2.
+Experiment 3 was pre-registered at [anonymized link] (https://osf.io/tk7sc/overview?view_only=874df1897c2e45e187aa71944bbc2392).
 
@@
 ### Participants
-Experiment 3 was run between XX and YY, at the same preschool as Experiments 1 and 2.
-Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and ZZ participants in experiment 2 were also in experiment 3.
+Experiment 3 was run between  and YY, at the same preschool as Experiments 1 and 2.
+Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and TODO participants in experiment 2 were also in experiment 3.
 TODO pairs of children completed all 16 critical trials, and TODO others completed ...
@@
 Our target age range was 4 and 5-year-olds.
-Of the TODO children, QQ were girls, and the children had a median age of ZZ months, and a range of XX-YY months.
-An additional XX games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.
+Of the TODO children, TODO were girls, and the children had a median age of TODO months, and a range of TODO-TODO months.
+An additional TODO games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.
 
 ### Materials
-A different set of critical images were used (see Figure TODO).
-These were chosen from a set of commonly used tangram images (CITATIONS), to be fairly distinct from each other.
-We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having convention names (TODO figure).
+A different set of critical images were used (see Figure \ref{game}).
+These were chosen from a set of commonly used tangram images [@clark1986;@hawkins2020dynamics;@boyce2024], to be fairly distinct from each other.
+We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having established, canonical names.
 
@@
 The procedure was the same as for Experiment 2.
-The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop.
+<!--The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop.
+-->
 
@@
 
+speed_mod_meta <- read_rds(here(mod_results, "speed_meta.rds"))
 description_mod_meta <- read_rds(here(mod_results, "description_meta.rds"))
@@
 As the experiments were all similar to one another, we re-ran models pooling the data across the three experiments, with experiment number as a random effect.
+
+### Accuracy and speed 
 Pooling across the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).
+Across all the experiments, children sped up over repetitions (effect of one trial later: `r stats_text(speed_mod_meta, 2)` seconds).
 
-TODO decide whether to include speed here again.
-
+### Description length
 Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
@@
 
+### Convergence
 For the other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks.
@@
 
-We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds.
-Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt.
+If children are sensitive to their partner's perspective, then whether a description was successful should influence whether the same description, or a variant, will be used in future rounds.
 To test whether accuracy is predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller.
@@
 
-Children in the late preschooler years are undergoing a time of rapidly developing linguistic and cognitive skills, so children may show age-based differences in how well they complete tasks that use language or executive functioning [citation needed].
+Children in the late preschooler years are undergoing a time of rapidly developing linguistic and cognitive skills, so children may show age-based differences in how well they complete tasks that use language or executive functioning [TODO citation needed].
 Working memory and other executive capacities are developing across childhood [@diamond2013].
 Additionally, there may be gender differences at this age.
-Young girls have on average an advantage in early linguistic skills, showing faster earlier vocabulary growth that boys [@marchman; @rinaldi2023], and this linguistic advantage continues until ....
-TODO
+Young girls have on average an advantage in early linguistic skills, showing faster earlier vocabulary growth that boys [@marchman; @rinaldi2023].
+TODO needs work
 
@@
 
-```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender.
-Each dot is a game, and linear fits are overlayed.
-\\label{agewords}" }
-joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
-  geom_smooth(method = "lm", show.legend=F)+
-  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
-  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
-  labs(x="Teller Age (months)", y="Words", color="Teller gender")
-```
-
 ```{r}
@@
 
+### Accuracy 
 The effects of children's age and gender on their performance are shown in Figure \ref{ageacc}.
 We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child on the accuracy of their selections^[`r form(acc_demog_form)`].
-Accuracy was higher for older children (OR for one month older speaker: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older listener: `r stats_text(acc_demog_meta, 2,2)`).
-Accuracy was also higher when the listener was a girl (OR: `r stats_text(acc_demog_meta, 4,2)`), but was not credibly higher when the speaker was a girl (OR: `r stats_text(acc_demog_meta, 5,2)`).
+Accuracy was higher for older children (OR for one month older teller: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older guesser: `r stats_text(acc_demog_meta, 2,2)`).
+Accuracy was also higher when the guesser was a girl (OR: `r stats_text(acc_demog_meta, 4,2)`), but was not credibly higher when the teller was a girl (OR: `r stats_text(acc_demog_meta, 5,2)`).
+The observed age and gender effects could be reflective of any of a number of developing capacities, including linguistic sophistication, improving executive functioning and impulse control, and better perspective taking and theory of mind.
 
-TODO be consistent about teller/guesser versus speaker/listener
-
-We also tested whether children's age or gender affected how many words they produced in their descriptions (Figure \ref{agewords}).
+### Description length
+We also tested whether children's age or gender affected how many words they produced in their descriptions.
 We ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors of description length^[`r form(description_demog_form)`].
-Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of listener (`r stats_text(description_demog_meta, 2,2)`).
+Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of guesser (`r stats_text(description_demog_meta, 2,2)`).
 For gender, there was no effect of teller gender on description length (girls produced `r stats_text(description_demog_meta, 5,2)` words compared to boys), but tellers produced longer descriptions for boy guessers (`r stats_text(description_demog_meta, 4,2)`).
@@
 
-It is curious that listener gender but not speaker gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic.
-At one end, longer descriptions may provide more information than short non-specific labels such as "person" or "diamond".
-On the other, succinct descriptions that reflect distinguishing features ("mountain", "man with backpack") are better than long descriptions about the component shapes ("there's a triangle and then a square and another square...").
+It is curious that guesser gender but not teller gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic, since short labels could be non-specific like "person", or adequately distinguishing like "mountain".
 
-For accuracy, the age and gender effects could be reflective of any of a number of developing capacities, including linguistic sophistication, improving executive functioning and impulse control, and better perspective taking and theory of mind.
+# General discussion
 
-# Qualitative variation in interaction structure
+Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
+However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
+Recent work has established that children as young as 4 years old are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents [@leung2024].
+Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
+
+Across 3 experiments and `r pairs` pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
+Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (`r pct`%), indicating that children were generally able to produce adequate descriptions.
+Additionally, children's utterances showed signs of converging toward conceptual pacts.
+While this task was substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
+
+Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game.
+Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds.
+TODO do we want to speculate on the accuracy levels?
+
+It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults.
+The low number of options and relatively distinct shapes may have obviated the need for long initial descriptions.
+Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020dynamics; @boyce2024].
+However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification.
+Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
+
+## Limitations
+The generalizability of our results is limited by the target population, the target images, and the task structure.
+We sampled a convenience population of children at a university nursery school.
+We used two sets of 4 tangram images, which may be easier to distinguish and have higher codability than other target images used in adult iterated reference games.
+We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner.
+Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.
+
+## Qualitative observations
+Even within this population, we observed substantial pair to pair differences in what types of descriptions were used, how readily they produced them, and how well children understood and cooperated with the "rules" of the game.
 
@@
 
+Occasionally, a guesser would provide descriptions of the images ("The walking person or the standing person that's holding something?") so the teller just had to indicate which one rather than self-generate a description.
+
 Some children asked clarifying questions of their partner, but knowing how much information to provide was not always consistent.
@@
 
-* maybe cut this section
-* do want to make the point that there's really wide variability (consistent with older kids and also adults for that matter) -- some pairs one kid does the describing including of options and the other chooses
+Other pairs of children relied on the experimenter to say whose turn it was or prompt them with options to make a guess or ask questions, and some children struggled to formulate useful questions (asking their partner just "what do you see?").
+Even for children who struggled, their behaviors such as gesturing, trying to turn their iPads to show their partner, or asking the researcher for help, are indicative of attempts to solve the communication problem (albeit in ways that did not follow the "rules" of the game).
 
-* may want to be clear that structuring the interaction in an effective way is a skill that not all the kiddos have -- some relied more on the experimenter saying whose turn and needed more prompting with "do you want to give more hints" or "do you want to ask a question".
-whereas other groups rapidly produced descriptions and then either clicked or asked for (and got) clarification (either completely on their own, or after the option was questions offered as a possibility).
-that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways that followed the "rules" of the game)
-where we see what looks like failures it may be hard to tell what part isn't working (is it knowing what i know, or what other knows, or cognitive control, or producing questions, or sampling)
+In instances where children did not communicate successfully, it is difficult to know what the limiting factor was.
+In general, our qualitative impression was that children were more likely to hit limits related to understanding the game structure (such as when tellers tapped their own iPad), cooperating with the game (such as guessers immediately pressing an option before the teller had a chance to say anything) or thinking the task was about correctly "naming" the shape (saying "I don't know what that's called"), than for children to specifically struggle with how to describe an unknown referent adequately.
+However, pairs of children also varied significantly in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.
 
-# General discussion
-
-Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966].
-However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low.
-Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents.
-Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.
-
-Across 3 experiments and TODO pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape.
-Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (TODO%), indicating that children were generally able to produce adequate descriptions.
-Additionally, children's utterances showed signs of converging toward conceptual pacts.
-While this task was substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
-
-Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game.
-Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds.
-These null findings are likely a result of initially high accuracy in the first block and initially short utterances that leave little room for reduction.
-
-It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults.
-In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions.
-Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020dynamics; @boyce2024].
-However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification.
-Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
-
-The generalizability of our results is limited by the target population, the target images, and the task structure.
-We sampled a convenience population of children at a university nursery school.
-We used two sets of 4 tangram images, which may be easier to distinguish and have higher codability than other target images used in adult reference games.
-We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner.
-Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.
-
-Even within this population, we observed some substantial pair to pair differences.
-Pairs varied in what types of descriptions used, how readily they produced them, as well as how well they understood the "rules" of the game and how much they cooperated with the rules.
-
-Our qualitative impression is that it was more common for children to hit limits around understanding the game structure (ex.
-that the "teller" doesn't press the iPad),  cooperating with the game (ex.
-not turning their iPad to show their partner, not pressing buttons before their partner has a chance to describe), and understanding that this wasn't a test about the "name" of the shape, than for children to specifically struggle with how to describe an unknown referent adequately.
-Pairs of children did vary more in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.
-Generally, it's hard to push at the lower limits of some of children's abilities because many skills are changing over the same time frame and variables that may not be of primary interest, such as patience and instruction-following, can influence the outcomes.
-[TODO do we want to explicitly comment that we aren't making claims about what the lower limit is]
+By reducing the task demands and scaffolding children's interaction with each other, we were able to see that many 4-5 year old children in our sample were able to coordinate with each other on informative utterances in a repeated matching game.
+In a post-hoc analysis, we also observed that within this age range, older children are somewhat more successful than younger children, although there is wide individual variability.
+We do not claim that this ability emerges at 4 years old, although we suspect that younger children may struggle more even with the reduced task demands, making it difficult to test child-child coordination in young children.
 

### 2025-11-26 code/expt_1/data_prep.Rmd
qc on transcription

--- 2025-10-28 code/expt_1/data_prep.Rmd
@@
   else{
-    if(file.exists(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))){
-        #file = read_csv(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))
-         file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+    if(file.exists(here("data/expt_1/raw_transcripts/exclude/", str_c(game_num, ".csv")))){
+        file = read_csv(here("data/expt_1/raw_transcripts/exclude/", str_c(game_num, ".csv")))
         }
     else{
-      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA, on_topic=NA)
     }

### 2025-11-26 code/expt_3/data_prep_3.Rmd
qc on transcription

--- 2025-10-28 code/expt_3/data_prep_3.Rmd
@@
   file = read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))
+  message(game_num)
   }
@@
     if(file.exists(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))){
+      message(game_num)
         file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
@@
     else{
-      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
+      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA,description=NA, on_topic=NA)
     }

### 2025-11-26 code/ms_analysis.Rmd
qc on transcription

@@
   filter(is.na(role)) |>
-  select(expt, game, trial, text) |> View() # to check on later!!!
+  select(expt, game, trial, text) # to check on later!!!
 
